### 2015-09-23 notes/maxmargin_formulation.tex
notes: update with the dropbox version.

%\documentclass{ecai2012}
\documentclass[a4paper,10pt]{article}
\usepackage{amsmath,amssymb,theorem,enumerate}
%\usepackage{latexsym}
%\usepackage{commath}
%\usepackage{mathcomp}
%\usepackage{amssymb}
\usepackage{subfigure}
\usepackage{epsfig}

\newcommand{\veps}{\varepsilon}
\newcommand{\vphi}{\varphi}

\newcommand{\calX}{\mathcal{X}}
\newcommand{\calD}{\mathcal{D}}

\newcommand{\bfa}{\mathbf{a}}
\newcommand{\bfA}{\mathbf{A}}
\newcommand{\bfb}{\mathbf{b}}
\newcommand{\bfB}{\mathbf{B}}
\newcommand{\bfx}{\mathbf{x}}
\newcommand{\bfX}{\mathbf{X}}
\newcommand{\bfy}{\mathbf{y}}
\newcommand{\bfY}{\mathbf{Y}}
\newcommand{\bfz}{\mathbf{z}}
\newcommand{\bfZ}{\mathbf{Z}}
\newcommand{\bfw}{\mathbf{w}}
\newcommand{\bfV}{\mathbf{V}}

\newcommand{\Feas}{\mathit{Feasible}}
\newcommand{\tuple}[1]{\langle #1 \rangle}
\newcommand{\denselist}{\itemsep 0pt\partopsep 0pt}
\newcommand{\mynote}[1]{\begin{center}\fbox{\parbox{4.7in}{#1}}\end{center}}
\newcommand{\mynoteSmaller}[1]{\begin{center}\fbox{\parbox{4.3in}{#1}}\end{center}}
\newcommand{\mynotetwocol}[1]{\begin{center}\fbox{\parbox{2.3in}{#1}}\end{center}}

%\title{A MILP Formulation for Setwise Max-Margin Learning}
%\author{Paolo Viappiani and friends}
\begin{document}
%\maketitle

\section*{Setwise Max-Margin Formulation}

{\em Preliminary assumptions:}

\begin{itemize}
 \item A multi attribute feature space $\bfX$; $n$ is the number of features. A configuration is written as $(x_1,\ldots,x_n)$.
  \item We suppose binary features: $x_i \in \{ 0, 1 \}$. 
 \item We assume possible options are specified by linear configuration constraints; as for example $x_1 + x_2 \geq 1$ to express an OR constraint.\\
 We will simply write $\bfx \in \Feas(\bfX)$ in the optimizations below.

 \item Vector of weights $\bfw = ( w_1, ..., w_n )$; the utility of an option $\bfx$ is then $\bfw \cdot \bfx = \sum_{z=1}^{n} w_z x_z$.
\end{itemize}
\vspace{-0.5cm} \mynote{Note: The weight vector $\bfw$ is unknown to the system
but we are given preference constraints (that encode a ``feasible region'').}
\vspace{0.2cm} \noindent {\em Input:}
\begin{itemize}
 \item The configuration constraints representing $\Feas(\bfX)$
 \item $\calD$ the ``learning set'', consisting of pairwise comparisons between a more preferred product $\bfy_{+}$ and a less preferred product $\bfy_{-}$ . 
 Each preference induces a constraint of the type $\bfw \cdot (\bfy_{+} - \bfy_{-}) \geq 0$.
\item Possibly other additional  constraints (representing some kind of ``prior'' knowledge, as for
example $w_2 > 0.3$)\\
We'll write $w^{\bot}_i \leq w_i \leq w^{\top}_i$ to represent such constraints.
\end{itemize}

\vspace{0.2cm} \noindent {\em Output:}
\begin{itemize}
 \item We want to find a {\em set} of vectors $\bfw^1, ..., \bfw^k$ (with $k$ given) of utility weights and associated configurations $\{ \bfx^1, ..., \bfx^k \}$ such that  
 \begin{itemize}
 \item All preferences are satisfied and does that by the largest margin; so for that a binary preference $\bfy_{+} \succeq \bfy_{-}$ induces the constraint $\bfw \cdot (\bfy_{+} - \bfy_{-}) \geq M$.
%  \item  We also use the same margin to enforce each
% vector $\bfw^i$ to be as ``farther away'' from each other as possible.
% This insight is realized by introducing a set of options $\{ \bfx^1, ..., \bfx^k \}$ as decision variables (of the same cardinality $k$).
\item For each $i$, we impose $\bfx^i$ to be the ``best'' option among the $\bfx^1,\ldots,\bfx^k$  when evaluated according to $\bfw^i$; we require these constraints to hold by at least the shared  margin $M$.\\
The imposed constraints are of the type: $\bfw^i \cdot ( \bfx^i -  \bfx^j ) \geq M$ with $i \neq j,\; i,j \in [ 1,k ]$.
 \end{itemize}
\end{itemize}

\vspace{0.2cm} \noindent {\em Additional requirements:}

\begin{itemize}
 \item {\bf Slack variables}: allow for non satisfied preference constraints and include a penalty term $\sum_{h=1}^{|\calD|} \veps_{h}$ in the objective function
 \item {\bf Sparsification}: we introduce a L1 norm term in the objective.
 This is achieved by adding the following term in the objective function
 $$
 \sum_{i=1}^{k} \sum_{z=1}^{n}|w^i_{z}|.
 $$
 \item {\bf Accordance of the $\bfx^i$s with $\bfw^i$s}: we would like to have 
 $$\bfx^i = \arg \max_{\bfx \in \Feas(\bfX)} \bfw^i \cdot \bfx.$$ 
 However since this is not possible (the solver will pick $\bfx^i$ high enough to meet the margin, but not necessarily better), we favor $\bfx^i$ with high utility  by adding a term $\sum_{i=1}^{k} \bfw^i \bfx^i$ in the objective function.
\end{itemize}

% 
% our goal is to find a number of utility weight vectors $\bfw^1, ..., \bfw^k$ (with $k$ given), consistent
% with out knowledge about the user,  but that are representative of different parts of the weights' feasible region.

\paragraph{Initial Non-linear Optimization}
We first provide a formulation with quadratic terms, that is not directly solvable.

{\footnotesize
\begin{align} 
\max_{M, \bfw^i, \bfx^i} & M \!-\!  \alpha \sum_{h} \veps_h - \beta \sum_{i=1}^{k} \sum_{z=1}^{n}|w^i_{z}| + \gamma \sum_{i=1}^{k} \sum_{z=1}^{n} w^i_z x^i_z \\ % \!-\! \alpha_2 \sum_{i,j} \veps^{\prime}_{i,j} & \! & \\
s.t. \;\; & \bfw^i \cdot (\bfy^{h}_{+} - \bfy^{h}_{-} ) \geq M \! - \! \veps_{h} &  \!  &
\forall (\bfy^{h}_{+}, \bfy^{h}_{-}) \! \in \! \calD, \forall i \! \in \! \{ 1, \ldots, k\} \\
 & \bfw^i \cdot ( \bfx^i -  \bfx^j ) \geq M \;\; &  \!  &  \forall j \! \neq \! i; \; i,j \! \in \{  1, \ldots, k \}
 \label{cnst-wx-m} \\
 & w^{\bot}_i \leq w_i \leq w^{\top}_i & & \forall i \in \{ 1, \ldots, k \} \\
 & \bfx^i \in Feasible(\bfX) &  \!  & \forall i \in \{ 1, \ldots, k\} \\
 & \veps_h \geq 0 & & \forall h \in \{ 1,\ldots, |\calD| \} \\
 & M \geq 0 & &
\end{align}}

\noindent {\em Decision variables:}
\begin{itemize}
 \item $\bfw^1,\ldots,\bfw^k$: set of $k$ utility weights; each $\bfw^i$ is a vector of $n$ binary attributes, for example $\bfw^i = (w^i_1,\ldots,w^i_n)$.
 \item $\bfx^1,\ldots,\bfx^k$: set of $k$ options; each  $\bfx^i$ is a vector of $n$ binary attributes, for example $\bfx^i = (x^i_1,\ldots,x^i_n)$.
 \item $M$: the shared margin.
 \item $\veps_1,\ldots,\veps_{|\calD|}$: slack variables representing penalties for violated preferences.
\end{itemize}

\vspace{-0.4cm}
\mynote{
Note that, we are choosing the options $\bfx^1, ..., \bfx^k$ and the weights $\bfw^1, ..., \bfw^k$
simultaneously: 
since we want to maximize $M$, the optimizer will be better off by choosing
a set of outcomes $\bfx^i$ that divide the weight space roughly equally,
and the utility functions such each $\bfw^i$ should lie (intuitively) near the centre of each subregion.}

This initial formulation is problematic, as we have quadratic terms. However, there is a solution:
in fact, by using integer programming tricks, the problem can be formulated as a mixed integer
linear program (MILP).

\paragraph{{\em Setwise} max margin}

We propose a Mixed Integer Linear Programming (MILP) model with {\em non-negative weights}.

{\footnotesize
\begin{align}
\max_{M, \bfw^i, \bfx^i, \bfA^{i,j}} &  \; \; M - \alpha \sum_{h} \veps_h - \beta \sum_{i=1}^{k} \sum_z w^i_z + \gamma \sum_{i=1}^{k} \sum_{z=1}^{n} A^{i,i}_{z}   \span   \\
s.t. \;\; & \bfw^i \cdot (\bfy^{h}_{+} - \bfy^{h}_{-} ) \geq M -\veps_{h}  & 
\forall (\bfy^{h}_{+}, \bfy^{h}_{-}) \in \calD, \; \forall i \in \{1,\ldots,k\}  \\
  & \sum_{z=1}^{n} A^{i,i}_z - A^{i,j}_z \geq M  \;\;\; &  \forall j \neq i , \; i,j \in \{1,\ldots,k\}  \label{cnst-ab-m} \\
  & A^{i,i}_{z} \leq w\!\uparrow x^i_z & \forall i \in \{1,\ldots,k\}, \; z \in \{1,\ldots,m\}  \label{cnst-a-1} \\
  & A^{i,i}_{z} \leq w^i_z & \forall i \in \{1,\ldots,k\}, \; z \in \{1,\ldots,m\} \label{cnst-a-2} \\
  & A^{i,j}_{z} \geq w^i_z - C \cdot (1 - x^j_z) & \forall j \neq i , \; i,j \in \{1,\ldots,k\}, \; z \in [1,m]  \label{cnst-b-1}   \\
  & A^{i,j}_{z} \geq 0 & \forall j \neq i , \; i,j \in \{1,\ldots,k\}, \; z \in \{1,\ldots,m\} \\
& w^{\bot}_i \leq w_i \leq w^{\top}_i & \forall i \! \in \! [1, k]  \\
 & \bfx^i \in Feasible(\bfX) & \forall i \in \{ 1, \ldots, k\} \\
 & x^i_z \in \{0,1\} & \forall i \in \{1,\ldots,k\}; z \in \{1,\ldots,m\}  \\
 & A^{i,j}_z \geq 0 & \forall i, j \in \{1,\ldots,k\}; \forall z \in \{1,\ldots,m\} \label{cnst-b-2} \\
 & \veps_h \geq 0 & \forall h \in \{ 1,\ldots, |\calD| \} \\ 
 & M \geq 0 
\end{align}
}

% {\footnotesize
% \begin{align} % _{M, \bfw^i, \bfx^i, \bfA^i, \bfB^{i,j}}
% \max & \;\; M \!-\! \alpha_1 \sum_{h} \veps_h \!-\! \alpha_2 \sum_{i,j} \veps^{\prime}_{i,j} & & \\
% s.t. & \;\; \bfw^i \cdot (\bfy^{h}_{+} \!-\! \bfy^{h}_{-} ) \geq M \!-\! \veps_{h} & & \forall (\bfy^{h}_{+}, \bfy^{h}_{-}) \in \calD, \forall i \in [1,k] \\
% & \sum_{z} A^{i}_z \!-\! B^{i,j}_z \geq M \!-\! \veps^{\prime}_{i,j} & & \forall j \! \neq i; \; i,j \in [1, k] \label{cnst-ab-m} \\
% & A^{i}_{z} \leq w\!\uparrow x^i_z & & \forall i \in [1, k], z \in [1,m]  \label{cnst-a-1} \\
% & A^{i}_{z} \leq w^i_z & & \forall i \in [1, k], z \in [1,m]  \label{cnst-a-2} \\
% & B^{i,j}_{z} \geq w^i_z - C \!\cdot\! (1 \!-\! x^j_z) & & \forall j \! \neq \! i  \in [1, k], z \in [1,m] \label{cnst-b-1}  \\
% & B^{i,j}_{z} \geq 0 & & \forall j \! \neq \! i \! \in \! [1, k], z \! \in \! [1,m] \label{cnst-b-2} \\
% & w^{\bot}_i \leq w_i \leq w^{\top}_i & & \forall i \! \in \! [1, k]\\
% & \bfx \in \Feas(\bfX) &  \!  &  \forall i \! \in \! [1,k] \nonumber
% \end{align} % \forall j \! \neq \! i , i,\! j \! \in \! [1, k], z 
% }
% 
% {\footnotesize
% \begin{align}
% \max_{M, \bfw^i, \bfx^i, \bfA^{i,j}, \bfB^j, \bfV^i} & \; \; M - \alpha_1 \sum_{h} \veps_h - \alpha_2 \sum_i \sum_z V^i_z + \beta \sum_i \sum_z A^{i}_{z} & \\
% s.t. \;\; & \bfw^i \cdot (\bfy^{h}_{+} - \bfy^{h}_{-} ) \geq M -\veps_{h}  & 
% \forall (\bfy^{h}_{+}, \bfy^{h}_{-}) \in D, \; \forall 1 \leq i \leq k \\
%   & \sum_{z} A^{i}_z - B^{i,j}_z \geq M  \;\;\; &  \forall j \neq i , \; i,j \in [1, k] \\
%   & A^{i}_{z} \leq w\!\uparrow x^i_z & \forall i \in [1, k], \; z \in [1,m]  \\
%   & A^{i}_{z} \leq w^i_z & \forall i \in [1, k], \; z \in [1,m] \\
%   & B^{i,j}_{z} \geq w^i_z - C \cdot (1 - x^j_z) & \forall j \neq i , \; i,j \in [1, k], \; z \in [1,m]  \\
%   & B^{i,j}_{z} \geq 0 & \forall j \neq i , \; i,j \in [1, k], \; z \in [1,m] \\
% & w^{\bot}_i \leq w_i \leq w^{\top}_i & & \forall i \! \in \! [1, k]\\  
%   & V^{i}_{z} \geq w^i_z & \forall i \in [1, k], \; z \in [1,m] \\
%   & V^{i}_{z} \geq -w^i_z & \forall i \in [1, k], \; z \in [1,m] \\ 
% \end{align}
% }

\noindent In the optimization, the decision variables are the following:
\begin{itemize}
\item $M$ is the shared margin.
\item $\bfw^1,...,\bfw^k$ is a set of utility vectors; each vector defined over $m$ attributes (features): $\bfw^i = (w^i_1, ..., w^i_m)$ .
\vspace{-0.4cm} 
\mynoteSmaller{{\em \bf Important:} The weights need to be non-negative in this formulation:\\ $w^{\bot}_z \geq 0 \;\, \forall z$.}
\item $\bfx^1,...,\bfx^k$ is a set of configurations (options) with each configuration $\bfx^i = (x^i_1, ..., x^i_m)$;
each element $x^i_z$ is binary.
\item $\veps$ slack variables to represent cost of unsatisfied constraints in $D$.
	%$\veps^{\prime}$ slack variables to represent cost of not respecting the margin $M$ when choosing the hyperplanes 
\item $\bfA^{i,j}$ encodes the vector $(w^i_1 x^j_1, ..., w^i_m x^j_m)$, the element-by-element product of $ \bfw^i$ and $\bfx^j$; $A^{i,j}_z$ is forced to take value $w^i_z x^i_z$.
%\item $\bfB^{i,j}$ encodes $(w^i_1 x^j_1, ..., w^i_m x^j_m)$, the element-by-element product of $\bfw^i$  and $\bfx^j$
\end{itemize}
\noindent We list all the parameters used, including newly introduced parameters: 
\vspace{-0.2cm}
\begin{itemize}
\item $n$ is the number of attributes.
\item $k$ is the size of the set.
\item $\alpha > 0$ control the tolerance with respect to violated constraints (we can differentiate
the tolerance for the two different types of constraints).
\item $\beta > 0$ is the weight associated to the L1 norm considering the absolute value of the weights (since this formulation assumes non negative weights $w_z^i$, we simply sum over the weights in the objective function).
\item $\gamma > 0$ in order to favour the choice of $\bfx^i$ that achieves high utility with respect to $\bfw^i$.
\item $\calD$ is the {\em learning set}: a set of pairwise comparisons known to the system.
\item Upper and lower bounds (from prior knowledge) on weights: $w^{\bot}_1,\ldots,w^{\bot}_n$ and $w^{\top}_1,\ldots,w^{\top}_n$,
\item $w\!\uparrow$ is any upper bound of the value of the utility weights (for instnace, it can bet set to $\max_z w^{\top}_z$).
\item $C$ is an arbitrary large number. 
\end{itemize}

\mynote{{\em Intuition:} \\
The solver aims at setting $M$, that is a decision variable, as large as possible.
\begin{itemize}
 \item Constraint \ref{cnst-wx-m} of the original program is replaced by constraint \ref{cnst-ab-m},
enforcing $M$ to be smaller than $\sum_{z} A^{i,i}_z \!-\! A^{i,j}_z$ for any $i,j$. \\
$\sum_{z} A^{i,i}_z \!-\! A^{i,j}_z$ is forced to evaluate to $\bfw^i \cdot ( \bfx^i -  \bfx^j )$
by setting some specific additional constraints.
 \item Note that, in order to maximize $M$, the solver will try to keep the $A^{i,i}_z $ as large as possible and the $A^{i,j}_z$ (with $i \neq j$) as small as possible.
 \item The fact that $A^{i,j}_z$ takes value $w^i_z x^j_z$ (and therefore $\bfw^i \cdot \bfx^j = \sum_{z=1}^{n} A^{i,j}_z$) is achieved by setting additional constraints.\\
 We differentiate between the cases 1) $A^{i,i}_z$ and 2) $A^{i,j}_z$ with $i \neq j$ (since they appear with opposite signs in the objective function):
 \begin{itemize}
  \item {\em [case $A^{i,i}_z$]} Constraints \ref{cnst-a-1} and \ref{cnst-a-2} together force each $\bfA^{i,i}$ to be the element-by-element  product of $ \bfw^i$ and $\bfx^i$  ($A^{i,i}_z = w^i_z x^i_z$): 
\begin{itemize}
\item If $x^i_z=0$ then (constraint \ref{cnst-a-1}) also $A^{i,i}_z$ must be $0$; 
\item Otherwise (if  $x^i_z=1$) $A^{i,i}_z \leq w^i_z$ (constraint \ref{cnst-a-2})  but because
of the objective function maximizing $M$, and each $A$ constrain $M$, the solver will set $A^{i,i}_z$ to $w^i_z$.
\end{itemize}
 \item {\em [case $A^{i,j}_z$ with $i \neq j$]} Similarly constraints \ref{cnst-b-1} and \ref{cnst-b-2} together make it so that $\bfA^{i,j}$ encodes $(w^i_1 x^j_1, ..., w^i_m x^j_m)$, the element-by-element product of $\bfw^i$  and $\bfx^j$
(as $C$ is an arbitrary large constant, constraint \ref{cnst-b-1} is binding only when $x_z^j$ is 1,
otherwise it is always satisfied).
 \end{itemize}
\end{itemize}
}

% {
% \tiny
% \bibliographystyle{plain}
% \bibliography{local,standard}
% }

\end{document}

### 2015-09-23 src/main.py
main: be more explicit about the queries we print.

@@ -64,7 +64,6 @@ def query_utility(w, xi, xj, rng, deterministic=False):
     return result
 
 def print_queries(queries, hidden_w):
-    print "queries ="
     for xi, xj, sign in queries:
         relation = {-1:"<", 0:"~", 1:">"}[sign]
         score_xi = np.dot(hidden_w, xi.T)[0]
@@ -138,6 +137,7 @@ def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
 
         if debug:
             print "\n\n\n==== ITERATION {} ====\n".format(t)
+            print "input queries ="
             print_queries(queries, hidden_w)
 
         old_time = time.time()
@@ -166,6 +166,7 @@ def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
         old_best_item = xs[0] if xs.shape[0] == 1 else None
 
         if debug:
+            print "updated queries ="
             print_queries(queries, hidden_w)
 
         # Compute the utility loss between the best item that we would

### 2015-09-23 src/run.sh
src/run: initial import.

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+N=5
+n=10
+
+for set_size in 2 3; do
+    for alpha in 1 10 100 1000; do
+        for beta in 0 1 10 100; do
+            for gamma in 1 10 100 1000; do
+                ./main.py synthetic -N $N -n $n -m $set_size -a $alpha -b $beta -c $gamma -d -s 0 --debug > log_${N}_${n}_${set_size}_${alpha}_${beta}_${gamma} &
+            done
+            wait
+        done
+    done
+done

### 2015-10-06 src/main.py
main: tweak the range of the y axis.
As suggested by Andrea.

@@ -324,7 +324,7 @@ def main():
     ax.set_title("Avg. loss over {} trials".format(args.num_trials))
     ax.set_xlabel("Iterations")
     ax.set_ylabel("Average loss")
-    ax.set_ylim([0.0, 1.0])
+    ax.set_ylim([0.0, 0.4])
     ax.errorbar(np.arange(1, args.num_iterations + 1), means, yerr=stddevs.reshape(-1,1))
     fig.savefig("results_{}_avgloss.svg".format(basename), bbox_inches="tight")
 

### 2015-10-06 src/table-synthetic.tex
experiment: add table-synthetic.tex.

\documentclass{standalone}
\usepackage{graphicx}
\begin{document}

    \begin{tabular}{|c|c|c|}
        {\bf 2 hyperplanes} &
        {\bf 3 hyperplanes} &
        {\bf 4 hyperplanes}
        \\
        \multicolumn{3}{|c|}{{\bf Deterministic, single margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_True_False_False_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Deterministic, multiple margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_True_False_True_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, single margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_False_False_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, multiple margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_False_True_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, single margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_True_False_0_avgloss}.png}
        \\
        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, multiple margin}}
        \\
        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png} &
        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_True_True_0_avgloss}.png}
    \end{tabular}
\end{document}

### 2015-10-06 src/main.py
main: implement alternative ranking_mode.

@@ -86,7 +86,7 @@ def is_onehot(domain_sizes, set_size, xs):
     return True
 
 def update_queries(hidden_w, ws, xs, old_best_item, rng, deterministic=False,
-                   no_indifference=False):
+                   no_indifference=False, ranking_mode="all_pairs"):
     """Computes the queries to ask the user for the given inputs.
 
     If there is only one candidate best item, then only one query is returned,
@@ -103,6 +103,7 @@ def update_queries(hidden_w, ws, xs, old_best_item, rng, deterministic=False,
     :param rng: an RNG object.
     :param deterministic: whether the user answers should be deterministic.
     :param no_indifference: disable 'whatever' answers.
+    :param ranking_mode: either ``"all_pairs"`` or ``"random_list"``.
     :returns: a pair (list of new queries, 
     """
     num_items, num_features = xs.shape
@@ -112,16 +113,51 @@ def update_queries(hidden_w, ws, xs, old_best_item, rng, deterministic=False,
         queries = [query_utility(hidden_w, xs[0], old_best_item, rng,
                                  deterministic=deterministic,
                                  no_indifference=no_indifference)]
-    else:
+        num_queries = 1
+    elif ranking_mode == "all_pairs":
+        # requires 1/2 * n * (n - 1) queries
+        # XXX note that in the non-deterministic setting we may actually lose
+        # information by only querying for ~half the pairs!
         queries = [query_utility(hidden_w, xi, xj, rng,
                                  deterministic=deterministic,
                                  no_indifference=no_indifference)
                    for (i, xi), (j, xj) in it.product(enumerate(xs), enumerate(xs)) if i < j]
+        num_queries = len(queries)
+    elif ranking_mode == "random_list":
+        # requires at most 1 + 2 + ... + (n-1) = 1/2 * (n - 1) * (n - 2)
+        # queries
+        # XXX note that in the non-deterministic setting we may actually lose
+        # information by only querying for a subset of the pairs!
+        permuted_xs = rng.permutation(xs)
+
+        num_queries = 0
+        sorted_xs, indifferent_queries = [permuted_xs[0]], []
+        for x in permuted_xs[1:]:
+            inserted = False
+            for i, other_x in enumerate(sorted_xs):
+                _, _, ans = query_utility(hidden_w, x, other_x, rng)
+                num_queries += 1
+                if ans == 0:
+                    indifferent_queries.append((x, other_x, 0))
+                    inserted = True
+                    break
+                elif ans == 1:
[…]

### 2015-10-06 src/main.py
main: set gurobi as the default solver.

@@ -265,8 +265,8 @@ def main():
                         help="whether the user can (not) be indifferent (default: False)")
     parser.add_argument("-M", "--multimargin", action="store_true",
                         help="whether the example and generated object margins should be independent (default: False)")
-    parser.add_argument("-S", "--solver", type=str, default="optimathsat",
-                        help="solver to use (default: 'optimathsat')")
+    parser.add_argument("-S", "--solver", type=str, default="gurobi",
+                        help="solver to use (default: 'gurobi')")
     parser.add_argument("-s", "--seed", type=int, default=None,
                         help="RNG seed (default: None)")
     parser.add_argument("--debug", action="store_true",

### 2015-10-07 src/main.py
main: replicate the loss by the number of queries asked.

@@ -152,7 +152,7 @@ def update_queries(hidden_w, ws, xs, old_best_item, rng, deterministic=False,
                   indifferent_queries
     else:
         raise ValueError("invalid ranking_mode '{}'".format(ranking_mode))
-    return queries
+    return queries, num_queries
 
 def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
         rng, ranking_mode="all_pairs", deterministic=False,
@@ -230,11 +230,11 @@ def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
             assert (np.abs(scores - debug_scores) < 1e-6).all(), "solver scores and debug scores mismatch:\n{}\n{}".format(scores, debug_scores)
 
         # Ask the user about the retrieved items
-        new_queries = update_queries(hidden_w, ws, xs,
-                                     old_best_item, rng,
-                                     ranking_mode=ranking_mode,
-                                     deterministic=deterministic,
-                                     no_indifference=no_indifference)
+        new_queries, num_queries = \
+            update_queries(hidden_w, ws, xs, old_best_item, rng,
+                           ranking_mode=ranking_mode,
+                           deterministic=deterministic,
+                           no_indifference=no_indifference)
         queries.extend(new_queries)
         old_best_item = xs[0] if xs.shape[0] == 1 else None
 
@@ -253,10 +253,8 @@ def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
         if any(np.linalg.norm(w) == 0 for w in ws):
             print "Warning: null weight vector found in the 1-item case:\n{}".format(ws)
 
-        # XXX replicate the result by the number of new_queries asked
-
         utility_loss = best_hidden_score - np.dot(hidden_w, xs[0])
-        avg_losses.append(utility_loss / np.linalg.norm(hidden_w))
+        avg_losses.extend([utility_loss / np.linalg.norm(hidden_w)] * num_queries)
 
         if debug:
             print "set_size=1 ws =", ws, "hidden_w =", hidden_w

### 2015-10-07 src/main.py
main: replicate the timing data by the number of queries asked.

@@ -215,8 +215,6 @@ def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
         if any(np.linalg.norm(w) == 0 for w in ws):
             print "Warning: null weight vector found in the m-item case:\n{}".format(ws)
 
-        times.append(time.time() - old_time)
-
         if debug:
             print "set_size=n ws =\n", ws
             print "set_size=n xs =\n", xs
@@ -241,6 +239,9 @@ def run(get_dataset, num_iterations, set_size, alphas, utility_sampling_mode,
         if debug:
             print "\nupdated queries ="
             print_queries(queries, hidden_w)
+
+        elapsed = time.time() - old_time
+        times.extend([elapsed / num_queries] * num_queries)
 
         # Compute the utility loss between the best item that we would
         # recommend given the queries collected so far and the best

### 2015-10-07 src/main.py
main: add support for loss/time sequences of varying length.

@@ -325,49 +325,57 @@ def main():
 
     rng = np.random.RandomState(args.seed)
 
-    avg_losses, times = [], []
+    losses, times = [], []
     for i in range(args.num_trials):
-
         print "==== TRIAL {} ====".format(i)
 
-        ls, ts = run(DATASETS[args.dataset], args.num_iterations,
-                     args.set_size, (args.alpha, args.beta, args.gamma),
-                     args.utility_sampling_mode, rng,
-                     ranking_mode=args.ranking_mode,
-                     deterministic=args.deterministic,
-                     no_indifference=args.no_indifference,
-                     multimargin=args.multimargin,
-                     solver_name=args.solver,
-                     debug=args.debug)
+        losses_for_trial, times_for_trial = \
+            run(DATASETS[args.dataset], args.num_iterations,
+                args.set_size, (args.alpha, args.beta, args.gamma),
+                args.utility_sampling_mode, rng,
+                ranking_mode=args.ranking_mode,
+                deterministic=args.deterministic,
+                no_indifference=args.no_indifference,
+                multimargin=args.multimargin,
+                solver_name=args.solver,
+                debug=args.debug)
 
-        avg_losses.extend(ls)
-        times.extend(ts)
+        losses.append(np.array(losses_for_trial).ravel())
+        times.append(np.array(times_for_trial))
 
-    avg_losses, times = np.array(avg_losses), np.array(times)
+    # Since distinct trials may have incurred a different number of queries
+    # each, here we resize the performance data to be uniformly shaped
+    max_queries = max(len(ls) for ls in losses)
+    loss_matrix = np.zeros((args.num_trials, max_queries))
+    time_matrix = np.zeros((args.num_trials, max_queries))
+    for i, (ls, ts) in enumerate(zip(losses, times)):
+        assert ls.shape == ts.shape
+        loss_matrix[i,:ls.shape[0]] = ls
+        time_matrix[i,:ts.shape[0]] = ts
 
-    if args.debug:
-        print "average losses:"
-        print avg_losses
+    np.savetxt("results_{}_time_matrix.txt".format(basename), time_matrix)
+    np.savetxt("results_{}_loss_matrix.txt".format(basename), loss_matrix)
[…]

### 2015-10-07 src/table-synthetic.tex
experiment: plot times per query as well.

@@
     \end{tabular}
+
+    \begin{tabular}{|c|c|c|}
+        {\bf 2 hyperplanes} &
+        {\bf 3 hyperplanes} &
+        {\bf 4 hyperplanes}
+        \\
+        \multicolumn{3}{|c|}{{\bf Deterministic, single margin}}
+        \\
+        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_True_False_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_True_False_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_True_False_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_True_False_False_0_avgtime}.png}
+        \\
+        \multicolumn{3}{|c|}{{\bf Deterministic, multiple margin}}
+        \\
+        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_True_False_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_True_False_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_True_False_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_True_False_True_0_avgtime}.png}
+        \\
+        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, single margin}}
+        \\
+        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_False_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_False_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_False_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_False_False_0_avgtime}.png}
+        \\
+        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, multiple margin}}
+        \\
+        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_False_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_False_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_False_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_False_True_0_avgtime}.png}
+        \\
+        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, single margin}}
+        \\
+        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_True_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_True_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_True_False_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_True_False_0_avgtime}.png}
+        \\
+        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, multiple margin}}
+        \\
+        \includegraphics[width=16em]{{results_synthetic_10_20_1_10.0_0.1_0.1_uniform_False_True_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_2_10.0_0.1_0.1_uniform_False_True_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_3_10.0_0.1_0.1_uniform_False_True_True_0_avgtime}.png} &
+        \includegraphics[width=16em]{{results_synthetic_10_20_4_10.0_0.1_0.1_uniform_False_True_True_0_avgtime}.png}
+    \end{tabular}
+
 \end{document}

### 2015-11-12 paper/ijcai16.tex
paper: add IJCAI-16 template.

%%%% ijcai16.tex

\typeout{IJCAI-16 Instructions for Authors}

% These are the instructions for authors for IJCAI-16.
% They are the same as the ones for IJCAI-11 with superficical wording
%   changes only.

\documentclass{article}
% The file ijcai16.sty is the style file for IJCAI-16 (same as ijcai07.sty).
\usepackage{ijcai16}

% Use the postscript times font!
\usepackage{times}

% the following package is optional:
%\usepackage{latexsym}

% Following comment is from ijcai97-submit.tex:
% The preparation of these files was supported by Schlumberger Palo Alto
% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
% preparation.

% These instructions can be modified and used in other conferences as long
% as credit to the authors and supporting agencies is retained, this notice
% is not changed, and further modification or reuse is not restricted.
% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
% contacts for providing assistance without their prior permission.

% To use for other conferences, change references to files and the
% conference appropriate and use other authors, contacts, publishers, and
% organizations.
% Also change the deadline and address for returning papers and the length and
% page charge instructions.
% Put where the files are available in the appropriate places.

\title{IJCAI--16 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
\author{Subbarao Kambhampati \\ 
Arizona State University, Tempe Arizona  \\
ijcai16programchair@gmail.com}

\begin{document}

\maketitle

\begin{abstract}
  The {\it IJCAI--16 Proceedings} will be printed from electronic
  manuscripts submitted by the authors. The electronic manuscript will
  also be included in the online version of the proceedings. This paper
  provides the style instructions.
\end{abstract}

\section{Introduction}

The {\it IJCAI--16 Proceedings} will be printed from electronic
manuscripts submitted by the authors. These must be PDF ({\em Portable
Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.

\subsection{Length of Papers}

Each accepted full paper is allocated six pages in the conference 
proceedings, excluded references. References can take up to one page.
Up to two additional pages may be purchased at a price 
of \$275 per page for any accepted paper. However, all 
{\em submissions} must 
be a maximum of six pages, plus at most one for references, in length.

\subsection{Word Processing Software}

As detailed below, IJCAI has prepared and made available a set of
\LaTeX{} macros and a Microsoft Word template for use in formatting
your paper. If you are using some other word processing software (such
as WordPerfect, etc.), please follow the format instructions given
below and ensure that your final paper looks as much like this sample
as possible.

\section{Style and Format}

\LaTeX{} and Word style files that implement these instructions
can be retrieved electronically. (See Appendix~\ref{stylefiles} for
instructions on how to obtain these files.)

\subsection{Layout}

Print manuscripts two columns to a page, in the manner in which these
instructions are printed. The exact dimensions for pages are:
\begin{itemize}
\item left and right margins: .75$''$
\item column width: 3.375$''$
\item gap between columns: .25$''$
\item top margin---first page: 1.375$''$
\item top margin---other pages: .75$''$
\item bottom margin: 1.25$''$
\item column height---first page: 6.625$''$
\item column height---other pages: 9$''$
\end{itemize}

All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
A4-size paper, use the given top and left margins, column width,
height, and gap, and modify the bottom and right margins as necessary.

\subsection{Format of Electronic Manuscript}

For the production of the electronic manuscript, you must use Adobe's
{\em Portable Document Format} (PDF). A PDF file can be generated, for
instance, on Unix systems using {\tt ps2pdf} or on Windows systems
using Adobe's Distiller. There is also a website with free software
and conversion services: {\tt http://www.ps2pdf.com/}. For reasons of
uniformity, use of Adobe's {\em Times Roman} font is strongly suggested. In
\LaTeX2e{}, this is accomplished by putting
\begin{quote} 
\mbox{\tt $\backslash$usepackage\{times\}}
\end{quote}
in the preamble.\footnote{You may want also to use the package {\tt
latexsym}, which defines all symbols known from the old \LaTeX{}
version.}

Additionally, it is of utmost importance to specify the American {\bf
letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
formatting the paper. When working with {\tt dvips}, for instance, one
should specify {\tt -t letter}.

\subsection{Title and Author Information}

Center the title on the entire width of the page in a 14-point bold
font. Below it, center the author name(s) in a 12-point bold font, and
then center the address(es) in a 12-point regular font. Credit to a
sponsoring agency can appear on the first page as a footnote.

\subsubsection{Blind Review}

In order to make blind reviewing possible, authors must omit their
names and affiliations when submitting the paper for review. In place
of names and affiliations, provide a list of content areas. When
referring to one's own work, use the third person rather than the
first person. For example, say, ``Previously,
Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
that\ldots'' Try to avoid including any information in the body of the
paper or references that would identify the authors or their
institutions. Such information can be added to the final camera-ready
version for publication.

\subsection{Abstract}

Place the abstract at the beginning of the first column 3$''$ from the
top of the page, unless that does not leave enough room for the title
and author information. Use a slightly smaller width than in the body
of the paper. Head the abstract with ``Abstract'' centered above the
body of the abstract in a 12-point bold font. The body of the abstract
should be in the same font as the body of the paper.

The abstract should be a concise, one-paragraph summary describing the
general thesis and conclusion of your paper. A reader should be able
to learn the purpose of the paper and the reason for its importance
from the abstract. The abstract should be no more than 200 words long.

\subsection{Text}

The main body of the text immediately follows the abstract. Use
10-point type in a clear, readable font with 1-point leading (10 on
11).

Indent when starting a new paragraph, except after major headings.

\subsection{Headings and Sections}

When necessary, headings should be used to separate major sections of
your paper. (These instructions use many headings to demonstrate their
appearance; your paper should have fewer headings.)

\subsubsection{Section Headings}

Print section headings in 12-point bold type in the style shown in
these instructions. Leave a blank space of approximately 10 points
above and 4 points below section headings.  Number sections with
arabic numerals.

\subsubsection{Subsection Headings}

Print subsection headings in 11-point bold type. Leave a blank space
of approximately 8 points above and 3 points below subsection
headings. Number subsections with the section number and the
subsection number (in arabic numerals) separated by a
period.

\subsubsection{Subsubsection Headings}

Print subsubsection headings in 10-point bold type. Leave a blank
space of approximately 6 points above subsubsection headings. Do not
number subsubsections.

\subsubsection{Special Sections}

You may include an unnumbered acknowledgments section, including
acknowledgments of help from colleagues, financial support, and
permission to publish.

Any appendices directly follow the text and look like sections, except
that they are numbered with capital letters instead of arabic
numerals.

The references section is headed ``References,'' printed in the same
style as a section heading but without a number. A sample list of
references is given at the end of these instructions. Use a consistent
format for references, such as that provided by Bib\TeX{}. The reference
list should not include unpublished work.

\subsection{Citations}

Citations within the text should include the author's last name and
the year of publication, for example~\cite{gottlob:nonmon}.  Append
lowercase letters to the year in cases of ambiguity.  Treat multiple
authors as in the following examples:~\cite{abelson-et-al:scheme}
or~\cite{bgf:Lixto} (for more than two authors) and
\cite{brachman-schmolze:kl-one} (for two authors).  If the author
portion of a citation is obvious, omit it, e.g.,
Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
\nocite{abelson-et-al:scheme}
\nocite{bgf:Lixto}
\nocite{brachman-schmolze:kl-one}
\nocite{gottlob:nonmon}
\nocite{gls:hypertrees}
\nocite{levesque:functional-foundations}
\nocite{levesque:belief}
\nocite{nebel:jair-2000}

\subsection{Footnotes}

Place footnotes at the bottom of the page in a 9-point font.  Refer to
them with superscript numbers.\footnote{This is how your footnotes
should appear.} Separate them from the text by a short
line.\footnote{Note the line separating these footnotes from the
text.} Avoid footnotes as much as possible; they interrupt the flow of
the text.

\section{Illustrations}

Place all illustrations (figures, drawings, tables, and photographs)
throughout the paper at the places where they are first discussed,
rather than at the end of the paper. If placed at the bottom or top of
a page, illustrations may run across both columns.

Illustrations must be rendered electronically or scanned and placed
directly in your document. All illustrations should be in black and
white, as color illustrations may cause problems. Line weights should
be 1/2-point or thicker. Avoid screens and superimposing type on
patterns as these effects may not reproduce well.

Number illustrations sequentially. Use references of the following
form: Figure 1, Table 2, etc. Place illustration numbers and captions
under illustrations. Leave a margin of 1/4-inch around the area
covered by the illustration and caption.  Use 9-point type for
captions, labels, and other text in illustrations.

\section*{Acknowledgments}

The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
files that implement them was supported by Schlumberger Palo Alto
Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
Preparation of the Microsoft Word file was supported by IJCAI.  An
early version of this document was created by Shirley Jowell and Peter
F. Patel-Schneider.  It was subsequently modified by Jennifer
Ballentine and Thomas Dean, Bernhard Nebel, and Daniel Pagenstecher.
These instructions are the same as the ones for IJCAI--05, prepared by
Kurt Steinkraus, Massachusetts Institute of Technology, Computer
Science and Artificial Intelligence Lab.

\appendix

\section{\LaTeX{} and Word Style Files}\label{stylefiles}

The \LaTeX{} and Word style files are available on the IJCAI--16
website, {\tt http://www.ijcai-16.org/}.
These style files implement the formatting instructions in this
document.

The \LaTeX{} files are {\tt ijcai16.sty} and {\tt ijcai16.tex}, and
the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai16.bib}. The
\LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
style file is for version 0.99c of Bib\TeX{} ({\em not} version
0.98i). The {\tt ijcai16.sty} file is the same as the {\tt
ijcai07.sty} file used for IJCAI--07.

The Microsoft Word style file consists of a single file, {\tt
ijcai16.doc}. This template is the same as the one used for
IJCAI--07.

These Microsoft Word and \LaTeX{} files contain the source of the
present document and may serve as a formatting sample.

Further information on using these styles for the preparation of
papers for IJCAI--16 can be obtained by contacting {\tt
ijcai16programchair@gmail.com}.

%% The file named.bst is a bibliography style file for BibTeX 0.99c
\bibliographystyle{named}
\bibliography{ijcai16}

\end{document}

### 2015-11-16 paper/ijcai16.tex
paper: jot down the main paper sections.

@@
 
-\typeout{IJCAI-16 Instructions for Authors}
-
-% These are the instructions for authors for IJCAI-16.
-% They are the same as the ones for IJCAI-11 with superficical wording
-%   changes only.
+\typeout{Setwise Max-margin Learning for Constructive Preference Elicitation or Somesuch}
 
 \documentclass{article}
-% The file ijcai16.sty is the style file for IJCAI-16 (same as ijcai07.sty).
 \usepackage{ijcai16}
-
-% Use the postscript times font!
 \usepackage{times}
-
-% the following package is optional:
 %\usepackage{latexsym}
 
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
-
-\title{IJCAI--16 Formatting Instructions\thanks{These match the formatting instructions of IJCAI-07. The support of IJCAI, Inc. is acknowledged.}}
-\author{Subbarao Kambhampati \\ 
-Arizona State University, Tempe Arizona  \\
-ijcai16programchair@gmail.com}
+\title{Setwise Max-margin Learning for Constructive Preference Elicitation or Somesuch}
+\author{ST and AP and PV}
 
@@
 \begin{abstract}
-  The {\it IJCAI--16 Proceedings} will be printed from electronic
-  manuscripts submitted by the authors. The electronic manuscript will
-  also be included in the online version of the proceedings. This paper
-  provides the style instructions.
+WRITEME
 \end{abstract}
@@
 
-The {\it IJCAI--16 Proceedings} will be printed from electronic
-manuscripts submitted by the authors. These must be PDF ({\em Portable
-Document Format}) files formatted for 8-1/2$''$ $\times$ 11$''$ paper.
+WRITEME
 
-\subsection{Length of Papers}
+\section{Related Work}
 
-Each accepted full paper is allocated six pages in the conference 
-proceedings, excluded references. References can take up to one page.
-Up to two additional pages may be purchased at a price 
-of \$275 per page for any accepted paper. However, all 
-{\em submissions} must 
-be a maximum of six pages, plus at most one for references, in length.
+WRITEME
 
-\subsection{Word Processing Software}
+\section{Setwise Max-margin Learning}
 
-As detailed below, IJCAI has prepared and made available a set of
-\LaTeX{} macros and a Microsoft Word template for use in formatting
-your paper. If you are using some other word processing software (such
-as WordPerfect, etc.), please follow the format instructions given
-below and ensure that your final paper looks as much like this sample
-as possible.
+WRITEME
 
-\section{Style and Format}
+\section{Experiments}
 
-\LaTeX{} and Word style files that implement these instructions
-can be retrieved electronically. (See Appendix~\ref{stylefiles} for
-instructions on how to obtain these files.)
+WRITEME
 
-\subsection{Layout}
+\section{Conclusion}
 
-Print manuscripts two columns to a page, in the manner in which these
-instructions are printed. The exact dimensions for pages are:
-\begin{itemize}
-\item left and right margins: .75$''$
-\item column width: 3.375$''$
-\item gap between columns: .25$''$
-\item top margin---first page: 1.375$''$
-\item top margin---other pages: .75$''$
-\item bottom margin: 1.25$''$
-\item column height---first page: 6.625$''$
-\item column height---other pages: 9$''$
-\end{itemize}
-
-All measurements assume an 8-1/2$''$ $\times$ 11$''$ page size. For
-A4-size paper, use the given top and left margins, column width,
-height, and gap, and modify the bottom and right margins as necessary.
-
-\subsection{Format of Electronic Manuscript}
-
-For the production of the electronic manuscript, you must use Adobe's
-{\em Portable Document Format} (PDF). A PDF file can be generated, for
-instance, on Unix systems using {\tt ps2pdf} or on Windows systems
-using Adobe's Distiller. There is also a website with free software
-and conversion services: {\tt http://www.ps2pdf.com/}. For reasons of
-uniformity, use of Adobe's {\em Times Roman} font is strongly suggested. In
-\LaTeX2e{}, this is accomplished by putting
-\begin{quote} 
-\mbox{\tt $\backslash$usepackage\{times\}}
-\end{quote}
-in the preamble.\footnote{You may want also to use the package {\tt
-latexsym}, which defines all symbols known from the old \LaTeX{}
-version.}
-
-Additionally, it is of utmost importance to specify the American {\bf
-letter} format (corresponding to 8-1/2$''$ $\times$ 11$''$) when
-formatting the paper. When working with {\tt dvips}, for instance, one
-should specify {\tt -t letter}.
-
-\subsection{Title and Author Information}
-
-Center the title on the entire width of the page in a 14-point bold
-font. Below it, center the author name(s) in a 12-point bold font, and
-then center the address(es) in a 12-point regular font. Credit to a
-sponsoring agency can appear on the first page as a footnote.
-
-\subsubsection{Blind Review}
-
-In order to make blind reviewing possible, authors must omit their
-names and affiliations when submitting the paper for review. In place
-of names and affiliations, provide a list of content areas. When
-referring to one's own work, use the third person rather than the
-first person. For example, say, ``Previously,
-Gottlob~\shortcite{gottlob:nonmon} has shown that\ldots'', rather
-than, ``In our previous work~\cite{gottlob:nonmon}, we have shown
-that\ldots'' Try to avoid including any information in the body of the
-paper or references that would identify the authors or their
-institutions. Such information can be added to the final camera-ready
-version for publication.
-
-\subsection{Abstract}
-
-Place the abstract at the beginning of the first column 3$''$ from the
-top of the page, unless that does not leave enough room for the title
-and author information. Use a slightly smaller width than in the body
-of the paper. Head the abstract with ``Abstract'' centered above the
-body of the abstract in a 12-point bold font. The body of the abstract
-should be in the same font as the body of the paper.
-
-The abstract should be a concise, one-paragraph summary describing the
-general thesis and conclusion of your paper. A reader should be able
-to learn the purpose of the paper and the reason for its importance
-from the abstract. The abstract should be no more than 200 words long.
-
-\subsection{Text}
-
-The main body of the text immediately follows the abstract. Use
-10-point type in a clear, readable font with 1-point leading (10 on
-11).
-
-Indent when starting a new paragraph, except after major headings.
-
-\subsection{Headings and Sections}
-
-When necessary, headings should be used to separate major sections of
-your paper. (These instructions use many headings to demonstrate their
-appearance; your paper should have fewer headings.)
-
-\subsubsection{Section Headings}
-
-Print section headings in 12-point bold type in the style shown in
-these instructions. Leave a blank space of approximately 10 points
-above and 4 points below section headings.  Number sections with
-arabic numerals.
-
-\subsubsection{Subsection Headings}
-
-Print subsection headings in 11-point bold type. Leave a blank space
-of approximately 8 points above and 3 points below subsection
-headings. Number subsections with the section number and the
-subsection number (in arabic numerals) separated by a
-period.
-
-\subsubsection{Subsubsection Headings}
-
-Print subsubsection headings in 10-point bold type. Leave a blank
-space of approximately 6 points above subsubsection headings. Do not
-number subsubsections.
-
-\subsubsection{Special Sections}
-
-You may include an unnumbered acknowledgments section, including
-acknowledgments of help from colleagues, financial support, and
-permission to publish.
-
-Any appendices directly follow the text and look like sections, except
-that they are numbered with capital letters instead of arabic
-numerals.
-
-The references section is headed ``References,'' printed in the same
-style as a section heading but without a number. A sample list of
-references is given at the end of these instructions. Use a consistent
-format for references, such as that provided by Bib\TeX{}. The reference
-list should not include unpublished work.
-
-\subsection{Citations}
-
-Citations within the text should include the author's last name and
-the year of publication, for example~\cite{gottlob:nonmon}.  Append
-lowercase letters to the year in cases of ambiguity.  Treat multiple
-authors as in the following examples:~\cite{abelson-et-al:scheme}
-or~\cite{bgf:Lixto} (for more than two authors) and
-\cite{brachman-schmolze:kl-one} (for two authors).  If the author
-portion of a citation is obvious, omit it, e.g.,
-Nebel~\shortcite{nebel:jair-2000}.  Collapse multiple citations as
-follows:~\cite{gls:hypertrees,levesque:functional-foundations}.
-\nocite{abelson-et-al:scheme}
-\nocite{bgf:Lixto}
-\nocite{brachman-schmolze:kl-one}
-\nocite{gottlob:nonmon}
-\nocite{gls:hypertrees}
-\nocite{levesque:functional-foundations}
-\nocite{levesque:belief}
-\nocite{nebel:jair-2000}
-
-\subsection{Footnotes}
-
-Place footnotes at the bottom of the page in a 9-point font.  Refer to
-them with superscript numbers.\footnote{This is how your footnotes
-should appear.} Separate them from the text by a short
-line.\footnote{Note the line separating these footnotes from the
-text.} Avoid footnotes as much as possible; they interrupt the flow of
-the text.
-
-\section{Illustrations}
-
-Place all illustrations (figures, drawings, tables, and photographs)
-throughout the paper at the places where they are first discussed,
-rather than at the end of the paper. If placed at the bottom or top of
-a page, illustrations may run across both columns.
-
-Illustrations must be rendered electronically or scanned and placed
-directly in your document. All illustrations should be in black and
-white, as color illustrations may cause problems. Line weights should
-be 1/2-point or thicker. Avoid screens and superimposing type on
-patterns as these effects may not reproduce well.
-
-Number illustrations sequentially. Use references of the following
-form: Figure 1, Table 2, etc. Place illustration numbers and captions
-under illustrations. Leave a margin of 1/4-inch around the area
-covered by the illustration and caption.  Use 9-point type for
-captions, labels, and other text in illustrations.
+WRITEME
 
@@
 
-The preparation of these instructions and the \LaTeX{} and Bib\TeX{}
-files that implement them was supported by Schlumberger Palo Alto
-Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-Preparation of the Microsoft Word file was supported by IJCAI.  An
-early version of this document was created by Shirley Jowell and Peter
-F. Patel-Schneider.  It was subsequently modified by Jennifer
-Ballentine and Thomas Dean, Bernhard Nebel, and Daniel Pagenstecher.
-These instructions are the same as the ones for IJCAI--05, prepared by
-Kurt Steinkraus, Massachusetts Institute of Technology, Computer
-Science and Artificial Intelligence Lab.
+WRITEME
 
-\appendix
-
-\section{\LaTeX{} and Word Style Files}\label{stylefiles}
-
-The \LaTeX{} and Word style files are available on the IJCAI--16
-website, {\tt http://www.ijcai-16.org/}.
-These style files implement the formatting instructions in this
-document.
-
-The \LaTeX{} files are {\tt ijcai16.sty} and {\tt ijcai16.tex}, and
-the Bib\TeX{} files are {\tt named.bst} and {\tt ijcai16.bib}. The
-\LaTeX{} style file is for version 2e of \LaTeX{}, and the Bib\TeX{}
-style file is for version 0.99c of Bib\TeX{} ({\em not} version
-0.98i). The {\tt ijcai16.sty} file is the same as the {\tt
-ijcai07.sty} file used for IJCAI--07.
-
-The Microsoft Word style file consists of a single file, {\tt
-ijcai16.doc}. This template is the same as the one used for
-IJCAI--07.
-
-These Microsoft Word and \LaTeX{} files contain the source of the
-present document and may serve as a formatting sample.
-
-Further information on using these styles for the preparation of
-papers for IJCAI--16 can be obtained by contacting {\tt
-ijcai16programchair@gmail.com}.
-
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
 \bibliographystyle{named}

### 2015-12-21 src/data/performSparsification_batch.m
data: add new data and scripts provided by Paolo.

@@ -1,0 +1,60 @@
+% This script performs sparsification of utility vectors on several files
+% at once; it make use of function 'sparsifyUtilityVecors'.
+%
+% Input files whose name contain the pattern 'spase' or 'sparsified' are 
+% ignored and as well those containing 'Guo'. Output is stored on different
+% files.
+%
+% We assume that vectors in input files are stored in utilityWeights.
+%
+% Use with caution, it may mess up your directory!
+sparsificationCoeff = 0.8;
+
+% define input directory
+baseDir = './randomUtility/';
+sparseLabel = 'sparse';
+GuoLabel = 'GuoFormat';
+extensionTXT = '.txt';
+extensionMAT = '.mat';
+
+% look at all MAT files in baseDir
+MATfiles=dir([baseDir,'*.mat']);
+
+fprintf('I am processing these files:\n');
+for file=MATfiles'
+    fullFileName = [baseDir, file.name];
+    
+    % avoid processing files that contain 'sparsification' in its name
+    if not(isempty(strfind(file.name,'sparse')))...
+            || not(isempty(strfind(file.name,'sparsified')))... 
+            || not(isempty(strfind(file.name,sparseLabel)))
+        continue
+    end
+    
+    % avoid processing files that contain 'Guo' in its name
+    if not(isempty(strfind(file.name,'Guo'))) 
+        continue
+    end
+    
+    % display the name of the file under consideration
+    disp(fullFileName);
+    load(fullFileName);
+    
+    % perform sparsification calling function 'sparsifyUtilityVectors'
+    % NOTE: assume that vectors are stored in variable utilityWeights
+    sparsifiedUtilityVectors=sparsifyUtilityVectors(utilityWeights, sparsificationCoeff);
+    
+    % define names of output files
+    [pathstr,nameWithoutExt,ext] = fileparts(fullFileName);
+    outputFileName_TXT = [baseDir, nameWithoutExt, '_',sparseLabel, extensionTXT];
+    outputFileName_MAT = [baseDir, nameWithoutExt, '_',sparseLabel, extensionMAT];
+    outputFileName_Guo_MAT = [baseDir, nameWithoutExt, '_',sparseLabel,'_',GuoLabel, extensionMAT];   
+    
+    % save to output files
+    utilityWeights = sparsifiedUtilityVectors;
+    save(outputFileName_MAT,'utilityWeights');
[…]

### 2015-12-21 src/data/sparsifyUtilityVectors.m

@@ -1,0 +1,23 @@
+function sparsifiedUtilityVectors=sparsifyUtilityVectors(utilityVectors, sparsificationCoefficient)
+% sparsifiedUtilityVectors=sparsifyUtilityVectors(utilityVectors, sparsificationCoefficint)
+%
+% INPUT
+% - utilityVectors:  matrix (n X m), with n=n.of vectors m=n.of binary 
+% featuresthis parameter the percentage of binary feature whose 
+% weight will be zero
+% - sparsificationCoefficint: fraction of features to be put to 0 in each
+% row
+%
+% OUTPUT
+% - sparsifiedUtilityVectors: matrix (n X m), with n=n.of vectors m=n.of 
+% binary features (same size as utilityVectors given in input) giving the
+% utility vectors obtained after sparsification
+nVectors=size(utilityVectors,1);
+nBinaryFeatures=size(utilityVectors,2);
+nIrrelevantFeatures = floor(sparsificationCoefficient*nBinaryFeatures);
+sparsifiedUtilityVectors = utilityVectors;
+for i=1:nVectors
+    % randomly sample nIrrelevantFeatures
+    irrelevantFeatures = randsample(nBinaryFeatures,nIrrelevantFeatures);
+    sparsifiedUtilityVectors(i,irrelevantFeatures)=0;
+end

### 2015-12-21 setmargin/table-synthetic.tex
setmargin: turn into package.

--- 2015-10-07 src/table-synthetic.tex
@@
 
-    \begin{tabular}{|c|c|c|}
+    \begin{tabular}{|c|c|c|c|}
+        {\bf 1 hyperplane} &
         {\bf 2 hyperplanes} &
@@
         \\
-        \multicolumn{3}{|c|}{{\bf Deterministic, single margin}}
+        \multicolumn{4}{|c|}{{\bf Deterministic, single margin}}
         \\
@@
         \\
-        \multicolumn{3}{|c|}{{\bf Deterministic, multiple margin}}
+        \multicolumn{4}{|c|}{{\bf Deterministic, multiple margin}}
         \\
@@
         \\
-        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, single margin}}
+        \multicolumn{4}{|c|}{{\bf Non-deterministic, with indifference, single margin}}
         \\
@@
         \\
-        \multicolumn{3}{|c|}{{\bf Non-deterministic, with indifference, multiple margin}}
+        \multicolumn{4}{|c|}{{\bf Non-deterministic, with indifference, multiple margin}}
         \\
@@
         \\
-        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, single margin}}
+        \multicolumn{4}{|c|}{{\bf Non-deterministic, no indifference, single margin}}
         \\
@@
         \\
-        \multicolumn{3}{|c|}{{\bf Non-deterministic, no indifference, multiple margin}}
+        \multicolumn{4}{|c|}{{\bf Non-deterministic, no indifference, multiple margin}}
         \\

### 2015-12-21 setmargin/__init__.py

@@ -1,0 +1,4 @@
+from datasets import *
+from user import *
+from solver import *
+from util import *

### 2016-01-06 paper/ijcai16.tex
paper: add algorithm and a few (expendable) definitions.
paper: draft the non-linear OP.
ijcai16: added draft of intro

@@
 
-\typeout{Setwise Max-margin Learning for Constructive Preference Elicitation or Somesuch}
+\typeout{Constructive Preference Elicitation by Setwise Max-margin Learning}
 
@@
 
-\title{Setwise Max-margin Learning for Constructive Preference Elicitation or Somesuch}
+\RequirePackage{amsmath,amssymb,amsfonts,amsthm}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+
+% ---------------------------------------------------------------------------
+
+\algrenewcommand{\algorithmicrequire}{\textbf{Input:}}
+\algrenewcommand{\algorithmicensure}{\textbf{Output:}}
+
+\renewcommand\[{\begin{equation}}
+\renewcommand\]{\end{equation}}
+
+% Bold math symbols
+\newcommand{\bbA}{\mathbb{A}}
+\newcommand{\bbB}{\mathbb{B}}
+\newcommand{\bbC}{\mathbb{C}}
+\newcommand{\bbD}{\mathbb{D}}
+\newcommand{\bbE}{\mathbb{E}}
+\newcommand{\bbF}{\mathbb{F}}
+\newcommand{\bbG}{\mathbb{G}}
+\newcommand{\bbH}{\mathbb{H}}
+\newcommand{\bbI}{\mathbb{I}}
+\newcommand{\bbJ}{\mathbb{J}}
+\newcommand{\bbK}{\mathbb{K}}
+\newcommand{\bbL}{\mathbb{L}}
+\newcommand{\bbM}{\mathbb{M}}
+\newcommand{\bbN}{\mathbb{N}}
+\newcommand{\bbO}{\mathbb{O}}
+\newcommand{\bbP}{\mathbb{P}}
+\newcommand{\bbQ}{\mathbb{Q}}
+\newcommand{\bbR}{\mathbb{R}}
+\newcommand{\bbS}{\mathbb{S}}
+\newcommand{\bbT}{\mathbb{T}}
+\newcommand{\bbU}{\mathbb{U}}
+\newcommand{\bbV}{\mathbb{V}}
+\newcommand{\bbW}{\mathbb{W}}
+\newcommand{\bbX}{\mathbb{X}}
+\newcommand{\bbY}{\mathbb{Y}}
+\newcommand{\bbZ}{\mathbb{Z}}
+
+% Calligraphic math symbols
+\newcommand{\calvar}[1]{\ensuremath{\mathcal{#1}}}
+\newcommand{\calA}{\calvar{A}}
+\newcommand{\calB}{\calvar{B}}
+\newcommand{\calC}{\calvar{C}}
+\newcommand{\calD}{\calvar{D}}
+\newcommand{\calE}{\calvar{E}}
+\newcommand{\calF}{\calvar{F}}
+\newcommand{\calG}{\calvar{G}}
+\newcommand{\calH}{\calvar{H}}
+\newcommand{\calI}{\calvar{I}}
+\newcommand{\calJ}{\calvar{J}}
+\newcommand{\calK}{\calvar{K}}
+\newcommand{\calL}{\calvar{L}}
+\newcommand{\calM}{\calvar{M}}
+\newcommand{\calN}{\calvar{N}}
+\newcommand{\calO}{\calvar{O}}
+\newcommand{\calP}{\calvar{P}}
+\newcommand{\calQ}{\calvar{Q}}
+\newcommand{\calR}{\calvar{R}}
+\newcommand{\calS}{\calvar{S}}
+\newcommand{\calT}{\calvar{T}}
+\newcommand{\calU}{\calvar{U}}
+\newcommand{\calV}{\calvar{V}}
+\newcommand{\calW}{\calvar{W}}
+\newcommand{\calX}{\calvar{X}}
+\newcommand{\calY}{\calvar{Y}}
+\newcommand{\calZ}{\calvar{Z}}
+
+% Vectors
+\newcommand{\vecvar}[1]{\ensuremath{\boldsymbol{#1}}}
+\newcommand{\va}{\vecvar{a}}
+\newcommand{\vb}{\vecvar{b}}
+\newcommand{\vc}{\vecvar{c}}
+\newcommand{\vd}{\vecvar{d}}
+\newcommand{\ve}{\vecvar{e}}
+\newcommand{\vf}{\vecvar{f}}
+\newcommand{\vg}{\vecvar{g}}
+\newcommand{\vh}{\vecvar{h}}
+\newcommand{\vi}{\vecvar{i}}
+\newcommand{\vj}{\vecvar{j}}
+\newcommand{\vk}{\vecvar{k}}
+\newcommand{\vl}{\vecvar{l}}
+\newcommand{\vm}{\vecvar{m}}
+\newcommand{\vn}{\vecvar{n}}
+\newcommand{\vo}{\vecvar{o}}
+\newcommand{\vp}{\vecvar{p}}
+\newcommand{\vq}{\vecvar{q}}
+\newcommand{\vr}{\vecvar{r}}
+\newcommand{\vs}{\vecvar{s}}
+\newcommand{\vt}{\vecvar{t}}
+\newcommand{\vu}{\vecvar{u}}
+\newcommand{\vv}{\vecvar{v}}
+\newcommand{\vw}{\vecvar{w}}
+\newcommand{\vx}{\vecvar{x}}
+\newcommand{\vy}{\vecvar{y}}
+\newcommand{\vz}{\vecvar{z}}
+\newcommand{\valpha}{\vecvar{\alpha}}
+\newcommand{\veps}{\vecvar{\varepsilon}}
+\newcommand{\vphi}{\vecvar{\varphi}}
+\newcommand{\vpsi}{\vecvar{\psi}}
+
+% Operators
+\DeclareMathOperator*{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
+
+% Debugging
+
+\usepackage{color}
+\newcommand{\andrea}[1]{{\bf \textcolor{blue}{{\fbox{Andrea:} #1}}}}
+\newcommand{\stefano}[1]{{\bf \textcolor{green}{{\fbox{Stefano:} #1}}}}
+\newcommand{\paolo}[1]{{\bf \textcolor{red}{{\fbox{Paolo:} #1}}}}
+
+% ---------------------------------------------------------------------------
+
+\title{Constructive Preference Elicitation by Setwise Max-margin Learning}
 \author{ST and AP and PV}
@@
 
+- importance of recommendation systems and preference elicitation
+
+- existing approaches and their limitations
+
+% constructive preference elicitation
+In this paper we take a {\em constructive} view on preference
+elicitation, enlarging its scope from the selection of items among a
+set of candidates to the synthesis of entirely novel
+instances. Instances are represented as combinations of basic elements
+(e.g. the components of a laptop) subject to a set of constraints
+(e.g. the laptop model determines the set of available CPUs). A
+utility function is learned over the feature representation of an
+instance, as customary in many preference elicitation approaches. The
+recommendation is then made by solving a constrained optimization
+problem in the space of feasible instances, guided by the learned
+utility. \andrea{other constructive preference elicitation approaches?}
+
+% setwise max-margin formulation to deal with uncertainty in user utility
+Learning a user utility function requires setting a preference
+elicitation strategy and dealing with uncertain and possibly
+inconsistent user feedback. Bayesian preference elicitation approaches
+deal with both problems by building a probability distribution on
+candidate functions and asking queries maximizing informativeness
+measures such as expected value of information (EVOI)~\cite{}. These
+approaches are however computationally expensive and cannot scale to
+fully constructive scenarios, as shown in our experimental results.
+We take a space decomposition perspective and jointly learn a set of
+weight vectors, each representing a candidate utility function,
+maximizing diversity between the vectors and consistency with the
+available feedback. These two conflicting objectives tend to generate
+equally plausible alternative hypotheses for the unknown
+utility. Preference elicitation works by combining weight vector
+learning with instance generation, so that each iteration of the
+algorithm produces two outcomes: a set of weight vectors and a set of
+instances, each maximizing its score according to one of the weight
+vectors.
+
+- experimental results
+
+- paper structure
+
 \section{Related Work}
@@
 
+\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
+capital letters $X$ for matrices, and $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
+
+\paragraph{Non-linear Formulation.}
+
+Let $k$ be set size and $n$ be the number of examples (answers).
+
+{\footnotesize
+
+\begin{align}
+    \max_{M, W, X}
+        & \;\; M - \frac{\alpha}{k |\calD|} \sum_{i=1}^k \| \veps^{(i)} \|_1 - \frac{\beta}{k} \sum_{i=1}^k \| \vw^{(i)} \|_1 + \frac{\gamma}{k} \sum_{i=1}^k \langle \vw^{(i)}, \vx^{(i)} \rangle
+        \nonumber
+    \\
+    \text{s.t.}
+        & \;\; \forall \; i \in [k] \nonumber, \forall \; h \in [n]
+    \\
+        & \;\; \qquad \langle \vw^{(i)}, \vy^{(h)}_+ - \vy^{(h)}_- \rangle \geq M - \varepsilon^{(i)}_h \nonumber
+    \\
+        & \;\; \forall \; i, j \in [k], i \neq j \nonumber
+    \\
+        & \;\; \qquad \langle \vw^{(i)}, \vx^{(i)} - \vx^{(j)} \rangle \geq M \label{cnst-ws-m}
+    \\
+        & \;\; \forall \; i \in [k] \nonumber
+    \\
+        & \;\; \qquad \vw^\bot \leq \vw^{(i)} \leq \vw^\top \nonumber
+    \\
+        & \;\; \forall \; i \in [k] \nonumber
+    \\
+        & \;\; \qquad \vx^i \in \calX_{\text{feasible}} \nonumber
+    \\
+        & \;\; \forall \; h \in 1, \ldots, n, \forall \; i \in [k] \nonumber
+    \\
+        & \;\; \qquad \varepsilon^{(i)}_h \geq 0 \nonumber
+    \\
+        & \;\; M \geq 0 \nonumber
+\end{align}
+
+}
+
+\paragraph{MILP Formulation.}
+
 WRITEME
+
+\begin{algorithm}
+\begin{algorithmic}[1]
+    \Procedure{PESetMargin}{$Y, m; \calW, \calX, \valpha, T$}
+        \State $\calA \gets \emptyset$
+        \For{$t = 1, \ldots, T$}
+            \State $\vw^*_1, \ldots, \vw^*_m, \vx^*_1, \ldots, \vx^*_m \gets \text{{\sc Solve}}(Y, \calA, m; \calW, \calX, \valpha)$
+            \For{$\vx^*_i,\vx^*_j \in \{ \vx^*_1, \ldots, \vx^*_m \} \; \text{{\bf s.t.}} \; i < j$}
+                \State $\calA \gets \calA \cup \text{{\sc QueryUser}}(\vx^*_i,\vx^*_j)$
+            \EndFor
+        \EndFor
+        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(Y, \calA, 1; \calW, \calX, \valpha)$
+        \State ${\bf return}\; \vw^*, \vx^*$
+    \EndProcedure
+\end{algorithmic}
+\end{algorithm}
 

### 2016-01-08 paper/ijcai16.tex
paper: beautify non-linear problem.
paper: beautify the setmargin algorithm.
paper: add a couple of bullet points.

@@
 \newcommand{\vpsi}{\vecvar{\psi}}
+\newcommand{\vtheta}{\vecvar{\theta}}
 
@@
 
-Let $k$ be set size and $n$ be the number of examples (answers).
+Let $k$ be the set size and $n$ be the number of examples (answers).
 
@@
     \max_{M, W, X}
-        & \;\; M - \frac{\alpha}{k |\calD|} \sum_{i=1}^k \| \veps^{(i)} \|_1 - \frac{\beta}{k} \sum_{i=1}^k \| \vw^{(i)} \|_1 + \frac{\gamma}{k} \sum_{i=1}^k \langle \vw^{(i)}, \vx^{(i)} \rangle
+        & \;\; M - \frac{\alpha}{k n} \sum_{i=1}^k \| \veps^{(i)} \|_1 - \frac{\beta}{k} \sum_{i=1}^k \| \vw^{(i)} \|_1 + \frac{\gamma}{k} \sum_{i=1}^k \langle \vw^{(i)}, \vx^{(i)} \rangle
         \nonumber
@@
     \text{s.t.}
-        & \;\; \forall \; i \in [k] \nonumber, \forall \; h \in [n]
+        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
     \\
-        & \;\; \qquad \langle \vw^{(i)}, \vy^{(h)}_+ - \vy^{(h)}_- \rangle \geq M - \varepsilon^{(i)}_h \nonumber
+        & \;\; \qquad \langle \vw^{(i)}, \vy^{(h)}_+ - \vy^{(h)}_- \rangle \geq M - \varepsilon^{(i)}_h
     \\
-        & \;\; \forall \; i, j \in [k], i \neq j \nonumber
+        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{(i)}, \vx^{(i)} - \vx^{(j)} \rangle \geq M \label{cnst-ws-m}
     \\
-        & \;\; \qquad \langle \vw^{(i)}, \vx^{(i)} - \vx^{(j)} \rangle \geq M \label{cnst-ws-m}
+        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \leq \vw^{(i)} \leq \vw^\top
     \\
-        & \;\; \forall \; i \in [k] \nonumber
+        & \;\; \forall \; i \in [k] \;.\; \vx^{(i)} \in \calX_{\text{feasible}}
     \\
-        & \;\; \qquad \vw^\bot \leq \vw^{(i)} \leq \vw^\top \nonumber
+        & \;\; \forall \; i \in [k] \;.\; \veps^{(i)}_h \geq 0
     \\
-        & \;\; \forall \; i \in [k] \nonumber
-    \\
-        & \;\; \qquad \vx^i \in \calX_{\text{feasible}} \nonumber
-    \\
-        & \;\; \forall \; h \in 1, \ldots, n, \forall \; i \in [k] \nonumber
-    \\
-        & \;\; \qquad \varepsilon^{(i)}_h \geq 0 \nonumber
-    \\
-        & \;\; M \geq 0 \nonumber
+        & \;\; M \geq 0
 \end{align}
@@
 
+\paragraph{Set-wise Preference Elicitation.}
+
+WRITEME
+
 \begin{algorithm}
+{\footnotesize
 \begin{algorithmic}[1]
-    \Procedure{PESetMargin}{$Y, m; \calW, \calX, \valpha, T$}
-        \State $\calA \gets \emptyset$
+    \Procedure{SetMargin}{$k, \vtheta, T$}
+        \State $\calQ \gets \emptyset$
         \For{$t = 1, \ldots, T$}
-            \State $\vw^*_1, \ldots, \vw^*_m, \vx^*_1, \ldots, \vx^*_m \gets \text{{\sc Solve}}(Y, \calA, m; \calW, \calX, \valpha)$
-            \For{$\vx^*_i,\vx^*_j \in \{ \vx^*_1, \ldots, \vx^*_m \} \; \text{{\bf s.t.}} \; i < j$}
-                \State $\calA \gets \calA \cup \text{{\sc QueryUser}}(\vx^*_i,\vx^*_j)$
+            \State $\vw^{(1)}, \ldots, \vw^{(k)}, \vx^{(1)}, \ldots, \vx^{(k)} \gets \text{{\sc Solve}}(\calQ, k, \vtheta)$
+            \For{$\vx^{(i)},\vx^{(j)} \in \{ \vx^{(1)}, \ldots, \vx^{(k)} \} \; \text{{\bf s.t.}} \; i < j$}
+                \State $\calQ \gets \calQ \cup \text{{\sc QueryUser}}(\vx^{(i)},\vx^{(j)})$
             \EndFor
         \EndFor
-        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(Y, \calA, 1; \calW, \calX, \valpha)$
+        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(\calQ, 1, \vtheta)$
         \State ${\bf return}\; \vw^*, \vx^*$
@@
 \end{algorithmic}
+}
+\caption{The {\sc SetMargin} algorithm. Here $k$ is the set size, $\vtheta$ are
+the $\alpha,\beta,\gamma$ hyperparameters, and $T$ is the maximum number of
+iterations.}
 \end{algorithm}
@@
 
-WRITEME
+\paragraph{Synthetic Dataset.} Compare against Guo's and Paolo's algorithms
+in terms of both convergence rate and running time. Do not forget the
+sparse variants.
+
+\paragraph{Constructive Dataset.} The PC dataset, with and without cost
+variables.
 

### 2016-01-08 ijcai16.py
ijcai16: complete IJCA16 script.

@@ -2,6 +2,7 @@
 # -*- coding: utf-8 -*-
 
 import numpy as np
+import matplotlib.pyplot as plt
 from itertools import product
 from os.path import join
 from textwrap import dedent
@@ -13,6 +14,8 @@ import setmargin
 class Grid(object):
     def __init__(self, d):
         self.__dict__.update(d)
+    def asdict(self):
+        return self.__dict__
     def iterate(self):
         values = []
         for value in self.__dict__.values():
@@ -30,7 +33,68 @@ def _load_utilities(num_attrs, sampling_mode):
     with open(join("data", "randomUtility", basename), "rb") as fp:
         for line in fp:
             utilities.append(map(float, line.split(",")))
-    return np.array(utilities) / 100.0
+    # rescale the range from [0,100) to [0,1)
+    utilities = np.array(utilities) / 100.0
+    # XXX turn negative values into zeros
+    utilities[utilities < 0] = 0
+    return utilities
+
+def dump_and_draw(dataset_name, config, infos):
+    basename = "__".join(map(str, [
+        dataset_name,
+        "k={}".format(config.set_size),
+        config.sampling_mode,
+        config.ranking_mode,
+        "deterministic={}".format(config.is_deterministic),
+        "indifferent={}".format(config.is_indifferent),
+        "theta={},{},{}".format(config.alpha, config.beta, config.gamma),
+        "multimargin={}".format(config.multimargin),
+        "threads={}".format(config.threads),
+        config.num_trials,
+        config.num_iterations,
+    ]))
+
+    num_trials = len(infos)
+    max_queries = max([sum(n for n, _, _ in info) for info in infos])
+
+    loss_matrix = np.zeros((num_trials, max_queries))
+    time_matrix = np.zeros((num_trials, max_queries))
+    for i, info in enumerate(infos):
+        base = 0
+        prev_loss = max(max(l for _, l, _ in info) for info in infos)
+        for num_queries, loss, time in info:
+            for j in range(num_queries):
+                alpha = 1 - (j + 1) / float(num_queries)
+                interpolated_loss = alpha*prev_loss + (1 - alpha)*loss
+                loss_matrix[i, base+j] = interpolated_loss
+                time_matrix[i, base+j] = time / num_queries
+            base += num_queries
+            prev_loss = loss
+
+    np.savetxt("results_{}_loss_matrix.txt".format(basename), loss_matrix)
+    np.savetxt("results_{}_time_matrix.txt".format(basename), time_matrix)
+
+    def ms(x):
[…]

### 2016-01-09 ijcai16.py
ijcai16: factorize out some code.

@@ -40,7 +40,7 @@ def _load_utilities(num_attrs, sampling_mode):
     utilities[utilities < 0] = 0
     return utilities
 
-def dump_and_draw(dataset_name, config, infos):
+def get_result_paths(dataset_name, config):
     basename = "__".join(map(str, [
         dataset_name,
         "k={}".format(config.set_size),
@@ -54,6 +54,15 @@ def dump_and_draw(dataset_name, config, infos):
         config.num_trials,
         config.num_iterations,
     ]))
+    path1 = "results_{}_loss_matrix.txt".format(basename)
+    path2 = "results_{}_time_matrix.txt".format(basename)
+    path3 = "results_{}_avgloss.svg".format(basename)
+    path4 = "results_{}_avgtime.svg".format(basename)
+    return path1, path2, path3, path4
+
+def dump_and_draw(dataset_name, config, infos):
+    loss_matrix_path, time_matrix_path, loss_svg_path, time_svg_path = \
+        get_result_paths(dataset_name, config)
 
     num_trials = len(infos)
     max_queries = max([sum(n for n, _, _ in info) for info in infos])
@@ -72,8 +81,8 @@ def dump_and_draw(dataset_name, config, infos):
             base += num_queries
             prev_loss = loss
 
-    np.savetxt("results_{}_loss_matrix.txt".format(basename), loss_matrix)
-    np.savetxt("results_{}_time_matrix.txt".format(basename), time_matrix)
+    np.savetxt(loss_matrix_path, loss_matrix)
+    np.savetxt(time_matrix_path, time_matrix)
 
     def ms(x):
         return np.mean(x, axis=0), np.std(x, ddof=1, axis=0).reshape(-1, 1)
@@ -87,7 +96,7 @@ def dump_and_draw(dataset_name, config, infos):
     ax.set_ylabel("Average loss over trials")
     ax.set_ylim([0.0, max(0.5, max(loss_means) + max(loss_stddevs) + 0.1)])
     ax.errorbar(np.arange(1, max_queries + 1), loss_means, yerr=loss_stddevs)
-    fig.savefig("results_{}_avgloss.svg".format(basename), bbox_inches="tight")
+    fig.savefig(loss_svg_path, bbox_inches="tight")
     del fig
     del ax
 
@@ -97,7 +106,7 @@ def dump_and_draw(dataset_name, config, infos):
     ax.set_ylabel("Average time over trials")
[…]

### 2016-01-09 ijcai16.py
ijcai16: pass the debug flag to the solver.

@@ -115,7 +115,7 @@ def solve(dataset, config, ws=None):
 
     solver = setmargin.Solver((config.alpha, config.beta, config.gamma),
                               multimargin=config.multimargin,
-                              threads=config.threads)
+                              threads=config.threads, debug=config.debug)
     infos = []
     for trial in range(config.num_trials):
         print dedent("""\

### 2016-01-09 setmargin/solver.py
setmargin/solver: update dumping/exception code.

@@ -9,6 +9,22 @@ from util import *
 
 MAX_W_Z = 1
 
+status_to_reason = {
+    1: "LOADED",
+    2: "OPTIMAL",
+    3: "INFEASIBLE",
+    4: "INF_OR_UNBD",
+    5: "UNBOUNDED",
+    6: "CUTOFF",
+    7: "ITERATION_LIMIT",
+    8: "NODE_LIMIT",
+    9: "TIME_LIMIT",
+    10: "SOLUTION_LIMIT",
+    11: "INTERRUPTED",
+    12: "NUMERIC",
+    13: "SUBOPTIMAL",
+}
+
 class Solver(object):
     """Set-margin solver based on the Gurobi MILP solver.
 
@@ -31,10 +47,12 @@ class Solver(object):
         self._debug = debug
 
     def _dump_model(self, model, prefix):
-        fp = tempfile.NamedTemporaryFile(prefix=prefix, suffix=".lp",
-                                         delete=False)
-        fp.close()
-        model.write(fp.name)
+        if self._debug:
+            fp = tempfile.NamedTemporaryFile(prefix=prefix, suffix=".lp",
+                                             delete=False)
+            fp.close()
+            print "dumping model to", fp.name
+            model.write(fp.name)
 
     def _check_best_score(self, dataset, user, best_score, best_item):
         if dataset.items is None:
@@ -108,7 +126,7 @@ class Solver(object):
             model.optimize()
             best_score = model.objVal
         except:
-            raise RuntimeError("optimization failed!")
+            raise RuntimeError("optimization failed! {}".format(status_to_reason[model.status]))
         best_item = np.array([var.x for var in x])
 
         self._check_best_score(dataset, user, best_score, best_item)
@@ -254,12 +272,12 @@ class Solver(object):
             x = [xs[(i,z)] for z in range(num_features)]
             self._add_item_constraints(model, dataset, x)
 
-        self._dump_model(model, "setmargin_gurobi")
+        self._dump_model(model, "setmargin_full")
         try:
             model.optimize()
             _ = model.objVal
         except:
-            raise RuntimeError("optimization failed!")
+            raise RuntimeError("optimization failed! {}".format(status_to_reason[model.status]))
 
         output_ws = np.zeros((set_size, num_features))
         output_xs = np.zeros((set_size, num_features))

### 2016-01-09 setmargin/solver.py
setmargin/solve: do not ignore alphas[2]...

@@ -184,9 +184,9 @@ class Solver(object):
         obj_weights = alpha * grb.quicksum(ws.values())
 
         alpha = self._alphas[2] / set_size
-        obj_scores = grb.quicksum([ps[i,i,z]
-                                   for i in range(set_size)
-                                   for z in range(num_features)])
+        obj_scores = alpha * grb.quicksum([ps[i,i,z]
+                                           for i in range(set_size)
+                                           for z in range(num_features)])
 
         model.setObjective(obj_margins - obj_slacks - obj_weights + obj_scores)
 

### 2016-01-09 setmargin/solver.py
setmargin/solver: print actual alphas when the problem is unsolvable.

@@ -173,20 +173,23 @@ class Solver(object):
         model.update()
 
         # Define the objective function
+        alphas = (
+            0.0 if len(slacks) == 0 else self._alphas[0] / (set_size * num_examples),
+            self._alphas[1] / set_size,
+            self._alphas[2] / set_size,
+        )
+
         obj_margins = grb.quicksum(margins)
 
         obj_slacks = 0
         if len(slacks) > 0:
-            alpha = self._alphas[0] / (set_size * num_examples)
-            obj_slacks = alpha * grb.quicksum(slacks.values())
+            obj_slacks = alphas[0] * grb.quicksum(slacks.values())
 
-        alpha = self._alphas[1] / set_size
-        obj_weights = alpha * grb.quicksum(ws.values())
+        obj_weights = alphas[1] * grb.quicksum(ws.values())
 
-        alpha = self._alphas[2] / set_size
-        obj_scores = alpha * grb.quicksum([ps[i,i,z]
-                                           for i in range(set_size)
-                                           for z in range(num_features)])
+        obj_scores = alphas[2] * grb.quicksum([ps[i,i,z]
+                                               for i in range(set_size)
+                                               for z in range(num_features)])
 
         model.setObjective(obj_margins - obj_slacks - obj_weights + obj_scores)
 
@@ -277,7 +280,17 @@ class Solver(object):
             model.optimize()
             _ = model.objVal
         except:
-            raise RuntimeError("optimization failed! {}".format(status_to_reason[model.status]))
+            message = dedent("""\
+                optimization failed!
+
+                answers =
+                {}
+
+                set_size = {}
+                status = {}
+                alphas = {}
+                """).format(answers, set_size, status_to_reason[model.status], alphas)
+            raise RuntimeError(message)
 
         output_ws = np.zeros((set_size, num_features))
         output_xs = np.zeros((set_size, num_features))

### 2016-01-09 ijcai16.py
ijcai16: do not ignore np warnings.

@@ -247,6 +247,7 @@ def run_from_command_line():
                   config, infos)
 
 if __name__ == "__main__":
+    np.seterr(all="raise")
     if len(sys.argv) == 1:
         run_synthetic()
         run_pc_nocost()

### 2016-01-09 setmargin/solver.py
setmargin/solver: set gurobi model seed for extra reproducibility.

@@ -105,6 +105,7 @@ class Solver(object):
         num_features = sum(dataset.domain_sizes)
 
         model = grb.Model("setmargin_dot")
+        model.params.Seed = 0
         model.params.OutputFlag = 0
 
         x = [model.addVar(vtype=GRB.BINARY, name="x_{}".format(z))
@@ -144,6 +145,7 @@ class Solver(object):
         num_features = sum(dataset.domain_sizes)
 
         model = grb.Model("facility")
+        model.params.Seed = 0
         model.params.Threads = self._threads
         model.params.OutputFlag = 0
 

### 2016-01-09 setmargin/solver.py
setmargin/solver: make sure to dump the updated model...

@@ -277,6 +277,7 @@ class Solver(object):
             x = [xs[(i,z)] for z in range(num_features)]
             self._add_item_constraints(model, dataset, x)
 
+        model.update()
         self._dump_model(model, "setmargin_full")
         try:
             model.optimize()

### 2016-01-09 paper/ijcai16.tex
paper: add a note to the equation, update slightly.

@@
 
-Let $k$ be the set size and $n$ be the number of examples (answers).
+Let $k$ be the set size and $n$ be the number of examples (answers). Here
+the $\vw$'s are non-negative real vectors, while the $\vx$'s are 0-1
+vectors.
 
@@
     \\
-        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{(i)}, \vx^{(i)} - \vx^{(j)} \rangle \geq M \label{cnst-ws-m}
+        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{(i)}, \vx^{(i)} - \vx^{(j)} \rangle \geq M \label{eq:wxconstr}
     \\
-        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \leq \vw^{(i)} \leq \vw^\top
+        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \leq \vw^{(i)} \leq \vw^\top \label{eq:wbounds}
     \\
@@
 }
+
+Note that due to the non-negativity of $\vw^\bot$, Eq.~\ref{eq:wbounds}
+subsumes the non-negativity constraint $\forall \; i \in [k] \;.\; \vw^{(i)}$.
+
+\stefano{for $k=1$, setting $\alpha < 1$  renders the problem unbounded, as
+Eq.~\ref{eq:wxconstr} vanishes.}
 

### 2016-01-09 alpha_lt_1_is_unbounded.sh
add a script demonstrating a problem with alpha < 1.

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+# these WORK
+python ijcai16.py synthetic -N 1 -n 10 -m 2 -a 1.0000 -b 0.0 -c 0.0 -i -s 0 --domain-sizes 2,2 --debug
+python ijcai16.py synthetic -N 1 -n 10 -m 2 -a 1.0000 -b 100.0 -c 100.0 -i -s 0 --domain-sizes 2,2 --debug
+
+# these FAIL (the problem is unbounded)
+python ijcai16.py synthetic -N 1 -n 10 -m 2 -a 0.9999 -b 0.0 -c 0.0 -i -s 0 --domain-sizes 2,2 --debug
+python ijcai16.py synthetic -N 1 -n 10 -m 2 -a 0.9999 -b 100.0 -c 100.0 -i -s 0 --domain-sizes 2,2 --debug

### 2016-01-09 ijcai16.py
ijcai16: loop over beta and gamma.

@@ -149,9 +149,9 @@ def run_synthetic():
         "is_deterministic": False,
         "is_indifferent": True,
         "set_size": range(1, 4+1),
-        "alpha": 10.0,
-        "beta": 0.1,
-        "gamma": 0.1,
+        "alpha": 1.0,
+        "beta": (10.0, 1.0, 0.1, 0.0),
+        "gamma": (10.0, 1.0, 0.1, 0.0),
         "multimargin": False,
         "threads": cpu_count(),
         "debug": False,

### 2016-01-10 paper/ijcai16.tex
paper: add MILP formulation, remove (obsolete?) hyperparameter normalization.
paper: draft the method and experiments sections.

--- 2016-01-06 paper/ijcai16.tex
@@
 \usepackage{ijcai16}
+\usepackage{graphicx}
 \usepackage{times}
@@
 \paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
-capital letters $X$ for matrices, and $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
+capital letters $X$ for matrices, and calligraphic capital letters $\calX$ for
+sets. The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$. We
+use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
 
-\paragraph{Non-linear Formulation.}
+\paragraph{Setting.} We assume to have a multi-attribute feature space $\calX$
+of configurations $\vx = (x_1, \ldots, x_n)$ over $n$ features. For the time
+being we will focus on binary features only, i.e. $x_z\in\{0,1\}$, to keep
+matters simple; this restriction will be partially lifted later on.
+\stefano{revisit the previous sentence once cost vars are in place}
+We further assume the set of {\em feasible} configurations can be expressed as
+a conjunction of linear constraints. This is not a severe restriction:
+arbitrary propositional constraints can be trivially encoded in this setting,
+e.g. the Boolean disjunction of two binary variables $x_1 \lor x_2$ can be
+rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations will be
+indicated as $\calX_\text{feasible}$.
 
-Let $k$ be set size and $n$ be the number of examples (answers).
+In adherence to previous work\stefano{add reference}, we model users by means
+of linear utility functions: the user's preferences are represented by a weight
+vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
+$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we will
+require all weights to be {\em non-negative} and {\em bounded}: the individual
+weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
+$[w^\bot_z, w^\top_z]$. These two requirements allow the translation of our core
+optimization problem into a mixed-integer linear problem, as done in the next
+section. \stefano{sentence is a bit bogus}
 
+During learning, the actual weight vector $\vw$ is {\em unknown} to the
+learning system, and must be estimated from pairwise choices elicited from the
+user. The set of examples consists of pairwise comparisons between a more
+preferred configuration $\vy_+$ and a less preferred on $\vy_-$. Each preference
+induces a ranking constraint $\langle \vw, \vy_+ - \vy_- \rangle \ge 0$.
+\stefano{add indifference}
+Details on the actual query selection strategy used will be provided in
+Section~\ref{sec:querystrategy}.
+
+\subsection{Problem Formulation}
+
+\paragraph{Non-linear Formulation.} We now introduce an initial formulation of
+the optimization problem, that we do {\em not} aim at solving directly.
+
+In the set-margin approach, the goal is twofold. First, for any given set size
+$k\geq 1$, we want to find a {\em set} of $k$ weight vectors $\vw^{1},
+\ldots, \vw^{k}$, chosen so that all preferences (examples) are satisfied by
+the largest possible margin. This is achieved by translating all binary
+preferences into inequalities of the form $\langle \vw, \vy_+ - \vy_- \rangle
+\geq M$, and maximizing the margin $M$. Non-separable datasets, which occur in
+practice due to user inconsistency, are handled by introducing slack variables.
+Our reasoning is summarized by the following constraint:
+%
+$$ \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h $$
+%
+where $\varepsilon^{i}_h$ is the penalty incurred by weight vector
+$\vw^{i}$ for violating the ranking of example $h$.
+
+Second, we want to produce a set of $k$ configurations $\vx^{1}, \ldots,
+\vx^{k}$, so that each configuration $\vx^{i}$ is the ``best'' possible
+option when evaluated according to the corresponding $\vw^{i}$. Ideally,
+this requirement would be implemented as:
+%
+$$ \vx^{i} = \argmax_{\vx \in \calX_\text{feasible}} \langle \vw^{i}, \vx^{i} \rangle $$
+%
+However this formulation is highly impractical. We therefore only require 
+each option $\vx^{i}$ to be the best {\em among} the various
+$\vx^{1}, \ldots, \vx^{k}$. This can be accomplished by imposing
+constraints of the form $\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$
+for all $i,j\in[k]$, $i \ne j$.
+This however does not guarantee that the $\vx^{i}$ have a high utility in
+absolute terms. We therefore favor high-quality configurations by introducing
+an additional term $\sum_{i=1}^k \langle \vw^{i} \vx^{i} \rangle$ in the
+objective function. \stefano{sparsity}
+
+A straightforward encoding of the above desiderata leads to the following {\em
+non-linear} optimization problem:
+%
 {\footnotesize
-
 \begin{align}
     \max_{M, W, X}
-        & \;\; M - \frac{\alpha}{k |\calD|} \sum_{i=1}^k \| \veps^{(i)} \|_1 - \frac{\beta}{k} \sum_{i=1}^k \| \vw^{(i)} \|_1 + \frac{\gamma}{k} \sum_{i=1}^k \langle \vw^{(i)}, \vx^{(i)} \rangle
+        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
         \nonumber
@@
     \text{s.t.}
-        & \;\; \forall \; i \in [k] \nonumber, \forall \; h \in [n]
+        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
     \\
-        & \;\; \qquad \langle \vw^{(i)}, \vy^{(h)}_+ - \vy^{(h)}_- \rangle \geq M - \varepsilon^{(i)}_h \nonumber
+        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \label{eq:wyconstr}
+    \\
+        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge M \label{eq:wxconstr}
+    \\
+        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
+    \\
+        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
+    \\
+        & \;\; \forall \; i \in [k] \;.\; \veps^{i} \ge 0 \nonumber
+    \\
+        & \;\; M \ge 0 \nonumber
+\end{align}
+}
+%
+Let us illustrate the above piece by piece. The objective is composed of four
+parts: we want to maximize the shared margin $M$ (first part) and minimize the
+total sum of the ranking errors incurred by each weight vector $\vw^{i}$
+(second part), while at the same time promoting sparsity of the individual
+weight vectors (third part) and quality of the configurations $\vx^{i}$ (last
+part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
+influence of the various components.
+
+The two major constraints are Eq~\ref{eq:wyconstr} and Eq~\ref{eq:wxconstr}.
+The former enforces the correct ranking of the observed user preferences, while
+the latter enforces the correct ranking among the generated configurations.
+\stefano{WRITEME}
+
+Equations \ref{eq:wbounds} and \ref{eq:xbounds} ensure that the obtained weight
+vectors and generated configurations are in fact feasible, while the remaining
+two guarantee the non-negativity of the slack and margin variables.  The
+non-negativity of the $\vw^{i}$'s can be subsumed by Eq~\ref{eq:wbounds}
+by choosing $w^\bot \ge 0$.
+
+Note that we are choosing the options $\vx^{1}, \ldots, \vx^{k}$ and the
+weight vectors $\vw^{1}, \ldots, \vw^{k}$ {\em simultaneously}. Since we
+maximize the margin $M$, the optimizer will prefer a set of outcomes
+$\vx^{i}$ that partition the weight space roughly equally, and corresponding
+utility functions such that each $\vw^{i}$ lies (intuitively) close to the
+centre of each subregion. See Figure~\ref{fig:setmargin} for a depiction of
+this intuition.
+
+\stefano{for $k=1$, setting $\alpha < 1$  renders the problem unbounded, as
+Eq.~\ref{eq:wxconstr} vanishes.}
+
+\begin{figure}[t]
+    \begin{center}
+        \includegraphics[width=16em]{setmargin}
+    \end{center}
+    \caption{\label{fig:setmargin} Intuition.}
+\end{figure}
+
+\paragraph{MILP Formulation.} This initial formulation is problematic, as
+Eq~\ref{eq:wxconstr} involves quadratic terms. Here we show how to reformulate
+the problem as a mixed integer linear program (MILP). Our goal is to replace
+Eq~\ref{eq:wxconstr} with a set of linear constraints through a suitable
+variable transformation.
+
+In order to do so, we introduce a set of fresh variables $p^{i,j}_z$ for every
+$i,j\in[k]$ and $z\in[n]$. Each $p^{i,j}_z$ is enforced to be equal to the
+product $w^{i}_z x^{i}_z$ through additional constraints. Assuming for the
+time being these constraints to be already in place, we rewrite the fourth
+component of the objective function in terms of the new variables as:
+%
+$$ \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z $$
+%
+and, similarly, Eq~\ref{eq:wxconstr} as:
+%
+\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M \label{eq:pxconstr} \]
+%
+The fact that $p^{i,j}_z$ takes the value $w^{i}_z x^{i}_z$ is achieved by
+setting the following additional constraints. We distinguish between two cases:
+(i) $p^{i,i}_z$, and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
+maximizing the margin $M$. Now, due to Eq~\ref{eq:pxconstr}, the optimizer will
+try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
+possible.
+
+(Case i) We add an explicit upper bound:
+%
+$$ p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \} $$
+%
+where $w_\text{max}$ is a sufficiently large constant.
+On one hand, if $x^i_z = 0$ the product $w^i_z x^i_z$ evaluates to $0$, and so does
+$p^{i,i}_z \le w_\text{max} x^{i}_z = 0$. On the other hand, if $x^i_z=1$
+then the product $w^i_z x^i_z$ amounts to $w^i_z$. In this case, the upper
+bound evaluates to $\min \{ w_\text{max}, w^{i}_z \}$. By taking a sufficiently
+large constant $w_\text{max}$ (e.g. $w_\text{max} := \max_z w^\top_z$) the upper bound
+boils down to $w^i_z$. Since $p^{i,i}_z$ is being maximized, it will attain the
+upper bound, and thus evaluate to $w^i_z x^i_z$.
+
+(Case ii) We add an explicit lower bound:
+%
+$$ p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} $$
+%
+If $x^i_z = 1$ the lower bound simplifies to $\max \{ 0, w^{i}_z \} = w^{i}_z$,
+due to the non-negativity of $w^i_z$, as expected. Otherwise, if $x^i_z = 0$
+then the lower bound becomes $\max \{ 0, w^{i}_z - w_\text{max} \}$, where
+the second term is at most $0$. Since $p^{i,j}_z$ is being minimized, in both
+cases it will attain the lower bound, and hence equate to $w^i_z x^i_z$.
+
+Substituting the above MILP constraints into the original optimization problem,
+we obtain:
+{\footnotesize
+\begin{align}
+    \max_{M, W, X}
+        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z
+        \nonumber
+    \\
+    \text{s.t.}
+        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
+    \\
+        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \nonumber
+    \\
+        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M
+    \\
+        & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
+    \\
+        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
     \\
@@
     \\
-        & \;\; \qquad \langle \vw^{(i)}, \vx^{(i)} - \vx^{(j)} \rangle \geq M \label{cnst-ws-m}
+        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
     \\
-        & \;\; \forall \; i \in [k] \nonumber
+        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \nonumber
     \\
-        & \;\; \qquad \vw^\bot \leq \vw^{(i)} \leq \vw^\top \nonumber
+        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \nonumber
     \\
-        & \;\; \forall \; i \in [k] \nonumber
+        & \;\; \forall \; i \in [k] \;.\; \veps^{i}_h \ge 0 \nonumber
     \\
-        & \;\; \qquad \vx^i \in \calX_{\text{feasible}} \nonumber
-    \\
-        & \;\; \forall \; h \in 1, \ldots, n, \forall \; i \in [k] \nonumber
-    \\
-        & \;\; \qquad \varepsilon^{(i)}_h \geq 0 \nonumber
-    \\
-        & \;\; M \geq 0 \nonumber
+        & \;\; M \ge 0 \nonumber
 \end{align}
@@
 
-\paragraph{MILP Formulation.}
+\subsection{Set-wise Preference Elicitation}
 
-WRITEME
+\stefano{WRITEME}
+
+See Algorithm~\ref{alg:setmargin}.
 
 \begin{algorithm}
+{\footnotesize
 \begin{algorithmic}[1]
-    \Procedure{PESetMargin}{$Y, m; \calW, \calX, \valpha, T$}
-        \State $\calA \gets \emptyset$
+    \Procedure{SetMargin}{$k, \alpha, \beta, \gamma, T$}
+        \State $\calQ \gets \emptyset$
         \For{$t = 1, \ldots, T$}
-            \State $\vw^*_1, \ldots, \vw^*_m, \vx^*_1, \ldots, \vx^*_m \gets \text{{\sc Solve}}(Y, \calA, m; \calW, \calX, \valpha)$
-            \For{$\vx^*_i,\vx^*_j \in \{ \vx^*_1, \ldots, \vx^*_m \} \; \text{{\bf s.t.}} \; i < j$}
-                \State $\calA \gets \calA \cup \text{{\sc QueryUser}}(\vx^*_i,\vx^*_j)$
+            \State \{$\vw^{i}, \vx^{i}\}_{i=1}^k \gets \text{{\sc Solve}}(\calQ, k, \alpha, \beta, \gamma)$
+            \For{$\vx^{i},\vx^{j} \in \{ \vx^{1}, \ldots, \vx^{k} \} \; \text{{\bf s.t.}} \; i < j$}
+                \State $\calQ \gets \calQ \cup \text{{\sc QueryUser}}(\vx^{i},\vx^{j})$
             \EndFor
         \EndFor
-        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(Y, \calA, 1; \calW, \calX, \valpha)$
+        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(\calQ, 1, \alpha, \beta, \gamma)$
         \State ${\bf return}\; \vw^*, \vx^*$
@@
 \end{algorithmic}
+}
+\caption{\label{alg:setmargin} The {\sc SetMargin} algorithm. Here $k$ is the
+set size, $\alpha,\beta,\gamma$ are the hyperparameters, and $T$ is the maximum
+number of iterations. The values of $\calX_\text{feasible}$, $\vw^\top$ and
+$\vw^\bot$ are left implicit.}
 \end{algorithm}
+
+\subsection{Query Strategy}
+\label{sec:querystrategy}
+
+\stefano{WRITEME}
+
+\paragraph{Generalization to real variables.} \stefano{WRITEME}
 
@@
 
-WRITEME
+In this section we provide an experimental evaluation of the {\sc SetMargin}
+algorithm. We compare the against two strong Bayesian baselines:
+\cite{guo2010real} and \cite{viappiani2010optimal}.
+
+In all experiments we use an internal cross-validation procedure to update the
+hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5 iterations. The
+hyperparameters are chosen as to maximize the ranking loss over the user
+answers collected so far. $\alpha$ is taken in $\{100, 10, 1\}$,
+while $\beta$ and $\gamma$ are taken in $\{100, 10, 1, 0.1, 0\}$. Note that
+$\beta=0$ implies that the weight vector is not regularized, nor sparse, while
+$\gamma=0$ implies that the absolute quality of the configurations is ignored.
+\stefano{is this desirable?}
+
+\paragraph{Synthetic Dataset.} As a first test, we evaluated the performance
+of the three methods on a synthetic dataset, where all configurations
+
+We ran all methods on $20$ different users. The hidden, true weight vectors
+were the same for all three methods.
+
+\stefano{plot average utility loss at the increase of dataset size}
+
+\stefano{plot average runtime at the increase of dataset size}
+
+\stefano{comment on usefulness of sparse weights}
+
+\stefano{what about an experiment with Horn constraints?}
+
+\paragraph{Constructive Dataset.} \stefano{WRITEME}
+
+- PC dataset without costs
+
+- PC dataset with costs
 

### 2016-01-11 setmargin/solver.py
setmargin/solver: the ps need not be positive.

@@ -248,13 +248,7 @@ class Solver(object):
             for z in range(num_features):
                 model.addConstr(ws[i,z] <= MAX_W_Z)
 
-        # Eq. 18a
-        for i in range(set_size):
-            for j in range(set_size):
-                for z in range(num_features):
-                    model.addConstr(ps[i,j,z] >= 0)
-
-        # Eq. 18b
+        # Eq. 18
         for i in range(set_size):
             for z in range(num_features):
                 model.addConstr(ws[i,z] >= 0)

### 2016-01-11 paper/ijcai16.tex
paper: add a fixme.
paper: draft the PC dataset attributes and constraints.

@@
 try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
-possible.
+possible. \stefano{this is not necessarily the case, $M$ is bounded by the
+$y$'s as well.}
 
@@
 
+The attributes are Manufacturer, CPU, Monitor size, PC type, Memory, and HD
+size. CPU type and CPU frequency as combined in a single variable, e.g. ``AMD
+Athlon @1000" stands for an AMD Athlon CPU clocked at 1 GHz. All other domains
+are left unchanged.
+
+Each attribute configuration is associated a corresponding {\em price}. The
+price of a configuration is the sum of the prices of the individual components.
+
+\begin{figure}
+    {\footnotesize
+    \begin{tabular}{cc}
+        {\bf Rule} & {\bf MILP constraint} \\ \hline\hline
+        & Manufacturer = Compacq $\Rightarrow$ Type = Laptop $\lor$ Type = Desktop \\
+        & Manufacturer = Apple $\Rightarrow$ CPU = PowerPC G3 @266 $\lor$ \ldots \\
+        & Type = Laptop $\Rightarrow$ Memory = 64 $\lor$ \ldots \\
+        & Type = HDSize $\Rightarrow$ HDSize = 20 $\lor$ \ldots \\
+        & Type = Desktop $\Rightarrow$ Monitor = 15 $\lor$ \ldots \\
+    \end{tabular}
+    }
+    \caption{\label{tab:pcconstraints} List of MILP constraints used in the
+    PC experiment.}
+\end{figure}
+
 - PC dataset without costs

### 2016-01-11 ijcai16.py
ijcai16: fix imports.

@@ -2,6 +2,7 @@
 # -*- coding: utf-8 -*-
 
 import sys
+import cPickle as pickle
 import numpy as np
 import matplotlib.pyplot as plt
 from itertools import product

### 2016-01-11 paper/ijcai16.tex
paper: add missing column.

@@
 absolute terms. We therefore favor high-quality configurations by introducing
-an additional term $\sum_{i=1}^k \langle \vw^{i} \vx^{i} \rangle$ in the
+an additional term $\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ in the
 objective function. \stefano{sparsity}

### 2016-01-11 setmargin/setmargin.py
ijcai16: implement proper cross-validation support. XXX it diverges!

@@ -53,8 +53,11 @@ def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
     user_w_norm = np.linalg.norm(user.w.ravel())
 
     do_crossval = alphas == "auto"
-    if crossval_set_size is None:
-        crossval_set_size = set_size
+    if do_crossval:
+        # XXX far from ideal
+        alphas = (1.0, 1.0, 1.0)
+        if crossval_set_size is None:
+            crossval_set_size = set_size
 
     answers, info, old_best_item = [], [], None
     for t in range(num_iterations):
@@ -69,25 +72,31 @@ def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
             """).format(t, num_iterations, pformat(answers))
 
         # Crossvalidate the hyperparameters if required
-        if do_crossval and t % crossval_interval == 0:
+        if do_crossval and t % crossval_interval == 0 and t > 0:
             loss_alphas = []
             for alphas in it.product(ALPHAS, BETAS, GAMMAS):
                 try:
-                    ws, xs = solver.compute_setmargin(dataset, answers,
-                                                      crossval_set_size,
-                                                      alphas)
+                    ws, _ = solver.compute_setmargin(dataset, answers,
+                                                     crossval_set_size, alphas)
                 except RuntimeError:
                     continue
-                mean_x = xs.mean(axis=0)
-                loss = np.dot(user.w.ravel(), best_item - mean_x) / user_w_norm
+
+                xis = np.array([xi for xi, _, _ in answers])
+                xjs = np.array([xj for _, xj, _ in answers])
+                ys = np.array([sign for _, _, sign in answers])
+
+                diff = np.abs(ys - np.sign(np.dot(ws, (xis - xjs).T)))
+                diff[diff > 0] = 1
+                loss = diff.sum(axis=1).mean()
                 loss_alphas.append((loss, alphas))
+
             assert len(loss_alphas) > 0
             alphas = sorted(loss_alphas)[0][1]
 
             if debug:
                 print "CROSS VALIDATION --> new alphas =", alphas
-                for loss, alpha in loss_alphas:
-                    print alpha, ": loss =", loss
+                for loss, temp in sorted(loss_alphas):
+                    print temp, ": loss =", loss
 
         old_time = time.time()
 

### 2016-01-11 paper/ijcai16.tex
paper: add rough sketch of the experimental section.

@@
 \section{Experiments}
+\label{sec:experiments}
 
@@
 \cite{guo2010real} and \cite{viappiani2010optimal}.
+
+\stefano{describe the baselines}
+
+We implemented the {\sc SetMargin} algorithm using Python, while the MILP
+problems were solved with Gurobi 6.5.0. The full source of the experiments is
+available at \stefano{add URL}.
 
@@
 
-\paragraph{Synthetic Dataset.} As a first test, we evaluated the performance
-of the three methods on a synthetic dataset, where all configurations
+\paragraph{Synthetic Dataset.} In order to evaluate the convergence and
+scalability properties of the proposed method, following~\cite{guo2010real} and
+\cite{viappiani2010optimal}, we evaluated it on a number of increasingly larger
+synthetic datasets. More specifically, we defined $7$ synthetic dataset over
+$n=2,\ldots,7$ attributes, with each attribute taking one of $n$ possible
+values.
+In other words, for $n=2$ we have that the feasible space
+$\calX_\text{feasible}$ is $[2] \times [2]$, while for $n=3$ we have
+$\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on.
+The cardinality of $\calX_\text{feasible}$ is $n^n$, and grows
+(super-)exponentially with $n$. For $n=3$, the dataset is comparable to
+the synthetic ones used in~\cite{guo2010real} and~\cite{viappiani2010optimal},
+while for larger $n$ it grows much larger than those typically used in the
+Bayesian preference elicitation literature, and as such represents a good
+test bed for evaluating the scalability of the various methods.
 
-We ran all methods on $20$ different users. The hidden, true weight vectors
-were the same for all three methods.
+In order to work, the competitors require the dataset to represented
+explicitly, so we had to ground the $7$ datasets. This is not the case for {\sc
+SetMargin}, in which case the set of feasible configurations
+$\calX_\text{feasible}$ can be encoded {\em implicitly} through simple linear
+integer constraints. All of the methods can deal with binary vectors only; we
+translated the integer-valued configurations to such format using a simple
+one-hot encoding. For {\sc SetMargin}, this requires to encode additional
+XOR constraints. \stefano{explain}
+
+The three methods were evaluated on user weight vectors sampled from two
+distinct distributions: a normal distribution with mean $25$ and
+variance $\frac{25}{3}$, and from a uniform distribution over $[0,
+100]$. This procedure mimics closely~\cite{guo2010real}. In order to assess our
+method in a truly sparse setting, we obtained a set of sparse normal and
+uniform weight vectors by randomly deleting $80\%$ of the non-zero entries from
+the already sampled vectors. We name the four settings {\sc normal}, {\sc
+uniform}, {\sc sparse\_normal} and {\sc sparse\_uniform}. We sampled a total
+of $100$ weight vectors for each setting. The three methods were evaluated
+on the very same weight vectors.
 
@@
 
-\stefano{what about an experiment with Horn constraints?}
-
-\paragraph{Constructive Dataset.} \stefano{WRITEME}
+\paragraph{Constructive Dataset.} Next, we tested {\sc SetMargin} on a truly
+constructive setting, which can not be readily handled by non-constructive
+methods.
 
@@
 are left unchanged.
+
+The configurations constraints:
+%
+$$ x_i \Leftarrow x_{j_1} \lor \ldots \lor x_{j_n} $$
+%
+have been translated into MILP constraints of the form:
+%
+$$ (1 - x_i) + x_{j_1} + \ldots + x_{j_n} \ge 1 $$
 

### 2016-01-12 paper/ijcai16.tex
paper: fix typeface.
paper: refine the experiments section.

--- 2016-01-11 paper/ijcai16.tex
@@
 non-negativity of the $\vw^{i}$'s can be subsumed by Eq~\ref{eq:wbounds}
-by choosing $w^\bot \ge 0$.
+by choosing $\vw^\bot \ge 0$.
 
@@
 
-\paragraph{Synthetic Dataset.} As a first test, we evaluated the performance
-of the three methods on a synthetic dataset, where all configurations
+\paragraph{Synthetic Dataset.} Following the experimental protocol
+in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
+we evaluate the behavior of the proposed method in a challenging artificial
+setting. We developed seven synthetic datasets: each dataset involves
+$n=2,\ldots,7$ attributes, where each attribute takes one of $n$ possible
+values. More explicitly, for $n=2$ the synthetic dataset is
+$\calX_\text{feasible} = [2] \times [2]$, for $n=3$ to $\calX_\text{feasible} =
+[3] \times [3] \times [3]$, and so on. The cardinality of
+$\calX_\text{feasible}$ is $n^n$, and grows (super-)exponentially with $n$. For
+$n=3$, the dataset is comparable to the synthetic one used
+in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
+of size of 2, 2 and 5, respectively, for a total of 20 feasible
+configurations), while for larger $n$ it grows much larger than those typically
+used in the Bayesian preference elicitation literature, and as such represents
+a good test bed for comparing the scalability of the various methods.
 
-We ran all methods on $20$ different users. The hidden, true weight vectors
-were the same for all three methods.
+The integer attributes were binarized using a standard one-hot encoding. The
+feasible configuration space was encoded in {\sc SetMargin} through appropriate
+MILP constraints (namely, mutual exclusivity of all binary variables
+corresponding to each attribute, to enforce the one-hot encoding), while the
+other methods required all datasets to be explicitly grounded.
+
+The user weight vectors were sampled from four different distributions to
+simulated different kinds of users. The first two were chosen as
+in~\cite{guo2010real}: a uniform distribution over $[0, 100]$ and a normal
+distribution with mean $25$ and variance $\frac{25}{3}$. In order to assess our
+method in a truly sparse setting, we obtained a set of sparse normal and
+uniform weight vectors by randomly deleting $80\%$ of the non-zero entries from
+the already sampled vectors. We will refer to the four settings as {\sc
+normal}, {\sc uniform}, {\sc sparse\_normal} and {\sc sparse\_uniform}. We
+sampled a total of $100$ weight vectors for each setting. All methods were
+evaluated on the very same weight vectors.
 

### 2016-01-13 setmargin/solver.py
setmargin/solver: set default # of threads to # of cores.

@@ -34,10 +34,10 @@ class Solver(object):
         again.
 
     :param multimargin: whether to use two distinct margins. (default: ``False``)
-    :param threads: how many threads Gurobi is allowed to use. (default: 1)
+    :param threads: how many threads Gurobi is allowed to use. (default: ``None``)
     :param debug: whether to spew debug output. (default: ``False``)
     """
-    def __init__(self, multimargin=False, threads=1, debug=False):
+    def __init__(self, multimargin=False, threads=None, debug=False):
         self._multimargin = multimargin
         self._threads = threads
         self._debug = debug
@@ -146,7 +146,8 @@ class Solver(object):
 
         model = grb.Model("facility")
         model.params.Seed = 0
-        model.params.Threads = self._threads
+        if self._threads is not None:
+            model.params.Threads = self._threads
         model.params.OutputFlag = 0
 
         # Declare the variables

### 2016-01-13 ijcai16.py
ijcai16: make the results path more uniform.

@@ -270,9 +270,12 @@ def run_from_command_line():
     if args.debug:
         print dataset
 
+    dataset_name = args.dataset
+    if args.dataset == "synthetic":
+        dataset_name += "_" + args.domain_sizes
+
     infos = solve(dataset, config)
-    dump_and_draw("{}_{}".format(args.dataset, ",".join(map(str, args.domain_sizes))),
-                  config, infos)
+    dump_and_draw(dataset_name, config, infos)
 
 if __name__ == "__main__":
     np.seterr(all="raise")

### 2016-01-13 draw.py
draw: add a simple stand-alone plotting script.

@@ -1,0 +1,33 @@
+#!/usr/bin/env python2
+
+import sys, os
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.font_manager import FontProperties
+import itertools as it
+from glob import glob
+
+fontP = FontProperties()
+fontP.set_size('small')
+
+fig, ax = plt.subplots(1, 1)
+ax.set_xlabel("Number of queries")
+ax.set_ylabel("Average loss over trials")
+
+paths = sorted(glob(os.path.join(sys.argv[1], "results_*_loss_matrix.txt")))
+
+for path in paths:
+    parts = os.path.basename(path).split("__")
+
+    loss_matrix = np.loadtxt(path)
+    num_trials, max_queries = loss_matrix.shape
+
+    loss_means = np.mean(loss_matrix, axis=0)
+
+    ax.set_ylim([0.0, max(0.5, max(loss_means) + 0.1)])
+    ax.errorbar(np.arange(1, max_queries + 1), loss_means, label=os.path.basename(path))
+
+lgd = ax.legend(prop=fontP, loc="upper center", bbox_to_anchor=(0.5, -0.1))
+
+path = os.path.join(sys.argv[1], "plot.svg")
+fig.savefig(path, bbox_extra_artists=(lgd,), bbox_inches="tight")

### 2016-01-13 setmargin/solver.py
setmargin/solver: no need to print weights that we KNOW are all zeros.

@@ -346,7 +346,7 @@ class Solver(object):
                         output_scores, output_slacks, output_margins)
 
         if any(np.linalg.norm(w) == 0 for w in output_ws):
-            print "Warning: null weight vector found:\n{}".format(output_ws)
+            print "Warning: null weight vector found!"
 
         assert all(dataset.is_item_valid(x) for x in output_xs)
 

### 2016-01-13 ijcai16.py
ijcai16: implement pre-crossvalidation.

@@ -5,6 +5,7 @@ import sys
 import cPickle as pickle
 import numpy as np
 import matplotlib.pyplot as plt
+from sklearn.cross_validation import KFold
 from itertools import product
 from os.path import join
 from textwrap import dedent
@@ -12,6 +13,12 @@ from pprint import pformat
 from multiprocessing import cpu_count
 
 import setmargin
+
+ALL_ALPHAS = list(product(
+         [100.0, 20.0, 15.0, 12.0, 11.0, 10.0, 9.0, 8.0, 7.5, 5.0, 1.0],
+         [20.0, 10.0, 5.0, 1.0, 0.5, 0.1, 0.05, 0.01],
+         [20.0, 10.0, 5.0, 1.0, 0.5, 0.1, 0.05, 0.01],
+))
 
 class Grid(object):
     def __init__(self, d):
@@ -120,16 +127,69 @@ def dump_and_draw(dataset_name, config, infos):
     del fig
     del ax
 
+def precrossvalidate(dataset, config, solver, users):
+    """Use the last 20 users for parameter selection."""
+    target_users = users[len(users) - 20:]
+    assert len(target_users) == 20
+
+    loss_alphas = []
+    for alphas in ALL_ALPHAS:
+
+        print "pre-crossvalidating", alphas
+
+        kfold = KFold(len(target_users), n_folds=5)
+
+        losses = []
+        for train_set, test_set in kfold:
+
+            for i in train_set:
+                infos = setmargin.run(dataset, target_users[i], solver,
+                                      config.num_iterations, config.set_size,
+                                      alphas, tol=config.tol, debug=config.debug)
+                # XXX we only consider the true utility loss at the last
+                # iteration
+                losses.append(infos[-1][1])
+
+        loss_alphas.append((sum(losses) / len(losses), alphas))
+
+    loss_alphas = sorted(loss_alphas)
+    best_alphas = loss_alphas[0][1]
+
+    print "pre-crossvalidation: best alphas = ", alphas
+    for loss, alphas in loss_alphas:
+        print alphas, ":", loss
+
+    return best_alphas
+
 def solve(dataset, config, ws=None):
     rng = np.random.RandomState(config.seed)
 
+    solver = setmargin.Solver(multimargin=config.multimargin,
+                              threads=config.threads, debug=config.debug)
+
+    num_users = config.num_trials if ws is None else ws.shape[0]
+
+    users = []
+    for i in range(num_users):
+        w = None if ws is None else ws[i].reshape(1, -1)
[…]

### 2016-01-13 ijcai16.py
ijcai16: do not plot std.dev., it will be redrawn later and often confuses matplotlib.

@@ -112,7 +112,7 @@ def dump_and_draw(dataset_name, config, infos):
     ax.set_xlabel("Number of queries")
     ax.set_ylabel("Average loss over trials")
     ax.set_ylim([0.0, max(0.5, max(loss_means) + max(loss_stddevs) + 0.1)])
-    ax.errorbar(np.arange(1, max_queries + 1), loss_means, yerr=loss_stddevs)
+    ax.errorbar(np.arange(1, max_queries + 1), loss_means)
     fig.savefig(loss_svg_path, bbox_inches="tight")
     del fig
     del ax
@@ -122,7 +122,7 @@ def dump_and_draw(dataset_name, config, infos):
     ax.set_xlabel("Number of queries")
     ax.set_ylabel("Average time over trials")
     ax.set_ylim([0.0, max(0.5, max(time_means) + max(time_stddevs) + 0.1)])
-    ax.errorbar(np.arange(1, max_queries + 1), time_means, yerr=time_stddevs)
+    ax.errorbar(np.arange(1, max_queries + 1), time_means)
     fig.savefig(time_svg_path, bbox_inches="tight")
     del fig
     del ax

### 2016-01-13 ijcai16.py
ijcai16: add a small test for infos_to_matrices(), just to be sure.

@@ -74,13 +74,7 @@ def get_result_paths(dataset_name, config):
     path4 = "results_{}_avgtime.svg".format(basename)
     return path0, path1, path2, path3, path4
 
-def dump_and_draw(dataset_name, config, infos):
-    infos_path, loss_matrix_path, time_matrix_path, loss_svg_path, time_svg_path = \
-        get_result_paths(dataset_name, config)
-
-    with open(infos_path, "wb") as fp:
-        pickle.dump(infos, fp)
-
+def infos_to_matrices(infos):
     num_trials = len(infos)
     max_queries = max([sum(n for n, _, _ in info) for info in infos])
 
@@ -98,6 +92,30 @@ def dump_and_draw(dataset_name, config, infos):
             base += num_queries
             prev_loss = loss
 
+    return loss_matrix, time_matrix
+
+# info = [
+#     (2, 1.0, 1.0),
+#     (2, 0.5, 1.0),
+#     (2, 0.0, 1.0),
+# ]
+# infos = [info, info]
+# print infos
+# lm, tm = infos_to_matrices(infos)
+# print lm.shape
+# print lm
+# print tm.shape
+# print tm
+# quit()
+
+def dump_and_draw(dataset_name, config, infos):
+    infos_path, loss_matrix_path, time_matrix_path, loss_svg_path, time_svg_path = \
+        get_result_paths(dataset_name, config)
+
+    with open(infos_path, "wb") as fp:
+        pickle.dump(infos, fp)
+
+    loss_matrix, time_matrix = infos_to_matrices(infos)
     np.savetxt(loss_matrix_path, loss_matrix)
     np.savetxt(time_matrix_path, time_matrix)
 

### 2016-01-13 ijcai16.py
ijcai16: fix unbound vars.

@@ -92,7 +92,7 @@ def infos_to_matrices(infos):
             base += num_queries
             prev_loss = loss
 
-    return loss_matrix, time_matrix
+    return num_trials, max_queries, loss_matrix, time_matrix
 
 # info = [
 #     (2, 1.0, 1.0),
@@ -115,7 +115,8 @@ def dump_and_draw(dataset_name, config, infos):
     with open(infos_path, "wb") as fp:
         pickle.dump(infos, fp)
 
-    loss_matrix, time_matrix = infos_to_matrices(infos)
+    num_trials, max_queries, loss_matrix, time_matrix = \
+        infos_to_matrices(infos)
     np.savetxt(loss_matrix_path, loss_matrix)
     np.savetxt(time_matrix_path, time_matrix)
 

### 2016-01-13 paper/ijcai16.tex
paper: update methods section.

@@
 
-\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
-capital letters $X$ for matrices, and calligraphic capital letters $\calX$ for
-sets. The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$. We
-use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
+\stefano{Paolo, feel free to improve the text and especially the citations
+as you see fit.}
+
+\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors, and
+calligraphic capital letters $\calX$ for sets. We frequently abbreviate the set
+$\{ x^i \}_{i=1}^n$ as $\{ x^i \}$ whenever the range of the index $i$ is clear
+from the context, and use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
+The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$.
 
@@
 being we will focus on binary features only, i.e. $x_z\in\{0,1\}$, to keep
-matters simple; this restriction will be partially lifted later on.
-\stefano{revisit the previous sentence once cost vars are in place}
-We further assume the set of {\em feasible} configurations can be expressed as
+matters simple. This is a common choice for preference elicitation methods,
+e.g.~\cite{guo2010real,viappiani2010optimal}; discrete features are handled by
+means of a simple one-hot encoding. Partial support for continuous features
+will be discussed later on.
+We further assume that the set of {\em feasible} configurations is expressed as
 a conjunction of linear constraints. This is not a severe restriction:
 arbitrary propositional constraints can be trivially encoded in this setting,
-e.g. the Boolean disjunction of two binary variables $x_1 \lor x_2$ can be
+e.g. the ``Boolean disjunction'' of two binary variables $x_1 \lor x_2$ can be
 rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations will be
@@
 
-In adherence to previous work\stefano{add reference}, we model users by means
+In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
 of linear utility functions: the user's preferences are represented by a weight
@@
 weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
-$[w^\bot_z, w^\top_z]$. These two requirements allow the translation of our core
-optimization problem into a mixed-integer linear problem, as done in the next
-section. \stefano{sentence is a bit bogus}
+$[w^\bot_z, w^\top_z]$. Both requirements are quite natural, and enable the
+translation of our core optimization problem into a mixed-integer linear
+problem (as done in Section~\ref{sec:formulation}).
 
@@
 \subsection{Problem Formulation}
+\label{sec:formulation}
 
@@
 A straightforward encoding of the above desiderata leads to the following {\em
-non-linear} optimization problem:
+non-linear} optimization problem over the variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$:
 %
@@
 \begin{align}
-    \max_{M, W, X}
+    \max
         & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
@@
 \begin{align}
-    \max_{M, W, X}
+    \max
         & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z

### 2016-01-13 ijcai16.py
ijcai: fix default # of threads (for the last time hopefully).

@@ -315,7 +315,7 @@ def run_from_command_line():
 
     parser.add_argument("-s", "--seed", type=int, default=None,
                         help="RNG seed")
-    parser.add_argument("--threads", type=int, default=1,
+    parser.add_argument("--threads", type=int, default=None,
                         help="Max number of threads to user")
     parser.add_argument("--debug", action="store_true",
                         help="Enable debug spew")

### 2016-01-13 paper/ijcai16.tex
paper: refine methods and experiments a bit, still unfinished.

@@
 will be discussed later on.
+
 We further assume that the set of {\em feasible} configurations is expressed as
@@
 weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
-$[w^\bot_z, w^\top_z]$. Both requirements are quite natural, and enable the
-translation of our core optimization problem into a mixed-integer linear
-problem (as done in Section~\ref{sec:formulation}).
+$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
+natural, and enable the translation of our core optimization problem into a
+mixed-integer linear problem (as done in Section~\ref{sec:formulation}).
 
@@
 induces a ranking constraint $\langle \vw, \vy_+ - \vy_- \rangle \ge 0$.
+Details on the actual query selection strategy will be provided in
+Section~\ref{sec:querystrategy}.
 \stefano{add indifference}
-Details on the actual query selection strategy used will be provided in
-Section~\ref{sec:querystrategy}.
 
@@
 the largest possible margin. This is achieved by translating all binary
-preferences into inequalities of the form $\langle \vw, \vy_+ - \vy_- \rangle
-\geq M$, and maximizing the margin $M$. Non-separable datasets, which occur in
-practice due to user inconsistency, are handled by introducing slack variables.
-Our reasoning is summarized by the following constraint:
+preferences into inequalities of the form $\langle \vw^i, \vy^h_+ - \vy^h_- \rangle
+\geq M$, where $h$ ranges over the examples, and maximizing the margin $M$.
+Non-separable datasets, which occur in practice due to user inconsistency, are
+handled by introducing slack variables. The updated constraints look as follows:
 %
@@
 %
-where $\varepsilon^{i}_h$ is the penalty incurred by weight vector
-$\vw^{i}$ for violating the ranking of example $h$.
+where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
+for violating the ranking of example $h$.
 
@@
 However this formulation is highly impractical. We therefore only require 
-each option $\vx^{i}$ to be the best {\em among} the various
-$\vx^{1}, \ldots, \vx^{k}$. This can be accomplished by imposing
+each option $\vx^{i}$ to be the best {\em among} the configurations
+$\{ \vx^i \}$. This can be accomplished by imposing
 constraints of the form $\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$
 for all $i,j\in[k]$, $i \ne j$.
-This however does not guarantee that the $\vx^{i}$ have a high utility in
-absolute terms. We therefore favor high-quality configurations by introducing
-an additional term $\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ in the
-objective function. \stefano{sparsity}
+This however does not guarantee that the produced $\{ \vx^{i} \}$ have a high
+{\em absolute} utility. We therefore favor high-quality configurations by
+introducing an additional term $\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$
+in the objective function.
 
@@
 Let us illustrate the above piece by piece. The objective is composed of four
-parts: we want to maximize the shared margin $M$ (first part) and minimize the
-total sum of the ranking errors incurred by each weight vector $\vw^{i}$
-(second part), while at the same time promoting sparsity of the individual
-weight vectors (third part) and quality of the configurations $\vx^{i}$ (last
+parts: we maximize the shared margin $M$ (first part) and minimize the total
+sum of the ranking errors $\veps^i$ incurred by each weight vector $\vw^{i}$
+(second part), while at the same time regularizing the magnitude of the
+weights (third part) and quality of the configurations $\{ \vx^{i} \}$ (last
 part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
@@
 
-The two major constraints are Eq~\ref{eq:wyconstr} and Eq~\ref{eq:wxconstr}.
-The former enforces the correct ranking of the observed user preferences, while
-the latter enforces the correct ranking among the generated configurations.
-\stefano{WRITEME}
+The third part requires requires some further explanation. We choose an
+$\ell_1$ regularizer in order to facilitate learning {\em sparse} weight
+vectors: it is reasonable to assume that a user will have strong opinions
+over some attributes, but will be indifferent over most of them.
+\stefano{Andrea, improve the above please}
 
-Equations \ref{eq:wbounds} and \ref{eq:xbounds} ensure that the obtained weight
-vectors and generated configurations are in fact feasible, while the remaining
-two guarantee the non-negativity of the slack and margin variables.  The
-non-negativity of the $\vw^{i}$'s can be subsumed by Eq~\ref{eq:wbounds}
-by choosing $\vw^\bot \ge 0$.
+The two core constraints as Equations~\ref{eq:wyconstr} and \ref{eq:wxconstr},
+which are taken directly from the discussion above. The former enforces the
+correct ranking of the observed user preferences, while the latter ensures that
+the generated configurations are ranked correctly with respect to the
+corresponding weight vectors. Equations \ref{eq:wbounds} and \ref{eq:xbounds}
+ensure that the weights and configurations are in fact feasible, while the
+remaining two guarantee the non-negativity of the slack and margin variables.
+Since we require $\vw^\bot \ge 0$, Eq.~\ref{eq:wbounds} also enforces the
+weights to be non-negative.
 
-Note that we are choosing the options $\vx^{1}, \ldots, \vx^{k}$ and the
-weight vectors $\vw^{1}, \ldots, \vw^{k}$ {\em simultaneously}. Since we
-maximize the margin $M$, the optimizer will prefer a set of outcomes
-$\vx^{i}$ that partition the weight space roughly equally, and corresponding
-utility functions such that each $\vw^{i}$ lies (intuitively) close to the
-centre of each subregion. See Figure~\ref{fig:setmargin} for a depiction of
-this intuition.
+Note that we are choosing the options $\{ \vx^i \}$ and the weight vectors $\{
+\vw^i \}$ {\em simultaneously}. Since we maximize the margin $M$, the optimizer
+will prefer a set of outcomes $\{ \vx^{i} \}$ that partition the weight space
+roughly equally, and corresponding utility functions such that each $\vw^{i}$
+lies (intuitively) close to the centre of each subregion. See
+Figure~\ref{fig:setmargin} for a depiction of this intuition.
 
@@
 In order to do so, we introduce a set of fresh variables $p^{i,j}_z$ for every
-$i,j\in[k]$ and $z\in[n]$. Each $p^{i,j}_z$ is enforced to be equal to the
-product $w^{i}_z x^{i}_z$ through additional constraints. Assuming for the
-time being these constraints to be already in place, we rewrite the fourth
+$i,j\in[k]$ and $z\in[n]$. Assuming for the time being that the new variables
+do satisfy the equation $p^{i,j}_z = w^i_z x^i_z$, we rewrite the fourth
 component of the objective function in terms of the new variables as:
@@
 %
-The fact that $p^{i,j}_z$ takes the value $w^{i}_z x^{i}_z$ is achieved by
+The fact that $p^{i,j}_z = w^{i}_z x^{i}_z$ is achieved by
 setting the following additional constraints. We distinguish between two cases:
@@
 On one hand, if $x^i_z = 0$ the product $w^i_z x^i_z$ evaluates to $0$, and so does
-$p^{i,i}_z \le w_\text{max} x^{i}_z = 0$. On the other hand, if $x^i_z=1$
-then the product $w^i_z x^i_z$ amounts to $w^i_z$. In this case, the upper
-bound evaluates to $\min \{ w_\text{max}, w^{i}_z \}$. By taking a sufficiently
-large constant $w_\text{max}$ (e.g. $w_\text{max} := \max_z w^\top_z$) the upper bound
-boils down to $w^i_z$. Since $p^{i,i}_z$ is being maximized, it will attain the
-upper bound, and thus evaluate to $w^i_z x^i_z$.
+the upper bound $w_\text{max} x^{i}_z = 0$. On the other hand, if $x^i_z=1$
+then the product $w^i_z x^i_z$ amounts to $w^i_z$, while the upper
+bound reduces to $\min \{ w_\text{max}, w^{i}_z \}$. By taking a sufficiently
+large constant $w_\text{max}$ (e.g. $w_\text{max} := \max_z w^\top_z$) the
+upper bound simplifies to $w^i_z$. Since $p^{i,i}_z$ is being maximized, in
+both cases it will attain the upper bound, and thus satisfy $p^{i,j}_z = w^i_z x^i_z$.
 
@@
 If $x^i_z = 1$ the lower bound simplifies to $\max \{ 0, w^{i}_z \} = w^{i}_z$,
-due to the non-negativity of $w^i_z$, as expected. Otherwise, if $x^i_z = 0$
+due to the non-negativity of $w^i_z$. Otherwise, if $x^i_z = 0$
 then the lower bound becomes $\max \{ 0, w^{i}_z - w_\text{max} \}$, where
 the second term is at most $0$. Since $p^{i,j}_z$ is being minimized, in both
-cases it will attain the lower bound, and hence equate to $w^i_z x^i_z$.
+cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z x^i_z$.
 
-Substituting the above MILP constraints into the original optimization problem,
-we obtain:
+Substituting the above MILP constraints into the original non-linear
+formulation, we obtain the following mixed-integer linear problem:
+%
 {\footnotesize
@@
 \end{align}
-
 }
+%
+which can be solved by any suitable MILP solver.
 
@@
 
-\paragraph{Generalization to real variables.} \stefano{WRITEME}
+The simplest alternative is to query the user $\frac{1}{2}n(n-1)$ pairs.
+
+\paragraph{Generalization to real variables.}
+
+\stefano{WRITEME}
 
@@
 
-In this section we provide an experimental evaluation of the {\sc SetMargin}
-algorithm. We compare the against two strong Bayesian baselines:
+In this section we provide extensive experimental evaluation of the {\sc
+SetMargin} algorithm. We implemented the {\sc SetMargin} algorithm using
+Python, and leverage Gurobi 6.5.0 for solving the MILP problem. Both the 
+{\sc SetMargin} source code and the full experimental setup are available at
+\stefano{add URL}.
+
+We compare {\sc SetMargin} against two strong Bayesian baselines:
 \cite{guo2010real} and \cite{viappiani2010optimal}.
+\stefano{Paolo, could you describe the baselines?}
 
-\stefano{describe the baselines}
-
-We implemented the {\sc SetMargin} algorithm using Python, while the MILP
-problems were solved with Gurobi 6.5.0. The full source of the experiments is
-available at \stefano{add URL}.
+We adopted the {\em indifference-augmented} Bradley-Torr user response model
+introduced in~\cite{guo2010real}. When queried about a pair of configurations,
+the user replies with the true utility of the two items. According to the
+classical Bradley-Torr model, which does not attempt to model {\em indifference},
+the proabability that a user $\vw$ will prefer configuration $\vx^i$ over
+the alternative $\vx^j$, barring indifference, amounts to:
+%
+$$ P(\vx^i > \vx^j \;|\; \vw, \vx^i \not\sim \vx^j) = (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
+%
+In other words, the probability grows with the difference between the utilities
+of the two items. Support for indifference has been introduced
+in~\cite{guo2010real} by postulating that the probability of the user being
+indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
+the two utilities are, as in:
+%
+$$ P(\vx^i \sim \vx^j \;|\; \vw) = \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
+%
+The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
+see the latter for more details
+\stefano{The above does not satisfy me at all.}
 
@@
 a good test bed for comparing the scalability of the various methods.
+The feasible configuration space was encoded in {\sc SetMargin} through
+appropriate MILP constraints, while the other methods required all datasets to
+be explicitly grounded.
 
-The integer attributes were binarized using a standard one-hot encoding. The
-feasible configuration space was encoded in {\sc SetMargin} through appropriate
-MILP constraints (namely, mutual exclusivity of all binary variables
-corresponding to each attribute, to enforce the one-hot encoding), while the
-other methods required all datasets to be explicitly grounded.
-
-The user weight vectors were sampled from four different distributions to
-simulated different kinds of users. The first two were chosen as
-in~\cite{guo2010real}: a uniform distribution over $[0, 100]$ and a normal
-distribution with mean $25$ and variance $\frac{25}{3}$. In order to assess our
-method in a truly sparse setting, we obtained a set of sparse normal and
-uniform weight vectors by randomly deleting $80\%$ of the non-zero entries from
-the already sampled vectors. We will refer to the four settings as {\sc
-normal}, {\sc uniform}, {\sc sparse\_normal} and {\sc sparse\_uniform}. We
-sampled a total of $100$ weight vectors for each setting. All methods were
-evaluated on the very same weight vectors.
+Users were simulated by drawing random utility vectors from four different
+distributions. The first two mimic those used in~\cite{guo2010real}: (1) a
+uniform distribution over $[1, 100]$ for each individual weight, and (2) a
+normal distribution with mean $25$ and covariance $\frac{25}{3}$\footnote{This
+procedure differs from the one reported in~\cite{guo2010real}; we opted for
+following the procedure implemented in their code. \stefano{more diplomacy needed}}.
+We further produced two novel {\em sparse} sets of weight vectors by deleting
+$80\%$ uniformly at random \stefano{Paolo, double check please} the entries
+from the vectors sampled in the uniform and normal settings. We sampled $100$
+vectors for each setting. All methods were evaluated on the very same weight
+vectors.
 

### 2016-01-13 setmargin/setmargin.py
setmargin/setmargin: warn when identical answers are added to the answer set.

@@ -140,6 +140,14 @@ def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
 
         # Update the user answers
         new_answers, num_queries = user.query_set(ws, xs, old_best_item)
+        num_identical_answers = 0
+        for xi, xj, sx in new_answers:
+            for zi, zj, sz in answers:
+                if (xi == zi).all() and (xj == zj).all():
+                    num_identical_answers += 1
+        if num_identical_answers > 0:
+            print "Warning: {} identical (up to sign) answers added!" \
+                      .format(num_identical_answers)
         answers.extend(new_answers)
         old_best_item = xs[0] if xs.shape[0] == 1 else None
 

### 2016-01-13 ijcai16.py
ijcai16,setmargin/datasets: add support for datasets with costs.

@@ -338,10 +338,12 @@ def run_from_command_line():
     domain_sizes = map(int, [ds for ds in args.domain_sizes.split(",") if len(ds)])
     if args.dataset == "synthetic":
         dataset = setmargin.SyntheticDataset(domain_sizes)
-    elif args.dataset == "random_constraints":
-        dataset = setmargin.RandomDataset(domain_sizes, rng=rng)
+    elif args.dataset == "debug_constraint":
+        dataset = setmargin.DebugConstraintDataset(domain_sizes, rng=rng)
+    elif args.dataset == "debug_cost":
+        dataset = setmargin.DebugCostDataset(domain_sizes, rng=rng)
     elif args.dataset == "pc":
-        dataset = setmargin.LiftedPCDataset()
+        dataset = setmargin.PCDataset()
     else:
         raise ValueError("invalid dataset.")
     if args.debug:

### 2016-01-13 setmargin/datasets.py

@@ -17,16 +17,31 @@ class Dataset(object):
     :param items: array of items as one-hot row vectors.
     :param x_constraints: constraints on the item configurations.
     """
-    def __init__(self, domain_sizes, items, x_constraints):
+    def __init__(self, domain_sizes, items, costs, x_constraints):
         self.domain_sizes = domain_sizes
         self.items = items
+        self.costs = costs
         self.x_constraints = x_constraints
 
     def __str__(self):
-        return "Dataset(domain_sizes={} len(items)={} constrs={}" \
-                   .format(self.domain_sizes,
-                           len(self.items) if self.items is not None else 0,
-                           self.x_constraints)
+        return "Dataset(domain_sizes={} bools={} reals={} constrs={}" \
+                   .format(self.domain_sizes, self.num_bools(),
+                           self.num_reals(), self.x_constraints)
+
+    def num_bools(self):
+        return sum(self.domain_sizes)
+
+    def num_reals(self):
+        return 0 if self.costs is None else self.costs.shape[0]
+
+    def num_vars(self):
+        return self.num_bools() + self.num_reals()
+
+    def get_domain_ranges(self):
+        base = 0
+        for size in self.domain_sizes:
+            yield base, base + size
+            base += size
 
     def _dom_var_to_bit(self, j, z):
         assert 0 <= j <= len(self.domain_sizes)
@@ -44,6 +59,10 @@ class Dataset(object):
         for zs_in_domain in get_zs_in_domains(self.domain_sizes):
             if sum(x[z] for z in zs_in_domain) != 1:
                 return False
+        if self.costs is not None:
+            x, c = x[:self.num_bools()], x[self.num_bools():]
+            if not (np.dot(self.costs, x) == c).all():
+                return False
         if self.x_constraints is not None:
             for head, body in self.x_constraints:
                 if x[head] and not any(x[atom] == 1 for atom in body):
@@ -51,15 +70,15 @@ class Dataset(object):
         return True
 
 class SyntheticDataset(Dataset):
-    def __init__(self, domain_sizes, x_constraints=None):
-        items = self._ground(domain_sizes, x_constraints)
-        super(SyntheticDataset, self).__init__(domain_sizes, items, x_constraints)
+    def __init__(self, domain_sizes):
+        items = self._ground(domain_sizes, None)
[…]

### 2016-01-13 paper/ijcai16.tex
paper: add some meat to the experiments + random updates.

@@
 SetMargin} algorithm. We implemented the {\sc SetMargin} algorithm using
-Python, and leverage Gurobi 6.5.0 for solving the MILP problem. Both the 
-{\sc SetMargin} source code and the full experimental setup are available at
-\stefano{add URL}.
+Python, leveraging Gurobi 6.5.0 as the MILP solver. Both the {\sc SetMargin}
+source code and the full experimental setup are available at \stefano{add URL}.
 
@@
 %
-$$ P(\vx^i > \vx^j \;|\; \vw, \vx^i \not\sim \vx^j) = (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
+$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
 %
@@
 %
-$$ P(\vx^i \sim \vx^j \;|\; \vw) = \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
+$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
 %
@@
 
-In all experiments we use an internal cross-validation procedure to update the
+In all experiments we use an internal 5-fold cross-validation procedure to update the
 hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5 iterations. The
-hyperparameters are chosen as to maximize the ranking loss over the user
-answers collected so far. $\alpha$ is taken in $\{100, 10, 1\}$,
-while $\beta$ and $\gamma$ are taken in $\{100, 10, 1, 0.1, 0\}$. Note that
-$\beta=0$ implies that the weight vector is not regularized, nor sparse, while
-$\gamma=0$ implies that the absolute quality of the configurations is ignored.
-\stefano{is this desirable?}
+hyperparameters are chosen as to minimize the ranking loss over the user
+answers collected so far. $\alpha$ is taken in $\{20, 15, 10, 5, 1\}$,
+while $\beta$ and $\gamma$ are taken in $\{10, 5, 1, 0.5, 0.1\}$.
+\stefano{explain that we can't turn off $\gamma$}
 
@@
 [3] \times [3] \times [3]$, and so on. The cardinality of
-$\calX_\text{feasible}$ is $n^n$, and grows (super-)exponentially with $n$. For
+$\calX_\text{feasible}$ is $n^n$, and grows (super) exponentially with $n$. For
 $n=3$, the dataset is comparable to the synthetic one used
@@
 We further produced two novel {\em sparse} sets of weight vectors by deleting
-$80\%$ uniformly at random \stefano{Paolo, double check please} the entries
-from the vectors sampled in the uniform and normal settings. We sampled $100$
+$80\%$ of the entries of the vectors sampled in the uniform and normal settings
+uniformly at random \stefano{Paolo, is this correct?}. We sampled $100$
 vectors for each setting. All methods were evaluated on the very same weight
@@
 
-\paragraph{Constructive Dataset.} Next, we tested {\sc SetMargin} on a truly
-constructive setting, which can not be readily handled by non-constructive
-methods.
+\paragraph{Constructive dataset.} Next, we tested {\sc SetMargin} on a truly
+constructive setting. We developed a constructive version of the PC dataset
+used in~\cite{guo2010real}: instead of explicitly enumerating all possible PC
+item, we definined the set of feasible configurations through MILP constraints.
 
-The attributes are Manufacturer, CPU, Monitor size, PC type, Memory, and HD
-size. CPU type and CPU frequency as combined in a single variable, e.g. ``AMD
-Athlon @1000" stands for an AMD Athlon CPU clocked at 1 GHz. All other domains
-are left unchanged.
+A PC configurations is defined by seven attributes: computer type (laptop,
+desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
+size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
+attributes. The price attribute is defined as a linear combination of the other
+attributes: this is a fair modeling choice, as often the price of a PC is
+well approximated by sum of the price of its components plus an offset due to
+branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
+type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
+G4 @733 MHz``), so to avoid multiplicative terms in the price function.
+\stefano{too detailed/handwavy?}
 
-The configurations constraints:
+Interactions between attributes are expressed as Horn clauses. An example
+costraint might look like this: ``if the manufacturer is Apple, then the
+CPU must be either a PowerPC G3 or a PowerPC G4'', which can be encoded
+into MILP form as:
 %
-$$ x_i \Leftarrow x_{j_1} \lor \ldots \lor x_{j_n} $$
+$$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $$
 %
-have been translated into MILP constraints of the form:
-%
-$$ (1 - x_i) + x_{j_1} + \ldots + x_{j_n} \ge 1 $$
+Note that mutual exclusivity between all binary variables corresponding to an
+attribute ensures that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1.
+The dataset includes constraints between the following attributes: manufacturer
+$\to$ type, manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
+size, type $\to$ monitor size. In total we encoded 16 Horn constraints, each
+with one head and multiple options in the body. We do not report the full
+list due to space limitations.
 
-Each attribute configuration is associated a corresponding {\em price}. The
-price of a configuration is the sum of the prices of the individual components.
-
-\begin{figure}
-    {\footnotesize
-    \begin{tabular}{cc}
-        {\bf Rule} & {\bf MILP constraint} \\ \hline\hline
-        & Manufacturer = Compacq $\Rightarrow$ Type = Laptop $\lor$ Type = Desktop \\
-        & Manufacturer = Apple $\Rightarrow$ CPU = PowerPC G3 @266 $\lor$ \ldots \\
-        & Type = Laptop $\Rightarrow$ Memory = 64 $\lor$ \ldots \\
-        & Type = HDSize $\Rightarrow$ HDSize = 20 $\lor$ \ldots \\
-        & Type = Desktop $\Rightarrow$ Monitor = 15 $\lor$ \ldots \\
+\begin{table}
+    \centering
+    \begin{tabular}{ccc}
+        {\bf Attribute} & {\bf Type} & {\bf Values} \\
+        \hline \hline
+        Type & discrete & 3 \\
+        Manufacturer & discrete & 8 \\
+        CPU model & discrete & 37 \\
+        RAM amount & discrete & 10 \\
+        HD size & discrete & 10 \\
+        Monitor size & discrete & 8 \\
+        Price & continuous & --
     \end{tabular}
-    }
-    \caption{\label{tab:pcconstraints} List of MILP constraints used in the
-    PC experiment.}
-\end{figure}
+    \caption{\label{tab:pcdataset} Statistics for the PC dataset.}
+\end{table}
 

### 2016-01-15 draw.py
update the drawing script.

@@ -7,6 +7,10 @@ from matplotlib.font_manager import FontProperties
 import itertools as it
 from glob import glob
 
+if len(sys.argv) != 3:
+    print "Usage: {} <results directory> <output image>".format(sys.argv[0])
+    sys.exit(1)
+
 fontP = FontProperties()
 fontP.set_size('small')
 
@@ -23,11 +27,11 @@ for path in paths:
     num_trials, max_queries = loss_matrix.shape
 
     loss_means = np.mean(loss_matrix, axis=0)
+    loss_stds = np.std(loss_matrix, axis=0).reshape(-1, 1)
 
-    ax.set_ylim([0.0, max(0.5, max(loss_means) + 0.1)])
-    ax.errorbar(np.arange(1, max_queries + 1), loss_means, label=os.path.basename(path))
+    ax.set_ylim([0.0, max(0.5, max(loss_means) + max(loss_stds) + 0.1)])
+    ax.errorbar(np.arange(1, max_queries + 1), loss_means, yerr=loss_stds, fmt="--o", label=os.path.basename(path))
 
-lgd = ax.legend(prop=fontP, loc="upper center", bbox_to_anchor=(0.5, -0.1))
-
-path = os.path.join(sys.argv[1], "plot.svg")
-fig.savefig(path, bbox_extra_artists=(lgd,), bbox_inches="tight")
+#lgd = ax.legend(prop=fontP, loc="upper center", bbox_to_anchor=(0.5, -0.1))
+#fig.savefig(sys.argv[2], bbox_extra_artists=(lgd,), bbox_inches="tight")
+fig.savefig(sys.argv[2], bbox_inches="tight")

### 2016-01-15 draw_all.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+
+for domain_size in `seq 2 5`; do
+    for sampling_mode in uniform uniform_sparse normal normal_sparse; do
+        echo "drawing $domain_size $sampling_mode"
+        ./draw.py results/synthetic_${domain_size}/${sampling_mode} paper/figure_synthetic_${domain_size}_${sampling_mode}.png
+    done
+done

### 2016-01-15 paper/ijcai16.tex
paper: add first figures, fixup text a bit.

@@
 \usepackage{ijcai16}
-\usepackage{graphicx}
 \usepackage{times}
 %\usepackage{latexsym}
+\usepackage{graphicx}
 
@@
 from the context, and use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
-The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$.
+The $\ell_1$ vector norm is denoted by $\|\vx\|_1 := \sum_z |x_z|$.
 
@@
 e.g. the ``Boolean disjunction'' of two binary variables $x_1 \lor x_2$ can be
-rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations will be
-indicated as $\calX_\text{feasible}$.
+rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations is
+denoted by $\calX_\text{feasible}$.
 
@@
 vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
-$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we will
+$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we
 require all weights to be {\em non-negative} and {\em bounded}: the individual
@@
 user. The set of examples consists of pairwise comparisons between a more
-preferred configuration $\vy_+$ and a less preferred on $\vy_-$. Each preference
+preferred configuration $\vy_+$ and a less preferred one $\vy_-$. Each preference
 induces a ranking constraint $\langle \vw, \vy_+ - \vy_- \rangle \ge 0$.
@@
 type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
-G4 @733 MHz``), so to avoid multiplicative terms in the price function.
+G4 @733 MHz''), so to avoid multiplicative terms in the price function.
 \stefano{too detailed/handwavy?}
@@
 
+\onecolumn
+
+\begin{figure}[b]
+    \centering
+    \begin{tabular}{cccc}
+        {\bf synthetic 2} & {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5}
+        \\
+        \includegraphics[width=12em]{figure_synthetic_2_uniform} &
+        \includegraphics[width=12em]{figure_synthetic_3_uniform} &
+        \includegraphics[width=12em]{figure_synthetic_4_uniform} &
+        \includegraphics[width=12em]{figure_synthetic_5_uniform}
+        \\
+        \includegraphics[width=12em]{figure_synthetic_2_uniform_sparse} &
+        \includegraphics[width=12em]{figure_synthetic_3_uniform_sparse} &
+        \includegraphics[width=12em]{figure_synthetic_4_uniform_sparse} &
+        \includegraphics[width=12em]{figure_synthetic_5_uniform_sparse}
+        \\
+        \includegraphics[width=12em]{figure_synthetic_2_normal} &
+        \includegraphics[width=12em]{figure_synthetic_3_normal} &
+        \includegraphics[width=12em]{figure_synthetic_4_normal} &
+        \includegraphics[width=12em]{figure_synthetic_5_normal}
+        \\
+        \includegraphics[width=12em]{figure_synthetic_2_normal_sparse} &
+        \includegraphics[width=12em]{figure_synthetic_3_normal_sparse} &
+        \includegraphics[width=12em]{figure_synthetic_4_normal_sparse} &
+        \includegraphics[width=12em]{figure_synthetic_5_normal_sparse}
+    \end{tabular}
+    \caption{Results on the synthetic dataset. Each row represents a different
+    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
+    the set size $k$: blue means $k=2$, greed $k=3$, and red $k=4$.}
+\end{figure}
+
+\twocolumn
+
 \end{document}

### 2016-01-16 paper/ijcai16.tex
paper: remove synthetic 2 figures, update the others.
paper: move figures to their own directory.
As requested by Andrea.

@@
     \begin{center}
-        \includegraphics[width=16em]{setmargin}
+        \includegraphics[width=16em]{figures/setmargin}
     \end{center}
@@
     \centering
-    \begin{tabular}{cccc}
-        {\bf synthetic 2} & {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5}
+    \begin{tabular}{ccc}
+        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5}
         \\
-        \includegraphics[width=12em]{figure_synthetic_2_uniform} &
-        \includegraphics[width=12em]{figure_synthetic_3_uniform} &
-        \includegraphics[width=12em]{figure_synthetic_4_uniform} &
-        \includegraphics[width=12em]{figure_synthetic_5_uniform}
+        \includegraphics[width=12em]{figures/synthetic_3_uniform} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform}
         \\
-        \includegraphics[width=12em]{figure_synthetic_2_uniform_sparse} &
-        \includegraphics[width=12em]{figure_synthetic_3_uniform_sparse} &
-        \includegraphics[width=12em]{figure_synthetic_4_uniform_sparse} &
-        \includegraphics[width=12em]{figure_synthetic_5_uniform_sparse}
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse}
         \\
-        \includegraphics[width=12em]{figure_synthetic_2_normal} &
-        \includegraphics[width=12em]{figure_synthetic_3_normal} &
-        \includegraphics[width=12em]{figure_synthetic_4_normal} &
-        \includegraphics[width=12em]{figure_synthetic_5_normal}
+        \includegraphics[width=12em]{figures/synthetic_3_normal} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal}
         \\
-        \includegraphics[width=12em]{figure_synthetic_2_normal_sparse} &
-        \includegraphics[width=12em]{figure_synthetic_3_normal_sparse} &
-        \includegraphics[width=12em]{figure_synthetic_4_normal_sparse} &
-        \includegraphics[width=12em]{figure_synthetic_5_normal_sparse}
+        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse}
     \end{tabular}

### 2016-01-16 draw_all.sh

@@ -3,6 +3,6 @@
 for domain_size in `seq 2 5`; do
     for sampling_mode in uniform uniform_sparse normal normal_sparse; do
         echo "drawing $domain_size $sampling_mode"
-        ./draw.py results/synthetic_${domain_size}/${sampling_mode} paper/figure_synthetic_${domain_size}_${sampling_mode}.png
+        ./draw.py results/synthetic_${domain_size}/${sampling_mode} paper/figures/synthetic_${domain_size}_${sampling_mode}.png
     done
 done

### 2016-01-16 draw.py
update drawing scripts.

@@ -20,18 +20,47 @@ ax.set_ylabel("Average loss over trials")
 
 paths = sorted(glob(os.path.join(sys.argv[1], "results_*_loss_matrix.txt")))
 
+# Colors taken from http://tango.freedesktop.org/Tango_Icon_Theme_Guidelines
+COLOR_OF = {
+#    ("#EDD400", "#FCE94F"), # yellow
+    2: ("#CE5C00", "#F57900"), # orange
+#    ("#4E9A06", "#73D216"), # green
+    3: ("#204A87", "#3465A4"), # blue
+#    ("#5C3566", "#75507B"), # violet
+#    2: ("#A40000", "#CC0000"), # red
+}
+
+max_max_x = 0
+max_max_y = 0
 for path in paths:
     parts = os.path.basename(path).split("__")
+    set_size = int(parts[1].split("=")[1])
+    if not set_size in (2, 3):
+        continue
 
     loss_matrix = np.loadtxt(path)
+    if loss_matrix.shape[1] > 100:
+        loss_matrix = loss_matrix[:,:100]
     num_trials, max_queries = loss_matrix.shape
 
-    loss_means = np.mean(loss_matrix, axis=0)
-    loss_stds = np.std(loss_matrix, axis=0).reshape(-1, 1)
+    xs = np.arange(max_queries)
+    ys = np.median(loss_matrix, axis=0)
+    yerrs = np.std(loss_matrix, axis=0)
 
-    ax.set_ylim([0.0, max(0.5, max(loss_means) + max(loss_stds) + 0.1)])
-    ax.errorbar(np.arange(1, max_queries + 1), loss_means, yerr=loss_stds, fmt="--o", label=os.path.basename(path))
+    if max_queries > max_max_x:
+        max_max_x = max_queries
+    max_y = max(ys + yerrs)
+    if max_y > max_max_y:
+        max_max_y = max_y
 
-#lgd = ax.legend(prop=fontP, loc="upper center", bbox_to_anchor=(0.5, -0.1))
-#fig.savefig(sys.argv[2], bbox_extra_artists=(lgd,), bbox_inches="tight")
+    fg, bg = COLOR_OF[set_size]
+
+    ax.plot(xs, ys, "k-", linewidth=2.0, color=fg)
+    ax.fill_between(xs, ys - yerrs, ys + yerrs, color=bg, alpha=0.35, linewidth=0)
+
+
[…]

### 2016-01-16 draw_all.sh

@@ -1,6 +1,6 @@
 #!/bin/bash
 
-for domain_size in `seq 2 5`; do
+for domain_size in `seq 3 6`; do
     for sampling_mode in uniform uniform_sparse normal normal_sparse; do
         echo "drawing $domain_size $sampling_mode"
         ./draw.py results/synthetic_${domain_size}/${sampling_mode} paper/figures/synthetic_${domain_size}_${sampling_mode}.png

### 2016-01-16 ijcai16.py
ijcai16,setmargin/setmargin: convert num_iterations into max_iterations.

@@ -65,7 +65,7 @@ def get_result_paths(dataset_name, config):
         "multimargin={}".format(config.multimargin),
         "threads={}".format(config.threads),
         config.num_trials,
-        config.num_iterations,
+        config.max_iterations,
     ]))
     path0 = "results_{}_infos.pickle".format(basename)
     path1 = "results_{}_loss_matrix.txt".format(basename)
@@ -163,8 +163,9 @@ def precrossvalidate(dataset, config, solver, users):
 
             for i in train_set:
                 infos = setmargin.run(dataset, target_users[i], solver,
-                                      config.num_iterations, config.set_size,
-                                      alphas, tol=config.tol, debug=config.debug)
+                                      config.set_size, alphas=alphas,
+                                      max_iterations=config.max_iterations,
+                                      tol=config.tol, debug=config.debug)
                 # XXX we only consider the true utility loss at the last
                 # iteration
                 losses.append(infos[-1][1])
@@ -217,10 +218,9 @@ def solve(dataset, config, ws=None):
             ===========
             """).format(trial, config.num_trials)
 
-        info = setmargin.run(dataset, users[trial], solver,
-                             config.num_iterations, config.set_size, alphas,
-                             tol=config.tol,
-                             crossval_set_size=config.crossval_set_size,
+        info = setmargin.run(dataset, users[trial], solver, config.set_size,
+                             max_iterations=config.max_iterations, tol=config.tol,
+                             alphas=alphas, crossval_set_size=config.crossval_set_size,
                              debug=config.debug)
         infos.append(info)
     return infos
@@ -228,7 +228,7 @@ def solve(dataset, config, ws=None):
 def run_synthetic():
     CONFIGS = Grid({
         "num_trials": 20,
-        "num_iterations": 100,
+        "max_iterations": 100,
         "sampling_mode": ("uniform", "uniform_sparse", "normal", "normal_sparse"),
         "ranking_mode": ("all_pairs",),
         "is_deterministic": False,
@@ -283,8 +283,8 @@ def run_from_command_line():
     parser.add_argument("--domain-sizes", type=str, default="2,2,5",
                         help="domain sizes for the synthetic dataset only")
 
-    parser.add_argument("-n", "--num_iterations", type=int, default=20,
-                        help="number of iterations")
+    parser.add_argument("-n", "--max-iterations", type=int, default=20,
+                        help="maximum number of iterations")
[…]

### 2016-01-16 setmargin/setmargin.py

@@ -72,19 +72,21 @@ def print_answers(queries, hidden_w):
                                    score_xi - score_xj))
     print "\n".join(message)
 
-def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
-        tol=1e-4, crossval_interval=5, crossval_set_size=None, debug=False):
+def run(dataset, user, solver, set_size, max_iterations=100, tol=1e-4,
+        alphas="auto", crossval_interval=5, crossval_set_size=None,
+        debug=False):
     """Runs the setmargin algorithm.
 
     :param dataset: the dataset.
     :param user: the user.
     :param solver: the setmargin solver.
-    :param num_iterations: number of iterations to run for.
     :param set_size: set size.
+    :param max_iterations: maximum number of iterations to run for.
+        (default: ``100``)
+    :param tol: user tolerance, used for termination. (default: ``1e-4``)
     :param alphas: either a triple of non-negative floats, or ``"auto"``,
         in which case the hyperparameters are determined automatically through
         a periodic cross-validation procedure. (default: ``"auto"``)
-    :param tol: user tolerance, used for termination. (default: ``1e-4``)
     :param crossval_interval: number of iterations between cross-validation
         calls. (default: ``5``)
     :param crossval_set_size: number of items. (default: ``set_size``)
@@ -92,8 +94,8 @@ def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
     :return: the number of queries, utility loss and elapsed time for
         each iteration.
     """
-    if not num_iterations > 0:
-        raise ValueError("num_iterations must be positive")
+    if not max_iterations > 0:
+        raise ValueError("max_iterations must be positive")
     if not crossval_interval > 0:
         raise ValueError("crossval_interval must be positive")
 
@@ -117,14 +119,15 @@ def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
         if crossval_set_size is None:
             crossval_set_size = set_size
 
-    answers, info, old_best_item = [], [], None
-    for t in range(num_iterations):
+    answers, info, old_best_item, t = [], [], None, 0
+    while True:
+
         if debug:
             print dedent("""\
-            ===============
-            ITERATION {}/{}
-            ===============
-            """).format(t, num_iterations)
+            ============
+            ITERATION {}
+            ============
+            """).format(t)
 
             print_answers(answers, user.w)
 
@@ -173,9 +176,17 @@ def run(dataset, user, solver, num_iterations, set_size, alphas="auto",
[…]

### 2016-01-16 ijcai16.py
ijcai16: rename -N option to -T.

@@ -278,7 +278,7 @@ def run_from_command_line():
     parser.add_argument("dataset", type=str,
                         help="dataset")
 
-    parser.add_argument("-N", "--num_trials", type=int, default=20,
+    parser.add_argument("-T", "--num_trials", type=int, default=20,
                         help="number of trials")
     parser.add_argument("--domain-sizes", type=str, default="2,2,5",
                         help="domain sizes for the synthetic dataset only")

### 2016-01-16 ijcai16.py
ijcai16,setmargin/setmargin: add support for --max-answers.

@@ -66,6 +66,7 @@ def get_result_paths(dataset_name, config):
         "threads={}".format(config.threads),
         config.num_trials,
         config.max_iterations,
+        config.max_answers,
     ]))
     path0 = "results_{}_infos.pickle".format(basename)
     path1 = "results_{}_loss_matrix.txt".format(basename)
@@ -165,6 +166,7 @@ def precrossvalidate(dataset, config, solver, users):
                 infos = setmargin.run(dataset, target_users[i], solver,
                                       config.set_size, alphas=alphas,
                                       max_iterations=config.max_iterations,
+                                      max_answers=config.max_answers,
                                       tol=config.tol, debug=config.debug)
                 # XXX we only consider the true utility loss at the last
                 # iteration
@@ -219,7 +221,8 @@ def solve(dataset, config, ws=None):
             """).format(trial, config.num_trials)
 
         info = setmargin.run(dataset, users[trial], solver, config.set_size,
-                             max_iterations=config.max_iterations, tol=config.tol,
+                             max_iterations=config.max_iterations,
+                             max_answers=config.max_answers, tol=config.tol,
                              alphas=alphas, crossval_set_size=config.crossval_set_size,
                              debug=config.debug)
         infos.append(info)
@@ -229,6 +232,7 @@ def run_synthetic():
     CONFIGS = Grid({
         "num_trials": 20,
         "max_iterations": 100,
+        "max_answers": 100,
         "sampling_mode": ("uniform", "uniform_sparse", "normal", "normal_sparse"),
         "ranking_mode": ("all_pairs",),
         "is_deterministic": False,
@@ -285,6 +289,8 @@ def run_from_command_line():
 
     parser.add_argument("-n", "--max-iterations", type=int, default=20,
                         help="maximum number of iterations")
+    parser.add_argument("-N", "--max-answers", type=int, default=20,
+                        help="number of iterations")
     parser.add_argument("-m", "--set-size", type=int, default=3,
                         help="number of hyperplanes/items to solve for")
     parser.add_argument("-t", "--tol", type=float, default=1e-4,

### 2016-01-16 setmargin/setmargin.py

@@ -72,8 +72,8 @@ def print_answers(queries, hidden_w):
                                    score_xi - score_xj))
     print "\n".join(message)
 
-def run(dataset, user, solver, set_size, max_iterations=100, tol=1e-4,
-        alphas="auto", crossval_interval=5, crossval_set_size=None,
+def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
+        tol=1e-4, alphas="auto", crossval_interval=5, crossval_set_size=None,
         debug=False):
     """Runs the setmargin algorithm.
 
@@ -82,6 +82,8 @@ def run(dataset, user, solver, set_size, max_iterations=100, tol=1e-4,
     :param solver: the setmargin solver.
     :param set_size: set size.
     :param max_iterations: maximum number of iterations to run for.
+        (default: ``100``)
+    :param max_answers: maximum number of answers to elicit in total.
         (default: ``100``)
     :param tol: user tolerance, used for termination. (default: ``1e-4``)
     :param alphas: either a triple of non-negative floats, or ``"auto"``,
@@ -96,6 +98,8 @@ def run(dataset, user, solver, set_size, max_iterations=100, tol=1e-4,
     """
     if not max_iterations > 0:
         raise ValueError("max_iterations must be positive")
+    if not max_answers > 0:
+        raise ValueError("max_answers must be positive")
     if not crossval_interval > 0:
         raise ValueError("crossval_interval must be positive")
 
@@ -180,6 +184,11 @@ def run(dataset, user, solver, set_size, max_iterations=100, tol=1e-4,
         if t >= max_iterations:
             if debug:
                 print "maximum iterations reached, terminating"
+            break
+
+        if len(answers) >= max_answers:
+            if debug:
+                print "maximum number of answers reached, terminating"
             break
 
         # If the user is satisfied (clicks the 'add to cart' button),

### 2016-01-16 ijcai16.py
ijcai16: organize arguments into groups.

@@ -279,52 +279,53 @@ def run_from_command_line():
     import argparse
 
     parser = argparse.ArgumentParser(formatter_class=argparse.ArgumentDefaultsHelpFormatter)
-    parser.add_argument("dataset", type=str,
-                        help="dataset")
-
+    parser.add_argument("dataset", type=str, help="dataset")
     parser.add_argument("-T", "--num_trials", type=int, default=20,
                         help="number of trials")
     parser.add_argument("--domain-sizes", type=str, default="2,2,5",
-                        help="domain sizes for the synthetic dataset only")
-
-    parser.add_argument("-n", "--max-iterations", type=int, default=20,
-                        help="maximum number of iterations")
-    parser.add_argument("-N", "--max-answers", type=int, default=20,
-                        help="number of iterations")
-    parser.add_argument("-m", "--set-size", type=int, default=3,
-                        help="number of hyperplanes/items to solve for")
-    parser.add_argument("-t", "--tol", type=float, default=1e-4,
-                        help="tolerance used for termination")
-    parser.add_argument("-a", "--alpha", type=float, default=0.1,
-                        help="hyperparameter controlling the importance of slacks")
-    parser.add_argument("-b", "--beta", type=float, default=0.1,
-                        help="hyperparameter controlling the importance of regularization")
-    parser.add_argument("-c", "--gamma", type=float, default=0.1,
-                        help="hyperparameter controlling the score of the output items")
-    parser.add_argument("-x", "--crossval", action="store_true",
-                        help="whether to perform automatic hyperparameter crossvalidation. If enabled, -a -b -c are ignored.")
-    parser.add_argument("-X", "--crossval-set-size", type=int, default=None,
-                        help="set_size for the hyperparameter crossvalidation.")
-    parser.add_argument("-y", "--precrossval", action="store_true",
-                        help="do parameter selection using crossvalidation prior to learning")
-    parser.add_argument("-M", "--multimargin", action="store_true",
-                        help="whether the example and generated object margins should be independent")
-
-    parser.add_argument("-u", "--sampling-mode", type=str, default="uniform",
-                        help="utility sampling mode, any of ('uniform', 'normal')")
-    parser.add_argument("-r", "--ranking-mode", type=str, default="all_pairs",
-                        help="ranking mode for set-wide queries, any of ('all_pairs', 'sorted_pairs')")
-    parser.add_argument("-d", "--is-deterministic", action="store_true",
[…]

### 2016-01-16 setmargin/setmargin.py
setmargin/setmargin: account for cross-validation time.

@@ -135,12 +135,12 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
 
             print_answers(answers, user.w)
 
+        old_time = time.time()
+
         # Crossvalidate the hyperparameters if required
         if do_crossval and t % crossval_interval == 0 and t > 0:
             alphas = crossvalidate(dataset, solver, answers, crossval_set_size,
                                    debug)
-
-        old_time = time.time()
 
         # Solve the set_size=k case
         ws, xs = solver.compute_setmargin(dataset, answers, set_size, alphas)

### 2016-01-16 setmargin/user.py
setmargin/user: split query_diff() out of query().

@@ -56,11 +56,10 @@ class User(object):
             w[mask == 0] = 0
         return w.reshape(1, -1)
 
-    def query(self, xi, xj):
+    def query_diff(self, diff):
         """Queries the user about a single pairwise choice.
 
-        :param xi: an item.
-        :param xj: the other item.
+        :param diff: different in utility loss: u(xi) - u(xj).
         :returns: 0 if the user is indifferent, 1 if xi is better than xj,
             and -1 if xi is worse than xj.
         """
@@ -68,8 +67,6 @@ class User(object):
         # uses ALPHA=1, BETA=1; here however we have weights in the range [0, 1]
         # so we must rescale ALPHA and BETA to obtain the same probabilities.
         ALPHA, BETA = 100, 100
-
-        diff = np.dot(self.w, xi.T - xj.T)
 
         if self.is_deterministic:
             return int(np.sign(diff))
@@ -84,6 +81,16 @@ class User(object):
         elif z < (eq + gt):
             return 1
         return -1
+
+    def query(self, xi, xj):
+        """Queries the user about a single pairwise choice.
+
+        :param xi: an item.
+        :param xj: the other item.
+        :returns: 0 if the user is indifferent, 1 if xi is better than xj,
+            and -1 if xi is worse than xj.
+        """
+        return self.query_diff(np.dot(self.w, xi.T - xj.T))
 
     def query_set(self, ws, xs, old_best_item):
         """Queries the user about the provided set of items.

### 2016-01-16 setmargin/setmargin.py
setmargin/setmargin: implement termination-by-indifference.

@@ -73,7 +73,7 @@ def print_answers(queries, hidden_w):
     print "\n".join(message)
 
 def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
-        tol=1e-4, alphas="auto", crossval_interval=5, crossval_set_size=None,
+        tol="auto", alphas="auto", crossval_interval=5, crossval_set_size=None,
         debug=False):
     """Runs the setmargin algorithm.
 
@@ -85,7 +85,8 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
         (default: ``100``)
     :param max_answers: maximum number of answers to elicit in total.
         (default: ``100``)
-    :param tol: user tolerance, used for termination. (default: ``1e-4``)
+    :param tol: user tolerance, used for termination, or ``"auto"``, in
+        which case user indifference is used instead. (default: ``"auto"``)
     :param alphas: either a triple of non-negative floats, or ``"auto"``,
         in which case the hyperparameters are determined automatically through
         a periodic cross-validation procedure. (default: ``"auto"``)
@@ -193,7 +194,11 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
 
         # If the user is satisfied (clicks the 'add to cart' button),
         # we are done
-        if loss < tol:
+        if tol == "auto" and user.query_diff(loss) == 0:
+            if debug:
+                print "user indifference reached, terminating"
+            break
+        elif tol != "auto" and loss < tol:
             if debug:
                 print "minimum tolerance reached, terminating"
             break

### 2016-01-16 setmargin/setmargin.py
setmargin/setmargin: do not normalize the utility loss, just like the competitors do.

@@ -170,7 +170,7 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
         # Compute the utility loss
         best_score = np.dot(user.w.ravel(), best_item)
         pred_score = np.dot(user.w.ravel(), xs[0])
-        loss = (best_score - pred_score) / user_w_norm
+        loss = best_score - pred_score
         if debug:
             print dedent("""\
                     best_item = (true score = {})

### 2016-01-16 ijcai16.py
ijcai16: allow to set the tolerance to 'auto'.

@@ -295,7 +295,7 @@ def run_from_command_line():
                        help="maximum number of iterations")
     group.add_argument("-N", "--max-answers", type=int, default=20,
                        help="number of iterations")
-    group.add_argument("-t", "--tol", type=float, default=1e-4,
+    group.add_argument("-t", "--tol", type=str, default="auto",
                        help="tolerance used for termination")
 
     group = parser.add_argument_group("setmargin hyperparameters")
@@ -330,6 +330,10 @@ def run_from_command_line():
 
     argsdict = vars(args)
     argsdict["dataset"] = args.dataset
+    try:
+        argsdict["tol"] = float(argsdict["tol"])
+    except:
+        pass
 
     config = Grid(argsdict)
 

### 2016-01-16 ijcai16.py
ijcai16: rename -m to -k, like in the paper.

@@ -299,7 +299,7 @@ def run_from_command_line():
                        help="tolerance used for termination")
 
     group = parser.add_argument_group("setmargin hyperparameters")
-    group.add_argument("-m", "--set-size", type=int, default=3,
+    group.add_argument("-k", "--set-size", type=int, default=3,
                        help="number of hyperplanes/items to solve for")
     group.add_argument("-a", "--alpha", type=float, default=0.1,
                        help="hyperparameter controlling the importance of slacks")

### 2016-01-16 ijcai16.py
ijcai16: stop rescaling the weight vectors.

@@ -42,8 +42,6 @@ def _load_utilities(num_attrs, sampling_mode):
     with open(join("data", "randomUtility", basename), "rb") as fp:
         for line in fp:
             utilities.append(map(float, line.split(",")))
-    # rescale the range from [0,100) to [0,1)
-    utilities = np.array(utilities) / 100.0
     # XXX turn negative values into zeros
     utilities[utilities < 0] = 0
     return utilities

### 2016-01-16 setmargin/user.py
setmargin/user: use same weight range and response alpha/beta as Guo&Sanner.

@@ -40,9 +40,9 @@ class User(object):
     def _sample(self, sampling_mode, domain_sizes, sparsity=0.2):
         num_attrs = sum(domain_sizes)
         if sampling_mode in ("uniform", "uniform_sparse"):
-            w = self._rng.uniform(0, 1, size=num_attrs)
+            w = self._rng.uniform(1, 100, size=num_attrs)
         elif sampling_mode in ("normal", "normal_sparse"):
-            w = self._rng.normal(0.25, 0.25 / 3, size=num_attrs)
+            w = self._rng.normal(25.0, 25.0 / 3, size=num_attrs)
         else:
             raise ValueError("invalid sampling_mode")
         if sampling_mode.endswith("sparse"):
@@ -63,10 +63,7 @@ class User(object):
         :returns: 0 if the user is indifferent, 1 if xi is better than xj,
             and -1 if xi is worse than xj.
         """
-        # The original problem has weights sampled in the range [0, 100], and
-        # uses ALPHA=1, BETA=1; here however we have weights in the range [0, 1]
-        # so we must rescale ALPHA and BETA to obtain the same probabilities.
-        ALPHA, BETA = 100, 100
+        ALPHA, BETA = 1.0, 1.0
 
         if self.is_deterministic:
             return int(np.sign(diff))

### 2016-01-17 setmargin/datasets.py
setmargin/{datasets,solver}: move get_zs_in_domains() to dataset and use it.

@@ -38,8 +38,16 @@ class Dataset(object):
             yield base, base + size
             base += size
 
+    def get_zs_in_domains(self):
+        zs_in_domains, last_z = [], 0
+        for domain_size in self.domain_sizes:
+            assert domain_size > 1
+            zs_in_domains.append(range(last_z, last_z + domain_size))
+            last_z += domain_size
+        return zs_in_domains
+
     def is_item_valid(self, x):
-        for zs_in_domain in get_zs_in_domains(self.domain_sizes):
+        for zs_in_domain in self.get_zs_in_domains():
             if sum(x[z] for z in zs_in_domain) != 1:
                 return False
         if self.costs is not None:

### 2016-01-17 setmargin/util.py
setmargin/util: delete.

@@ -1,24 +1,0 @@
-import numpy as np
-
-def _cls(obj):
-    return type(obj).__name__
-
-def prod(l):
-    return l[0] if len(l) == 1 else l[0]*prod(l[1:])
-
-def onehot(domain_size, value):
-    assert 0 <= value < domain_size, "out-of-bounds value: 0/{}/{}".format(value, domain_size)
-    value_onehot = np.zeros(domain_size, dtype=np.int8)
-    value_onehot[value] = 1
-    return value_onehot
-
-def vonehot(domain_sizes, item):
-    return np.hstack((onehot(domain_sizes[i], v) for i, v in enumerate(item)))
-
-def get_zs_in_domains(domain_sizes):
-    zs_in_domains, last_z = [], 0
-    for domain_size in domain_sizes:
-        assert domain_size > 1
-        zs_in_domains.append(range(last_z, last_z + domain_size))
-        last_z += domain_size
-    return zs_in_domains

### 2016-01-17 setmargin/__init__.py
setmargin: remove obsolete imports.

@@ -1,5 +1,4 @@
 from datasets import *
 from user import *
 from solver import *
-from util import *
 from setmargin import *

### 2016-01-17 setmargin/datasets.py

@@ -3,7 +3,6 @@
 import numpy as np
 import itertools as it
 from sklearn.utils import check_random_state
-from util import *
 
 class Dataset(object):
     """A dataset over the Cartesian product of all attribute domains.

### 2016-01-17 setmargin/setmargin.py

@@ -8,8 +8,6 @@ from sklearn.utils import check_random_state
 from textwrap import dedent
 from pprint import pformat
 
-from util import *
-
 NUM_FOLDS = 5
 ALL_ALPHAS = list(it.product(
          [20.0, 10.0, 5.0, 1.0],

### 2016-01-17 setmargin/solver.py

@@ -5,7 +5,6 @@ import tempfile
 import gurobipy as grb
 from gurobipy import GRB
 from textwrap import dedent
-from util import *
 
 MAX_W_Z = 1
 ENABLE_LP_DUMP = False

### 2016-01-17 ijcai16.py
ijcai16: fix thinko.

@@ -204,7 +204,7 @@ def solve(dataset, config, ws=None, num_repeats=1):
         print user
         users.extend([user] * num_repeats)
         i += num_repeats
-    assert len(users) == num_users * num_repeats
+    assert len(users) == num_users
 
     if config.precrossval:
         alphas = precrossvalidate(dataset, config, solver, users)

### 2016-01-17 ijcai16.py
ijcai16: implement user repeats properly.

@@ -182,7 +182,7 @@ def precrossvalidate(dataset, config, solver, users):
 
     return best_alphas
 
-def solve(dataset, config, ws=None, num_repeats=1):
+def solve(dataset, config, ws=None):
     rng = np.random.RandomState(config.seed)
 
     solver = setmargin.Solver(multimargin=config.multimargin,
@@ -190,8 +190,8 @@ def solve(dataset, config, ws=None, num_repeats=1):
 
     num_users = config.num_trials if ws is None else ws.shape[0]
 
-    users, i = [], 0
-    while i < num_users:
+    users = []
+    for i in range(num_users):
         w = None if ws is None else ws[i].reshape(1, -1)
         user = setmargin.User(dataset,
                               sampling_mode=config.sampling_mode,
@@ -200,11 +200,12 @@ def solve(dataset, config, ws=None, num_repeats=1):
                               is_indifferent=config.is_indifferent,
                               w=w,
                               rng=rng)
-        print "created user ({} times) =".format(num_repeats)
-        print user
-        users.extend([user] * num_repeats)
-        i += num_repeats
-    assert len(users) == num_users
+        users.append(user)
+
+    if config.debug:
+        print "users ="
+        for user in users:
+            print user
 
     if config.precrossval:
         alphas = precrossvalidate(dataset, config, solver, users)
@@ -267,10 +268,12 @@ def run_synthetic(same_user):
             key = (num_attrs, config.sampling_mode)
             if not key in utilities:
                 utilities[key] = _load_utilities(*key)
-            ws = utilities[key]
 
-            num_repeats = config.num_trials if same_user else 1
-            infos = solve(dataset, config, ws=ws, num_repeats=num_repeats)
+            ws = utilities[key][:config.num_trials]
+            if same_user:
+                ws = np.tile(ws[0], (config.num_trials, 1))
+
+            infos = solve(dataset, config, ws=ws)
             dump_and_draw("synthetic_{}".format(num_attrs), config, infos)
 
 def run_pc_nocost():

### 2016-01-17 ijcai16.py
ijcai16: add proper PC experiments.

@@ -276,11 +276,39 @@ def run_synthetic(same_user):
             infos = solve(dataset, config, ws=ws)
             dump_and_draw("synthetic_{}".format(num_attrs), config, infos)
 
-def run_pc_nocost():
-    pass
+def run_pc(has_costs):
+    CONFIGS = Grid({
+        "num_trials": 20,
+        "max_iterations": 100,
+        "max_answers": 100,
+        "sampling_mode": ("uniform_sparse", "normal_sparse", "uniform", "normal"),
+        "ranking_mode": ("all_pairs",),
+        "is_deterministic": False,
+        "is_indifferent": True,
+        "set_size": range(2, 4+1),
+        "precrossval": False,
+        "crossval": True,
+        "crossval_set_size": 1,
+        "crossval_interval": 5,
+        "multimargin": False,
+        "tol": 1e-2,
+        "threads": cpu_count(),
+        "debug": True,
+        "seed": 0,
+    })
 
-def run_pc():
-    pass
+    dataset = PCDataset(has_costs=has_costs)
+
+    for config in CONFIGS.iterate():
+
+        print dedent("""\
+            =====================
+            RUNNING CONFIGURATION
+            {}
+            """).format(pformat(config.asdict()))
+
+        infos = solve(dataset, config)
+        dump_and_draw("pc_with_costs" if has_cost else "pc_no_costs", config, infos)
 
 def run_from_command_line():
     import argparse
@@ -374,7 +402,7 @@ if __name__ == "__main__":
     if len(sys.argv) == 1:
         run_synthetic(False)
         run_synthetic(True)
-        run_pc_nocost()
-        run_pc()
+        run_pc(False)
+        run_pc(True)
     else:
         run_from_command_line()

### 2016-01-17 ijcai16.py
ijcai16: add experiment selection.

@@ -399,10 +399,16 @@ def run_from_command_line():
 
 if __name__ == "__main__":
     np.seterr(all="raise")
-    if len(sys.argv) == 1:
-        run_synthetic(False)
-        run_synthetic(True)
-        run_pc(False)
-        run_pc(True)
+    if len(sys.argv) == 2:
+        if sys.argv[1] == "synthetic":
+            run_synthetic(False)
+        elif sys.argv[1] == "synthetic-variance":
+            run_synthetic(True)
+        elif sys.argv[1] == "pc-no-costs":
+            run_pc(False)
+        elif sys.argv[1] == "pc-with-costs":
+            run_pc(True)
+        else:
+            raise ValueError("invalid IJCAI experiment name.")
     else:
         run_from_command_line()

### 2016-01-17 ijcai16.py
ijcai16: clean-up solve() a bit.

@@ -183,25 +183,22 @@ def precrossvalidate(dataset, config, solver, users):
     return best_alphas
 
 def solve(dataset, config, ws=None):
-    rng = np.random.RandomState(config.seed)
-
     solver = setmargin.Solver(multimargin=config.multimargin,
                               threads=config.threads, debug=config.debug)
 
-    num_users = config.num_trials if ws is None else ws.shape[0]
+    rng = np.random.RandomState(config.seed)
 
     users = []
+    num_users = config.num_trials if ws is None else ws.shape[0]
     for i in range(num_users):
         w = None if ws is None else ws[i].reshape(1, -1)
-        user = setmargin.User(dataset,
-                              sampling_mode=config.sampling_mode,
-                              ranking_mode=config.ranking_mode,
-                              is_deterministic=config.is_deterministic,
-                              is_indifferent=config.is_indifferent,
-                              w=w,
-                              rng=rng)
-        users.append(user)
-
+        users.append(setmargin.User(dataset,
+                                    sampling_mode=config.sampling_mode,
+                                    ranking_mode=config.ranking_mode,
+                                    is_deterministic=config.is_deterministic,
+                                    is_indifferent=config.is_indifferent,
+                                    w=w,
+                                    rng=rng))
     if config.debug:
         print "users ="
         for user in users:

### 2016-01-17 paper/ijcai16.tex
paper: add section on real attributes.

@@
     \\
-        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \nonumber
+        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
     \\
@@
 which can be solved by any suitable MILP solver.
+
+\paragraph{Linearly dependent real attributes.} In many domains of interest,
+items are composed of both Boolean and real-valued attributes, where the latter
+depend entirely on the former. For instance, the price, weight and power
+consumption of a laptop may be modelled as real-valued attributes, and further
+they depend entirely on the choice of components.
+%
+In this setting, item configurations are composed of two parts: $\vx = (\vx_B;\vx_R)$, where $\vx_B$ is
+Boolean and $\vx_R$ is real-valued, and the relation between the two is
+entirely described by a non-negative\footnote{Non-negativity is required for
+the MILP translation to work.\stefano{better explanation please}} cost matrix
+$C\in\bbR^{\text{num costs} \times \text{num bools}}$, such that $\vx_R = C
+\vx_B$.  It is straightforward to extend the MILP formulation to this setting,
+as follows.
+
+First, we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility becomes
+%
+$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C^\top \vw_B, \vx_B \rangle $$
+%
+We let $\vv^i := \vw_B^i + C^\top \vv_B^i$, and $q^{ij}_z := (\vw^i_B + C^\top \vw^i_R)_z x^j_z$,
+and substitute them in the MILP problem to $\vw$ and $p^{ij}_z$, respectively.
+All constraints remain the same, except Eq.~\ref{eq:wbounds2}, which must be
+rewritten in terms of $\vv^i$:
+%
+$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C^\bot \vw_R) \le \vv^i \le (\vw_B^\top + C^\top \vw_R^\top)$$
+
+\stefano{improve please}
 
@@
 The simplest alternative is to query the user $\frac{1}{2}n(n-1)$ pairs.
-
-\paragraph{Generalization to real variables.}
-
-\stefano{WRITEME}
 

### 2016-01-17 ijcai16.py
ijcai16: fixup minor issue with the PC dataset.

@@ -294,7 +294,7 @@ def run_pc(has_costs):
         "seed": 0,
     })
 
-    dataset = PCDataset(has_costs=has_costs)
+    dataset = setmargin.PCDataset(has_costs=has_costs)
 
     for config in CONFIGS.iterate():
 

### 2016-01-17 setmargin/solver.py
setmargin/solver: reconstruct full items for the cost case.

@@ -95,7 +95,6 @@ class Solver(object):
         assert item.shape == (num_bools + num_reals,)
         return item
 
-
     def compute_best_score(self, dataset, user):
         """Finds the highest-scoring item in the dataset according to the user.
 
@@ -334,6 +333,12 @@ class Solver(object):
                 output_ws[i,z] = ws[i,z].x
                 output_xs[i,z] = xs[i,z].x
 
+        full_output_xs = []
+        for i in range(set_size):
+            x = [xs[i,z] for z in range(num_bools)]
+            full_output_xs.append(self._compose_item(dataset, x))
+        full_output_xs = np.array(full_output_xs)
+
         output_ps = np.zeros((set_size, set_size, num_bools))
         output_scores = np.zeros((set_size, set_size))
         for i in range(set_size):
@@ -360,6 +365,8 @@ class Solver(object):
             {}
             xs =
             {}
+            full xs =
+            {}
             ps =
             {}
             scores =
@@ -367,13 +374,13 @@ class Solver(object):
             slacks =
             {}
             margins = {}
-            """).format(set_size, output_ws, output_xs, output_ps,
-                        output_scores, output_slacks, output_margins)
+            """).format(set_size, output_ws, output_xs, full_output_xs,
+                        output_ps, output_scores, output_slacks, output_margins)
 
         if any(np.linalg.norm(w) == 0 for w in output_ws):
             print "Warning: null weight vector found!"
 
-        assert all(dataset.is_item_valid(x) for x in output_xs)
+        assert all(dataset.is_item_valid(x) for x in full_output_xs)
 
         debug_scores = np.dot(output_ws, output_xs.T)
         if (np.abs(output_scores - debug_scores) >= 1e-10).any():
@@ -385,4 +392,4 @@ class Solver(object):
                 {}
                 """).format(output_scores, debug_scores)
 
-        return output_ws, output_xs
+        return output_ws, full_output_xs

### 2016-01-17 setmargin/datasets.py
setmargin/datasets: check for negative values in is_item_valid().

@@ -46,8 +46,10 @@ class Dataset(object):
         return zs_in_domains
 
     def is_item_valid(self, x):
+        if (x < -1e-10).any():
+            return False
         for zs_in_domain in self.get_zs_in_domains():
-            if sum(x[z] for z in zs_in_domain) != 1:
+            if sum(x[zs_in_domain]) != 1:
                 return False
         if self.costs is not None:
             x, c = x[:self.num_bools()], x[self.num_bools():]

### 2016-01-17 setmargin/solver.py
setmargin/solver: compute w_max for real.

@@ -6,7 +6,6 @@ import gurobipy as grb
 from gurobipy import GRB
 from textwrap import dedent
 
-MAX_W_Z = 1
 ENABLE_LP_DUMP = False
 
 status_to_reason = {
@@ -160,6 +159,11 @@ class Solver(object):
         num_bools = dataset.num_bools()
         num_reals = dataset.num_reals()
 
+        w_top = np.ones(num_bools)
+        if num_reals > 0:
+            w_top += np.dot(dataset.costs.T, np.ones(num_reals))
+        w_max = np.max(w_top)
+
         model = grb.Model("setmargin")
         model.params.Seed = 0
         if self._threads is not None:
@@ -255,7 +259,7 @@ class Solver(object):
         # Eq. 11
         for i in range(set_size):
             for z in range(num_bools):
-                model.addConstr(ps[i,i,z] <= (MAX_W_Z * xs[i,z]))
+                model.addConstr(ps[i,i,z] <= (w_max * xs[i,z]))
 
         # Eq. 12
         for i in range(set_size):
@@ -266,12 +270,12 @@ class Solver(object):
         for i in range(set_size):
             for j in range(i) + range(i+1, set_size):
                 for z in range(num_bools):
-                    model.addConstr(ps[i,j,z] >= (ws[i,z] - 2 * MAX_W_Z * (1 - xs[j,z])))
+                    model.addConstr(ps[i,j,z] >= (ws[i,z] - 2 * w_max * (1 - xs[j,z])))
 
         # Eq. 15
         for i in range(set_size):
             for z in range(num_bools):
-                model.addConstr(ws[i,z] <= MAX_W_Z)
+                model.addConstr(ws[i,z] <= w_max)
 
         # Eq. 18a
         for i in range(set_size):

### 2016-01-17 setmargin/setmargin.py
setmargin/{setmargin,user}: remove useless parameter.

@@ -144,7 +144,7 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
         ws, xs = solver.compute_setmargin(dataset, answers, set_size, alphas)
 
         # Update the user answers
-        new_answers, num_queries = user.query_set(ws, xs, old_best_item)
+        new_answers, num_queries = user.query_set(xs, old_best_item)
         num_identical_answers = 0
         for xi, xj, sx in new_answers:
             for zi, zj, sz in answers:

### 2016-01-17 setmargin/user.py

@@ -95,11 +95,10 @@ class User(object):
         """
         return self.query_diff(np.dot(self.w, xi.T - xj.T))
 
-    def query_set(self, ws, xs, old_best_item):
+    def query_set(self, xs, old_best_item):
         """Queries the user about the provided set of items.
 
         :param user: the user.
-        :param ws: the estimated user preference(s) at the current iteration.
         :param xs: the estimated best item(s) at the current iteration.
         :param old_best_item: the estimated best item at the previous iteration.
         :returns: WRITEME

### 2016-01-17 setmargin/datasets.py
setmargin/datasets: add compose_item() method.

@@ -61,6 +61,13 @@ class Dataset(object):
                     return False
         return True
 
+    def compose_item(self, x):
+        assert x.shape == (self.num_bools(),)
+        if self.num_reals() > 0:
+            x = np.hstack((x, np.dot(self.costs, x)))
+        assert x.shape == (self.num_bools() + self.num_reals(),)
+        return x
+
 class SyntheticDataset(Dataset):
     def __init__(self, domain_sizes):
         super(SyntheticDataset, self).__init__(domain_sizes, None, None)

### 2016-01-17 setmargin/solver.py
setmargin/solver: use the new compose() method.

@@ -80,20 +80,6 @@ class Solver(object):
             for body, head in dataset.x_constraints:
                 model.addConstr((1 - x[body]) + grb.quicksum([x[atom] for atom in head]) >= 1)
 
-    def _compose_item(self, dataset, x):
-        """
-        :param dataset:
-        :param x: the Boolean part of an item as a list of Gurobi variables.
-        :returns:
-        """
-        num_reals, num_bools = dataset.num_reals(), dataset.num_bools()
-        item = np.array([x[z].x for z in range(num_bools)])
-        assert item.shape == (num_bools,)
-        if num_reals > 0:
-            item = np.hstack((item, np.dot(dataset.costs, item)))
-        assert item.shape == (num_bools + num_reals,)
-        return item
-
     def compute_best_score(self, dataset, user):
         """Finds the highest-scoring item in the dataset according to the user.
 
@@ -138,7 +124,8 @@ class Solver(object):
         except:
             raise RuntimeError("optimization failed! {}".format(status_to_reason[model.status]))
 
-        best_item = self._compose_item(dataset, x)
+        best_item = np.array([x[z].x for z in range(num_bools)])
+        best_item = dataset.compose_item(best_item)
         assert dataset.is_item_valid(best_item)
 
         return best_score, best_item
@@ -339,8 +326,8 @@ class Solver(object):
 
         full_output_xs = []
         for i in range(set_size):
-            x = [xs[i,z] for z in range(num_bools)]
-            full_output_xs.append(self._compose_item(dataset, x))
+            x = np.array([xs[i,z].x for z in range(num_bools)])
+            full_output_xs.append(dataset.compose_item(x))
         full_output_xs = np.array(full_output_xs)
 
         output_ps = np.zeros((set_size, set_size, num_bools))

### 2016-01-17 setmargin/setmargin.py
setmargin/setmargin: finish off support for cost variables!

@@ -58,15 +58,15 @@ def crossvalidate(dataset, solver, answers, set_size, debug):
 
     return alphas
 
-def print_answers(queries, hidden_w):
+def print_answers(user, answers):
     message = ["answers ="]
-    for xi, xj, sign in queries:
+    for xi, xj, sign in answers:
         relation = {-1:"<", 0:"~", 1:">"}[sign]
-        score_xi = np.dot(hidden_w, xi.T)[0]
-        score_xj = np.dot(hidden_w, xj.T)[0]
+        utility_xi = user.utility(xi)
+        utility_xj = user.utility(xj)
         message.append("  {} ({:6.3f}) {} ({:6.3f}) {} -- diff {:6.3f}" \
-                           .format(xi, score_xi, relation, score_xj, xj,
-                                   score_xi - score_xj))
+                           .format(xi, utility_xi, relation, utility_xj, xj,
+                                   utility_xi - utility_xj))
     print "\n".join(message)
 
 def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
@@ -102,6 +102,7 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
         raise ValueError("crossval_interval must be positive")
 
     best_score, best_item = solver.compute_best_score(dataset, user)
+    assert best_item.shape == (dataset.num_bools(),)
     user_w_norm = np.linalg.norm(user.w.ravel())
 
     if debug:
@@ -131,7 +132,7 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
             ============
             """).format(t)
 
-            print_answers(answers, user.w)
+            print_answers(user, answers)
 
         old_time = time.time()
 
@@ -141,7 +142,8 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
                                    debug)
 
         # Solve the set_size=k case
-        ws, xs = solver.compute_setmargin(dataset, answers, set_size, alphas)
+        _, xs = solver.compute_setmargin(dataset, answers, set_size, alphas)
+        assert xs.shape == (set_size, dataset.num_bools())
 
         # Update the user answers
         new_answers, num_queries = user.query_set(xs, old_best_item)
@@ -159,14 +161,15 @@ def run(dataset, user, solver, set_size, max_iterations=100, max_answers=100,
         elapsed = time.time() - old_time
 
         if debug:
-            print_answers(answers, user.w)
[…]

### 2016-01-18 setmargin/user.py
setmargin/user: finish off support for cost variables!

@@ -25,19 +25,20 @@ class User(object):
     def __init__(self, dataset, sampling_mode="uniform", ranking_mode="all_pairs",
                  is_deterministic=False, is_indifferent=False, w=None,
                  rng=None):
+        self.dataset = dataset
         self.ranking_mode = ranking_mode
         self.is_deterministic = is_deterministic
         self.is_indifferent = is_indifferent
         self._rng = check_random_state(rng)
-        self.w = self._sample(dataset, sampling_mode) if w is None else w
+        self.w = self._sample(sampling_mode) if w is None else w
 
     def __str__(self):
         return "User({} D={} I={} RM={})".format(self.w, self.is_deterministic,
                                                  self.is_indifferent,
                                                  self.ranking_mode)
 
-    def _sample(self, dataset, sampling_mode, sparsity=0.2):
-        num_vars = dataset.num_vars()
+    def _sample(self, sampling_mode, sparsity=0.2):
+        num_vars = self.dataset.num_vars()
         if sampling_mode in ("uniform", "uniform_sparse"):
             w = self._rng.uniform(1, 100, size=num_vars)
         elif sampling_mode in ("normal", "normal_sparse"):
@@ -46,7 +47,7 @@ class User(object):
             raise ValueError("invalid sampling_mode")
         if sampling_mode.endswith("sparse"):
             mask = np.zeros(w.shape)
-            for attr0, attr1 in dataset.get_domain_ranges():
+            for attr0, attr1 in self.dataset.get_domain_ranges():
                 domain_size = attr1 - attr0
                 domain_mask = np.zeros(domain_size)
                 num_ones = max(1, int(domain_size * sparsity))
@@ -55,12 +56,6 @@ class User(object):
             # XXX the cost weights are not masked at all
             w[mask == 0] = 0
         return w.reshape(1, -1)
-
-    def utility(self, dataset, x):
-        return np.dot(self.w.ravel(), x)
-
-    def utility_loss(self, dataset, xi, xj):
-        return self.utility(dataset, xi) - self.utility(dataset, xj)
 
     def query_diff(self, diff):
         """Queries the user about a single pairwise choice.
@@ -85,6 +80,10 @@ class User(object):
             return 1
         return -1
 
+    def utility(self, x):
+        assert x.shape == (self.dataset.num_bools(),)
+        return np.dot(self.w.ravel(), self.dataset.compose_item(x))
+
     def query(self, xi, xj):
[…]

### 2016-01-18 setmargin/solver.py
setmargin/solver: finish off support for cost variables!

@@ -8,7 +8,7 @@ from textwrap import dedent
 
 ENABLE_LP_DUMP = False
 
-status_to_reason = {
+STATUS_TO_REASON = {
     1: "LOADED",
     2: "OPTIMAL",
     3: "INFEASIBLE",
@@ -24,7 +24,7 @@ status_to_reason = {
     13: "SUBOPTIMAL",
 }
 
-def gudot(x, z):
+def _gudot(x, z):
     """
     :param x: a list of scalars or Gurobi expressions.
     :param z: a list of scalars or Gurobi expressions.
@@ -33,7 +33,7 @@ def gudot(x, z):
     assert len(x) == len(z)
     return grb.quicksum([x[i] * z[i] for i in range(len(x))])
 
-def gubilinear(x, a, z):
+def _gubilinear(x, a, z):
     """
     :param x: a list of scalars or Gurobi expressions.
     :param a: a numpy array.
@@ -42,8 +42,8 @@ def gubilinear(x, a, z):
     """
     assert len(x) == a.shape[0]
     assert len(z) == a.shape[1]
-    q = [gudot(a[i], z) for i in range(len(x))]
-    return gudot(x, q)
+    q = [_gudot(a[i], z) for i in range(len(x))]
+    return _gudot(x, q)
 
 class Solver(object):
     """Set-margin solver based on the Gurobi MILP solver.
@@ -71,9 +71,13 @@ class Solver(object):
             model.write(fp.name)
 
     def _add_item_constraints(self, model, dataset, x):
-        """Add one-hot and other item constraints."""
-        zs_in_domains = dataset.get_zs_in_domains()
-        for zs_in_domain in zs_in_domains:
+        """Add one-hot and Horn item constraints.
+
+        :param model: the Gurobi model.
+        :param dataset: the dataset.
+        :param x: a list of Guorbi expressions.
+        """
+        for zs_in_domain in dataset.get_zs_in_domains():
             model.addConstr(grb.quicksum([x[z] for z in zs_in_domain]) == 1)
 
         if dataset.x_constraints is not None:
@@ -90,7 +94,7 @@ class Solver(object):
 
         :param dataset: the dataset.
         :param user: the user.
-        :returns: the best score and the corresponding best item.
+        :returns: the best score and the corresponding best bool item.
         """
         num_bools, num_reals = dataset.num_bools(), dataset.num_reals()
 
@@ -107,10 +111,10 @@ class Solver(object):
[…]

### 2016-01-18 setmargin/datasets.py
setmargin/datasets: normalize cost matrix to 1.

@@ -100,7 +100,7 @@ class DebugConstraintDataset(Dataset):
 class DebugCostDataset(Dataset):
     def __init__(self, domain_sizes, num_costs=2, rng=None):
         rng = check_random_state(rng)
-        costs = rng.uniform(-1, 1, size=(num_costs, sum(domain_sizes)))
+        costs = rng.uniform(0, 1, size=(num_costs, sum(domain_sizes)))
         super(DebugCostDataset, self).__init__(domain_sizes, costs, None)
 
 class PCDataset(Dataset):
@@ -138,8 +138,8 @@ class PCDataset(Dataset):
         self.domain_of = domain_of
 
         costs_of = {
-            "Manufacturer": [100, 0, 100, 50, 0, 0, 50, 50],
-            "CPU": map(int, [
+            "Manufacturer": map(float, [100, 0, 100, 50, 0, 0, 50, 50]),
+            "CPU": [
                 # AMD Athlon
                 1.4*100, 1.4*130,
                 # AMD Duron
@@ -156,31 +156,33 @@ class PCDataset(Dataset):
                 1.4*27, 1.4*30, 1.4*40, 1.4*45, 1.4*50, 1.4*55, 1.4*60, 1.4*70,
                 # PowerPC G4
                 1.6*70, 1.6*73
-            ]),
-            "Monitor": [
+            ],
+            "Monitor": map(float, [
                 0.6*100, 0.6*104, 0.6*120, 0.6*133, 0.6*140, 0.6*150, 0.6*170,
                 0.6*210
-            ],
-            "Type": [80, 0, 120],
+            ]),
+            "Type": map(float, [50, 0, 80]),
             "Memory": [
                 0.8*64, 0.8*128, 0.8*160, 0.8*192, 0.8*256, 0.8*320, 0.8*384,
                 0.8*512, 0.8*1024, 0.8*2048
             ],
-            "HDSize": [
+            "HDSize": map(float, [
                 4*8, 4*10, 4*12, 4*15, 4*20, 4*30, 4*40, 4*60, 4*80, 4*120
-            ],
+            ]),
         }
 
         self.attributes = sorted(domain_of.keys())
 
         domain_sizes = [len(domain_of[attr]) for attr in self.attributes]
 
-        cost = []
+        cost, max_cost = [], 0.0
         assert len(domain_of) == len(costs_of)
         for attr in self.attributes:
             assert len(domain_of[attr]) == len(costs_of[attr])
[…]

### 2016-01-18 draw.py
draw: update.

@@ -14,53 +14,68 @@ if len(sys.argv) != 3:
 fontP = FontProperties()
 fontP.set_size('small')
 
-fig, ax = plt.subplots(1, 1)
-ax.set_xlabel("Number of queries")
-ax.set_ylabel("Average loss over trials")
+figs, axs = {}, {}
 
-paths = sorted(glob(os.path.join(sys.argv[1], "results_*_loss_matrix.txt")))
+figs[True], axs[True] = plt.subplots(1, 1)
+axs[True].set_xlabel("Number of queries")
+axs[True].set_ylabel("Average loss per answer")
+
+figs[False], axs[False] = plt.subplots(1, 1)
+axs[False].set_xlabel("Number of queries")
+axs[False].set_ylabel("Average time per answer (in seconds)")
+
+paths = sorted(glob(os.path.join(sys.argv[1], "results_*_matrix.txt")))
 
 # Colors taken from http://tango.freedesktop.org/Tango_Icon_Theme_Guidelines
 COLOR_OF = {
-#    ("#EDD400", "#FCE94F"), # yellow
-    2: ("#CE5C00", "#F57900"), # orange
-#    ("#4E9A06", "#73D216"), # green
-    3: ("#204A87", "#3465A4"), # blue
-#    ("#5C3566", "#75507B"), # violet
-#    2: ("#A40000", "#CC0000"), # red
+    2: ("#CE5C00", "#F57900"),
+    3: ("#204A87", "#3465A4"),
+    4: ("#5C3566", "#75507B"),
 }
 
-max_max_x = 0
-max_max_y = 0
+max_max = {}
 for path in paths:
     parts = os.path.basename(path).split("__")
     set_size = int(parts[1].split("=")[1])
-    if not set_size in (2, 3):
+    if not set_size in (2, 3, 4):
         continue
+    is_loss = parts[-1].split("_")[1] == "loss"
 
-    loss_matrix = np.loadtxt(path)
-    if loss_matrix.shape[1] > 100:
-        loss_matrix = loss_matrix[:,:100]
-    num_trials, max_queries = loss_matrix.shape
+    matrix = np.loadtxt(path)
+    if matrix.shape[1] > 100:
+        matrix = matrix[:,:100]
+    num_trials, max_queries = matrix.shape
 
     xs = np.arange(max_queries)
-    ys = np.median(loss_matrix, axis=0)
-    yerrs = np.std(loss_matrix, axis=0)
-
[…]

### 2016-01-18 draw_all.sh

@@ -2,7 +2,13 @@
 
 for domain_size in `seq 3 6`; do
     for sampling_mode in uniform uniform_sparse normal normal_sparse; do
-        echo "drawing $domain_size $sampling_mode"
-        ./draw.py results/synthetic_${domain_size}/${sampling_mode} paper/figures/synthetic_${domain_size}_${sampling_mode}.png
+        ipath="results/final_cv@1/synthetic_${domain_size}/${sampling_mode}"
+        opath="paper/figures/synthetic_${domain_size}_${sampling_mode}_cv1"
+        echo "drawing $domain_size $sampling_mode ($ipath $opath)"
+        ./draw.py $ipath $opath
+        ipath="results/final_cv@5/synthetic_${domain_size}/${sampling_mode}"
+        opath="paper/figures/synthetic_${domain_size}_${sampling_mode}_cv5"
+        echo "drawing $domain_size $sampling_mode ($ipath $opath)"
+        ./draw.py $ipath $opath
     done
 done

### 2016-01-18 paper/ijcai16.tex
paper: update the plots with the latest figures.

@@
     \centering
-    \begin{tabular}{ccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5}
+    \begin{tabular}{cccc}
+        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform}
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv1_loss}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse}
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv1_loss}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal}
+        \includegraphics[width=12em]{figures/synthetic_3_normal_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_cv1_loss}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse}
+        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv1_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv1_loss}
     \end{tabular}
-    \caption{Results on the synthetic dataset. Each row represents a different
+    \caption{Results on the synthetic dataset, with crossvalidation at each step. Each row represents a different
     setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: blue means $k=2$, greed $k=3$, and red $k=4$.}
+    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+\end{figure}
+
+\begin{figure}[b]
+    \centering
+    \begin{tabular}{cccc}
+        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv5_loss}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv5_loss}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_normal_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_cv5_loss}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv5_loss}
+    \end{tabular}
+    \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
+    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
+    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+\end{figure}
+
+\begin{figure}[b]
+    \centering
+    \begin{tabular}{cccc}
+        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv1_time}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv1_time}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_normal_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_cv1_time}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv1_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv1_time}
+    \end{tabular}
+    \caption{Results on the synthetic dataset, with crossvalidation at each step. Each row represents a different
+    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
+    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+\end{figure}
+
+\begin{figure}[b]
+    \centering
+    \begin{tabular}{cccc}
+        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv5_time}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv5_time}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_normal_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_cv5_time}
+        \\
+        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv5_time} &
+        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv5_time}
+    \end{tabular}
+    \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
+    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
+    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
 \end{figure}

### 2016-01-18 ijcai16.py
ijcai16: simplify.

@@ -307,10 +307,13 @@ def run_pc(has_costs):
         infos = solve(dataset, config)
         dump_and_draw("pc_with_costs" if has_cost else "pc_no_costs", config, infos)
 
-def run_from_command_line():
+def main():
     import argparse
 
+    np.seterr(all="raise")
+
     parser = argparse.ArgumentParser(formatter_class=argparse.ArgumentDefaultsHelpFormatter)
+    parser.add_argument("command", type=str, help="run")
     parser.add_argument("dataset", type=str, help="dataset")
     parser.add_argument("-T", "--num_trials", type=int, default=20,
                         help="number of trials")
@@ -362,7 +365,21 @@ def run_from_command_line():
 
     args = parser.parse_args()
 
+    if args.command == "run-synthetic":
+        run_synthetic(False)
+    elif args.command == "run-synthetic-variance":
+        run_synthetic(True)
+    elif args.command == "run-pc-no-costs":
+        run_pc(False)
+    elif args.command == "run-pc-with-costs":
+        run_pc(True)
+    elif args.command == "run":
+        pass
+    else:
+        raise ValueError("invalid IJCAI experiment name.")
+
     argsdict = vars(args)
+    del argsdict["command"]
     argsdict["dataset"] = args.dataset
     try:
         argsdict["tol"] = float(argsdict["tol"])
@@ -395,17 +412,4 @@ def run_from_command_line():
     dump_and_draw(dataset_name, config, infos)
 
 if __name__ == "__main__":
-    np.seterr(all="raise")
-    if len(sys.argv) == 2:
-        if sys.argv[1] == "synthetic":
-            run_synthetic(False)
-        elif sys.argv[1] == "synthetic-variance":
-            run_synthetic(True)
-        elif sys.argv[1] == "pc-no-costs":
-            run_pc(False)
-        elif sys.argv[1] == "pc-with-costs":
-            run_pc(True)
-        else:
-            raise ValueError("invalid IJCAI experiment name.")
-    else:
-        run_from_command_line()
+    main()

### 2016-01-18 ijcai16.py
ijcai16: remove precrossvalidation, it's useless.

@@ -14,12 +14,6 @@ from multiprocessing import cpu_count
 
 import setmargin
 
-ALL_ALPHAS = list(product(
-         [100.0, 20.0, 15.0, 12.0, 11.0, 10.0, 9.0, 8.0, 7.5, 5.0, 1.0],
-         [20.0, 10.0, 5.0, 1.0, 0.5, 0.1, 0.05, 0.01],
-         [20.0, 10.0, 5.0, 1.0, 0.5, 0.1, 0.05, 0.01],
-))
-
 class Grid(object):
     def __init__(self, d):
         self.__dict__.update(d)
@@ -146,42 +140,6 @@ def dump_and_draw(dataset_name, config, infos):
     del fig
     del ax
 
-def precrossvalidate(dataset, config, solver, users):
-    """Use the last 20 users for parameter selection."""
-    target_users = users[len(users) - 20:]
-    assert len(target_users) == 20
-
-    loss_alphas = []
-    for alphas in ALL_ALPHAS:
-
-        print "pre-crossvalidating", alphas
-
-        kfold = KFold(len(target_users), n_folds=5)
-
-        losses = []
-        for train_set, test_set in kfold:
-
-            for i in train_set:
-                infos = setmargin.run(dataset, target_users[i], solver,
-                                      config.set_size, alphas=alphas,
-                                      max_iterations=config.max_iterations,
-                                      max_answers=config.max_answers,
-                                      tol=config.tol, debug=config.debug)
-                # XXX we only consider the true utility loss at the last
-                # iteration
-                losses.append(infos[-1][1])
-
-        loss_alphas.append((sum(losses) / len(losses), alphas))
-
-    loss_alphas = sorted(loss_alphas)
-    best_alphas = loss_alphas[0][1]
-
-    print "pre-crossvalidation: best alphas = ", alphas
-    for loss, alphas in loss_alphas:
-        print alphas, ":", loss
-
-    return best_alphas
-
 def solve(dataset, config, ws=None):
     solver = setmargin.Solver(multimargin=config.multimargin,
                               threads=config.threads, debug=config.debug)
@@ -204,8 +162,6 @@ def solve(dataset, config, ws=None):
         for user in users:
             print user
 
-    if config.precrossval:
-        alphas = precrossvalidate(dataset, config, solver, users)
     if config.crossval:
         alphas = "auto"
     else:
@@ -238,7 +194,6 @@ def run_synthetic(same_user):
[…]

### 2016-01-18 paper/ijcai16.tex
paper: update.

@@
 
-\stefano{Paolo, feel free to improve the text and especially the citations
-as you see fit.}
+\stefano{Paolo, Andrea, feel free to improve the text and especially the
+citations as you see fit.}
 
-\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors, and
-calligraphic capital letters $\calX$ for sets. We frequently abbreviate the set
-$\{ x^i \}_{i=1}^n$ as $\{ x^i \}$ whenever the range of the index $i$ is clear
-from the context, and use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
-The $\ell_1$ vector norm is denoted by $\|\vx\|_1 := \sum_z |x_z|$.
+\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
+uppercase letters $X$ for matrices, and calligraphic capital letters $\calX$
+for sets. We frequently abbreviate the set $\{ x^i \}_{i=1}^n$ as $\{ x^i \}$
+whenever the range of the index $i$ is clear from the context, and use $[n]$ as
+a shorthand for the set $\{1, \ldots, n\}$. We write $\|\vx\|_1 := \sum_z |x_z|$
+to indicate the $\ell_1$ vector norm, and $\langle \cdot, \cdot \rangle$ to
+denote the usual dot product.
 
@@
 of configurations $\vx = (x_1, \ldots, x_n)$ over $n$ features. For the time
-being we will focus on binary features only, i.e. $x_z\in\{0,1\}$, to keep
-matters simple. This is a common choice for preference elicitation methods,
-e.g.~\cite{guo2010real,viappiani2010optimal}; discrete features are handled by
-means of a simple one-hot encoding. Partial support for continuous features
-will be discussed later on.
+being we will focus on binary features only, i.e. $x_z\in\{0,1\}$ for all
+$z\in[n]$, to keep matters simple. This is a common choice for preference
+elicitation methods~\cite{guo2010real,viappiani2010optimal}; discrete features
+can be handled for instance by means of a simple one-hot encoding. Support for
+linearly dependent continuous features will be discussed later on.
 
-We further assume that the set of {\em feasible} configurations is expressed as
+We further assume that the set of {\em feasible} configurations, denoted by
+$\calX_\text{feasible}$, is expressed as
 a conjunction of linear constraints. This is not a severe restriction:
 arbitrary propositional constraints can be trivially encoded in this setting,
-e.g. the ``Boolean disjunction'' of two binary variables $x_1 \lor x_2$ can be
-rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations is
-denoted by $\calX_\text{feasible}$.
+e.g. under the canonical mapping of $True$ to $1$ and $False$ to $0$, the
+Boolean disjunction of two binary variables $x_1 \lor x_2$ can be rewritten as
+$x_1 + x_2 \ge 1$.
 
@@
 vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
-$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we
-require all weights to be {\em non-negative} and {\em bounded}: the individual
-weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
+$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the remainder of the paper we
+require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
+weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
 $[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
@@
 learning system, and must be estimated from pairwise choices elicited from the
-user. The set of examples consists of pairwise comparisons between a more
-preferred configuration $\vy_+$ and a less preferred one $\vy_-$. Each preference
-induces a ranking constraint $\langle \vw, \vy_+ - \vy_- \rangle \ge 0$.
-Details on the actual query selection strategy will be provided in
-Section~\ref{sec:querystrategy}.
-\stefano{add indifference}
+user.  We consider three possible outcomes for a pairwise comparison between
+two configurations $\vx$ and $\vx'$: either $\vx$ is preferred to $\vx'$ (written
+$\vx \succ \vx'$), $\vx'$ is preferred to $\vx$ ($\vx \prec \vx'$), or there is
+no clear preference between the two items ($\vx \approx \vx'$). We write
+$\calD$ to denote the full set of answers elicited from the user. Queries are
+typically selected so as to maximize some measure of informativeness of the
+resulting responses.  Details on our actual query selection strategy will be
+provided in Section~\ref{sec:formulation}.
 
@@
 In the set-margin approach, the goal is twofold. First, for any given set size
-$k\geq 1$, we want to find a {\em set} of $k$ weight vectors $\vw^{1},
-\ldots, \vw^{k}$, chosen so that all preferences (examples) are satisfied by
-the largest possible margin. This is achieved by translating all binary
-preferences into inequalities of the form $\langle \vw^i, \vy^h_+ - \vy^h_- \rangle
-\geq M$, where $h$ ranges over the examples, and maximizing the margin $M$.
-Non-separable datasets, which occur in practice due to user inconsistency, are
-handled by introducing slack variables. The updated constraints look as follows:
+$k\geq 1$, we want to find a {\em set} of $k$ weight vectors $\vw^{1}, \ldots,
+\vw^{k}$, chosen so that all user-provided preferences are satisfied by the
+largest possible margin (i.e. all weight vector are consistent with respect to
+the user responses $\calD$, modulo inconsistencies) and so that they are
+maximally diverse.  Second, we want to construct a {\em set} of $k$
+configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration $\vx^{i}$
+is the ``best'' possible option when evaluated according to the corresponding
+$\vw^{i}$, and consequently maximally diverse as well. These options will be
+later employed to formulate the user queries.
+
+The first goal achieved by translating all pairwise preferences $\calD$ into
+ranking constraints: preferences of the form $\vy^h_+ \succ \vy^h_-$ become
+linear inequalities of the form $\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq
+M$, where $M$ is the {\em margin} variable (which we aim at maximizing) and $h$
+ranges over the responses.  Non-separable datasets, which occur in practice due
+to user inconsistency, are handled by introducing slack variables (whose sum we
+aim at minimizing). When augmented with the slacks, the above inequalities take
+the form:
 %
@@
 where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
-for violating the ranking of example $h$.
+for violating the ranking of example $h$. Indifference preferences, i.e.
+$\vy^h_1 \approx \vy^h_2$, are translated as $|\langle \vw^i, \vy^h_1 - \vy^h_2 \rangle| < \varepsilon^i_h$:
+the slack increases with the difference between the estimated utility of the
+two options.
 
-Second, we want to produce a set of $k$ configurations $\vx^{1}, \ldots,
-\vx^{k}$, so that each configuration $\vx^{i}$ is the ``best'' possible
-option when evaluated according to the corresponding $\vw^{i}$. Ideally,
-this requirement would be implemented as:
+Ideally, the second goal would be implemented as:
 %
@@
 
-The third part requires requires some further explanation. We choose an
-$\ell_1$ regularizer in order to facilitate learning {\em sparse} weight
-vectors: it is reasonable to assume that a user will have strong opinions
-over some attributes, but will be indifferent over most of them.
-\stefano{Andrea, improve the above please}
+The third part of the objective does require a more in depth explanation. In
+many scenarios the user will have strong preferences about some attributes, but
+will be indifferent to most of them~\cite{}. The $\ell_1$ penalty is frequently
+used to improve the sparsity of learned models~\cite{}, with consequent gains
+in generalization ability and efficiency, as confirmed by our empirical
+findings (see Section~\ref{sec:experiments}). \stefano{add refs}
 
@@
 try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
-possible. \stefano{this is not necessarily the case, $M$ is bounded by the
-$y$'s as well.}
+possible.
 
@@
 the second term is at most $0$. Since $p^{i,j}_z$ is being minimized, in both
-cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z x^i_z$.
+cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
+x^i_z$\footnote{Since $M$ is upper-bounded by Eq.~\ref{eq:wyconstr}, in some
+cases the $p^{i,j}_z$ variables do not attain the lower bound. As a
+consequence, the MILP reformulation of Eq.~\ref{eq:wxconstr} is marginally
+inexact. Despite this issue, the MILP formulation acceptably in practice.}.
 
@@
 
-\paragraph{Linearly dependent real attributes.} In many domains of interest,
-items are composed of both Boolean and real-valued attributes, where the latter
-depend entirely on the former. For instance, the price, weight and power
-consumption of a laptop may be modelled as real-valued attributes, and further
-they depend entirely on the choice of components.
-%
-In this setting, item configurations are composed of two parts: $\vx = (\vx_B;\vx_R)$, where $\vx_B$ is
-Boolean and $\vx_R$ is real-valued, and the relation between the two is
-entirely described by a non-negative\footnote{Non-negativity is required for
-the MILP translation to work.\stefano{better explanation please}} cost matrix
-$C\in\bbR^{\text{num costs} \times \text{num bools}}$, such that $\vx_R = C
-\vx_B$.  It is straightforward to extend the MILP formulation to this setting,
-as follows.
-
-First, we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility becomes
-%
-$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C^\top \vw_B, \vx_B \rangle $$
-%
-We let $\vv^i := \vw_B^i + C^\top \vv_B^i$, and $q^{ij}_z := (\vw^i_B + C^\top \vw^i_R)_z x^j_z$,
-and substitute them in the MILP problem to $\vw$ and $p^{ij}_z$, respectively.
-All constraints remain the same, except Eq.~\ref{eq:wbounds2}, which must be
-rewritten in terms of $\vv^i$:
-%
-$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C^\bot \vw_R) \le \vv^i \le (\vw_B^\top + C^\top \vw_R^\top)$$
-
-\stefano{improve please}
-
-\subsection{Set-wise Preference Elicitation}
-
-\stefano{WRITEME}
-
-See Algorithm~\ref{alg:setmargin}.
+\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm follows the
+usual preference elicitation loop. Starting from an initially empty set of user
+responses $\calD$, it repeatedly solves the MILP problem above using the set
+$\calD$ to enforce ranking constraints on the weight vectors $\{\vw^i\}$. The
+generated configurations $\{\vx^i\}$, which are chosen to be as good as
+possible with respect to the estimated user preferences, and as diverse as
+possible, are then employed to formulate a set of user queries. The new replies
+are added to $\calD$ and the whole procedure is repeated. The algorithm
+terminates after a fixed number of iterations, or once another suitable
+condition is met. The procedure is sketched in Algorithm~\ref{alg:setmargin}.
 
@@
     \Procedure{SetMargin}{$k, \alpha, \beta, \gamma, T$}
-        \State $\calQ \gets \emptyset$
+        \State $\calD \gets \emptyset$
         \For{$t = 1, \ldots, T$}
-            \State \{$\vw^{i}, \vx^{i}\}_{i=1}^k \gets \text{{\sc Solve}}(\calQ, k, \alpha, \beta, \gamma)$
+            \State \{$\vw^{i}, \vx^{i}\}_{i=1}^k \gets \text{{\sc Solve}}(\calD, k, \alpha, \beta, \gamma)$
             \For{$\vx^{i},\vx^{j} \in \{ \vx^{1}, \ldots, \vx^{k} \} \; \text{{\bf s.t.}} \; i < j$}
-                \State $\calQ \gets \calQ \cup \text{{\sc QueryUser}}(\vx^{i},\vx^{j})$
+                \State $\calD \gets \calD \cup \text{{\sc QueryUser}}(\vx^{i},\vx^{j})$
             \EndFor
         \EndFor
-        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(\calQ, 1, \alpha, \beta, \gamma)$
+        \State $\vw^*, \vx^* \gets \text{{\sc Solve}}(\calD, 1, \alpha, \beta, \gamma)$
         \State ${\bf return}\; \vw^*, \vx^*$
@@
 
-\subsection{Query Strategy}
-\label{sec:querystrategy}
+\paragraph{Query selection.} Contrarily to most other preference elicitation
+methods, which generate a single candidate option per iteration, {\sc
+SetMargin} constructs a whole {\em set} of options. We thus can not leverage
+standard query selection strategies \stefano{which ones?}.
 
-\stefano{WRITEME}
+The simplest solution is to query the user about all $k \choose 2$ pairs of
+configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
+number of queries does not grow large in practice.  An appealing alternative
+involves using a modified sorting algorithm, such as merge-sort, to order the
+set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
+approach is very susceptible to noise: a single inconsistent user response may
+affect all successive comparisons, leading to a cascading effect. While more
+robust strategies may be conceived, empirically we found that the simpler
+quadratic scheme tends to work reasonably well.
 
-The simplest alternative is to query the user $\frac{1}{2}n(n-1)$ pairs.
+\paragraph{Linearly dependent real attributes.} In many domains of interest,
+items are composed of both Boolean and real-valued attributes, where the latter
+depend linearly  on the former. This is for instance the case for the price,
+weight and power consumption of a laptop model, which depend linearly on the
+choice of components.
+%
+In this setting, item configurations are composed of two parts: $\vx =
+(\vx_B;\vx_R)$, where $\vx_B$ is Boolean and $\vx_R$ is real-valued, and the
+relation between the two is entirely described by an appropriately sized
+non-negative cost matrix $C$, such that $\vx_R = C \vx_B$. It is
+straightforward to extend the MILP formulation to this setting, as follows.
+
+First, we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
+becomes:
+%
+$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C^\top \vw_B, \vx_B \rangle $$
+%
+The generalized problem is obtained by substituting $\vw^i$ with $\vv^i := \vw_B^i + C^\top \vv_B^i$ and $p^{i,j}_z$
+with $q^{i,j}_z := (\vw^i_B + C^\top \vw^i_R)_z x^j_z$ in the MILP problem.
+All constraints remain the same, except Eq.~\ref{eq:wbounds2}, which must be
+rewritten in terms of $\vv^i$:
+%
+$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C^\bot \vw_R) \le \vv^i \le (\vw_B^\top + C^\top \vw_R^\top)$$
+%
+As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
+to be non-negative.
 
@@
 
-In this section we provide extensive experimental evaluation of the {\sc
-SetMargin} algorithm. We implemented the {\sc SetMargin} algorithm using
-Python, leveraging Gurobi 6.5.0 as the MILP solver. Both the {\sc SetMargin}
-source code and the full experimental setup are available at \stefano{add URL}.
+We implemented the {\sc SetMargin} algorithm using Python, leveraging Gurobi
+6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
+and the full experimental setup are available at \stefano{add URL}.
 
@@
 
-We adopted the {\em indifference-augmented} Bradley-Torr user response model
+We adopt the {\em indifference-augmented} Bradley-Torr user response model
 introduced in~\cite{guo2010real}. When queried about a pair of configurations,
@@
 
-In all experiments we use an internal 5-fold cross-validation procedure to update the
-hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5 iterations. The
-hyperparameters are chosen as to minimize the ranking loss over the user
-answers collected so far. $\alpha$ is taken in $\{20, 15, 10, 5, 1\}$,
-while $\beta$ and $\gamma$ are taken in $\{10, 5, 1, 0.5, 0.1\}$.
+In all experiments we use an internal 5-fold cross-validation procedure to
+update the hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5
+iterations. The hyperparameters are chosen as to minimize the ranking loss over
+the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
+while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
 \stefano{explain that we can't turn off $\gamma$}
@@
 
-Users were simulated by drawing random utility vectors from four different
-distributions. The first two mimic those used in~\cite{guo2010real}: (1) a
-uniform distribution over $[1, 100]$ for each individual weight, and (2) a
-normal distribution with mean $25$ and covariance $\frac{25}{3}$\footnote{This
-procedure differs from the one reported in~\cite{guo2010real}; we opted for
-following the procedure implemented in their code. \stefano{more diplomacy needed}}.
-We further produced two novel {\em sparse} sets of weight vectors by deleting
-$80\%$ of the entries of the vectors sampled in the uniform and normal settings
-uniformly at random \stefano{Paolo, is this correct?}. We sampled $100$
-vectors for each setting. All methods were evaluated on the very same weight
-vectors.
+Users were simulated by drawing $20$ random utility vectors from each of four
+different distributions. The first two mimic those used
+in~\cite{guo2010real}\footnote{The normal distribution reported
+in~\cite{guo2010real} differs slightly from the one implemented in the code of
+their experimental setting. In this paper we use the latter.}:
+(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
+a normal distribution with mean $25$ and covariance
+$\frac{25}{3}$.  We further produced two novel
+{\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
+vectors sampled in the uniform and normal settings uniformly at random
+\stefano{Paolo, is this correct?}. All methods were evaluated on the very same
+weight vectors.
+
+In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
+loss:
+%
+$$ \left(\max_{\vx\in\calX_\text{feasible}} u(\vx)) - u(\vx^{i}) \right) $$
+%
+varying the number of collected user responses. The thick line is the median
+utility loss taken over all 20 users, while the shaded area represents the
+variance.
 
@@
 
+\begin{figure}[b]
+    \centering
+    \begin{tabular}{cc}
+        {\bf loss} & {\bf time}
+        \\
+        \multicolumn{2}{c}{{\bf uniform sparse (no costs)}}
+        \\
+        \includegraphics[width=12em]{figures/pc_no_costs_uniform_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/pc_no_costs_uniform_sparse_cv5_time}
+        \\
+        \multicolumn{2}{c}{{\bf normal sparse (no costs)}}
+        \\
+        \includegraphics[width=12em]{figures/pc_no_costs_normal_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/pc_no_costs_normal_sparse_cv5_time}
+        \\
+        \multicolumn{2}{c}{{\bf uniform sparse (with costs)}}
+        \\
+        \includegraphics[width=12em]{figures/pc_with_costs_uniform_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/pc_with_costs_uniform_sparse_cv5_time}
+        \\
+        \multicolumn{2}{c}{{\bf normal sparse (with costs)}}
+        \\
+        \includegraphics[width=12em]{figures/pc_with_costs_normal_sparse_cv5_loss} &
+        \includegraphics[width=12em]{figures/pc_with_costs_normal_sparse_cv5_time}
+    \end{tabular}
+    \caption{Results for the PC dataset with and without costs.}
+\end{figure}
+
 \twocolumn

### 2016-01-21 paper/ijcai16.tex
paper: remove all PNGs, let's commit them once they are done...

@@
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv1_loss}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv1_loss}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_cv1_loss}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv1_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv1_loss}
-    \end{tabular}
-    \caption{Results on the synthetic dataset, with crossvalidation at each step. Each row represents a different
-    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
-\end{figure}
-
-\begin{figure}[b]
-    \centering
-    \begin{tabular}{cccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
-        \\
         \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_loss} &
@@
     \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
-    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
-\end{figure}
-
-\begin{figure}[b]
-    \centering
-    \begin{tabular}{cccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv1_time}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv1_time}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_cv1_time}
-        \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv1_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv1_time}
-    \end{tabular}
-    \caption{Results on the synthetic dataset, with crossvalidation at each step. Each row represents a different
     setting: uniform, uniform sparse, normal, normal sparse. Color indicates

### 2016-01-21 ijcai16.py
ijcai16: set IJCAI-16 num of queries and iterations.

@@ -232,7 +232,7 @@ def run_pc(has_costs):
     CONFIGS = Grid({
         "num_trials": 20,
         "max_iterations": 100,
-        "max_answers": 100,
+        "max_answers": 300,
         "sampling_mode": ("uniform_sparse", "normal_sparse", "uniform", "normal"),
         "ranking_mode": ("all_pairs",),
         "is_deterministic": False,

### 2016-01-21 paper/ijcai16.tex
paper: update figures.

@@
     \centering
+    {\footnotesize
     \begin{tabular}{cccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \hline
+        {\sc synthetic 3} & {\sc synthetic 4} & {\sc synthetic 5} & {\sc synthetic 6}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv5_loss}
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv5_loss}
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_per_iter_loss}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_cv5_loss}
+        \hline
+        \multicolumn{4}{c}{{\sc Uniform Sparse}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv5_loss}
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_sparse_per_iter_loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_per_iter_loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_sparse_per_iter_loss}
+        \\
+        \hline
     \end{tabular}
-    \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
-    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+    }
 \end{figure}
@@
     \centering
+    {\footnotesize
     \begin{tabular}{cccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \hline
+        {\sc synthetic 3} & {\sc synthetic 4} & {\sc synthetic 5} & {\sc synthetic 6}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv5_time}
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_per_query_loss}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_cv5_time}
+        \hline
+        \multicolumn{4}{c}{{\sc Uniform Sparse}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_sparse_per_query_loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_normal_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_per_query_loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_sparse_per_query_loss}
+        \\
+        \hline
     \end{tabular}
-    \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
-    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+    }
 \end{figure}
@@
     \centering
-    \begin{tabular}{cc}
-        {\bf loss} & {\bf time}
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        {\sc synthetic 3} & {\sc synthetic 4} & {\sc synthetic 5} & {\sc synthetic 6}
         \\
-        \multicolumn{2}{c}{{\bf uniform sparse (no costs)}}
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
         \\
-        \includegraphics[width=12em]{figures/pc_no_costs_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_no_costs_uniform_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_per_iter_time}
         \\
-        \multicolumn{2}{c}{{\bf normal sparse (no costs)}}
+        \hline
+        \multicolumn{4}{c}{{\sc Uniform Sparse}}
         \\
-        \includegraphics[width=12em]{figures/pc_no_costs_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_no_costs_normal_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_sparse_per_iter_time}
         \\
-        \multicolumn{2}{c}{{\bf uniform sparse (with costs)}}
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
         \\
-        \includegraphics[width=12em]{figures/pc_with_costs_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_with_costs_uniform_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/synthetic_3_normal_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_per_iter_time}
         \\
-        \multicolumn{2}{c}{{\bf normal sparse (with costs)}}
+        \hline
+        \multicolumn{4}{c}{{\sc Normal Sparse}}
         \\
-        \includegraphics[width=12em]{figures/pc_with_costs_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_with_costs_normal_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/synthetic_3_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_sparse_per_iter_time}
+        \\
+        \hline
     \end{tabular}
-    \caption{Results for the PC dataset with and without costs.}
+    }
+\end{figure}
+
+\begin{figure}[b]
+    \centering
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        {\sc synthetic 3} & {\sc synthetic 4} & {\sc synthetic 5} & {\sc synthetic 6}
+        \\
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_per_query_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Uniform Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_uniform_sparse_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_uniform_sparse_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_uniform_sparse_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_uniform_sparse_per_query_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_normal_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_per_query_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_3_normal_sparse_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_4_normal_sparse_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_5_normal_sparse_per_query_time} &
+        \includegraphics[width=10em]{figures/synthetic_6_normal_sparse_per_query_time}
+        \\
+        \hline
+    \end{tabular}
+    }
+\end{figure}
+
+\begin{figure}[b]
+    \centering
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        \multicolumn{4}{c}{{\sc uniform sparse}}
+        \\
+        \multicolumn{2}{c}{{\sc loss/time per iteration}} &
+        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc normal sparse}}
+        \\
+        \multicolumn{2}{c}{{\sc loss/time per iteration}} &
+        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
+        \\
+        \hline
+    \end{tabular}
+    }
+    \caption{Results for the PC dataset.}
 \end{figure}

### 2016-01-21 draw.py
draw: update.

@@ -3,24 +3,25 @@
 import sys, os
 import cPickle as pickle
 import numpy as np
+from scipy.io import loadmat
 import matplotlib.pyplot as plt
 from matplotlib.font_manager import FontProperties
 from glob import glob
 
 # Colors taken from http://tango.freedesktop.org/Tango_Icon_Theme_Guidelines
 COLOR_OF = {
+    1: ("#000000", "#000000"),
     2: ("#CE5C00", "#F57900"),
     3: ("#204A87", "#3465A4"),
-#    4: ("#5C3566", "#75507B"),
     4: ("#4E9A06", "#73D216"),
 }
 
-def infos_to_matrices(infos, elongate=True):
+def infos_to_matrices(infos, per_query):
     num_trials = len(infos)
     max_queries = max([sum(n for n, _, _ in info) for info in infos])
     max_iterations = max(len(info) for info in infos)
 
-    if elongate:
+    if per_query:
         loss_matrix = np.zeros((num_trials, max_queries))
         time_matrix = np.zeros((num_trials, max_queries))
     else:
@@ -31,7 +32,7 @@ def infos_to_matrices(infos, elongate=True):
         base = 0
         prev_loss = max(max(l for _, l, _ in info) for info in infos)
         for iteration, (num_queries, loss, time) in enumerate(info):
-            if elongate:
+            if per_query:
                 for j in range(num_queries):
                     alpha = 1 - (j + 1) / float(num_queries)
                     interpolated_loss = alpha*prev_loss + (1 - alpha)*loss
@@ -43,9 +44,12 @@ def infos_to_matrices(infos, elongate=True):
                 loss_matrix[trial,iteration] = loss
                 time_matrix[trial,iteration] = time
 
-    return max_queries if elongate else max_iterations, loss_matrix, time_matrix
+    if per_query:
+        return loss_matrix[:,:100], time_matrix[:,:100]
+    else:
+        return loss_matrix[:,:50], time_matrix[:,:50]
 
-def draw(paths, dest_path):
+def draw(paths, dest_path, per_query):
     loss_fig, loss_ax = plt.subplots(1, 1)
     time_fig, time_ax = plt.subplots(1, 1)
 
@@ -53,10 +57,34 @@ def draw(paths, dest_path):
     max_loss_max_y = None
     max_time_max_y = None
     for path in paths:
-        with open(path, "rb") as fp:
-            infos = pickle.load(fp)
 
[…]

### 2016-01-22 paper/ijcai16.tex
paper: mock-up the final figures.

--- 2016-01-21 paper/ijcai16.tex
@@
     \centering
+    {\footnotesize
     \begin{tabular}{cccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \hline
+        {\sc synthetic 3 loss} & {\sc synthetic 4 loss} & {\sc synthetic 5 loss} & {\sc synthetic 5 time}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv5_loss}
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv5_loss}
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_cv5_loss}
+        \hline
+        \multicolumn{4}{c}{{\sc Uniform Sparse}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv5_loss}
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time}
+        \\
+        \hline
     \end{tabular}
-    \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
-    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+    }
+    \caption{\label{fig:comparison} {\sc SetMargin} vs Guo vs Viappiani.}
 \end{figure}
@@
     \centering
+    {\footnotesize
     \begin{tabular}{cccc}
-        {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5} & {\bf synthetic 6}
+        \hline
+        \multicolumn{2}{c}{{\sc per query loss}} &
+        \multicolumn{2}{c}{{\sc per iter. loss}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_cv5_time}
+        {\sc synthetic 4} & {\sc synthetic 6} & {\sc synthetic 4} & {\sc synthetic 6}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_uniform_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_uniform_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_uniform_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_uniform_sparse_cv5_time}
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_cv5_time}
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss}
         \\
-        \includegraphics[width=12em]{figures/synthetic_3_normal_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_4_normal_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_5_normal_sparse_cv5_time} &
-        \includegraphics[width=12em]{figures/synthetic_6_normal_sparse_cv5_time}
+        \hline
+        \multicolumn{4}{c}{{\sc Uniform Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal Sparse}}
+        \\
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/loss}
+        \\
+        \hline
     \end{tabular}
-    \caption{Results on the synthetic dataset, with crossvalidation every 5 steps. Each row represents a different
-    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
-    the set size $k$: orange means $k=2$, blue $k=3$, and violet $k=4$.}
+    \caption{\label{fig:selfcomparison} Self-comparison experiment. Only
+    losses, per-query and per-iteration. Runtime described in the text. Each
+    row is a user sampling mode. Three curves per plot: $k=2$ vs $k=3$ vs
+    $k=4$.}
+    }
 \end{figure}
@@
     \centering
-    \begin{tabular}{cc}
-        {\bf loss} & {\bf time}
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        \multicolumn{4}{c}{{\sc uniform sparse}}
         \\
-        \multicolumn{2}{c}{{\bf uniform sparse (no costs)}}
+        \multicolumn{2}{c}{{\sc loss/time per iteration}} &
+        \multicolumn{2}{c}{{\sc loss/time per query}}
         \\
-        \includegraphics[width=12em]{figures/pc_no_costs_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_no_costs_uniform_sparse_cv5_time}
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time}
         \\
-        \multicolumn{2}{c}{{\bf normal sparse (no costs)}}
+        \hline
+        \multicolumn{4}{c}{{\sc normal sparse}}
         \\
-        \includegraphics[width=12em]{figures/pc_no_costs_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_no_costs_normal_sparse_cv5_time}
+        \multicolumn{2}{c}{{\sc loss/time per iteration}} &
+        \multicolumn{2}{c}{{\sc loss/time per query}}
         \\
-        \multicolumn{2}{c}{{\bf uniform sparse (with costs)}}
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time} &
+        \includegraphics[width=10em]{figures/loss} &
+        \includegraphics[width=10em]{figures/time}
         \\
-        \includegraphics[width=12em]{figures/pc_with_costs_uniform_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_with_costs_uniform_sparse_cv5_time}
-        \\
-        \multicolumn{2}{c}{{\bf normal sparse (with costs)}}
-        \\
-        \includegraphics[width=12em]{figures/pc_with_costs_normal_sparse_cv5_loss} &
-        \includegraphics[width=12em]{figures/pc_with_costs_normal_sparse_cv5_time}
+        \hline
     \end{tabular}
-    \caption{Results for the PC dataset with and without costs.}
+    }
+    \caption{Results for the PC dataset.}
 \end{figure}

### 2016-01-25 paper/ijcai16.tex
paper: went through methods section, corrected errors and added fixmes

@@
 
+\andrea{manca un titolo a questa sezione?}
+
 \paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
@@
 linearly dependent continuous features will be discussed later on. 
+\andrea{mmh sembra che gestiamo solo booleani.. riscriverei che ci concentriamo solo su feature categoriche, e 
+che le rappresentiamo con one-hot encoding come tipico dei metodi preference elicitation..}
 
@@
 Boolean disjunction of two binary variables $x_1 \lor x_2$ can be rewritten as
-$x_1 + x_2 \ge 1$.
+$x_1 + x_2 \ge 1$. \andrea{mmh cosa ci da in piu'? altrimenti non e' una restrizione ma un modo 
+per scrivere vincoli logici}
 
@@
 
-\paragraph{Non-linear Formulation.} We now introduce an initial formulation of
-the optimization problem, that we do {\em not} aim at solving directly.
+\paragraph{Non-linear Formulation.} We first introduce the problem
+formulation as a mixed integer non-linear optimization problem, and
+then show how to reduce it to a mixed integer linear optimization one.
 
-In the set-margin approach, the goal is twofold. First, for any given set size
-$k\geq 1$, we want to find a {\em set} of $k$ weight vectors $\vw^{1}, \ldots,
-\vw^{k}$, chosen so that all user-provided preferences are satisfied by the
-largest possible margin (i.e. all weight vector are consistent with respect to
-the user responses $\calD$, modulo inconsistencies) and so that they are
-maximally diverse.  Second, we want to construct a {\em set} of $k$
-configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration $\vx^{i}$
-is the ``best'' possible option when evaluated according to the corresponding
-$\vw^{i}$, and consequently maximally diverse as well. These options will be
-later employed to formulate the user queries.
+The goal of our setwise max-margin approach is twofold. First, for any
+given set size $k\geq 1$, we want to find a {\em set} of $k$ weight
+vectors $\vw^{1}, \ldots, \vw^{k}$, chosen so that all user-provided
+preferences are satisfied by the largest possible margin (i.e. all
+weight vectors are consistent with respect to the user responses
+$\calD$, modulo inconsistencies) and so that they are maximally
+diverse.  Second, we want to construct a {\em set} of $k$
+configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration
+$\vx^{i}$ is the ``best'' possible option when evaluated according to
+the corresponding $\vw^{i}$ and configurations are maximally diverse
+among each other. These options will be later employed to formulate
+the user queries.
 
-The first goal achieved by translating all pairwise preferences $\calD$ into
-ranking constraints: preferences of the form $\vy^h_+ \succ \vy^h_-$ become
-linear inequalities of the form $\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq
-M$, where $M$ is the {\em margin} variable (which we aim at maximizing) and $h$
-ranges over the responses.  Non-separable datasets, which occur in practice due
-to user inconsistency, are handled by introducing slack variables (whose sum we
-aim at minimizing). When augmented with the slacks, the above inequalities take
-the form:
+The first goal is achieved by translating all pairwise preferences
+$\calD$ into ranking constraints: preferences of the form
+$\vy^h_+ \succ \vy^h_-$ become linear inequalities of the form
+$\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq M$, where $M$ is the
+{\em margin} variable (which we aim at maximizing) and $h$ ranges over
+the responses.  Non-separable datasets, which occur in practice due to
+occasional inconsistencies in user feedback, are handled by
+introducing slack variables (whose sum we aim at minimizing). When
+augmented with the slacks, the above inequalities take the form:
 %
@@
 where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
-for violating the ranking of example $h$. Indifference preferences, i.e.
-$\vy^h_1 \approx \vy^h_2$, are translated as $|\langle \vw^i, \vy^h_1 - \vy^h_2 \rangle| < \varepsilon^i_h$:
+for violating the margin separation of pair $h$. Indifference preferences, i.e.
+$\vy^h_1 \approx \vy^h_2$, are translated as $|\langle \vw^i, \vy^h_1 - \vy^h_2 \rangle| < \varepsilon^i_h$;
 the slack increases with the difference between the estimated utility of the
@@
 
-Ideally, the second goal would be implemented as:
-%
-$$ \vx^{i} = \argmax_{\vx \in \calX_\text{feasible}} \langle \vw^{i}, \vx^{i} \rangle $$
-%
-However this formulation is highly impractical. We therefore only require 
-each option $\vx^{i}$ to be the best {\em among} the configurations
-$\{ \vx^i \}$. This can be accomplished by imposing
-constraints of the form $\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$
-for all $i,j\in[k]$, $i \ne j$.
-This however does not guarantee that the produced $\{ \vx^{i} \}$ have a high
-{\em absolute} utility. We therefore favor high-quality configurations by
-introducing an additional term $\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$
-in the objective function.
+The second goal requires to jointly maximize the utility of each
+$\vx^{i}$ according to its corresponding weight vector $\vw^i$ and its
+scoring difference with respect to the other configurations $\vx^j$ in
+the set. We achieve this by maximizing the sum of utilities
+$\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ and adding ranking
+constraints of the form
+$\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$ for all
+$i,j\in[k]$, $i \ne j$.
 
-A straightforward encoding of the above desiderata leads to the following {\em
-non-linear} optimization problem over the variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$:
+% Ideally, the second goal would be implemented as:
+% %
+% $$ \vx^{i} = \argmax_{\vx \in \calX_\text{feasible}} \langle \vw^{i}, \vx^{i} \rangle $$
+% %
+% However this formulation is highly impractical. We therefore only require 
+% each option $\vx^{i}$ to be the best {\em among} the configurations
+% $\{ \vx^i \}$. This can be accomplished by imposing
+% constraints of the form $\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$
+% for all $i,j\in[k]$, $i \ne j$.
+% This however does not guarantee that the produced $\{ \vx^{i} \}$ have a high
+% {\em absolute} utility. We therefore favor high-quality configurations by
+% introducing an additional term $\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$
+% in the objective function.
+
+A straightforward encoding of the above desiderata leads to the
+following mixed integer {\em non-linear} optimization problem over the
+variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$: \andrea{dovremmo forse specificare i domini delle variabili per chiarire la natura mixed}
 %
@@
 \end{align}
-}
+} \andrea{perche' un punto tra il quantificatore e la formula? inoltre mi sa che ``n'' e' usato per due cose, queries e features}
 %
@@
 (second part), while at the same time regularizing the magnitude of the
-weights (third part) and quality of the configurations $\{ \vx^{i} \}$ (last
+weights (third part) and the quality of the configurations $\{ \vx^{i} \}$ (last
 part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
@@
 will be indifferent to most of them~\cite{}. The $\ell_1$ penalty is frequently
-used to improve the sparsity of learned models~\cite{}, with consequent gains
+used to improve the sparsity of learned models~\cite{lasso}, with consequent gains
 in generalization ability and efficiency, as confirmed by our empirical
@@
 
-The two core constraints as Equations~\ref{eq:wyconstr} and \ref{eq:wxconstr},
+The two core constraints are Equations~(\ref{eq:wyconstr}) and~(\ref{eq:wxconstr}),
 which are taken directly from the discussion above. The former enforces the
 correct ranking of the observed user preferences, while the latter ensures that
-the generated configurations are ranked correctly with respect to the
-corresponding weight vectors. Equations \ref{eq:wbounds} and \ref{eq:xbounds}
+the generated configurations are diverse in terms of the weight vectors they
+maximize. Equations~(\ref{eq:wbounds}) and~(\ref{eq:xbounds})
 ensure that the weights and configurations are in fact feasible, while the
 remaining two guarantee the non-negativity of the slack and margin variables.
-Since we require $\vw^\bot \ge 0$, Eq.~\ref{eq:wbounds} also enforces the
+Since we require $\vw^\bot \ge 0$, Eq.~(\ref{eq:wbounds}) also enforces the
 weights to be non-negative.
 
-Note that we are choosing the options $\{ \vx^i \}$ and the weight vectors $\{
+Note that we are choosing the configurations $\{ \vx^i \}$ and the weight vectors $\{
 \vw^i \}$ {\em simultaneously}. Since we maximize the margin $M$, the optimizer
-will prefer a set of outcomes $\{ \vx^{i} \}$ that partition the weight space
+will prefer a set of configurations $\{ \vx^{i} \}$ that partitions the weight space
 roughly equally, and corresponding utility functions such that each $\vw^{i}$
 lies (intuitively) close to the centre of each subregion. See
-Figure~\ref{fig:setmargin} for a depiction of this intuition.
+Figure~\ref{fig:setmargin} for a depiction of this intuition.\andrea{mmh non chiarissimo, da rivedere insieme, legenda inclusa}
 
@@
 
-\paragraph{MILP Formulation.} This initial formulation is problematic, as
-Eq~\ref{eq:wxconstr} involves quadratic terms. Here we show how to reformulate
-the problem as a mixed integer linear program (MILP). Our goal is to replace
-Eq~\ref{eq:wxconstr} with a set of linear constraints through a suitable
-variable transformation.
+\paragraph{MILP Formulation.} This initial formulation is problematic,
+as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed
+continuous integer variables.\andrea{check} Here we show how to reformulate the
+problem as a mixed integer linear program (MILP). Our goal is to
+replace Eq.~(\ref{eq:wxconstr}) with a set of linear constraints through
+a suitable variable transformation.
 
@@
 $i,j\in[k]$ and $z\in[n]$. Assuming for the time being that the new variables
-do satisfy the equation $p^{i,j}_z = w^i_z x^i_z$, we rewrite the fourth
+do satisfy the equation $p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth
 component of the objective function in terms of the new variables as:
@@
 %
-and, similarly, Eq~\ref{eq:wxconstr} as:
+and, similarly, Eq.~(\ref{eq:wxconstr}) as:
 %
@@
 %
-The fact that $p^{i,j}_z = w^{i}_z x^{i}_z$ is achieved by
+The fact that $p^{i,j}_z = w^{i}_z x^{j}_z$ is achieved by
 setting the following additional constraints. We distinguish between two cases:
-(i) $p^{i,i}_z$, and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
-maximizing the margin $M$. Now, due to Eq~\ref{eq:pxconstr}, the optimizer will
+(i) $p^{i,i}_z$ and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
+maximizing the margin $M$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
 try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
@@
 %
-If $x^i_z = 1$ the lower bound simplifies to $\max \{ 0, w^{i}_z \} = w^{i}_z$,
-due to the non-negativity of $w^i_z$. Otherwise, if $x^i_z = 0$
+If $x^j_z = 1$ the lower bound simplifies to $\max \{ 0, w^{i}_z \} = w^{i}_z$,
+due to the non-negativity of $w^i_z$. Otherwise, if $x^j_z = 0$
 then the lower bound becomes $\max \{ 0, w^{i}_z - w_\text{max} \}$, where
@@
 cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
-x^i_z$\footnote{Since $M$ is upper-bounded by Eq.~\ref{eq:wyconstr}, in some
+x^j_z$.\footnote{Since $M$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
 cases the $p^{i,j}_z$ variables do not attain the lower bound. As a
-consequence, the MILP reformulation of Eq.~\ref{eq:wxconstr} is marginally
-inexact. Despite this issue, the MILP formulation acceptably in practice.}.
+consequence, the MILP reformulation of Eq.~(\ref{eq:wxconstr}) is marginally
+inexact. Despite this issue, the MILP formulation acceptably in practice. \andrea{acceptably what?}}
 
@@
     \\
-        & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
+    %     & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
+    % \\
+        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [n] \nonumber
     \\
         & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
-    \\
-        & \;\; \forall \; i, j \in [k], i \neq j \nonumber
     \\
@@
 
-\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm follows the
-usual preference elicitation loop. Starting from an initially empty set of user
-responses $\calD$, it repeatedly solves the MILP problem above using the set
-$\calD$ to enforce ranking constraints on the weight vectors $\{\vw^i\}$. The
-generated configurations $\{\vx^i\}$, which are chosen to be as good as
-possible with respect to the estimated user preferences, and as diverse as
-possible, are then employed to formulate a set of user queries. The new replies
-are added to $\calD$ and the whole procedure is repeated. The algorithm
-terminates after a fixed number of iterations, or once another suitable
-condition is met. The procedure is sketched in Algorithm~\ref{alg:setmargin}.
+\paragraph{Set-wise max-margin.} The full {\sc SetMargin} algorithm
+follows the usual preference elicitation loop. Starting from an
+initially empty set of user responses $\calD$, it repeatedly solves
+the MILP problem above using the set $\calD$ to enforce ranking
+constraints on the weight vectors $\{\vw^i\}$. The generated
+configurations $\{\vx^i\}$, which are chosen to be as good as possible
+with respect to the estimated user preferences, and as diverse as
+possible, are then employed to formulate a set of user queries. The
+new replies are added to $\calD$ and the whole procedure is
+repeated. The algorithm terminates after a fixed number of iterations,
+or once another suitable condition is met. \andrea{another suitable
+  condition like?}  The procedure is sketched in
+Algorithm~\ref{alg:setmargin}. Note that at the end of the preference
+elicitation procedure, a final recommendation is made by solving the
+MILP problem for $k=1$.
 
@@
 
+\andrea{questa sezione e' da rivedere, con query selection si itende un'altra cosa di solito}
+
 \paragraph{Linearly dependent real attributes.} In many domains of interest,
@@
 As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
-to be non-negative.
+to be non-negative. \andrea{check formulation}
 

### 2016-01-25 paper/ijcai16.tex
paper: fixed clash between feature and query number

@@
 
-\paragraph{Setting.} We assume to have a multi-attribute feature space $\calX$
-of configurations $\vx = (x_1, \ldots, x_n)$ over $n$ features. For the time
-being we will focus on binary features only, i.e. $x_z\in\{0,1\}$ for all
-$z\in[n]$, to keep matters simple. This is a common choice for preference
-elicitation methods~\cite{guo2010real,viappiani2010optimal}; discrete features
-can be handled for instance by means of a simple one-hot encoding. Support for
-linearly dependent continuous features will be discussed later on. 
-\andrea{mmh sembra che gestiamo solo booleani.. riscriverei che ci concentriamo solo su feature categoriche, e 
-che le rappresentiamo con one-hot encoding come tipico dei metodi preference elicitation..}
+\paragraph{Setting.} We assume to have a multi-attribute feature space
+$\calX$ of configurations $\vx = (x_1, \ldots, x_m)$ over $m$
+features. For the sake of simplicity we focus on binary features only,
+i.e. $x_z\in\{0,1\}$ for all $z\in[m]$, assuming a one-hot encoding of
+categorical features. This is a common choice for preference
+elicitation methods~\cite{guo2010real,viappiani2010optimal}. Support
+for linearly dependent continuous features will be discussed later on.
 
@@
 of linear utility functions: the user's preferences are represented by a weight
-vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
-$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the remainder of the paper we
+vector $\vw\in\bbR^m$, and the utility of a configuration $\vx$ is given by
+$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. In the remainder of the paper we
 require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
@@
 \end{align}
-} \andrea{perche' un punto tra il quantificatore e la formula? inoltre mi sa che ``n'' e' usato per due cose, queries e features}
+} \andrea{perche' un punto tra il quantificatore e la formula?}
 %
@@
 In order to do so, we introduce a set of fresh variables $p^{i,j}_z$ for every
-$i,j\in[k]$ and $z\in[n]$. Assuming for the time being that the new variables
+$i,j\in[k]$ and $z\in[m]$. Assuming for the time being that the new variables
 do satisfy the equation $p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth
@@
 %
-$$ \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z $$
+$$ \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z $$
 %
@@
 %
-\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M \label{eq:pxconstr} \]
+\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge M \label{eq:pxconstr} \]
 %
@@
     \max
-        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^n p^{i,i}_z
+        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
         \nonumber
@@
     \\
-        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^n p^{i,i}_z - p^{i,j}_z \ge M
+        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge M
     \\
-    %     & \;\; \forall \; i \in [k], \forall \; z \in [n] \nonumber
+    %     & \;\; \forall \; i \in [k], \forall \; z \in [m] \nonumber
     % \\
-        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [n] \nonumber
+        & \;\; \forall \; i, j \in [k], i \neq j, \forall \; z \in [m] \nonumber
     \\
@@
 setting. We developed seven synthetic datasets: each dataset involves
-$n=2,\ldots,7$ attributes, where each attribute takes one of $n$ possible
-values. More explicitly, for $n=2$ the synthetic dataset is
-$\calX_\text{feasible} = [2] \times [2]$, for $n=3$ to $\calX_\text{feasible} =
+$m=2,\ldots,7$ attributes, where each attribute takes one of $m$ possible
+values. More explicitly, for $m=2$ the synthetic dataset is
+$\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to $\calX_\text{feasible} =
 [3] \times [3] \times [3]$, and so on. The cardinality of
-$\calX_\text{feasible}$ is $n^n$, and grows (super) exponentially with $n$. For
-$n=3$, the dataset is comparable to the synthetic one used
+$\calX_\text{feasible}$ is $m^m$, and grows (super) exponentially with $m$. For
+$m=3$, the dataset is comparable to the synthetic one used
 in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
 of size of 2, 2 and 5, respectively, for a total of 20 feasible
-configurations), while for larger $n$ it grows much larger than those typically
+configurations), while for larger $m$ it grows much larger than those typically
 used in the Bayesian preference elicitation literature, and as such represents

### 2016-01-26 paper/ijcai16.tex
paper: simplify slightly.

@@
 The generalized problem is obtained by substituting $\vw^i$ with $\vv^i := \vw_B^i + C^\top \vv_B^i$ and $p^{i,j}_z$
-with $q^{i,j}_z := (\vw^i_B + C^\top \vw^i_R)_z x^j_z$ in the MILP problem.
+with $q^{i,j}_z := \vv^i_z x^j_z$ in the MILP problem.
 All constraints remain the same, except Eq.~\ref{eq:wbounds2}, which must be

### 2016-01-26 paper/ijcai16.tex
paper: nota per paolo

--- 2016-01-25 paper/ijcai16.tex
@@
 learning system, and must be estimated from pairwise choices elicited from the
-user.  We consider three possible outcomes for a pairwise comparison between
+user.\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  We consider three possible outcomes for a pairwise comparison between
 two configurations $\vx$ and $\vx'$: either $\vx$ is preferred to $\vx'$ (written

### 2016-01-26 paper/ijcai16.tex
paper: fix methods and notation with help from Andrea.

@@
 
-\section{Setwise Max-margin Learning}
-
-\stefano{Paolo, Andrea, feel free to improve the text and especially the
-citations as you see fit.}
-
-\andrea{manca un titolo a questa sezione?}
+\section{Preference Elicitation}
 
@@
 a shorthand for the set $\{1, \ldots, n\}$. We write $\|\vx\|_1 := \sum_z |x_z|$
-to indicate the $\ell_1$ vector norm, and $\langle \cdot, \cdot \rangle$ to
-denote the usual dot product.
+to indicate the $\ell_1$ vector norm, $\langle \cdot, \cdot \rangle$ to
+denote the usual dot product, and $X'$ to denote matrix transposition.
 
@@
 $\calX_\text{feasible}$, is expressed as
-a conjunction of linear constraints. This is not a severe restriction:
-arbitrary propositional constraints can be trivially encoded in this setting,
+a conjunction of linear constraints. This allows to formulate both arithmetic
+and logical constraints,
 e.g. under the canonical mapping of $True$ to $1$ and $False$ to $0$, the
 Boolean disjunction of two binary variables $x_1 \lor x_2$ can be rewritten as
-$x_1 + x_2 \ge 1$. \andrea{mmh cosa ci da in piu'? altrimenti non e' una restrizione ma un modo 
-per scrivere vincoli logici}
+$x_1 + x_2 \ge 1$.
 
@@
 
-\subsection{Problem Formulation}
+\section{Setwise Max-margin Learning}
 \label{sec:formulation}
@@
 $\vy^h_+ \succ \vy^h_-$ become linear inequalities of the form
-$\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq M$, where $M$ is the
+$\langle \vw^i, \vy^h_+ - \vy^h_- \rangle \geq \mu$, where $\mu$ is the
 {\em margin} variable (which we aim at maximizing) and $h$ ranges over
@@
 %
-$$ \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h $$
+$$ \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h $$
 %
@@
 constraints of the form
-$\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$ for all
+$\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq \mu$ for all
 $i,j\in[k]$, $i \ne j$.
@@
 % $\{ \vx^i \}$. This can be accomplished by imposing
-% constraints of the form $\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq M$
+% constraints of the form $\langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \geq \mu$
 % for all $i,j\in[k]$, $i \ne j$.
@@
 following mixed integer {\em non-linear} optimization problem over the
-variables $M$, $\{ \vw^i \}$, $\{ \vx^i \}$: \andrea{dovremmo forse specificare i domini delle variabili per chiarire la natura mixed}
+variables $\mu \in \bbR$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
 %
@@
     \max
-        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
+        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle
         \nonumber
@@
     \\
-        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \label{eq:wyconstr}
+        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
     \\
-        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge M \label{eq:wxconstr}
+        & \;\; \forall \; i, j \in [k], i \neq j \quad \langle \vw^{i}, \vx^{i} - \vx^{j} \rangle \ge \mu \label{eq:wxconstr}
     \\
-        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
+        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds}
     \\
-        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
+        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
     \\
-        & \;\; \forall \; i \in [k] \;.\; \veps^{i} \ge 0 \nonumber
+        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
     \\
-        & \;\; M \ge 0 \nonumber
+        & \;\; \mu \ge 0 \nonumber
 \end{align}
-} \andrea{perche' un punto tra il quantificatore e la formula?}
+}
 %
 Let us illustrate the above piece by piece. The objective is composed of four
-parts: we maximize the shared margin $M$ (first part) and minimize the total
+parts: we maximize the shared margin $\mu$ (first part) and minimize the total
 sum of the ranking errors $\veps^i$ incurred by each weight vector $\vw^{i}$
@@
 many scenarios the user will have strong preferences about some attributes, but
-will be indifferent to most of them~\cite{}. The $\ell_1$ penalty is frequently
+will be indifferent to most of them~\cite{} \stefano{@Paolo: add refs}. The $\ell_1$ penalty is frequently
 used to improve the sparsity of learned models~\cite{lasso}, with consequent gains
 in generalization ability and efficiency, as confirmed by our empirical
-findings (see Section~\ref{sec:experiments}). \stefano{add refs}
+findings (see Section~\ref{sec:experiments}).
 
@@
 Note that we are choosing the configurations $\{ \vx^i \}$ and the weight vectors $\{
-\vw^i \}$ {\em simultaneously}. Since we maximize the margin $M$, the optimizer
+\vw^i \}$ {\em simultaneously}. Since we maximize the margin $\mu$, the optimizer
 will prefer a set of configurations $\{ \vx^{i} \}$ that partitions the weight space
@@
 lies (intuitively) close to the centre of each subregion. See
-Figure~\ref{fig:setmargin} for a depiction of this intuition.\andrea{mmh non chiarissimo, da rivedere insieme, legenda inclusa}
-
-\stefano{for $k=1$, setting $\alpha < 1$  renders the problem unbounded, as
-Eq.~\ref{eq:wxconstr} vanishes.}
+Figure~\ref{fig:setmargin} for a depiction of this intuition.\andrea{@Paolo: mmh non chiarissimo, da rivedere insieme, legenda inclusa}
 
@@
 as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed
-continuous integer variables.\andrea{check} Here we show how to reformulate the
+continuous integer variables. Here we show how to reformulate the
 problem as a mixed integer linear program (MILP). Our goal is to
@@
 %
-\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge M \label{eq:pxconstr} \]
+\[ \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu \label{eq:pxconstr} \]
 %
@@
 (i) $p^{i,i}_z$ and (ii) $p^{i,j}_z$ for $i \ne j$.  Recall that we are
-maximizing the margin $M$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
+maximizing the margin $\mu$. Now, due to Eq.~(\ref{eq:pxconstr}), the optimizer will
 try to keep $p^{i,i}_z$ as large as possible and $p^{i,j}_z$ as small as
@@
 cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
-x^j_z$.\footnote{Since $M$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
+x^j_z$.\footnote{Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
 cases the $p^{i,j}_z$ variables do not attain the lower bound. As a
-consequence, the MILP reformulation of Eq.~(\ref{eq:wxconstr}) is marginally
-inexact. Despite this issue, the MILP formulation acceptably in practice. \andrea{acceptably what?}}
+consequence, the MILP reformulation of Eq.~(\ref{eq:wxconstr}) is a (tight)
+approximation of the original one. This has no impact on the quality of the
+solutions.}
 
@@
     \max
-        & \;\; M - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
+        & \;\; \mu - \alpha \sum_{i=1}^k \| \veps^{i} \|_1 - \beta \sum_{i=1}^k \| \vw^{i} \|_1 + \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z
         \nonumber
@@
     \\
-        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge M - \varepsilon^{i}_h \nonumber
+        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
     \\
-        & \;\; \forall \; i, j \in [k], i \neq j \;.\; \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge M
+        & \;\; \forall \; i, j \in [k], i \neq j \quad \sum_{z=1}^m p^{i,i}_z - p^{i,j}_z \ge \mu
     \\
@@
     \\
-        & \;\; \forall \; i \in [k] \;.\; \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
+        & \;\; \forall \; i \in [k] \quad \vw^\bot \le \vw^{i} \le \vw^\top \label{eq:wbounds2}
     \\
-        & \;\; \forall \; i \in [k] \;.\; \vx^{i} \in \calX_{\text{feasible}} \nonumber
+        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
     \\
-        & \;\; \forall \; i \in [k] \;.\; \veps^{i}_h \ge 0 \nonumber
+        & \;\; \forall \; i \in [k] \quad \veps^{i}_h \ge 0 \nonumber
     \\
-        & \;\; M \ge 0 \nonumber
+        & \;\; \mu \ge 0 \nonumber
 \end{align}
@@
 repeated. The algorithm terminates after a fixed number of iterations,
-or once another suitable condition is met. \andrea{another suitable
-  condition like?}  The procedure is sketched in
+or once another suitable condition is met. \andrea{@Paolo: what termination
+condition for pref elicitation?}  The procedure is sketched in
 Algorithm~\ref{alg:setmargin}. Note that at the end of the preference
@@
 
-\paragraph{Query selection.} Contrarily to most other preference elicitation
-methods, which generate a single candidate option per iteration, {\sc
-SetMargin} constructs a whole {\em set} of options. We thus can not leverage
-standard query selection strategies \stefano{which ones?}.
-
-The simplest solution is to query the user about all $k \choose 2$ pairs of
-configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
-number of queries does not grow large in practice.  An appealing alternative
-involves using a modified sorting algorithm, such as merge-sort, to order the
-set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
-approach is very susceptible to noise: a single inconsistent user response may
-affect all successive comparisons, leading to a cascading effect. While more
-robust strategies may be conceived, empirically we found that the simpler
-quadratic scheme tends to work reasonably well.
-
-\andrea{questa sezione e' da rivedere, con query selection si itende un'altra cosa di solito}
-
 \paragraph{Linearly dependent real attributes.} In many domains of interest,
@@
 %
-$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C^\top \vw_B, \vx_B \rangle $$
+$$ \langle \vw, \vx \rangle = \langle \vw_B, \vx_B \rangle + \langle \vw_R, C \vx_B \rangle = \langle \vw_B + C' \vw_R, \vx_B \rangle $$
 %
-The generalized problem is obtained by substituting $\vw^i$ with $\vv^i := \vw_B^i + C^\top \vv_B^i$ and $p^{i,j}_z$
-with $q^{i,j}_z := \vv^i_z x^j_z$ in the MILP problem.
-All constraints remain the same, except Eq.~\ref{eq:wbounds2}, which must be
-rewritten in terms of $\vv^i$:
+The generalized problem is obtained by substituting $\vw^i$ with $\vv^i :=
+\vw_B^i + C' \vw_R^i$.  All constraints remain the same. The only notable
+change occurs in Eq.~(\ref{eq:wbounds2}), which becomes:
 %
-$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C^\bot \vw_R) \le \vv^i \le (\vw_B^\top + C^\top \vw_R^\top)$$
+$$ \forall \; i \in [k] \;.\; (\vw_B^\bot + C' \vw_R^\bot) \le \vv^i \le (\vw_B^\top + C' \vw_R^\top)$$
 %
 As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
-to be non-negative. \andrea{check formulation}
+to be non-negative.
 
@@
 \stefano{explain that we can't turn off $\gamma$}
+
+\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
+methods, which generate a single candidate option per iteration, {\sc
+SetMargin} constructs a whole {\em set} of options. We thus can not leverage
+standard query selection strategies \stefano{which ones?}.
+
+The simplest solution is to query the user about all $k \choose 2$ pairs of
+configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
+number of queries does not grow large in practice.  An appealing alternative
+involves using a modified sorting algorithm, such as merge-sort, to order the
+set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
+approach is very susceptible to noise: a single inconsistent user response may
+affect all successive comparisons, leading to a cascading effect. While more
+robust strategies may be conceived, empirically we found that the simpler
+quadratic scheme tends to work reasonably well.
+
+\andrea{rewrite as footnote}
 

### 2016-01-26 paper/ijcai16.tex
paper: remove authors for the time being.
paper: worked on experiment section

@@
 \title{Constructive Preference Elicitation by Setwise Max-margin Learning}
-\author{ST and AP and PV}
+\author{}
 
@@
 
-We adopt the {\em indifference-augmented} Bradley-Torr user response model
-introduced in~\cite{guo2010real}. When queried about a pair of configurations,
-the user replies with the true utility of the two items. According to the
-classical Bradley-Torr model, which does not attempt to model {\em indifference},
-the proabability that a user $\vw$ will prefer configuration $\vx^i$ over
-the alternative $\vx^j$, barring indifference, amounts to:
+We adopt the {\em indifference-augmented} Bradley-Terry user response
+model introduced in~\cite{guo2010real}. According to the classical
+(without indifference) Bradley-Terry model~\cite{BraTer52}, when queried about a pair
+of configurations, the probability that a user will prefer
+configuration $\vx^i$ over the alternative $\vx^j$ amounts to:
 %
-$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
+$$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
 %
-In other words, the probability grows with the difference between the utilities
-of the two items. Support for indifference has been introduced
-in~\cite{guo2010real} by postulating that the probability of the user being
-indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
-the two utilities are, as in:
+where $\vw$ is the weight vector of the true underlying user utility.
+In other words, the probability grows with the difference between the
+utilities of the two items. Support for indifference has been
+introduced in~\cite{guo2010real} by postulating that the probability
+of the user being indifferent about the two configurations depends on
+how close their utilities are, as in:
 %
-$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
+$$ \exp(\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
 %
-The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
-see the latter for more details
-\stefano{The above does not satisfy me at all.}
+The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
+simulatins, as in~\cite{guo2010real}. Please see the latter for more
+details. \stefano{The above does not satisfy me at all.}\andrea{sounds reasonable now, I would remove Please see..?}
 
@@
 while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
-\stefano{explain that we can't turn off $\gamma$}
+\stefano{explain that we can't turn off $\gamma$}\andrea{not so needed, but explain why ranges are different}
+
+\paragraph{Synthetic Dataset.} Following the experimental protocol
+in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first
+experiment we evaluate the behavior of the proposed method in an
+artificial setting with increasingly complex problems. We developed
+synthetic datasets with $r$ attributes, for increasing values of $r$.
+Each attribute takes one of $r$ possible values, so that the one-hot
+encoding of attributes results in $m=r^2$ features. In terms of space
+of configurations, for $r=2$ the synthetic dataset corresponds to 
+$\calX_\text{feasible} = [2] \times [2]$, for $r=3$ to
+$\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
+cardinality of $\calX_\text{feasible}$ is $r^r$, and grows (super)
+exponentially with $r$.
+% We developed
+% seven synthetic datasets: each dataset involves $m=2,\ldots,7$
+% attributes, where each attribute takes one of $m$ possible
+% values. More explicitly, for $m=2$ the synthetic dataset is
+% $\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to
+% $\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
+% cardinality of $\calX_\text{feasible}$ is $m^m$, and grows (super)
+% exponentially with $m$. 
+For $r=3$, the dataset is comparable to the synthetic one used
+in~\cite{guo2010real} and~\cite{viappiani2010optimal}, which have
+three attributes with 2, 2 and 5 values respectively, for a total of
+20 feasible configurations.  For larger $r$ the size of the space
+grows much larger than the one typically used in the Bayesian
+preference elicitation literature, and as such represents a good test
+bed for comparing the scalability of the various methods. The feasible
+configuration space was encoded in {\sc SetMargin} through appropriate
+MILP constraints, while the other methods require all datasets to be
+explicitly grounded.
+
+Users were simulated by drawing $20$ random utility vectors from each of four
+different distributions. The first two mimic those used
+in~\cite{guo2010real}\footnote{The normal distribution reported
+in~\cite{guo2010real} differs slightly from the one implemented in the code of
+their experimental setting. In this paper we use the latter.}:
+(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
+a normal distribution with mean $25$ and covariance
+$\frac{25}{3}$.  We further produced two novel
+{\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
+vectors sampled in the uniform and normal settings uniformly at random
+\stefano{Paolo, is this correct?}. All methods were evaluated on the very same
+weight vectors.
+
+In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
+loss:
+%
+$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
+%
+varying the number of collected user responses. Here $u(\cdot)$ is the
+true unknown user utility, and $\vx^*$ is the solution recommended to
+the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
+
+The thick line is the median
+utility loss taken over all 20 users, while the shaded area represents the
+variance. \andrea{standard deviation?}
+
+\stefano{plot average utility loss at the increase of dataset size}
+
+\stefano{plot average runtime at the increase of dataset size}
+
+\stefano{comment on usefulness of sparse weights}
 
@@
 
-\paragraph{Synthetic Dataset.} Following the experimental protocol
-in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
-we evaluate the behavior of the proposed method in a challenging artificial
-setting. We developed seven synthetic datasets: each dataset involves
-$m=2,\ldots,7$ attributes, where each attribute takes one of $m$ possible
-values. More explicitly, for $m=2$ the synthetic dataset is
-$\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to $\calX_\text{feasible} =
-[3] \times [3] \times [3]$, and so on. The cardinality of
-$\calX_\text{feasible}$ is $m^m$, and grows (super) exponentially with $m$. For
-$m=3$, the dataset is comparable to the synthetic one used
-in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
-of size of 2, 2 and 5, respectively, for a total of 20 feasible
-configurations), while for larger $m$ it grows much larger than those typically
-used in the Bayesian preference elicitation literature, and as such represents
-a good test bed for comparing the scalability of the various methods.
-The feasible configuration space was encoded in {\sc SetMargin} through
-appropriate MILP constraints, while the other methods required all datasets to
-be explicitly grounded.
-
-Users were simulated by drawing $20$ random utility vectors from each of four
-different distributions. The first two mimic those used
-in~\cite{guo2010real}\footnote{The normal distribution reported
-in~\cite{guo2010real} differs slightly from the one implemented in the code of
-their experimental setting. In this paper we use the latter.}:
-(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
-a normal distribution with mean $25$ and covariance
-$\frac{25}{3}$.  We further produced two novel
-{\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
-vectors sampled in the uniform and normal settings uniformly at random
-\stefano{Paolo, is this correct?}. All methods were evaluated on the very same
-weight vectors.
-
-In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
-loss:
-%
-$$ \left(\max_{\vx\in\calX_\text{feasible}} u(\vx)) - u(\vx^{i}) \right) $$
-%
-varying the number of collected user responses. The thick line is the median
-utility loss taken over all 20 users, while the shaded area represents the
-variance.
-
-\stefano{plot average utility loss at the increase of dataset size}
-
-\stefano{plot average runtime at the increase of dataset size}
-
-\stefano{comment on usefulness of sparse weights}
-
 \paragraph{Constructive dataset.} Next, we tested {\sc SetMargin} on a truly
@@
 G4 @733 MHz''), so to avoid multiplicative terms in the price function.
-\stefano{too detailed/handwavy?}
+\stefano{too detailed/handwavy?} \andrea{si'. irrilevante}
 
 Interactions between attributes are expressed as Horn clauses. An example
-costraint might look like this: ``if the manufacturer is Apple, then the
+constraint might look like this: ``if the manufacturer is Apple, then the
 CPU must be either a PowerPC G3 or a PowerPC G4'', which can be encoded
-into MILP form as:
+into MILP form as:\andrea{sbaglio o e' vecchiotto il dataset? footnote?}
 %
@@
 %
-Note that mutual exclusivity between all binary variables corresponding to an
-attribute ensures that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1.
-The dataset includes constraints between the following attributes: manufacturer
-$\to$ type, manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
-size, type $\to$ monitor size. In total we encoded 16 Horn constraints, each
-with one head and multiple options in the body. We do not report the full
-list due to space limitations.
+Note that mutual exclusivity between all binary variables of the
+one-hot encoding of an attribute ensures that only one of
+$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
+constraints between the following attributes: manufacturer $\to$ type,
+manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
+size, type $\to$ monitor size. In total we encoded 16 Horn
+constraints, each with one head and multiple options in the body. We
+do not report the full list due to space limitations.
 

### 2016-01-26 paper/ijcai16.tex
paper: update experiments section under Andrea's guidance.

@@
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
-simulatins, as in~\cite{guo2010real}. Please see the latter for more
-details. \stefano{The above does not satisfy me at all.}\andrea{sounds reasonable now, I would remove Please see..?}
+simulations, as in~\cite{guo2010real}.
 
@@
 the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
-while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
-\stefano{explain that we can't turn off $\gamma$}\andrea{not so needed, but explain why ranges are different}
+while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$\footnote{Note
+that $\alpha$ can not be taken to be less than $1$, as in this case, the
+objective can be increased arbitrarily while keeping the right-hand side of
+Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}.
 
@@
 encoding of attributes results in $m=r^2$ features. In terms of space
-of configurations, for $r=2$ the synthetic dataset corresponds to 
-$\calX_\text{feasible} = [2] \times [2]$, for $r=3$ to
-$\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
+of configurations, for $r=3$ the synthetic dataset corresponds to 
+$\calX_\text{feasible} = [3] \times [3] \times [3]$, for $r=4$ to
+$\calX_\text{feasible} = [4] \times [4] \times [4] \times [4]$, and so on. The
 cardinality of $\calX_\text{feasible}$ is $r^r$, and grows (super)
@@
 utility loss taken over all 20 users, while the shaded area represents the
-variance. \andrea{standard deviation?}
+standard deviation.
 
@@
 well approximated by sum of the price of its components plus an offset due to
-branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
-type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
-G4 @733 MHz''), so to avoid multiplicative terms in the price function.
-\stefano{too detailed/handwavy?} \andrea{si'. irrilevante}
+branding.
 
@@
 constraint might look like this: ``if the manufacturer is Apple, then the
-CPU must be either a PowerPC G3 or a PowerPC G4'', which can be encoded
-into MILP form as:\andrea{sbaglio o e' vecchiotto il dataset? footnote?}
+CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
+included in the dataset, which are a bit outdated, the dataset itself is pretty
+realistic.}, which can be encoded
+into MILP form as:
 %
@@
 
-- PC dataset without costs
-
-- PC dataset with costs
+\stefano{Add figures and comments.}
 

### 2016-01-26 paper/ijcai16.tex
edits in the abstract a# Please enter the commit message for your changes. Lines starting

--- 2016-01-26 paper/ijcai16.tex
@@
 
-\title{Constructive Preference Elicitation by Setwise Max-margin Learning}
-\author{ST and AP and PV}
+\title{Constructive Preference Elicitation by Setwise Max-margin Learning} %with?
+\author{ID xxxx}
 
@@
 \begin{abstract}
-WRITEME
+In this paper we propose an approach to preference elicitation that is suitable to large
+configuration problem domains where stated preferences can be noisy.
+Our setwise max-margin method can be viewed as a generalization of max-margin learning to sets,
+and can produce a set of ``diverse'' items that can be used to ask informative queries to the user.
+Moreover, the approach can ensuring sparsity in the parameter space, in order to favor the assessment of utility towards combinations of weights that concentrate on just few features.
+We present an optimization formulation and show, with simulations, how our approach is able to efficiently elicit the user's utility function even in very large datasets.
+
 \end{abstract}
@@
 
-WRITEME
+%The problem is that of learning or assessing the preferences of the user (or decision maker).
+Preferences \cite{Peintner2008} play an important role in artificial intelligence,
+including fields spanning from recommender systems to automatic planning, 
+from non monotonic reasoning to computational social choice and algorithmic decision theory.
+The elicitation of learning of preferences is therefore a crucial task, .
+The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision.
 
-- importance of recommendation systems and preference elicitation
-
-- existing approaches and their limitations
+While preference elicitation has been often cast as that of adapting using ad-hoc rules \cite{},
+recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
+One option is to be Bayesian and to maintain a probability distribution over the possible realizations of the utility's parameters \cite{chajewska2000,boutilier2002,guo2010real,viappiani2010optimal}.
+Alternatively, when distribution information about the parameters is not available, one can reason about all feasible utility functions consistent with the currently known information about the user; from the user's response we are able to infer constraints on the parameters implicated in the utility model \cite{Wang2003,boutilier2006,braziunasmmr:uai07,viappiani2009}.
+This latter approach has also the advantage of avoiding the computationally demanding task of maintaining distribution information and performing Bayesian updates.
 
@@
 problem in the space of feasible instances, guided by the learned
-utility. \andrea{other constructive preference elicitation approaches?}
+utility.
+
+%\andrea{other constructive preference elicitation approaches?}
 
@@
 equally plausible alternative hypotheses for the unknown
-utility. Preference elicitation works by combining weight vector
+utility. 
+Our approach to preference elicitation works by combining weight vector
 learning with instance generation, so that each iteration of the
@@
 
-- experimental results
+We evaluate the effectiveness of our approach by testing our elicitation method
+in both synthetic and configuration problems, and comparing it to state-of-the-art
+methods.
 
-- paper structure
+The paper is structured as follows: after discussing related works in next Section,
+we present some brackground material in Section\ref{sec:background} 
+and then proceed to our setwise maxmargin method in Section\ref{sec:formulationm}; 
+we discuss experimental results in Section\ref{sec:experiments}
+and conclude with final remarks (Section\ref{sec:conclusions}).
 
@@
 
+Preference elicitation for the customization of user interfaces \cite{gajos2005}.
+
 WRITEME
 
-\section{Preference Elicitation}
+\section{Background}
+\ref{sec:background}
+%\section{Preference Elicitation}
 
@@
 
-In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
+Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
 of linear utility functions: the user's preferences are represented by a weight
@@
 learning system, and must be estimated from pairwise choices elicited from the
-user.\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  We consider three possible outcomes for a pairwise comparison between
+user.
+
+%\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  
+Comparison queries are especially natural, asking a user if she prefers one option x to another y;
+these comparisons can be extended to choice sets of more than two options \cite{viappiani2009,viappiani2010optimal}.
+Comparison and choice queries are common in conjoint analysis \cite{louviere2000,toubia2004}.
+
+We consider three possible outcomes for a pairwise comparison between
 two configurations $\vx$ and $\vx'$: either $\vx$ is preferred to $\vx'$ (written
@@
 no clear preference between the two items ($\vx \approx \vx'$). We write
-$\calD$ to denote the full set of answers elicited from the user. Queries are
-typically selected so as to maximize some measure of informativeness of the
+$\calD$ to denote the set of preferences (answers to comparison queries) 
+elicited from the user. Queries are typically selected so as to maximize 
+some measure of informativeness of the
 resulting responses.  Details on our actual query selection strategy will be
@@
 configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration
-$\vx^{i}$ is the ``best'' possible option when evaluated according to
+$\vx^{i}$ is the ``best'' possref:keeney-mautible option when evaluated according to
 the corresponding $\vw^{i}$ and configurations are maximally diverse
@@
 
-\paragraph{MILP Formulation.} This initial formulation is problematic,
-as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed
-continuous integer variables. Here we show how to reformulate the
+\paragraph{MILP Formulation.} This initial formulation is problematic to solve,
+as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed continuous integer variables. 
+Here we show how to reformulate the
 problem as a mixed integer linear program (MILP). Our goal is to
@@
 classical Bradley-Torr model, which does not attempt to model {\em indifference},
-the proabability that a user $\vw$ will prefer configuration $\vx^i$ over
+the probability that a user $\vw$ will prefer configuration $\vx^i$ over
 the alternative $\vx^j$, barring indifference, amounts to:
@@
 standard query selection strategies \stefano{which ones?}. 
+\paolo{I would remove the last sentence}
 
@@
 
-\section*{Acknowledgments}
-
-WRITEME
+%\section*{Acknowledgments}
+%WRITEME
 

### 2016-01-26 paper/ijcai16.tex
paper: added lasso refs

--- 2016-01-26 paper/ijcai16.tex
@@
 will be indifferent to most of them~\cite{} \stefano{@Paolo: add refs}. The $\ell_1$ penalty is frequently
-used to improve the sparsity of learned models~\cite{lasso}, with consequent gains
+used to improve the sparsity of learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent gains
 in generalization ability and efficiency, as confirmed by our empirical
@@
 the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
-while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$\footnote{Note
+while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
 that $\alpha$ can not be taken to be less than $1$, as in this case, the
 objective can be increased arbitrarily while keeping the right-hand side of
-Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}.
+Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}
 

### 2016-01-26 paper/ijcai16.tex

--- 2016-01-26 paper/ijcai16.tex
@@
 
-We adopt the {\em indifference-augmented} Bradley-Torr user response model
-introduced in~\cite{guo2010real}. When queried about a pair of configurations,
-the user replies with the true utility of the two items. According to the
-classical Bradley-Torr model, which does not attempt to model {\em indifference},
-the probability that a user $\vw$ will prefer configuration $\vx^i$ over
-the alternative $\vx^j$, barring indifference, amounts to:
+We adopt the {\em indifference-augmented} Bradley-Terry user response
+model introduced in~\cite{guo2010real}. According to the classical
+(without indifference) Bradley-Terry model~\cite{BraTer52}, when queried about a pair
+of configurations, the probability that a user will prefer
+configuration $\vx^i$ over the alternative $\vx^j$ amounts to:
+
 %
-$$ (1 + \exp(-\alpha \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
+$$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
 %
-In other words, the probability grows with the difference between the utilities
-of the two items. Support for indifference has been introduced
-in~\cite{guo2010real} by postulating that the probability of the user being
-indifferent about the two configurations $\vx^i, \vx^j$ depends on how close
-the two utilities are, as in:
+where $\vw$ is the weight vector of the true underlying user utility.
+In other words, the probability grows with the difference between the
+utilities of the two items. Support for indifference has been
+introduced in~\cite{guo2010real} by postulating that the probability
+of the user being indifferent about the two configurations depends on
+how close their utilities are, as in:
 %
-$$ \exp(\beta |\langle\vw,\vx^i - \vx^j\rangle|) $$
+$$ \exp(\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
 %
-The parameters $\alpha$ and $\beta$ were set as in~\cite{guo2010real}. Please
-see the latter for more details
-\stefano{The above does not satisfy me at all.}
+The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
+simulations, as in~\cite{guo2010real}.
 
@@
 the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
-while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
-\stefano{explain that we can't turn off $\gamma$}
-
-\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
-methods, which generate a single candidate option per iteration, {\sc
-SetMargin} constructs a whole {\em set} of options. We thus can not leverage
-standard query selection strategies \stefano{which ones?}. 
-\paolo{I would remove the last sentence}
-
-The simplest solution is to query the user about all $k \choose 2$ pairs of
-configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
-number of queries does not grow large in practice.  An appealing alternative
-involves using a modified sorting algorithm, such as merge-sort, to order the
-set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
-approach is very susceptible to noise: a single inconsistent user response may
-affect all successive comparisons, leading to a cascading effect. While more
-robust strategies may be conceived, empirically we found that the simpler
-quadratic scheme tends to work reasonably well.
-
-\andrea{rewrite as footnote}
+% <<<<<<< HEAD
+% while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
+% \stefano{explain that we can't turn off $\gamma$}
+% 
+% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
+% methods, which generate a single candidate option per iteration, {\sc
+% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
+% standard query selection strategies \stefano{which ones?}. 
+% \paolo{I would remove the last sentence}
+% 
+% The simplest solution is to query the user about all $k \choose 2$ pairs of
+% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
+% number of queries does not grow large in practice.  An appealing alternative
+% involves using a modified sorting algorithm, such as merge-sort, to order the
+% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
+% approach is very susceptible to noise: a single inconsistent user response may
+% affect all successive comparisons, leading to a cascading effect. While more
+% robust strategies may be conceived, empirically we found that the simpler
+% quadratic scheme tends to work reasonably well.
+% 
+% \andrea{rewrite as footnote}
+% =======
+while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$\footnote{Note
+that $\alpha$ can not be taken to be less than $1$, as in this case, the
+objective can be increased arbitrarily while keeping the right-hand side of
+Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}.
 
 \paragraph{Synthetic Dataset.} Following the experimental protocol
-in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first experiment
-we evaluate the behavior of the proposed method in a challenging artificial
-setting. We developed seven synthetic datasets: each dataset involves
-$m=2,\ldots,7$ attributes, where each attribute takes one of $m$ possible
-values. More explicitly, for $m=2$ the synthetic dataset is
-$\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to $\calX_\text{feasible} =
-[3] \times [3] \times [3]$, and so on. The cardinality of
-$\calX_\text{feasible}$ is $m^m$, and grows (super) exponentially with $m$. For
-$m=3$, the dataset is comparable to the synthetic one used
-in~\cite{guo2010real} and~\cite{viappiani2010optimal} (which have three domains
-of size of 2, 2 and 5, respectively, for a total of 20 feasible
-configurations), while for larger $m$ it grows much larger than those typically
-used in the Bayesian preference elicitation literature, and as such represents
-a good test bed for comparing the scalability of the various methods.
-The feasible configuration space was encoded in {\sc SetMargin} through
-appropriate MILP constraints, while the other methods required all datasets to
-be explicitly grounded.
+in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first
+experiment we evaluate the behavior of the proposed method in an
+artificial setting with increasingly complex problems. We developed
+synthetic datasets with $r$ attributes, for increasing values of $r$.
+Each attribute takes one of $r$ possible values, so that the one-hot
+encoding of attributes results in $m=r^2$ features. In terms of space
+of configurations, for $r=3$ the synthetic dataset corresponds to 
+$\calX_\text{feasible} = [3] \times [3] \times [3]$, for $r=4$ to
+$\calX_\text{feasible} = [4] \times [4] \times [4] \times [4]$, and so on. The
+cardinality of $\calX_\text{feasible}$ is $r^r$, and grows (super)
+exponentially with $r$.
+% We developed
+% seven synthetic datasets: each dataset involves $m=2,\ldots,7$
+% attributes, where each attribute takes one of $m$ possible
+% values. More explicitly, for $m=2$ the synthetic dataset is
+% $\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to
+% $\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
+% cardinality of $\calX_\text{feasible}$ is $m^m$, and grows (super)
+% exponentially with $m$. 
+For $r=3$, the dataset is comparable to the synthetic one used
+in~\cite{guo2010real} and~\cite{viappiani2010optimal}, which have
+three attributes with 2, 2 and 5 values respectively, for a total of
+20 feasible configurations.  For larger $r$ the size of the space
+grows much larger than the one typically used in the Bayesian
+preference elicitation literature, and as such represents a good test
+bed for comparing the scalability of the various methods. The feasible
+configuration space was encoded in {\sc SetMargin} through appropriate
+MILP constraints, while the other methods require all datasets to be
+explicitly grounded.
 
@@
 %
-$$ \left(\max_{\vx\in\calX_\text{feasible}} u(\vx)) - u(\vx^{i}) \right) $$
+$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
 %
-varying the number of collected user responses. The thick line is the median
+varying the number of collected user responses. Here $u(\cdot)$ is the
+true unknown user utility, and $\vx^*$ is the solution recommended to
+the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
+
+The thick line is the median
 utility loss taken over all 20 users, while the shaded area represents the
-variance.
+standard deviation.
 
@@
 \stefano{comment on usefulness of sparse weights}
+
+\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
+methods, which generate a single candidate option per iteration, {\sc
+SetMargin} constructs a whole {\em set} of options. We thus can not leverage
+standard query selection strategies \stefano{which ones?}.
+
+The simplest solution is to query the user about all $k \choose 2$ pairs of
+configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
+number of queries does not grow large in practice.  An appealing alternative
+involves using a modified sorting algorithm, such as merge-sort, to order the
+set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
+approach is very susceptible to noise: a single inconsistent user response may
+affect all successive comparisons, leading to a cascading effect. While more
+robust strategies may be conceived, empirically we found that the simpler
+quadratic scheme tends to work reasonably well.
+
+\andrea{rewrite as footnote}
 
@@
 well approximated by sum of the price of its components plus an offset due to
-branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
-type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
-G4 @733 MHz''), so to avoid multiplicative terms in the price function.
-\stefano{too detailed/handwavy?}
+branding.
 
 Interactions between attributes are expressed as Horn clauses. An example
-costraint might look like this: ``if the manufacturer is Apple, then the
-CPU must be either a PowerPC G3 or a PowerPC G4'', which can be encoded
+constraint might look like this: ``if the manufacturer is Apple, then the
+CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
+included in the dataset, which are a bit outdated, the dataset itself is pretty
+realistic.}, which can be encoded
 into MILP form as:

### 2016-01-26 paper/ijcai16.tex
paper: add download URL.

@@
 the user responses collected so far. $\alpha$ is taken in $\{20, 10, 5, 1\}$,
-% <<<<<<< HEAD
-% while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.
-% \stefano{explain that we can't turn off $\gamma$}
-% 
-% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
-% methods, which generate a single candidate option per iteration, {\sc
-% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
-% standard query selection strategies \stefano{which ones?}. 
-% \paolo{I would remove the last sentence}
-% 
-% The simplest solution is to query the user about all $k \choose 2$ pairs of
-% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
-% number of queries does not grow large in practice.  An appealing alternative
-% involves using a modified sorting algorithm, such as merge-sort, to order the
-% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
-% approach is very susceptible to noise: a single inconsistent user response may
-% affect all successive comparisons, leading to a cascading effect. While more
-% robust strategies may be conceived, empirically we found that the simpler
-% quadratic scheme tends to work reasonably well.
-% 
-% \andrea{rewrite as footnote}
-% =======
-while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$\footnote{Note
+while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
 that $\alpha$ can not be taken to be less than $1$, as in this case, the
 objective can be increased arbitrarily while keeping the right-hand side of
-Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}.
+Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}
 

### 2016-01-26 paper/ijcai16.tex
Introduction

@@
 
-%The problem is that of learning or assessing the preferences of the user (or decision maker).
 Preferences \cite{Peintner2008} play an important role in artificial intelligence,
@@
 from non monotonic reasoning to computational social choice and algorithmic decision theory.
-The elicitation of learning of preferences is therefore a crucial task, .
-The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision.
+The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be usually high (from a cognitive and/or computational point of view).
+The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision or a course of action.
 
-While preference elicitation has been often cast as that of adapting using ad-hoc rules \cite{},
+The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{}.
+In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
+While preferences are often assessed using some ad-hoc rules \cite{},
 recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
-One option is to be Bayesian and to maintain a probability distribution over the possible realizations of the utility's parameters \cite{chajewska2000,boutilier2002,guo2010real,viappiani2010optimal}.
-Alternatively, when distribution information about the parameters is not available, one can reason about all feasible utility functions consistent with the currently known information about the user; from the user's response we are able to infer constraints on the parameters implicated in the utility model \cite{Wang2003,boutilier2006,braziunasmmr:uai07,viappiani2009}.
-This latter approach has also the advantage of avoiding the computationally demanding task of maintaining distribution information and performing Bayesian updates.
+
+%One option is to be Bayesian and to maintain a probability distribution over the possible realizations of the utility's parameters \cite{chajewska2000,boutilier2002,guo2010real,viappiani2010optimal}.
+%Alternatively, when distribution information about the parameters is not available, one can reason about all feasible utility functions consistent with the currently known information about the user; from the user's response we are able to infer constraints on the parameters implicated in the utility model \cite{Wang2003,boutilier2006,braziunasmmr:uai07,viappiani2009}.
+%This latter approach has also the advantage of avoiding the computationally demanding task of maintaining distribution information and performing Bayesian updates.
 
 % constructive preference elicitation
-In this paper we take a {\em constructive} view on preference
+While most works assume that items or decisions are available in a (possibly large)
+dataset, in this paper we propose an adaptive elicitation framework that takes a {\em constructive} view on preference
 elicitation, enlarging its scope from the selection of items among a
-set of candidates to the synthesis of entirely novel
-instances. Instances are represented as combinations of basic elements
+set of candidates to the synthesis of entirely novelinstances. 
+Instances are solutions to a given optimization problem; they are represented as combinations of basic elements
 (e.g. the components of a laptop) subject to a set of constraints
-(e.g. the laptop model determines the set of available CPUs). A
-utility function is learned over the feature representation of an
-instance, as customary in many preference elicitation approaches. The
-recommendation is then made by solving a constrained optimization
+(e.g. the laptop model determines the set of available CPUs). 
+A utility function is learned over the feature representation of an
+instance, as customary in many preference elicitation approaches. 
+The recommendation is then made by solving a constrained optimization
  problem in the space of feasible instances, guided by the learned
@@
 
-%\andrea{other constructive preference elicitation approaches?}
+Our approach is related to \cite{boutilier2006}, where minimax regret is used both
+as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
+The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that item x is preferred to y, even though the latter has higher utility).
 
 % setwise max-margin formulation to deal with uncertainty in user utility
-Learning a user utility function requires setting a preference
+Indeed, learning a user utility function requires setting a preference
 elicitation strategy and dealing with uncertain and possibly
-inconsistent user feedback. Bayesian preference elicitation approaches
+inconsistent user feedback. 
+Bayesian preference elicitation approaches
 deal with both problems by building a probability distribution on
-candidate functions and asking queries maximizing informativeness
-measures such as expected value of information (EVOI)~\cite{}. These
+candidate functions (endowed with a response or error model to be used for inference) and asking queries maximizing informativeness
+measures such as expected value of information (EVOI)~\cite{chajewska2000,viappiani2010optimal}.
+
+These
 approaches are however computationally expensive and cannot scale to
@@
 The paper is structured as follows: after discussing related works in next Section,
-we present some brackground material in Section\ref{sec:background} 
-and then proceed to our setwise maxmargin method in Section\ref{sec:formulationm}; 
-we discuss experimental results in Section\ref{sec:experiments}
-and conclude with final remarks (Section\ref{sec:conclusions}).
+we present some brackground material in Section \ref{sec:background} 
+and then proceed to our setwise maxmargin method in Section \ref{sec:formulation}; 
+we discuss experimental results in Section \ref{sec:experiments}
+and conclude with final remarks (Section \ref{sec:conclusions}).
 
-\section{Related Work}
+%\section{Related Work}
 
-Preference elicitation for the customization of user interfaces \cite{gajos2005}.
+%Preference elicitation for the customization of user interfaces \cite{gajos2005}.
 
-WRITEME
+%WRITEME
 
 \section{Background}
-\ref{sec:background}
+\label{sec:background}
 %\section{Preference Elicitation}
@@
     \end{center}
-    \caption{\label{fig:setmargin} Intuition.}
+    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utilty vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
 \end{figure}
@@
 or once another suitable condition is met. \andrea{@Paolo: what termination
-condition for pref elicitation?}  The procedure is sketched in
+condition for pref elicitation?}  
+Termination might be left to the user to decide (see for example \ref{} for discussion on user-involved recommender systems); the system can however propose
+to terminate when the difference between the utility weights is small enough.
+
+The procedure is sketched in
 Algorithm~\ref{alg:setmargin}. Note that at the end of the preference
@@
 6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
-and the full experimental setup are available at \stefano{add URL}.
+and the full experimental setup are available at {\tt https://db.tt/oqY62odW}.
+
+ \stefano{add URL}.
 
@@
 \section{Conclusion}
+\label{sec:conclusions}
 
-WRITEME
+This paper presented a maxmargin approach to generating a set of diverse examples.
+Our approach, that can be seen as an extension of maxmargin learning to sets, is effective in the generation of a diverse set of items that can be used to ask informative queries to the user.
+The main advantages of our elicitation method are 1) its aptitude for eliciting preferences and providing recommendations in large configuration problems  2) robustness with respect to erroneous preference input, and 3) its ability to support sparse utility functions.
+In a number of experiments we showed how our method compare with state-of-the-art approaches.
+
+\paolo{any idea about possible future works?}
+% FUTURE WORKS
 

### 2016-01-27 paper/ijcai16.tex
paper: adjust single word in the abstract.

@@
 and can produce a set of ``diverse'' items that can be used to ask informative queries to the user.
-Moreover, the approach can ensuring sparsity in the parameter space, in order to favor the assessment of utility towards combinations of weights that concentrate on just few features.
+Moreover, the approach can encourage sparsity in the parameter space, in order to favor the assessment of utility towards combinations of weights that concentrate on just few features.
 We present an optimization formulation and show, with simulations, how our approach is able to efficiently elicit the user's utility function even in very large datasets.

### 2016-01-27 paper/ijcai16.tex
paper: remove obsolete TODO.
paper: shorten/improve PC section a bit.

@@
 
- \stefano{add URL}.
-
 We compare {\sc SetMargin} against two strong Bayesian baselines:
@@
 
-A PC configurations is defined by seven attributes: computer type (laptop,
-desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
-size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
-attributes. The price attribute is defined as a linear combination of the other
+A PC configurations is defined by eight attributes: computer type (laptop,
+desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)
+size, and price.
+%See Table~\ref{tab:pcdataset} for an overivew of the attributes.
+The price attribute is defined as a linear combination of the other
 attributes: this is a fair modeling choice, as often the price of a PC is
@@
 branding.
-
+%
 Interactions between attributes are expressed as Horn clauses. An example
@@
 realistic.}, which can be encoded
-into MILP form as:
+into MILP form as
 %
-$$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $$
+$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
 %
@@
 manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
-size, type $\to$ monitor size. In total we encoded 16 Horn
-constraints, each with one head and multiple options in the body. We
+size, type $\to$ monitor size, for 16 Horn constraints total. We
 do not report the full list due to space limitations.
 
-\begin{table}
-    \centering
-    \begin{tabular}{ccc}
-        {\bf Attribute} & {\bf Type} & {\bf Values} \\
-        \hline \hline
-        Type & discrete & 3 \\
-        Manufacturer & discrete & 8 \\
-        CPU model & discrete & 37 \\
-        RAM amount & discrete & 10 \\
-        HD size & discrete & 10 \\
-        Monitor size & discrete & 8 \\
-        Price & continuous & --
-    \end{tabular}
-    \caption{\label{tab:pcdataset} Statistics for the PC dataset.}
-\end{table}
+% \begin{table}
+%     \centering
+%     \begin{tabular}{ccc}
+%         {\bf Attribute} & {\bf Type} & {\bf Values} \\
+%         \hline \hline
+%         Type & discrete & 3 \\
+%         Manufacturer & discrete & 8 \\
+%         CPU model & discrete & 37 \\
+%         RAM amount & discrete & 10 \\
+%         HD size & discrete & 10 \\
+%         Monitor size & discrete & 8 \\
+%         Price & continuous & --
+%     \end{tabular}
+%     \caption{\label{tab:pcdataset} Statistics for the PC dataset.}
+% \end{table}
 

### 2016-01-27 paper/ijcai16.tex
paper: s/cannot/can not/.
paper: minor spelling fixes.

@@
 elicitation, enlarging its scope from the selection of items among a
-set of candidates to the synthesis of entirely novelinstances. 
+set of candidates to the synthesis of entirely novel instances. 
 Instances are solutions to a given optimization problem; they are represented as combinations of basic elements
@@
 These
-approaches are however computationally expensive and cannot scale to
+approaches are however computationally expensive and can not scale to
 fully constructive scenarios, as shown in our experimental results.
@@
 The paper is structured as follows: after discussing related works in next Section,
-we present some brackground material in Section \ref{sec:background} 
+we present some background material in Section \ref{sec:background} 
 and then proceed to our setwise maxmargin method in Section \ref{sec:formulation}; 
@@
     \end{center}
-    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utilty vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
+    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utility vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
 \end{figure}
@@
 used in~\cite{guo2010real}: instead of explicitly enumerating all possible PC
-item, we definined the set of feasible configurations through MILP constraints.
+item, we defined the set of feasible configurations through MILP constraints.
 

### 2016-01-27 draw.py
draw: update for Guo/Viappiani comparison.

@@ -6,15 +6,16 @@ import numpy as np
 from scipy.io import loadmat
 import matplotlib.pyplot as plt
 from matplotlib.font_manager import FontProperties
+from collections import defaultdict
 from glob import glob
 
 # Colors taken from http://tango.freedesktop.org/Tango_Icon_Theme_Guidelines
-COLOR_OF = {
-    1: ("#000000", "#000000"),
-    2: ("#CE5C00", "#F57900"),
-    3: ("#204A87", "#3465A4"),
-    4: ("#4E9A06", "#73D216"),
-}
+COLORS = [
+    ("#CE5C00", "#F57900"), # orange
+    ("#204A87", "#3465A4"), # sky blue
+    ("#4E9A06", "#73D216"), # chameleon
+    ("#2E3436", "#555753"), # aluminium
+]
 
 def infos_to_matrices(infos, per_query):
     num_trials = len(infos)
@@ -49,109 +50,110 @@ def infos_to_matrices(infos, per_query):
     else:
         return loss_matrix[:,:50], time_matrix[:,:50]
 
-def draw(paths, dest_path, per_query):
-    loss_fig, loss_ax = plt.subplots(1, 1)
-    time_fig, time_ax = plt.subplots(1, 1)
+def load_setmargin(path):
+    with open(path, "rb") as fp:
+        infos = pickle.load(fp)
+    return infos
 
-    max_max_x = None
-    max_loss_max_y = None
-    max_time_max_y = None
-    for path in paths:
+def load_viappiani(path):
+    with open(path, "rb") as fp:
+        lines = map(str.strip, fp.readlines())
+    infos = defaultdict(dict)
+    for line in lines[1:]:
+        trial, iteration, loss, time = line.split(", ")
+        infos[int(trial) - 1][int(iteration)] = (1, float(loss), float(time))
+    info_list = []
+    assert infos.keys() == range(20)
+    for trial in range(20):
+        assert infos[trial].keys() == range(101)
+        info_list.append([infos[trial][iteration] for iteration in range(100)])
+    return info_list
 
-        try:
-            with open(path, "rb") as fp:
-                infos = pickle.load(fp)
-            loss_matrix, time_matrix = infos_to_matrices(infos, per_query=per_query)
-            # XXX we have multiple sets of experiments that end up being
-            # drawn together; remove this hack post-IJCAI
[…]

### 2016-01-27 draw_all.sh

@@ -1,19 +1,71 @@
 #!/bin/bash
 
-for sampling_mode in uniform uniform_sparse normal normal_sparse; do
-    for domain_size in `seq 3 6`; do
-        ipath="results/ijcai16/synthetic_${domain_size}/${sampling_mode}"
-        opath="paper/figures/synthetic_${domain_size}_${sampling_mode}"
-        echo "drawing $ipath -> $opath"
-        ./draw.py $ipath ${opath}_per_iter 0
-        ./draw.py $ipath ${opath}_per_query 1
+# setmargin vs guo vs viappiani
+
+for domain_size in `seq 3 4`; do # XXX 5
+    for sampling_mode in uniform uniform_sparse normal normal_sparse; do
+        echo "setmargin vs guo vs viappiani, per iteration, $domain_size $sampling_mode"
+        setmargin_files=`ls results/ijcai16/synthetic_${domain_size}/${sampling_mode}/*k=2*50__300*.pickle`
+        guo_files=`ls results/ijcai16/guo/synthetic_${domain_size}/${sampling_mode}/*.mat`
+        viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
+        viappiani_eus_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*EUS*.txt`
+        output="paper/figures/synthetic_vs_others_${domain_size}_${sampling_mode}_per_iter"
+        ./draw.py 0 $output "$setmargin_files" "$guo_files" "$viappiani_qi_files" "$viappiani_eus_files"
     done
 done
 
+for domain_size in `seq 3 4`; do # XXX 5
+    for sampling_mode in uniform uniform_sparse normal normal_sparse; do
+        echo "setmargin vs guo vs viappiani, per query, $domain_size $sampling_mode"
+        setmargin_files=`ls results/ijcai16/synthetic_${domain_size}/${sampling_mode}/*k=2*100__100*.pickle`
+        guo_files=`ls results/ijcai16/guo/synthetic_${domain_size}/${sampling_mode}/*.mat`
+        viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
[…]

### 2016-01-27 draw_all.sh
draw: use the threads=1 runs for the setmargin vs guo vs viappiani figures.

@@ -5,7 +5,7 @@
 for domain_size in `seq 3 4`; do # XXX 5
     for sampling_mode in uniform uniform_sparse normal normal_sparse; do
         echo "setmargin vs guo vs viappiani, per iteration, $domain_size $sampling_mode"
-        setmargin_files=`ls results/ijcai16/synthetic_${domain_size}/${sampling_mode}/*k=2*50__300*.pickle`
+        setmargin_files=`ls results/ijcai16/threads\=1/synthetic_${domain_size}/${sampling_mode}/*k=2*.pickle`
         guo_files=`ls results/ijcai16/guo/synthetic_${domain_size}/${sampling_mode}/*.mat`
         viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
         viappiani_eus_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*EUS*.txt`
@@ -17,7 +17,7 @@ done
 for domain_size in `seq 3 4`; do # XXX 5
     for sampling_mode in uniform uniform_sparse normal normal_sparse; do
         echo "setmargin vs guo vs viappiani, per query, $domain_size $sampling_mode"
-        setmargin_files=`ls results/ijcai16/synthetic_${domain_size}/${sampling_mode}/*k=2*100__100*.pickle`
+        setmargin_files=`ls results/ijcai16/threads\=1/synthetic_${domain_size}/${sampling_mode}/*k=2*.pickle`
         guo_files=`ls results/ijcai16/guo/synthetic_${domain_size}/${sampling_mode}/*.mat`
         viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
         viappiani_eus_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*EUS*.txt`

### 2016-01-27 draw_all.sh
paper,draw: fig figure filenames.

@@ -21,7 +21,7 @@ for domain_size in `seq 3 4`; do # XXX 5
         guo_files=`ls results/ijcai16/guo/synthetic_${domain_size}/${sampling_mode}/*.mat`
         viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
         viappiani_eus_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*EUS*.txt`
-        output="paper/figures/synthetic_vs_others_${domain_size}_${sampling_mode}_per_iter"
+        output="paper/figures/synthetic_vs_others_${domain_size}_${sampling_mode}_per_query"
         ./draw.py 1 $output "$setmargin_files" "$guo_files" "$viappiani_qi_files" "$viappiani_eus_files"
     done
 done

### 2016-01-27 paper/ijcai16.tex
paper: update to the final (?) figures.

@@
         \hline
-        {\sc synthetic 3 loss} & {\sc synthetic 4 loss} & {\sc synthetic 5 loss} & {\sc synthetic 5 time}
+        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
         \\
@@
         \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time}
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
         \\
@@
         \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Normal}}
-        \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time}
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
         \\
@@
         \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time}
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
         \\
@@
     }
-    \caption{\label{fig:comparison} {\sc SetMargin} vs Guo vs Viappiani.}
+    \caption{\label{fig:comparison} Comparison between {\sc SetMargin} and the
+    Bayesian competitors.}
 \end{figure}
@@
         \hline
-        \multicolumn{2}{c}{{\sc per query loss}} &
-        \multicolumn{2}{c}{{\sc per iter. loss}}
+        \multicolumn{2}{c}{{\sc loss per iter.}} &
+        \multicolumn{2}{c}{{\sc loss per query}}
         \\
-        {\sc synthetic 4} & {\sc synthetic 6} & {\sc synthetic 4} & {\sc synthetic 6}
+        {\sc synthetic 4} & {\sc synthetic 5} & {\sc synthetic 4} & {\sc synthetic 5}
         \\
         \hline \hline
-        \multicolumn{4}{c}{{\sc Uniform}}
+        \multicolumn{4}{c}{{\sc uniform}}
         \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss}
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
         \\
         \hline
-        \multicolumn{4}{c}{{\sc Uniform Sparse}}
+        \multicolumn{4}{c}{{\sc normal sparse}}
         \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Normal}}
-        \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Normal Sparse}}
-        \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/loss}
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
         \\
@@
      \end{tabular}
-    \caption{\label{fig:selfcomparison} Self-comparison experiment. Only
-    losses, per-query and per-iteration. Runtime described in the text. Each
-    row is a user sampling mode. Three curves per plot: $k=2$ vs $k=3$ vs
-    $k=4$.}
     }
+    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin} with
+    $k=2,3,4$.}
 \end{figure}
+
+% times for the figure above
+% synthetic_vs_self_4_normal_sparse_per_iter_time
+% synthetic_vs_self_4_normal_sparse_per_query_time
+% synthetic_vs_self_4_uniform_per_iter_time
+% synthetic_vs_self_4_uniform_per_query_time
+% synthetic_vs_self_5_normal_sparse_per_iter_time
+% synthetic_vs_self_5_normal_sparse_per_query_time
+% synthetic_vs_self_5_uniform_per_iter_time
+% synthetic_vs_self_5_uniform_per_query_time
 
@@
         \hline
+        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
+        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
         \multicolumn{4}{c}{{\sc uniform sparse}}
         \\
-        \multicolumn{2}{c}{{\sc loss/time per iteration}} &
-        \multicolumn{2}{c}{{\sc loss/time per query}}
-        \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time}
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
         \\
@@
         \\
-        \multicolumn{2}{c}{{\sc loss/time per iteration}} &
-        \multicolumn{2}{c}{{\sc loss/time per query}}
-        \\
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time} &
-        \includegraphics[width=10em]{figures/loss} &
-        \includegraphics[width=10em]{figures/time}
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
         \\
@@
     }
-    \caption{Results for the PC dataset.}
+    \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
+    dataset.}
 \end{figure}

### 2016-01-27 paper/ijcai16.tex
paper: reorder the second exp figure for clarity.
small fixes

@@
 to indicate the $\ell_1$ vector norm, $\langle \cdot, \cdot \rangle$ to
-denote the usual dot product, and $X'$ to denote matrix transposition.
+denote the usual dot product, and $X'$ for matrix transposition.
 
@@
 user.
-
 %\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  
@@
 \label{sec:formulation}
-
 \paragraph{Non-linear Formulation.} We first introduce the problem
-formulation as a mixed integer non-linear optimization problem, and
-then show how to reduce it to a mixed integer linear optimization one.
+formulation as a %mixed integer 
+non-linear optimization problem, and
+then show how to reduce it to a mixed integer linear program.% optimization one.
 
@@
 configurations $\vx^{1}, \ldots, \vx^{k}$, so that each configuration
-$\vx^{i}$ is the ``best'' possref:keeney-mautible option when evaluated according to
+$\vx^{i}$ is the ``best'' possible option when evaluated according to
 the corresponding $\vw^{i}$ and configurations are maximally diverse
@@
 augmented with the slacks, the above inequalities take the form:
-%
-$$ \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h $$
-%
+\begin{align*}
+\langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h 
+\end{align*}
 where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
@@
         \hline
-        \multicolumn{2}{c}{{\sc loss per iter.}} &
-        \multicolumn{2}{c}{{\sc loss per query}}
+        \multicolumn{2}{c}{{\sc synthetic 4}} &
+        \multicolumn{2}{c}{{\sc synthetic 5}}
         \\
-        {\sc synthetic 4} & {\sc synthetic 5} & {\sc synthetic 4} & {\sc synthetic 5}
+        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
         \\
@@
         \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
         \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
         \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
@@
         \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
         \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
         \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}

### 2016-01-27 paper/ijcai16.tex
paper: use hyperref package for URL link.
paper: add a couple of notes for Andrea and myself.

@@
 \usepackage{graphicx}
+\usepackage{hyperref}
 
@@
 
+\stefano{add early stopping in algorithm and description and experiments.}
+
 \begin{algorithm}
@@
 6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
-and the full experimental setup are available at {\tt https://db.tt/oqY62odW}.
+and the full experimental setup are available at \url{https://db.tt/oqY62odW}.
 
@@
 
-The thick line is the median
-utility loss taken over all 20 users, while the shaded area represents the
-standard deviation.
+\begin{itemize}
 
-\stefano{plot average utility loss at the increase of dataset size}
+    \item \ref{fig:comparison} setmargin vs Bayes. similar loss wrt Bayes.
+    variance slighly higher on synthetic 4. funzioniamo meglio su sparso al
+    contrario degli altri. Guo e' piantato al primo utente da tre giorni
+    su synthetic 5.  Paolo procede su S-5 da qualche ora (~8) (1290s su
+    synthetic-5 uniforme per completamento di un utente).
 
-\stefano{plot average runtime at the increase of dataset size}
+    \stefano{@Paolo: varianza tempi?}
 
-\stefano{comment on usefulness of sparse weights}
+    \item \ref{fig:selfcomparison} crossval ogni 5 iterazioni, il che vuol
+    dire diverso numero di queri per $k$ diversi.
+
+\end{itemize}
 
@@
 do not report the full list due to space limitations.
+
+% 76 features (somma)
+% >>> 3*8*37*8*10*10
+% 710400
+
+la crossval ogni 5 iterazioni richiede tempo considerevole, che nel
+grafico cumulativo del tempo si assomma alle crossval successivie, ampliando
+la varianza; in alcuni casi degeneri per alcuni parametri la cosa prende
+tempi biblici. in tali casi i risultati sono tipicamente indecente, quindi mettendo
+un timeout sui tempi di cross validation e' possibile ridurre questo problema.
+
+\ref{fig:pc}
 
@@
     \caption{\label{fig:comparison} Comparison between {\sc SetMargin} and the
-    Bayesian competitors.}
+    Bayesian competitors. The thick lines represent the median utility loss
+    over all users, while the shaded area is the standard deviation. {\sc
+    SetMargin} orange, Guo blue, Viappiani QI green, Viappiani EUS gray.}
 \end{figure}
@@
     \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin} with
-    $k=2,3,4$.}
+    $k=2,3,4$, in orange, blue and green respectively.}
 \end{figure}

### 2016-01-27 paper/ijcai16.tex
paper: modificato abstract e esperimenti

@@
 \begin{abstract}
-In this paper we propose an approach to preference elicitation that is suitable to large
-configuration problem domains where stated preferences can be noisy.
-Our setwise max-margin method can be viewed as a generalization of max-margin learning to sets,
-and can produce a set of ``diverse'' items that can be used to ask informative queries to the user.
-Moreover, the approach can encourage sparsity in the parameter space, in order to favor the assessment of utility towards combinations of weights that concentrate on just few features.
-We present an optimization formulation and show, with simulations, how our approach is able to efficiently elicit the user's utility function even in very large datasets.
-
+  In this paper we propose an approach to preference elicitation that
+  is suitable to large configuration spaces beyond the reach of
+  existing state-of-the-art approaches. Our setwise max-margin method
+  can be viewed as a generalization of max-margin learning to sets,
+  and can produce a set of ``diverse'' items that can be used to ask
+  informative queries to the user.  Moreover, the approach can
+  encourage sparsity in the parameter space, in order to favor the
+  assessment of utility towards combinations of weights that
+  concentrate on just few features.  We present a mixed integer linear
+  programming formulation and show how our approach compares
+  favourably with Bayesian preference elicitation alternatives and
+  easily scales to realistic datasets.
 \end{abstract}
@@
 
-In Figure~\ref{fig:} we plot, for each setting, the behavior of the utility
-loss:
+\andrea{give names to competitors}
+
+In Figure~\ref{fig:comparison} we report solution quality and timing
+values for increasing number of collected user responses, for the
+different competitors on each of the four different utility vector
+distributions. Solution quality is measured in terms of utility loss:
 %
@@
 %
-varying the number of collected user responses. Here $u(\cdot)$ is the
-true unknown user utility, and $\vx^*$ is the solution recommended to
-the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
+where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
+solution recommended to the user after the elicitation phase (see
+Algorithm~\ref{alg:setmargin}). Computational cost is measured in
+terms of cumulative time. Given that Guo and Viappiani are
+single-threaded, we disabled multi-threading when running our
+algorithm in these comparisons. All experiments were run on a XXXXXX.
+In all figures, thick lines indicate median values over the different
+users, shaded areas indicate standard deviations. \andrea{@stefano:
+  possiamo riportare la mediana anche per i tempi?}  Our {\sc
+  SetMargin} algorithm for $k=2$ is reported in orange, Guo in blue,
+Viappiani QI in green and Viappiani EUS in gray. For all algorithms,
+one iteration corresponds to a single pairwise query to the user.
+Results for $r=3$ and $r=4$ are reported.  For dense weight vector
+distributions (first two rows), our approach achieves results which
+are indistinguishable from the competitors in a fraction of their
+time.  Indeed, all Bayesian approaches cannot scale over $r=4$ (for
+$r=5$, Guo did not finish a single user after 24 hours for, Viappiani
+QI takes 20 minutes per user), \andrea{20 minuti mi paiono pochi, sei
+  sicuro?}  while our algorithm can easily scale to much larger
+datasets, as will be shown later on. For sparse weight vector
+distributions (last two rows) our approach, in addition to being
+substantially faster on each iteration, requires less queries in order
+to reach optimal solutions. This is an expected result as the
+sparsification norm in our formulation ($\| \vw^{i} \|_1$) is
+enforcing sparsity in the weight vectors, while none of the other
+approaches are explicitly designed to favour sparse functions.
 
-\begin{itemize}
+% Guo e' piantato al primo utente da tre giorni su synthetic 5.
+% Paolo procede su S-5 da qualche ora (~8) (1290s su synthetic-5
+% uniforme per completamento di un utente).
 
-    \item \ref{fig:comparison} setmargin vs Bayes. similar loss wrt Bayes.
-    variance slighly higher on synthetic 4. funzioniamo meglio su sparso al
-    contrario degli altri. Guo e' piantato al primo utente da tre giorni
-    su synthetic 5.  Paolo procede su S-5 da qualche ora (~8) (1290s su
-    synthetic-5 uniforme per completamento di un utente).
+%\stefano{@Paolo: varianza tempi?}
 
-    \stefano{@Paolo: varianza tempi?}
+In order to study the effect of increasing the number of weight
+vectors in our formulation, we also ran {\sc SetMargin} varying the
+parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
+results on $r=4$ and $r=5$ datasets for the uniform and normal sparse
+distributions\footnote{We picked the most difficult dense distribution
+  (uniform) and the simplest sparse distribution (normal) to provide
+  the broadest possible picture, as space constraints prevent us from
+  including all results. Results for the other distributions are
+  however qualitative similar to the ones reported.}, for $k=2$
+(orange), $k=3$ (blue) and k=4 (green). The first and third columns
+report results in terms of number of iterations. It can be seen that
+increasing the number of weight vectors tends to favour earlier
+convergence, especially for the more complex dataset ($r=5$). However,
+as in each iteration the user is asked to compare $k$ items, different
+values of $k$ imply a different cognitive effort for the user. The
+second and fourth columns report results in terms of number of
+queries, where we count all $k \choose 2$ pairs of queries when
+comparing $k$ items. In this case, $k=2$ seems to be the best
+option. The cognitive cost for the user will likely lay in between
+these two extremes, but formalizing this concept in an efficient query
+ordering strategy needs to face the effect of noise.  A modified
+sorting algorithm asking only $O(k\log k)$ queries to the user
+resulted in a performance worsening, likely because of a cascading
+effect of inconsistence feedback. \andrea{@stefano: l'ho rifrasato un
+  po', vediamo se ci piace}
 
-    \item \ref{fig:selfcomparison} crossval ogni 5 iterazioni, il che vuol
-    dire diverso numero di queri per $k$ diversi.
+% \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
+% methods, which generate a single candidate option per iteration, {\sc
+% SetMargin} constructs a whole {\em set} of options. We thus can not leverage
+% standard query selection strategies \stefano{which ones?}.
 
-\end{itemize}
-
-\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
-methods, which generate a single candidate option per iteration, {\sc
-SetMargin} constructs a whole {\em set} of options. We thus can not leverage
-standard query selection strategies \stefano{which ones?}.
-
-The simplest solution is to query the user about all $k \choose 2$ pairs of
-configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
-number of queries does not grow large in practice.  An appealing alternative
-involves using a modified sorting algorithm, such as merge-sort, to order the
-set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
-approach is very susceptible to noise: a single inconsistent user response may
-affect all successive comparisons, leading to a cascading effect. While more
-robust strategies may be conceived, empirically we found that the simpler
-quadratic scheme tends to work reasonably well.
-
-\andrea{rewrite as footnote}
+% The simplest solution is to query the user about all $k \choose 2$ pairs of
+% configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
+% number of queries does not grow large in practice.  An appealing alternative
+% involves using a modified sorting algorithm, such as merge-sort, to order the
+% set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
+% approach is very susceptible to noise: a single inconsistent user response may
+% affect all successive comparisons, leading to a cascading effect. While more
+% robust strategies may be conceived, empirically we found that the simpler
+% quadratic scheme tends to work reasonably well.
 

### 2016-01-27 paper/ijcai16.tex
paolo:finished pass section 2

--- 2016-01-27 paper/ijcai16.tex
@@
 While preferences are often assessed using some ad-hoc rules \cite{},
-recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas-mmr:uai07,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
+recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
+
+%braziunas-mmr:uai07,
 
@@
 
-Our approach is related to \cite{boutilier2006}, where minimax regret is used both
+Preference elicitation in configuration problems has been previously tackled with regret-based elicitation \cite{boutilier2006,braziunas2007}, where minimax regret is used both
 as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
@@
 candidate functions (endowed with a response or error model to be used for inference) and asking queries maximizing informativeness
-measures such as expected value of information (EVOI)~\cite{chajewska2000,viappiani2010optimal}.
+measures such as expected value of information (EVOI)~\cite{chajewska2000,guo2010real,viappiani2010optimal}.
 
@@
 
-Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
-of linear utility functions: the user's preferences are represented by a weight
-vector $\vw\in\bbR^m$, and the utility of a configuration $\vx$ is given by
-$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. In the remainder of the paper we
+Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
+we model users by means of utility functions of a given structure; the user's preferences are represented by a latent weight
+vector $\vw\in\bbR^m$.
+We assume linear utility functions; the utility of a configuration $\vx$ is given by
+$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$
+(see \cite{keeney1976} for theoretical conditions justifying additive utility functions). 
+In the remainder of the paper we
 require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
 weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
-$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
-natural, and enable the translation of our core optimization problem into a
+$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. 
+Both requirements are quite natural\footnote{Utility values are defined on an
+interval scale, therefore the assumption of non negativity is not restrictive as it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
 mixed-integer linear problem (as done in Section~\ref{sec:formulation}).
@@
 in~\cite{guo2010real} differs slightly from the one implemented in the code of
-their experimental setting. In this paper we use the latter.}:
+their experimental setting. 
+In our experiments we use the latter.}:
 (1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
-a normal distribution with mean $25$ and covariance
-$\frac{25}{3}$.  We further produced two novel
+a normal distribution with mean $25$ and %covariance
+standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).  
+We further produced two novel
 {\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
@@
 \stefano{comment on usefulness of sparse weights}
+
+\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
 

### 2016-01-27 paper/ijcai16.tex

--- 2016-01-27 paper/ijcai16.tex
@@
 %
+%<<<<<<< HEAD
+%varying the number of collected user responses. Here $u(\cdot)$ is the
+%true unknown user utility, and $\vx^*$ is the solution recommended to
+%the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
+%
+%The thick line is the median
+%utility loss taken over all 20 users, while the shaded area represents the
+%standard deviation.
+%
+%\stefano{plot average utility loss at the increase of dataset size}
+%
+%\stefano{plot average runtime at the increase of dataset size}
+%
+%\stefano{comment on usefulness of sparse weights}
+%
+%\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
+%
+%\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
+%methods, which generate a single candidate option per iteration, {\sc
+%SetMargin} constructs a whole {\em set} of options. We thus can not leverage
+%standard query selection strategies \stefano{which ones?}. 
+%
+%The simplest solution is to query the user about all $k \choose 2$ pairs of
+%configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
+%number of queries does not grow large in practice.  An appealing alternative
+%involves using a modified sorting algorithm, such as merge-sort, to order the
+%set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
+%approach is very susceptible to noise: a single inconsistent user response may
+%affect all successive comparisons, leading to a cascading effect. While more
+%robust strategies may be conceived, empirically we found that the simpler
+%quadratic scheme tends to work reasonably well.
+%
+%\andrea{rewrite as footnote}
+%=======
 where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
@@
 % quadratic scheme tends to work reasonably well.
+
+\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}
 

### 2016-01-27 paper/ijcai16.tex
paper: modificato esperimenti

--- 2016-01-27 paper/ijcai16.tex
@@
 Algorithm~\ref{alg:setmargin}). Computational cost is measured in
-terms of cumulative time. Given that Guo and Viappiani are
+terms of cumulative time. Given that Guo and Viappiani (both QI and EUS) are
 single-threaded, we disabled multi-threading when running our
@@
 parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
-results on $r=4$ and $r=5$ datasets for the uniform and normal sparse
+results on $r=4$ and $r=5$ datasets for the uniform and sparse normal
 distributions\footnote{We picked the most difficult dense distribution
@@
 manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
-size, type $\to$ monitor size, for 16 Horn constraints total. We
-do not report the full list due to space limitations.
+size, type $\to$ monitor size, for 16 Horn constraints total. We do
+not report the full list due to space limitations. Note that the
+search space is of the order of hundreds of thousands of candidate
+configurations, and is far beyond reach of existing Bayesian
+approaches. Furthermore, existing Bayesian preference elicitation
+approaches are designed for categorical attributes and cannot deal
+with (even dependent) continuous features like the price. \andrea{@andrea: mention Bonilla}
 
-% 76 features (somma)
-% >>> 3*8*37*8*10*10
-% 710400
-
-la crossval ogni 5 iterazioni richiede tempo considerevole, che nel
-grafico cumulativo del tempo si assomma alle crossval successivie, ampliando
-la varianza; in alcuni casi degeneri per alcuni parametri la cosa prende
-tempi biblici. in tali casi i risultati sono tipicamente indecente, quindi mettendo
-un timeout sui tempi di cross validation e' possibile ridurre questo problema.
-
-\ref{fig:pc}
+Figure~\ref{fig:pc} reports results of {\sc SetMargin} for $k=2$
+(red), $k=3$ (blue) and $k=4$ (green) using the sparse uniform (top
+row) and sparse normal (bottom row) distributions. The first and third
+column report utility loss for increasing number of iterations and
+queries respectively, showing a behaviour which is similar to the one
+in Figure~\ref{fig:selfcomparison}. Overall, between 50 and 70 queries
+on average are needed in order to find out a solution which is only
+10\% worse than the optimal one, out of the more than 700,000
+thousands available. \andrea{something about the lack of smart
+  constraints, as a vendor would have..}  The second and fourth
+columns report cumulative times. Note that in some cases, standard
+deviations have a bump. This is due to cases in which some of the
+hyperparameters of the internal cross validation result in
+ill-conditioned optimization problems which are hard to solve. These
+exceptions can be easily dealt with by setting an appropriate timeout
+on the cross validation without affecting the results, as these
+hyperparameters typically end up having bad performance and being
+discarded.
 
@@
 % \end{table}
-
-\stefano{Add figures and comments.}
 
@@
         \hline
-        \multicolumn{4}{c}{{\sc Uniform Sparse}}
+        \multicolumn{4}{c}{{\sc Sparse Uniform}}
         \\
@@
         \hline
-        \multicolumn{4}{c}{{\sc Normal Sparse}}
+        \multicolumn{4}{c}{{\sc Sparse Normal}}
         \\
@@
     }
-    \caption{\label{fig:comparison} Comparison between {\sc SetMargin} and the
-    Bayesian competitors. The thick lines represent the median utility loss
-    over all users, while the shaded area is the standard deviation. {\sc
-    SetMargin} orange, Guo blue, Viappiani QI green, Viappiani EUS gray.}
+    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
+      and the Bayesian competitors on the $r=3$ (left) and $r=4$
+      (right) datasets. Each row represents a different distribution
+      for sampling the underlying user utility.  The x-axis reports
+      the number of iterations (i.e. pairwise queries), while the
+      y-axis reports the utility loss (first and third columns) and
+      the cumulative time in seconds (second and fourth
+      columns). Thick lines indicate median values over all users,
+      while standard deviations are shown as shaded areas. Curves
+      represent {\sc SetMargin} (orange), Guo (blue), Viappiani QI
+      (green) and Viappiani EUS (gray) respectively. \andrea{@stefano remove median and average from y-axis}}
 \end{figure}
@@
         \hline
-        \multicolumn{4}{c}{{\sc normal sparse}}
+        \multicolumn{4}{c}{{\sc sparse normal}}
         \\
@@
     }
-    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin} with
-    $k=2,3,4$, in orange, blue and green respectively.}
+    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
+      with $k=2,3,4$, in orange, blue and green respectively for the
+      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions for user utility.
+Median (thick lines) and standard deviation (shaded areas) utility loss values are reported for increasing number of iterations (first and third columns) and pairwise queries (second and fourth columns).}
 \end{figure}
@@
         \\
-        \multicolumn{4}{c}{{\sc uniform sparse}}
+        \multicolumn{4}{c}{{\sc sparse uniform}}
         \\

### 2016-01-27 paper/ijcai16.tex
paper: tolti tempi per r=5

@@
 Algorithm~\ref{alg:setmargin}). Computational cost is measured in
-terms of cumulative time. Given that Guo and Viappiani (both QI and EUS) are
-single-threaded, we disabled multi-threading when running our
-algorithm in these comparisons. All experiments were run on a XXXXXX.
+terms of cumulative time. Given that Guo and Viappiani (both QI and
+EUS) are single-threaded, we disabled multi-threading when running our
+algorithm in these comparisons. All experiments were run on a XXXXXX\andrea{@stefano}.
 In all figures, thick lines indicate median values over the different
-users, shaded areas indicate standard deviations. \andrea{@stefano:
-  possiamo riportare la mediana anche per i tempi?}  Our {\sc
-  SetMargin} algorithm for $k=2$ is reported in orange, Guo in blue,
-Viappiani QI in green and Viappiani EUS in gray. For all algorithms,
-one iteration corresponds to a single pairwise query to the user.
-Results for $r=3$ and $r=4$ are reported.  For dense weight vector
-distributions (first two rows), our approach achieves results which
-are indistinguishable from the competitors in a fraction of their
-time.  Indeed, all Bayesian approaches cannot scale over $r=4$ (for
-$r=5$, Guo did not finish a single user after 24 hours for, Viappiani
-QI takes 20 minutes per user), \andrea{20 minuti mi paiono pochi, sei
-  sicuro?}  while our algorithm can easily scale to much larger
-datasets, as will be shown later on. For sparse weight vector
-distributions (last two rows) our approach, in addition to being
-substantially faster on each iteration, requires less queries in order
-to reach optimal solutions. This is an expected result as the
-sparsification norm in our formulation ($\| \vw^{i} \|_1$) is
-enforcing sparsity in the weight vectors, while none of the other
-approaches are explicitly designed to favour sparse functions.
+users, shaded areas indicate standard deviations. Our {\sc SetMargin}
+algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
+in green and Viappiani EUS in gray. For all algorithms, one iteration
+corresponds to a single pairwise query to the user.  Results for $r=3$
+and $r=4$ are reported.  For dense weight vector distributions (first
+two rows), our approach achieves results which are indistinguishable
+from the competitors in a fraction of their time.  Indeed, all
+Bayesian approaches become quickly intractable for growing values of $r$,
+%cannot scale over $r=4$ (for $r=5$, Guo did not
+%finish a single user after 24 hours for, Viappiani QI takes 20 minutes
+%per user), \andrea{20 minuti mi paiono pochi, sei sicuro?}  
+while our algorithm can easily scale to much larger datasets, as will
+be shown later on. For sparse weight vector distributions (last two
+rows) our approach, in addition to being substantially faster on each
+iteration, requires less queries in order to reach optimal
+solutions. This is an expected result as the sparsification norm in
+our formulation ($\| \vw \|_1$) is enforcing sparsity in the weight
+vectors, while none of the other approaches are explicitly designed to
+favour sparse functions.
 
@@
 
-This paper presented a maxmargin approach to generating a set of diverse examples.
-Our approach, that can be seen as an extension of maxmargin learning to sets, is effective in the generation of a diverse set of items that can be used to ask informative queries to the user.
-The main advantages of our elicitation method are 1) its aptitude for eliciting preferences and providing recommendations in large configuration problems  2) robustness with respect to erroneous preference input, and 3) its ability to support sparse utility functions.
-In a number of experiments we showed how our method compare with state-of-the-art approaches.
+We presented a max-margin approach for generating a set of diverse
+examples.  Our approach, that can be seen as an extension of maxmargin
+learning to sets, is effective in the generation of a diverse set of
+items that can be used to ask informative queries to the user.  The
+main advantages of our elicitation method are 1) its aptitude for
+eliciting preferences and providing recommendations in large
+configuration problems 2) robustness with respect to erroneous
+preference input, and 3) its ability to support sparse utility
+functions.  In a number of experiments we showed how our method
+compare with state-of-the-art approaches.
 
@@
       represent {\sc SetMargin} (orange), Guo (blue), Viappiani QI
-      (green) and Viappiani EUS (gray) respectively. \andrea{@stefano remove median and average from y-axis}}
+      (green) and Viappiani EUS (gray) respectively.}
 \end{figure}

### 2016-01-27 paper/ijcai16.tex
modifiche mercoledi sera

@@
 \title{Constructive Preference Elicitation by Setwise Max-margin Learning} %with?
-\author{ID xxxx}
+\author{ID 2243}
 
@@
 the generated configurations are diverse in terms of the weight vectors they
-maximize. Equations~(\ref{eq:wbounds}) and~(\ref{eq:xbounds})
+maximize. Equation~(\ref{eq:wbounds}), where the inequality is intended to be component-wise, and Eq.~(\ref{eq:xbounds})
 ensure that the weights and configurations are in fact feasible, while the
 remaining two guarantee the non-negativity of the slack and margin variables.
-Since we require $\vw^\bot \ge 0$, Eq.~(\ref{eq:wbounds}) also enforces the
-weights to be non-negative.
+Since we require $\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the weights to be non-negative.
 
@@
 %
-$$ \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z $$
+$$ \gamma \sum_{i=1}^k \sum_{z=1}^m p^{i,i}_z$$
 %
@@
     \\
-        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \}
-    \\
-        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
+        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \}
+    \\
+        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
     \\
@@
 repeated. The algorithm terminates after a fixed number of iterations,
-or once another suitable condition is met. \andrea{@Paolo: what termination
-condition for pref elicitation?}  
+or once another suitable condition is met. 
+%\andrea{@Paolo: what termination condition for pref elicitation?}  
 Termination might be left to the user to decide (see for example \ref{} for discussion on user-involved recommender systems); the system can however propose
@@
 
-In all experiments we use an internal 5-fold cross-validation procedure to
+In all experiments {\sc SetMargin} uses an internal 5-fold cross-validation procedure to
 update the hyperparameters $\alpha$, $\beta$, and $\gamma$ after every 5
@@
 We further produced two novel
-{\em sparse} sets of weight vectors by deleting $80\%$ of the entries of the
-vectors sampled in the uniform and normal settings uniformly at random
-\stefano{Paolo, is this correct?}. All methods were evaluated on the very same
-weight vectors.
+{\em sparse} sets of weight vectors by setting to zero $80\%$ of the entries of the
+vectors sampled in the uniform and normal settings uniformly at random. 
+All methods were evaluated on the very same weight vectors.
 
@@
 used in~\cite{guo2010real}: instead of explicitly enumerating all possible PC
-item, we defined the set of feasible configurations through MILP constraints.
+items, we defined the set of feasible configurations with MILP constraints.
 
@@
 columns report cumulative times. Note that in some cases, standard
-deviations have a bump. This is due to cases in which some of the
+deviations have a bump; this is due to cases in which some of the
 hyperparameters of the internal cross validation result in
@@
 configuration problems 2) robustness with respect to erroneous
-preference input, and 3) its ability to support sparse utility
+preference input, and 3) its ability to encourage sparse utility
 functions.  In a number of experiments we showed how our method

### 2016-01-27 paper/ijcai16.tex
paper: riportato fig. 2 e 3 in una pagina

@@
 
+%\onecolumn
+
+\begin{figure*}[h!]
+    \centering
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
+        \\
+        \hline \hline
+        \multicolumn{4}{c}{{\sc Uniform}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Sparse Uniform}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc Sparse Normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
+        \\
+        \hline
+    \end{tabular}
+    }
+    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
+      (orange), Guo (blue), Viappiani QI (green) and Viappiani EUS
+      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
+      represents a different distribution for sampling the underlying
+      user utility.  The number of iterations is plotted against the
+      utility loss (first and third columns) and the cumulative time
+      (second and fourth columns).  Thick lines indicate median values
+      over users, while standard deviations are shown as shaded
+      areas.}
+\end{figure*}
+
+\begin{figure*}[h!]
+    \centering
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        \multicolumn{2}{c}{{\sc synthetic 4}} &
+        \multicolumn{2}{c}{{\sc synthetic 5}}
+        \\
+        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
+        \\
+        \hline \hline
+        \multicolumn{4}{c}{{\sc uniform}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc sparse normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
+        \\
+        \hline
+     \end{tabular}
+    }
+    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
+      with $k=2,3,4$, in orange, blue and green respectively for the
+      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions.
+Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
+\end{figure*}
+
+%\twocolumn
+
 In Figure~\ref{fig:comparison} we report solution quality and timing
@@
 
-\onecolumn
-
-\begin{figure}[b]
-    \centering
-    {\footnotesize
-    \begin{tabular}{cccc}
-        \hline
-        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
-        \\
-        \hline \hline
-        \multicolumn{4}{c}{{\sc Uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Sparse Uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Sparse Normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
-        \\
-        \hline
-    \end{tabular}
-    }
-    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
-      and the Bayesian competitors on the $r=3$ (left) and $r=4$
-      (right) datasets. Each row represents a different distribution
-      for sampling the underlying user utility.  The x-axis reports
-      the number of iterations (i.e. pairwise queries), while the
-      y-axis reports the utility loss (first and third columns) and
-      the cumulative time in seconds (second and fourth
-      columns). Thick lines indicate median values over all users,
-      while standard deviations are shown as shaded areas. Curves
-      represent {\sc SetMargin} (orange), Guo (blue), Viappiani QI
-      (green) and Viappiani EUS (gray) respectively.}
-\end{figure}
-
-\begin{figure}[b]
-    \centering
-    {\footnotesize
-    \begin{tabular}{cccc}
-        \hline
-        \multicolumn{2}{c}{{\sc synthetic 4}} &
-        \multicolumn{2}{c}{{\sc synthetic 5}}
-        \\
-        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
-        \\
-        \hline \hline
-        \multicolumn{4}{c}{{\sc uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc sparse normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
-        \\
-        \hline
-     \end{tabular}
-    }
-    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
-      with $k=2,3,4$, in orange, blue and green respectively for the
-      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions for user utility.
-Median (thick lines) and standard deviation (shaded areas) utility loss values are reported for increasing number of iterations (first and third columns) and pairwise queries (second and fourth columns).}
-\end{figure}
-
 % times for the figure above
@@
 
-\begin{figure}[b]
-    \centering
-    {\footnotesize
-    \begin{tabular}{cccc}
-        \hline
-        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
-        \multicolumn{2}{c}{{\sc loss/time per query}}
-        \\
-        \multicolumn{4}{c}{{\sc sparse uniform}}
-        \\
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc sparse normal}}
-        \\
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
-        \\
-        \hline
-    \end{tabular}
-    }
-    \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
-    dataset.}
-\end{figure}
+% \begin{figure}[b]
+%     \centering
+%     {\footnotesize
+%     \begin{tabular}{cccc}
+%         \hline
+%         \multicolumn{2}{c}{{\sc loss/time per iter.}} &
+%         \multicolumn{2}{c}{{\sc loss/time per query}}
+%         \\
+%         \multicolumn{4}{c}{{\sc sparse uniform}}
+%         \\
+%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+%         \\
+%         \hline
+%         \multicolumn{4}{c}{{\sc sparse normal}}
+%         \\
+%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
+%         \\
+%         \hline
+%     \end{tabular}
+%     }
+%     \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
+%     dataset.}
+% \end{figure}
 

### 2016-01-28 paper/ijcai16.tex
paper: modificato conclusioni

@@
 
-\begin{figure*}[h!]
-    \centering
-    {\footnotesize
-    \begin{tabular}{cccc}
-        \hline
-        \multicolumn{2}{c}{{\sc synthetic 4}} &
-        \multicolumn{2}{c}{{\sc synthetic 5}}
-        \\
-        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
-        \\
-        \hline \hline
-        \multicolumn{4}{c}{{\sc uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc sparse normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
-        \\
-        \hline
-     \end{tabular}
-    }
-    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
-      with $k=2,3,4$, in orange, blue and green respectively for the
-      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions.
-Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
-\end{figure*}
-
 %\twocolumn
@@
 
+\begin{figure*}[h!]
+    \centering
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        \multicolumn{2}{c}{{\sc synthetic 4}} &
+        \multicolumn{2}{c}{{\sc synthetic 5}}
+        \\
+        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
+        \\
+        \hline \hline
+        \multicolumn{4}{c}{{\sc uniform}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc sparse normal}}
+        \\
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
+        \\
+        \hline
+     \end{tabular}
+    }
+    \caption{\label{fig:selfcomparison} Comparison for {\sc SetMargin}
+      with $k=2,3,4$, in orange, blue and green respectively for the
+      $r=4$ (left) and $r=5$ (right) datasets using uniform (top row) and sparse normal (bottom row) distributions.
+Median and standard deviation utility loss values are reported for increasing number of iterations ($1^{st}$ and $3^{rd}$ columns) and pairwise queries ($2^{nd}$ and $4^{th}$ columns).}
+\end{figure*}
+
 In order to study the effect of increasing the number of weight
@@
 configurations, and is far beyond reach of existing Bayesian
-approaches. Furthermore, existing Bayesian preference elicitation
-approaches are designed for categorical attributes and cannot deal
-with (even dependent) continuous features like the price. \andrea{@andrea: mention Bonilla}
+approaches.
 
@@
 
-We presented a max-margin approach for generating a set of diverse
-examples.  Our approach, that can be seen as an extension of maxmargin
-learning to sets, is effective in the generation of a diverse set of
-items that can be used to ask informative queries to the user.  The
-main advantages of our elicitation method are 1) its aptitude for
-eliciting preferences and providing recommendations in large
-configuration problems 2) robustness with respect to erroneous
-preference input, and 3) its ability to encourage sparse utility
-functions.  In a number of experiments we showed how our method
-compare with state-of-the-art approaches.
+We presented a max-margin approach for efficient preference
+elicitation in large configuration spaces. Our approach relies on an
+extension of max-margin learning to sets of hyperplanes, and is
+effective in the generation of a diverse set of items that can be used
+to ask informative queries to the user.  The main advantages of our
+elicitation method are 1) its ability to elicit preferences and
+provide recommendations in large configuration problems 2) its
+robustness with respect to erroneous feedback and 3) its ability to
+encourage sparse utility functions. Experimental comparisons against
+state-of-the-art Bayesian preference elicitation strategies confirm
+its substantial computational advantages. For sparse utility
+functions, the method compares favourably also in terms of the number
+of queries asked to the user.
 
@@
 
-% \begin{figure}[b]
-%     \centering
-%     {\footnotesize
-%     \begin{tabular}{cccc}
-%         \hline
-%         \multicolumn{2}{c}{{\sc loss/time per iter.}} &
-%         \multicolumn{2}{c}{{\sc loss/time per query}}
-%         \\
-%         \multicolumn{4}{c}{{\sc sparse uniform}}
-%         \\
-%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-%         \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
-%         \\
-%         \hline
-%         \multicolumn{4}{c}{{\sc sparse normal}}
-%         \\
-%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
-%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
-%         \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
-%         \\
-%         \hline
-%     \end{tabular}
-%     }
-%     \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
-%     dataset.}
-% \end{figure}
-
-\twocolumn
+\begin{figure*}
+    \centering
+    {\footnotesize
+    \begin{tabular}{cccc}
+        \hline
+        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
+        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
+        \multicolumn{4}{c}{{\sc sparse uniform}}
+        \\
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        \\
+        \hline
+        \multicolumn{4}{c}{{\sc sparse normal}}
+        \\
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
+        \\
+        \hline
+    \end{tabular}
+    }
+    \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
+    dataset.}
+\end{figure*}
 

### 2016-01-28 draw.py
draw: change axes labels.

@@ -129,9 +129,9 @@ def draw_groups(basename, groups, upper_max_x, per_query=False):
     ax.set_xlim([0.0, max_x])
     ax.set_xticks(np.arange(0, max_x, 10))
 
-    ax.set_ylabel("Median utility loss")
+    ax.set_ylabel("Utility loss")
     ax.set_ylim([0.0, max_y + 0.1])
-    ax.set_yticks(np.arange(0, max_y, 10))
+    ax.set_yticks(np.arange(0, max_y, 10 if max_y < 130 else 20))
 
     fig.savefig(basename + "_loss.png", bbox_inches="tight")
 
@@ -143,9 +143,9 @@ def draw_groups(basename, groups, upper_max_x, per_query=False):
     ax.set_xlim([0.0, max_x])
     ax.set_xticks(np.arange(0, max_x, 10))
 
-    ax.set_ylabel("Cumulative average time (in seconds)")
+    ax.set_ylabel("Cumulative time (in seconds)")
     ax.set_ylim([0.0, max_y + 0.1])
-    ax.set_yticks(np.arange(0, max_y, 50))
+    ax.set_yticks(np.arange(0, max_y, 50 if max_y < 130 else 100))
 
     fig.savefig(basename + "_time.png", bbox_inches="tight")
 

### 2016-01-28 paper/ijcai16.tex
paper: update all figures with the latest axis labels.
paper: remove commented block.
paper: add machine desc.
paper: fix minor wart, remove todo.

@@
 
-\stefano{add early stopping in algorithm and description and experiments.}
-
 \begin{algorithm}
@@
 %
-%<<<<<<< HEAD
-%varying the number of collected user responses. Here $u(\cdot)$ is the
-%true unknown user utility, and $\vx^*$ is the solution recommended to
-%the user after the elicitation phase (see Algorithm~\ref{alg:setmargin}).
-%
-%The thick line is the median
-%utility loss taken over all 20 users, while the shaded area represents the
-%standard deviation.
-%
-%\stefano{plot average utility loss at the increase of dataset size}
-%
-%\stefano{plot average runtime at the increase of dataset size}
-%
-%\stefano{comment on usefulness of sparse weights}
-%
-%\paolo{Maybe a plot of an experiment with increased noise in user reponse? (for example, $\beta=0.1$ instead of $1$.}
-%
-%\paragraph{Cognitive effort.} Contrarily to most other preference elicitation
-%methods, which generate a single candidate option per iteration, {\sc
-%SetMargin} constructs a whole {\em set} of options. We thus can not leverage
-%standard query selection strategies \stefano{which ones?}. 
-%
-%The simplest solution is to query the user about all $k \choose 2$ pairs of
-%configurations in $\{\vx^i\}$. Since $k$ is a small constant, the resulting
-%number of queries does not grow large in practice.  An appealing alternative
-%involves using a modified sorting algorithm, such as merge-sort, to order the
-%set $\{\vx^i\}$ by asking only $O(n\log n)$ queries to the user. However, this
-%approach is very susceptible to noise: a single inconsistent user response may
-%affect all successive comparisons, leading to a cascading effect. While more
-%robust strategies may be conceived, empirically we found that the simpler
-%quadratic scheme tends to work reasonably well.
-%
-%\andrea{rewrite as footnote}
-%=======
 where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
@@
 EUS) are single-threaded, we disabled multi-threading when running our
-algorithm in these comparisons. All experiments were run on a XXXXXX\andrea{@stefano}.
+algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
+Xeon CPU with 8 cores and 32 GiB of RAM.
 In all figures, thick lines indicate median values over the different
@@
   however qualitative similar to the ones reported.}, for $k=2$
-(orange), $k=3$ (blue) and k=4 (green). The first and third columns
+(orange), $k=3$ (blue) and $k=4$ (green). The first and third columns
 report results in terms of number of iterations. It can be seen that
@@
 resulted in a performance worsening, likely because of a cascading
-effect of inconsistence feedback. \andrea{@stefano: l'ho rifrasato un
-  po', vediamo se ci piace}
+effect of inconsistence feedback.
 

### 2016-01-28 paper/ijcai16.tex
paper: fix todo.

@@
 attributes: this is a fair modeling choice, as often the price of a PC is
-well approximated by sum of the price of its components plus an offset due to
+well approximated by the sum of the price of its components plus a bias due to
 branding.
@@
 10\% worse than the optimal one, out of the more than 700,000
-thousands available. \andrea{something about the lack of smart
-  constraints, as a vendor would have..}  The second and fourth
+thousands available. Note that a vendor may ensure a considerably smaller
+number of queries by cleverly constraining the feasible configuration space;
+since our primary aim is benchmarking, we chose not to pursue this direction
+further.
+The second and fourth
 columns report cumulative times. Note that in some cases, standard

### 2016-01-28 paper/ijcai16.tex
paper: remove spurious blank line.
paper: describe early stopping for setmargin.
paper: shrink intuition.
paper: corretto errore su formula indifferenza
piccoli cambiamenti

@@
 In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
-While preferences are often assessed using some ad-hoc rules \cite{},
-recently a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
+%While preferences are often assessed using some ad-hoc rules \cite{},
+Recently, in the last decade or so, a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
 
@@
 configuration $\vx^i$ over the alternative $\vx^j$ amounts to:
-
 %
@@
 vectors sampled in the uniform and normal settings uniformly at random. 
-All methods were evaluated on the very same weight vectors.
+All methods were evaluated on the very same weight vectors. All users had
+a budget of 100 iterations; for {\sc SetMargin} we used a $0.01$
+threshold on the utility loss for early stopping (that does not affect the
+cross-validation procedure). \stefano{I have no idea if the other methods
+have any kind of early stopping logic.}
 

### 2016-01-28 paper/ijcai16.tex
descrizione dei metodi bayesiani

@@
 from non monotonic reasoning to computational social choice and algorithmic decision theory.
-The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be usually high (from a cognitive and/or computational point of view).
-The goal in preference elicitation is that of gaining enough information about the user's preferences (typically represented by an utility function $u$, unknown to the system) in order to recommend a decision or a course of action.
+The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be high (from a cognitive and/or computational point of view).
+The goal in preference elicitation is that of gaining enough information about the user's preferences %(typically represented by an utility function $u$, unknown to the system) 
+in order to recommend a decision or a course of action.
 
-The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{}.
-In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective; within the machine learning field the term preference learning has been catching up.
+The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{White1984,JacquetLagreze95}.
+%In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
 %While preferences are often assessed using some ad-hoc rules \cite{},
-Recently, in the last decade or so, a number of researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
+Recently, in the last decade or so, a number of artificial intelligence researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
 
@@
 as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
-The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that item x is preferred to y, even though the latter has higher utility).
+The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that an item is preferred to another one, even though the latter has higher utility).
 
@@
     \begin{center}
-        \includegraphics[width=16em]{figures/setmargin}
+        \includegraphics[width=10em]{figures/setmargin}
     \end{center}
@@
 %\andrea{@Paolo: what termination condition for pref elicitation?}  
-Termination might be left to the user to decide (see for example \ref{} for discussion on user-involved recommender systems); the system can however propose
-to terminate when the difference between the utility weights is small enough.
+Termination might be left to the user to decide (see for example \cite{Reilly2007}); 
+alternatively we can terminate when the difference between the utility weights is small enough.
 
@@
 
-We compare {\sc SetMargin} against two strong Bayesian baselines:
-\cite{guo2010real} and \cite{viappiani2010optimal}.
-\stefano{Paolo, could you describe the baselines?}
+We compare {\sc SetMargin} against the following competitive Bayesian baselines:
+% and \cite{viappiani2010optimal}.
+
+\paolo{Se bisogno di spazio, posso accorciare la descrizione.}
+\begin{itemize}
+ \item The Bayesian approach from \cite{guo2010real}, selecting queries according to 
+ {\em restricted uninformed VOI}, an heuristic approximation of value-of-information a query, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
+ \item The Bayesian framework of \cite{viappiani2010optimal}, using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called EUS), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+ \item {\em Query Iteration} (referred as QI below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector, forming a first proposal for a query set.
+Then, a new set is computed by considering the optimal items in the associated  beliefs, and the process is repeated until convergence.
+\end{itemize}
 
@@
 %
-$$ \exp(\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
+$$ \exp(-\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
 %
@@
 threshold on the utility loss for early stopping (that does not affect the
-cross-validation procedure). \stefano{I have no idea if the other methods
-have any kind of early stopping logic.}
+cross-validation procedure).
+The Bayesian methods can use the estimated loss to decide when to stop (the expectation
+of the utility loss with respect to the current probabilistic belief)
+\footnote{In our experiments, we left the Bayesian methods to run the full amount of iterations,
+even when the estimated loss was very low.}.
 
@@
 solutions. This is an expected result as the sparsification norm in
-our formulation ($\| \vw \|_1$) is enforcing sparsity in the weight
-vectors, while none of the other approaches are explicitly designed to
-favour sparse functions.
+our formulation ($\| \vw \|_1$) is enforcing sparsity in the weights, 
+while none of the other approaches is %explicitly 
+designed to favour sparse functions.
 
@@
 
-\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}
+%\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}
 

### 2016-01-28 paper/ijcai16.tex
altre piccole modifiche parte esperimenti

@@
 many scenarios the user will have strong preferences about some attributes, but
-will be indifferent to most of them~\cite{} \stefano{@Paolo: add refs}. The $\ell_1$ penalty is frequently
+will be indifferent to most of them. %~\cite{} \stefano{@Paolo: add refs}. 
+The $\ell_1$ penalty is frequently
 used to improve the sparsity of learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent gains
@@
 \begin{itemize}
- \item The Bayesian approach from \cite{guo2010real}, selecting queries according to 
- {\em restricted uninformed VOI}, an heuristic approximation of value-of-information a query, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
+ \item The Bayesian approach from \cite{guo2010real}, selecting queries according to {\em restricted uninformed VOI} (ruVOI), an heuristic approximation of value-of-information a query, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
+
  \item The Bayesian framework of \cite{viappiani2010optimal}, using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called EUS), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+
  \item {\em Query Iteration} (referred as QI below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector, forming a first proposal for a query set.
@@
 of the user being indifferent about the two configurations depends on
-how close their utilities are, as in:
+how close their utilities are, as in the exponential distribution:
 %
-$$ \exp(-\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|) $$
+$$ \exp(-\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|). $$
 %
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
-simulations, as in~\cite{guo2010real}.
+simulations\footnote{When experiment with EUS and QI, the user is forced to select
+its preffered item, so indifference is not possible. While support for indifference
+could be added, we chose to .}, as in~\cite{guo2010real}.
 
@@
 grows much larger than the one typically used in the Bayesian
-preference elicitation literature, and as such represents a good test
-bed for comparing the scalability of the various methods. The feasible
+preference elicitation literature, and as such represents a good testbed
+for comparing the scalability of the various methods. The feasible
 configuration space was encoded in {\sc SetMargin} through appropriate
@@
 users, shaded areas indicate standard deviations. Our {\sc SetMargin}
-algorithm for $k=2$ is reported in orange, Guo in blue, Viappiani QI
-in green and Viappiani EUS in gray. For all algorithms, one iteration
+algorithm for $k=2$ is reported in orange, Guo's ruVOI in blue, Viappiani's QI
+in green and EUS in gray. For all algorithms, one iteration
 corresponds to a single pairwise query to the user.  Results for $r=3$

### 2016-01-28 paper/ijcai16.tex
update modifica intuizione setwise margin

@@
 Note that we are choosing the configurations $\{ \vx^i \}$ and the weight vectors $\{
-\vw^i \}$ {\em simultaneously}. Since we maximize the margin $\mu$, the optimizer
-will prefer a set of configurations $\{ \vx^{i} \}$ that partitions the weight space
-roughly equally, and corresponding utility functions such that each $\vw^{i}$
-lies (intuitively) close to the centre of each subregion. See
-Figure~\ref{fig:setmargin} for a depiction of this intuition.\andrea{@Paolo: mmh non chiarissimo, da rivedere insieme, legenda inclusa}
+\vw^i \}$ {\em simultaneously}.
+Look at Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a pair ($k=2$). 
+The constraint \ref{eq:wxconstr} is represented by a red line, that partitions the space of
+feasible utility weights in two parts (in general, there will be $k$ subregions).
+Since we maximize the margin $\mu$, the optimizer
+will prefer a set of configurations $\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space
+in an even way.
+In each subregion, we have corresponding utility weight vector $\vw^{i}$
+lying ``close'' to the centre of each subregion. 
+Note that if we present $\vx^{1}$ and $\vx^{2}$ to the user, and the user states that he prefers
+the former to the latter, the feasible region for the weights will then become the part of the polytope to the left of the red line; moreover $\vw^{1}$ will be vector maximizing the margin in the classic sense.
+Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max margin learner a posteriori.
+%Therefore, setwise maxmargin can be thought as the set that, given any possible preference about the items, ensure that the reduction of utility uncertainty is the greatest
+\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}
 
@@
     \end{center}
-    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utility vectors $\vw^{i}$ $\vw^{j}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
+    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utility vectors $\vw^{1}$ $\vw^{2}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
 \end{figure}

### 2016-01-28 paper/ijcai16.tex
typo

@@
 \begin{itemize}
- \item The Bayesian approach from \cite{guo2010real}, selecting queries according to {\em restricted uninformed VOI} (ruVOI), an heuristic approximation of value-of-information a query, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
+ \item The Bayesian approach from \cite{guo2010real}, selecting queries according to {\em restricted uninformed VOI} (ruVOI), an heuristic approximation of value-of-information, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
 
- \item The Bayesian framework of \cite{viappiani2010optimal}, using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called EUS), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+ \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called EUS), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
 

### 2016-01-28 paper/ijcai16.tex
paper: aggiornato nomi

@@
 \begin{itemize}
- \item The Bayesian approach from \cite{guo2010real}, selecting queries according to {\em restricted uninformed VOI} (ruVOI), an heuristic approximation of value-of-information, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
+ \item The Bayesian approach from \cite{guo2010real}, selecting queries according to {\em restricted uninformed VOI} ({\sc ruVOI}), an heuristic approximation of value-of-information, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
 
- \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called EUS), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+ \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
 
- \item {\em Query Iteration} (referred as QI below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector, forming a first proposal for a query set.
+ \item {\em Query Iteration} (referred as {\sc QI} below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector, forming a first proposal for a query set.
 Then, a new set is computed by considering the optimal items in the associated  beliefs, and the process is repeated until convergence.
@@
 
-\andrea{give names to competitors}
-
-%\onecolumn
-
 \begin{figure*}[h!]
@@
     \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
-      (orange), Guo (blue), Viappiani QI (green) and Viappiani EUS
+      (orange), {\sc ruVOI} (blue), {\sc QI} (green) and {\sc EUS}
       (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
@@
 Algorithm~\ref{alg:setmargin}). Computational cost is measured in
-terms of cumulative time. Given that Guo and Viappiani (both QI and
-EUS) are single-threaded, we disabled multi-threading when running our
+terms of cumulative time. Given that {\sc ruVOI},  {\sc QI} and
+{\sc EUS} are single-threaded, we disabled multi-threading when running our
 algorithm in these comparisons. All experiments were run on a 2.8 GHz Intel
@@
 users, shaded areas indicate standard deviations. Our {\sc SetMargin}
-algorithm for $k=2$ is reported in orange, Guo's ruVOI in blue, Viappiani's QI
-in green and EUS in gray. For all algorithms, one iteration
+algorithm for $k=2$ is reported in orange, {\sc ruVOI} in blue, {\sc QI}
+in green and {\sc EUS} in gray. For all algorithms, one iteration
 corresponds to a single pairwise query to the user.  Results for $r=3$
@@
 Bayesian approaches become quickly intractable for growing values of $r$,
-%cannot scale over $r=4$ (for $r=5$, Guo did not
+%cannot scale over $r=4$ (for $r=5$, {\sc ruVOI} did not
 %finish a single user after 24 hours for, Viappiani QI takes 20 minutes
@@
 
-% Guo e' piantato al primo utente da tre giorni su synthetic 5.
+% {\sc ruVOI} e' piantato al primo utente da tre giorni su synthetic 5.
 % Paolo procede su S-5 da qualche ora (~8) (1290s su synthetic-5

### 2016-01-28 paper/ijcai16.tex
paper: updated Guo name

@@
 \begin{itemize}
- \item The Bayesian approach from \cite{guo2010real}, selecting queries according to {\em restricted uninformed VOI} ({\sc ruVOI}), an heuristic approximation of value-of-information, and performing inference using  TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference scheme based on expectation propagation \cite{Minka01}).
+\item The best Bayesian approach from \cite{guo2010real}, selecting
+  queries according to {\em restricted informed VOI} ({\sc riVOI}), a
+  heuristic approximation of value-of-information, and performing
+  inference using TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference
+  scheme based on expectation propagation \cite{Minka01}).
 
@@
     \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
-      (orange), {\sc ruVOI} (blue), {\sc QI} (green) and {\sc EUS}
+      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
       (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
@@
 Algorithm~\ref{alg:setmargin}). Computational cost is measured in
-terms of cumulative time. Given that {\sc ruVOI},  {\sc QI} and
+terms of cumulative time. Given that {\sc riVOI},  {\sc QI} and
 {\sc EUS} are single-threaded, we disabled multi-threading when running our
@@
 users, shaded areas indicate standard deviations. Our {\sc SetMargin}
-algorithm for $k=2$ is reported in orange, {\sc ruVOI} in blue, {\sc QI}
+algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
 in green and {\sc EUS} in gray. For all algorithms, one iteration
@@
 Bayesian approaches become quickly intractable for growing values of $r$,
-%cannot scale over $r=4$ (for $r=5$, {\sc ruVOI} did not
+%cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
 %finish a single user after 24 hours for, Viappiani QI takes 20 minutes
@@
 
-% {\sc ruVOI} e' piantato al primo utente da tre giorni su synthetic 5.
+% {\sc riVOI} e' piantato al primo utente da tre giorni su synthetic 5.
 % Paolo procede su S-5 da qualche ora (~8) (1290s su synthetic-5

### 2016-01-28 paper/ijcai16.tex
typo

@@
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
-simulations\footnote{When experiment with EUS and QI, the user is forced to select
+simulations\footnote{When experimenting with EUS and QI, the user is forced to select
 its preffered item, so indifference is not possible. While support for indifference
-could be added, we chose to .}, as in~\cite{guo2010real}.
+could be added, we chose to be consistent with the methods as originally published.}, as in~\cite{guo2010real}.
 

### 2016-01-28 paper/ijcai16.tex
added two footnotes about bayesian methods

@@
 
- \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+ \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods
+ \footnote{We use 50000 particles to represent the probability distribution.}for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
 
@@
 from the competitors in a fraction of their time.  Indeed, all
-Bayesian approaches become quickly intractable for growing values of $r$,
+Bayesian approaches\footnote{Note that for {\sc QI} and {\sc EUS}, the time
+spent for Bayesian inference greatly dominates the time spent for choosing the next
+query to ask.} become quickly intractable for growing values of $r$,
 %cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not

### 2016-01-28 paper/ijcai16.tex
paper: minor modifications

--- 2016-01-28 paper/ijcai16.tex
@@
 
-Preferences \cite{Peintner2008} play an important role in artificial intelligence,
-including fields spanning from recommender systems to automatic planning, 
-from non monotonic reasoning to computational social choice and algorithmic decision theory.
-The task of eliciting or learning preferences is therefore a crucial; typically only limited information about the user's preferences will be available and the cost of obtaining additional preference information will be high (from a cognitive and/or computational point of view).
-The goal in preference elicitation is that of gaining enough information about the user's preferences %(typically represented by an utility function $u$, unknown to the system) 
+Preferences \cite{Peintner2008} play an important role in artificial
+intelligence, including fields spanning from recommender systems to
+automatic planning, from non monotonic reasoning to computational
+social choice and algorithmic decision theory.  The task of eliciting
+or learning preferences is therefore a crucial one; typically only limited
+information about the user's preferences will be available and the
+cost of obtaining additional preference information will be high (from
+a cognitive and/or computational point of view).  The goal in
+preference elicitation is that of gaining enough information about the
+user's
+preferences %(typically represented by an utility function $u$, unknown to the system)
 in order to recommend a decision or a course of action.
@@
 %While preferences are often assessed using some ad-hoc rules \cite{},
-Recently, in the last decade or so, a number of artificial intelligence researchers have proposed principled methods that elicit utility in an adaptive way \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}; the aims is that of focusing on learning the ``important'' part of the utility, allowing to recommend good (or even optimal) decision with only a partial information about the utility function.
+Recently, in the last decade or so, a number of artificial
+intelligence researchers have proposed principled methods that elicit
+utility in an adaptive way
+\cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}. The
+aim is that of focusing on learning the ``important'' part of the
+utility, allowing to recommend good (or even optimal) decisions with
+only partial knowledge of the utility function.
 
@@
 
-Preference elicitation in configuration problems has been previously tackled with regret-based elicitation \cite{boutilier2006,braziunas2007}, where minimax regret is used both
-as a robust recommendation criterion and as a technique to drive elicitation in configuration problems.
-The main limitation of the minimax regret approach is the lack of tolerance with respect to user inconsistency (the user might state that an item is preferred to another one, even though the latter has higher utility).
-
+Preference elicitation in configuration problems has been previously
+tackled with regret-based elicitation
+\cite{boutilier2006,braziunas2007}, where minimax regret is used both
+as a robust recommendation criterion and as a technique to drive
+elicitation in configuration problems.  The main limitation of the
+minimax regret approach is the lack of tolerance with respect to user
+inconsistency (the user might state that an item is preferred to
+another one, even though the latter has higher utility).
 % setwise max-margin formulation to deal with uncertainty in user utility
+
 Indeed, learning a user utility function requires setting a preference
 elicitation strategy and dealing with uncertain and possibly
-inconsistent user feedback. 
-Bayesian preference elicitation approaches
-deal with both problems by building a probability distribution on
-candidate functions (endowed with a response or error model to be used for inference) and asking queries maximizing informativeness
-measures such as expected value of information (EVOI)~\cite{chajewska2000,guo2010real,viappiani2010optimal}.
-
-These
+inconsistent user feedback.  Bayesian preference elicitation
+approaches deal with both problems by building a probability
+distribution on candidate functions (endowed with a response or error
+model to be used for inference) and asking queries maximizing
+informativeness measures such as expected value of information
+(EVOI)~\cite{chajewska2000,guo2010real,viappiani2010optimal}.  These
 approaches are however computationally expensive and can not scale to
 fully constructive scenarios, as shown in our experimental results.
+
 We take a space decomposition perspective and jointly learn a set of
@@
 
-We evaluate the effectiveness of our approach by testing our elicitation method
-in both synthetic and configuration problems, and comparing it to state-of-the-art
-methods.
-
-The paper is structured as follows: after discussing related works in next Section,
-we present some background material in Section \ref{sec:background} 
-and then proceed to our setwise maxmargin method in Section \ref{sec:formulation}; 
-we discuss experimental results in Section \ref{sec:experiments}
-and conclude with final remarks (Section \ref{sec:conclusions}).
+We evaluate the effectiveness of our approach by testing our
+elicitation method in both synthetic and real-world problems, and
+comparing it to state-of-the-art methods.
+
+The paper is structured as follows: we present some background
+material in Section \ref{sec:background} and then proceed to our
+setwise max-margin method in Section \ref{sec:formulation}; we discuss
+experimental results in Section \ref{sec:experiments} and conclude
+with final remarks (Sec.~\ref{sec:conclusions}).
 
@@
 mixed-integer linear problem (as done in Section~\ref{sec:formulation}). 
-
-During learning, the actual weight vector $\vw$ is {\em unknown} to the
-learning system, and must be estimated from pairwise choices elicited from the
-user.
-%\andrea{@PAOLO: qualche parola e reference per giustificare la scelta di pairwise queries?}  
-Comparison queries are especially natural, asking a user if she prefers one option x to another y;
-these comparisons can be extended to choice sets of more than two options \cite{viappiani2009,viappiani2010optimal}.
-Comparison and choice queries are common in conjoint analysis \cite{louviere2000,toubia2004}.
-
-We consider three possible outcomes for a pairwise comparison between
-two configurations $\vx$ and $\vx'$: either $\vx$ is preferred to $\vx'$ (written
-$\vx \succ \vx'$), $\vx'$ is preferred to $\vx$ ($\vx \prec \vx'$), or there is
-no clear preference between the two items ($\vx \approx \vx'$). We write
-$\calD$ to denote the set of preferences (answers to comparison queries) 
-elicited from the user. Queries are typically selected so as to maximize 
-some measure of informativeness of the
-resulting responses.  Details on our actual query selection strategy will be
+%\andrea{@paolo: sulla giustificazione delle utilita' additive, vogliamo metterla? c'e' anche molto lavoro che contesta il fatto che basti no?}
+
+During learning, the actual weight vector $\vw$ is {\em unknown} to
+the learning system, and must be estimated by interacting with the
+user. We mostly focus on pairwise comparison queries, which are the
+simplest of the comparative queries. These can be extended to choice
+sets of more than two options
+\cite{viappiani2009,viappiani2010optimal}.  Comparison and choice
+queries are common in conjoint analysis
+\cite{louviere2000,toubia2004}. We consider three possible outcomes
+for a pairwise comparison between two configurations $\vx$ and $\vx'$:
+either $\vx$ is preferred to $\vx'$ (written $\vx \succ \vx'$), $\vx'$
+is preferred to $\vx$ ($\vx \prec \vx'$), or there is no clear
+preference between the two items ($\vx \approx \vx'$). We write
+$\calD$ to denote the set of preferences (answers to comparison
+queries) elicited from the user. Queries are typically selected so as
+to maximize some measure of informativeness of the resulting
+responses.  Details on our actual query selection strategy will be
 provided in Section~\ref{sec:formulation}.
@@
 vectors $\vw^{1}, \ldots, \vw^{k}$, chosen so that all user-provided
-preferences are satisfied by the largest possible margin (i.e. all
-weight vectors are consistent with respect to the user responses
-$\calD$, modulo inconsistencies) and so that they are maximally
+preferences are satisfied by the largest possible margin (%
+%i.e. all
+%weight vectors are consistent with respect to the user responses
+%$\calD$, 
+modulo inconsistencies) and so that they are maximally
 diverse.  Second, we want to construct a {\em set} of $k$
@@
 introducing slack variables (whose sum we aim at minimizing). When
-augmented with the slacks, the above inequalities take the form:
-\begin{align*}
-\langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h 
-\end{align*}
+augmented with the slacks, the above inequalities take the form
+$\langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h$
 where $\varepsilon^{i}_h$ is the penalty incurred by weight vector $\vw^{i}$
@@
 %
-Let us illustrate the above piece by piece. The objective is composed of four
-parts: we maximize the shared margin $\mu$ (first part) and minimize the total
-sum of the ranking errors $\veps^i$ incurred by each weight vector $\vw^{i}$
-(second part), while at the same time regularizing the magnitude of the
-weights (third part) and the quality of the configurations $\{ \vx^{i} \}$ (last
-part). The non-negative hyperparameters $\alpha,\beta,\gamma$ control the
-influence of the various components.
-
-The third part of the objective does require a more in depth explanation. In
-many scenarios the user will have strong preferences about some attributes, but
-will be indifferent to most of them. %~\cite{} \stefano{@Paolo: add refs}. 
-The $\ell_1$ penalty is frequently
-used to improve the sparsity of learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent gains
-in generalization ability and efficiency, as confirmed by our empirical
-findings (see Section~\ref{sec:experiments}).
+Let us illustrate the above piece by piece. The objective is composed
+of four parts: we maximize the shared margin $\mu$ (first part) and
+minimize the total sum of the ranking errors $\veps^i$ incurred by
+each weight vector $\vw^{i}$ (second part), while at the same time
+regularizing the magnitude of the weights (third part) and the quality
+of the configurations $\{ \vx^{i} \}$ (last part). The non-negative
+hyperparameters $\alpha,\beta,\gamma$ control the influence of the
+various components. The weight regularization term copes with the
+common scenario in which the user has strong preferences about
+some attributes, but is indifferent to most of
+them. %~\cite{} \stefano{@Paolo: add refs}.
+The $\ell_1$ penalty is frequently used to improve the sparsity of
+learned models~\cite{lasso,zhang2008,Hensinger2010}, with consequent
+gains in generalization ability and efficiency, as confirmed by our
+empirical findings (see Section~\ref{sec:experiments}).
 
@@
 the former to the latter, the feasible region for the weights will then become the part of the polytope to the left of the red line; moreover $\vw^{1}$ will be vector maximizing the margin in the classic sense.
-Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max margin learner a posteriori.
+Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner a posteriori.
 %Therefore, setwise maxmargin can be thought as the set that, given any possible preference about the items, ensure that the reduction of utility uncertainty is the greatest
-\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}
+%\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}
+%\andrea{cosa intendi per a posteriori?}
 
@@
 
-\paolo{Se bisogno di spazio, posso accorciare la descrizione.}
 \begin{itemize}
@@
   queries according to {\em restricted informed VOI} ({\sc riVOI}), a
-  heuristic approximation of value-of-information, and performing
-  inference using TrueSkill$^{TM}$ \cite{HerbrichMG06} (an inference
-  scheme based on expectation propagation \cite{Minka01}).
-
- \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
-
- \item {\em Query Iteration} (referred as {\sc QI} below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector, forming a first proposal for a query set.
+  heuristic approximation (equally effective but more efficient) of
+  value-of-information, and inference using TrueSkill$^{TM}$
+  \cite{HerbrichMG06} (an inference scheme based on expectation
+  propagation \cite{Minka01}).
+
+ \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of displayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+
+ \item {\em Query Iteration} (referred as {\sc QI} below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector and forming a first proposal for a query set.
 Then, a new set is computed by considering the optimal items in the associated  beliefs, and the process is repeated until convergence.
@@
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
-simulations\footnote{When experiment with EUS and QI, the user is forced to select
-its preffered item, so indifference is not possible. While support for indifference
-could be added, we chose to .}, as in~\cite{guo2010real}.
+simulations, as in~\cite{guo2010real}.
+
+% \footnote{When experiment with EUS and QI, the user is forced to select
+% its preffered item, so indifference is not possible. While support for indifference
+% could be added, we chose to .}
+%\andrea{@paolo: e' necessaria la footnote? dobbiamo tagliare..}
 
@@
 % exponentially with $m$. 
-For $r=3$, the dataset is comparable to the synthetic one used
-in~\cite{guo2010real} and~\cite{viappiani2010optimal}, which have
-three attributes with 2, 2 and 5 values respectively, for a total of
-20 feasible configurations.  For larger $r$ the size of the space
-grows much larger than the one typically used in the Bayesian
-preference elicitation literature, and as such represents a good testbed
-for comparing the scalability of the various methods. The feasible
-configuration space was encoded in {\sc SetMargin} through appropriate
-MILP constraints, while the other methods require all datasets to be
-explicitly grounded.
-
-Users were simulated by drawing $20$ random utility vectors from each of four
-different distributions. The first two mimic those used
+For $r=3$, the dataset is comparable in size to the synthetic one used
+in~\cite{guo2010real} and~\cite{viappiani2010optimal}.
+%, which have
+%three attributes with 2, 2 and 5 values respectively, for a total of
+%20 feasible configurations.  
+For larger $r$ the size of the space grows much larger than the one
+typically used in the Bayesian preference elicitation literature, and
+as such represents a good testbed for comparing the scalability of the
+various methods. The feasible configuration space was encoded in {\sc
+  SetMargin} through appropriate MILP constraints, while the other
+methods require all datasets to be explicitly grounded.
+
+Users were simulated by drawing $20$ random utility vectors from each
+of four different distributions. The first two mimic those used
 in~\cite{guo2010real}\footnote{The normal distribution reported
-in~\cite{guo2010real} differs slightly from the one implemented in the code of
-their experimental setting. 
-In our experiments we use the latter.}:
-(1) a uniform distribution over $[1, 100]$ for each individual weight, and (2)
-a normal distribution with mean $25$ and %covariance
+  in~\cite{guo2010real} differs slightly from the one implemented in
+  the code of their experimental setting.  In our experiments we use
+  the latter.}: (1) a uniform distribution over $[1, 100]$ for each
+individual weight, and (2) a normal distribution with mean $25$
+and %covariance
 standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).
-We further produced two novel
-{\em sparse} sets of weight vectors by setting to zero $80\%$ of the entries of the
-vectors sampled in the uniform and normal settings uniformly at random. 
-All methods were evaluated on the very same weight vectors. All users had
-a budget of 100 iterations; for {\sc SetMargin} we used a $0.01$
-threshold on the utility loss for early stopping (that does not affect the
-cross-validation procedure).
-The Bayesian methods can use the estimated loss to decide when to stop (the expectation
-of the utility loss with respect to the current probabilistic belief)
-\footnote{In our experiments, we left the Bayesian methods to run the full amount of iterations,
-even when the estimated loss was very low.}.
+We further produced two novel {\em sparse} versions of the uniform and
+normal distributions setting to zero $80\%$ of the entries (sampled
+uniformly at random). All methods were evaluated on the same weight
+vectors. All users had a budget of 100 iterations; for {\sc SetMargin}
+we used a $0.01$ threshold on the utility loss for early stopping
+(that does not affect the cross-validation procedure).  The Bayesian
+methods can use the estimated loss to decide when to stop (the
+expectation of the utility loss with respect to the current
+probabilistic belief) \footnote{In our experiments, we left the
+  Bayesian methods to run the full amount of iterations, even when the
+  estimated loss was very low.}. \andrea{@stefano: cosi' non va. O tutti early stopping o nessuno..}
 

### 2016-01-28 paper/ijcai16.tex

@@
 
- \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of displayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
+ \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods
+ \footnote{We use 50000 particles to represent the probability distribution.}for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
 
@@
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
-simulations, as in~\cite{guo2010real}.
-
-% \footnote{When experiment with EUS and QI, the user is forced to select
-% its preffered item, so indifference is not possible. While support for indifference
-% could be added, we chose to .}
+simulations\footnote{When experimenting with EUS and QI, the user is
+  forced to select its preffered item, so indifference is not
+  possible. While support for indifference could be added, we chose to
+  be consistent with the methods as originally published.}, as
+in~\cite{guo2010real}.
 %\andrea{@paolo: e' necessaria la footnote? dobbiamo tagliare..}

### 2016-01-28 paper/ijcai16.tex
paper: removed early stopping, fixed Guo

@@
 mixed-integer linear problem (as done in Section~\ref{sec:formulation}). 
-%\andrea{@paolo: sulla giustificazione delle utilita' additive, vogliamo metterla? c'e' anche molto lavoro che contesta il fatto che basti no?}
+\andrea{@paolo: sulla giustificazione delle utilita' additive, vogliamo metterla? c'e' anche molto lavoro che contesta il fatto che basti no?}
 
@@
 %\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}
-%\andrea{cosa intendi per a posteriori?}
+\andrea{cosa intendi per a posteriori?}
 
@@
 repeated. The algorithm terminates after a fixed number of iterations,
-or once another suitable condition is met. 
-%\andrea{@Paolo: what termination condition for pref elicitation?}  
-Termination might be left to the user to decide (see for example \cite{Reilly2007}); 
-alternatively we can terminate when the difference between the utility weights is small enough.
+or once another suitable condition is met.  Termination might be left
+to the user to decide (see for example \cite{Reilly2007});
+alternatively we can terminate when the difference between the utility
+weights is small enough.
 
@@
 \begin{itemize}
-\item The best Bayesian approach from \cite{guo2010real}, selecting
-  queries according to {\em restricted informed VOI} ({\sc riVOI}), a
-  heuristic approximation (equally effective but more efficient) of
+\item The Bayesian approach from \cite{guo2010real}, selecting queries
+  according to {\em restricted informed VOI} ({\sc riVOI}), a
+  computationally efficient heuristic approximation of
   value-of-information, and inference using TrueSkill$^{TM}$
@@
 in~\cite{guo2010real}.
-%\andrea{@paolo: e' necessaria la footnote? dobbiamo tagliare..}
+\andrea{@paolo: e' necessaria la footnote? dobbiamo tagliare..}
 
@@
 normal distributions setting to zero $80\%$ of the entries (sampled
-uniformly at random). All methods were evaluated on the same weight
-vectors. All users had a budget of 100 iterations; for {\sc SetMargin}
-we used a $0.01$ threshold on the utility loss for early stopping
-(that does not affect the cross-validation procedure).  The Bayesian
-methods can use the estimated loss to decide when to stop (the
-expectation of the utility loss with respect to the current
-probabilistic belief) \footnote{In our experiments, we left the
-  Bayesian methods to run the full amount of iterations, even when the
-  estimated loss was very low.}. \andrea{@stefano: cosi' non va. O tutti early stopping o nessuno..}
+uniformly at random). We set a budget of 100 iterations for all
+methods for simplicity.\andrea{@stefano: vediamo se rimane a hundred}
+
+% ; for {\sc SetMargin}a
+% we used a $0.01$ threshold on the utility loss for early stopping
+% (that does not affect the cross-validation procedure).  The Bayesian
+% methods can use the estimated loss to decide when to stop (the
+% expectation of the utility loss with respect to the current
+% probabilistic belief) \footnote{In our experiments, we left the
+%   Bayesian methods to run the full amount of iterations, even when the
+%   estimated loss was very low.}. \andrea{@stefano: cosi' non va. O tutti early stopping o nessuno..}
 

### 2016-01-28 paper/ijcai16.tex
tagli nella prima parte

@@
 
-Preferences \cite{Peintner2008} play an important role in artificial
-intelligence, including fields spanning from recommender systems to
-automatic planning, from non monotonic reasoning to computational
-social choice and algorithmic decision theory.  The task of eliciting
-or learning preferences is therefore a crucial one; typically only limited
+Preferences \cite{Peintner2008} play an important role in a variety of artificial
+intelligence applications and the task of eliciting or learning preferences is a crucial one; 
+typically only limited
 information about the user's preferences will be available and the
-cost of obtaining additional preference information will be high (from
-a cognitive and/or computational point of view).  The goal in
-preference elicitation is that of gaining enough information about the
-user's
-preferences %(typically represented by an utility function $u$, unknown to the system)
-in order to recommend a decision or a course of action.
-
-The assessment of preferences have received considerable attention in the literature, with early works confined to the operation research and management literature mostly \cite{White1984,JacquetLagreze95}.
+cost (cognitive or computational) obtaining additional preference information will be high.  
+%The goal in
+%preference elicitation is that of gaining enough information about the
+%user's
+%preferences %(typically represented by an utility function $u$, unknown to the system)
+%in order to recommend a decision or a course of action.
+
+The automated assessment of preferences has received considerable attention,
+starting with pioneering works \cite{White1984,JacquetLagreze95} in the OR community.
 %In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
 %While preferences are often assessed using some ad-hoc rules \cite{},
-Recently, in the last decade or so, a number of artificial
-intelligence researchers have proposed principled methods that elicit
-utility in an adaptive way
-\cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,braziunas2007,guo2010real,viappiani2010optimal}. The
-aim is that of focusing on learning the ``important'' part of the
-utility, allowing to recommend good (or even optimal) decisions with
-only partial knowledge of the utility function.
-
+More recently, a number of AI researchers have proposed interactive methods that elicit
+preferences in an adaptive way
+\cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,guo2010real,viappiani2010optimal},
+observing that, by asking informative questions, it is often possible to make near-optimal decisions with only partial preference information.
+
+%focusing on learning the ``important'' part of the
+%preferences, 
 %braziunas-mmr:uai07,
@@
 as a robust recommendation criterion and as a technique to drive
-elicitation in configuration problems.  The main limitation of the
-minimax regret approach is the lack of tolerance with respect to user
-inconsistency (the user might state that an item is preferred to
-another one, even though the latter has higher utility).
+elicitation.  The main limitation of their
+approach is the lack of tolerance with respect to user
+inconsistency. %(the user might state that an item is preferred to
+%another one, even though the latter has higher utility).
 % setwise max-margin formulation to deal with uncertainty in user utility
-
-Indeed, learning a user utility function requires setting a preference
-elicitation strategy and dealing with uncertain and possibly
-inconsistent user feedback.  Bayesian preference elicitation
-approaches deal with both problems by building a probability
+Indeed, learning a user utility function requires 
+%elicitation strategy and 
+to deal with uncertain and possibly inconsistent user feedback.
+
+Bayesian preference elicitation
+approaches deal with this problem by 
+building a probability
 distribution on candidate functions (endowed with a response or error
 model to be used for inference) and asking queries maximizing
-informativeness measures such as expected value of information
-(EVOI)~\cite{chajewska2000,guo2010real,viappiani2010optimal}.  These
+informativeness measures such as {\em expected value of information
+(EVOI)}~\cite{chajewska2000,guo2010real,viappiani2010optimal}.  
+These
 approaches are however computationally expensive and can not scale to
@@
 utility. 
-Our approach to preference elicitation works by combining weight vector
+Our approach to elicitation works by combining weight vector
 learning with instance generation, so that each iteration of the
@@
 vectors. 
-
 We evaluate the effectiveness of our approach by testing our
@@
 
-The paper is structured as follows: we present some background
-material in Section \ref{sec:background} and then proceed to our
-setwise max-margin method in Section \ref{sec:formulation}; we discuss
-experimental results in Section \ref{sec:experiments} and conclude
-with final remarks (Sec.~\ref{sec:conclusions}).
-
-%\section{Related Work}
+% The paper is structured as follows: we present some background
+% material in Section \ref{sec:background} and then proceed to our
+% setwise max-margin method in Section \ref{sec:formulation}; we discuss
+% experimental results in Section \ref{sec:experiments} and conclude
+% with final remarks (Sec.~\ref{sec:conclusions}).
 
@@
 
-\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors,
+%\paragraph{Notation.}
+We first introduce some notation.
+We use boldface letters $\vx$ to indicate vectors,
 uppercase letters $X$ for matrices, and calligraphic capital letters $\calX$
-for sets. We frequently abbreviate the set $\{ x^i \}_{i=1}^n$ as $\{ x^i \}$
+for sets. We %frequently 
+abbreviate the set $\{ x^i \}_{i=1}^n$ as $\{ x^i \}$
 whenever the range of the index $i$ is clear from the context, and use $[n]$ as
-a shorthand for the set $\{1, \ldots, n\}$. We write $\|\vx\|_1 := \sum_z |x_z|$
-to indicate the $\ell_1$ vector norm, $\langle \cdot, \cdot \rangle$ to
-denote the usual dot product, and $X'$ for matrix transposition.
-
-\paragraph{Setting.} We assume to have a multi-attribute feature space
+a shorthand for %the set 
+$\{1, \ldots, n\}$. We write $\|\vx\|_1 := \sum_z |x_z|$
+to indicate the $\ell_1$ vector norm, $\langle \cdot, \cdot \rangle$ for the usual dot product,  $X'$ for matrix transposition.
+
+%\paragraph{Setting.} 
+We assume to have a multi-attribute feature space
 $\calX$ of configurations $\vx = (x_1, \ldots, x_m)$ over $m$
@@
 Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
-we model users by means of utility functions of a given structure; the user's preferences are represented by a latent weight
-vector $\vw\in\bbR^m$.
-We assume linear utility functions; the utility of a configuration $\vx$ is given by
-$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$
-(see \cite{keeney1976} for theoretical conditions justifying additive utility functions). 
+we model users with additive utility functions \cite{keeney1976}; 
+the user's preferences are represented by a  weight
+vector $\vw\in\bbR^m$ and
+%We assume linear utility functions \cite{keeney1976}; 
+the utility of a configuration $\vx$ is given by
+$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. 
 In the remainder of the paper we
@@
 mixed-integer linear problem (as done in Section~\ref{sec:formulation}).
-\andrea{@paolo: sulla giustificazione delle utilita' additive, vogliamo metterla? c'e' anche molto lavoro che contesta il fatto che basti no?}
 
@@
 sets of more than two options
-\cite{viappiani2009,viappiani2010optimal}.  Comparison and choice
-queries are common in conjoint analysis
-\cite{louviere2000,toubia2004}. We consider three possible outcomes
-for a pairwise comparison between two configurations $\vx$ and $\vx'$:
+\cite{viappiani2009,viappiani2010optimal} and are common in conjoint analysis
+\cite{louviere2000,toubia2004}. %We consider three possible outcomes
+For a pairwise comparison between two configurations $\vx$ and $\vx'$:
 either $\vx$ is preferred to $\vx'$ (written $\vx \succ \vx'$), $\vx'$
@@
 $\calD$ to denote the set of preferences (answers to comparison
-queries) elicited from the user. Queries are typically selected so as
-to maximize some measure of informativeness of the resulting
-responses.  Details on our actual query selection strategy will be
-provided in Section~\ref{sec:formulation}.
+queries) elicited from the user.
+
+In the next Section we describe how informative queries can be generated
+using our setwise maxmargin learning.
+% Queries are typically selected so as
+% to maximize some measure of informativeness of the resulting
+% responses.  In the next Section we describe how Details on our actual query selection strategy will be
+% provided in Section~\ref{sec:formulation}.
 
@@
 the corresponding $\vw^{i}$ and configurations are maximally diverse
-among each other. These options will be later employed to formulate
-the user queries.
+among each other. These options will be later used to formulate
+queries.
 
@@
 
-The two core constraints are Equations~(\ref{eq:wyconstr}) and~(\ref{eq:wxconstr}),
+The two core constraints are~(\ref{eq:wyconstr}) and~(\ref{eq:wxconstr}),
 which are taken directly from the discussion above. The former enforces the
@@
 the generated configurations are diverse in terms of the weight vectors they
-maximize. Equation~(\ref{eq:wbounds}), where the inequality is intended to be component-wise, and Eq.~(\ref{eq:xbounds})
-ensure that the weights and configurations are in fact feasible, while the
-remaining two guarantee the non-negativity of the slack and margin variables.
+maximize. Constraint~(\ref{eq:wbounds}), intended to be component-wise, and~(\ref{eq:xbounds})
+ensure that the weights and configurations are feasible, while the
+remaining two guarantee the non-negativity of the slack and of the margin.
 Since we require $\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the weights to be non-negative.
@@
 lying ``close'' to the centre of each subregion. 
-Note that if we present $\vx^{1}$ and $\vx^{2}$ to the user, and the user states that he prefers
-the former to the latter, the feasible region for the weights will then become the part of the polytope to the left of the red line; moreover $\vw^{1}$ will be vector maximizing the margin in the classic sense.
-Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner a posteriori.
+If we then present $\vx^{1}$ and $\vx^{2}$ to the user, and the user states that he prefers
+the former to the latter, the feasible region for the weights will then become the part of the polytope to the left of the red line; moreover $\vw^{1}$ will maximize the margin in the classic sense ($k=1$).
+%Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner.
 %Therefore, setwise maxmargin can be thought as the set that, given any possible preference about the items, ensure that the reduction of utility uncertainty is the greatest
 %\paolo{ho fatto un tentativo di migliorare l'intuizione, ma e' ancora possibile migliorare...}
-\andrea{cosa intendi per a posteriori?}
 
@@
     \end{center}
-    \caption{\label{fig:setmargin} The intuition behind setwise max-margin; the red points are utility vectors $\vw^{1}$ $\vw^{2}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
+    \caption{\label{fig:setmargin} Optimization of setwise max-margin; the black lines corresponds to preference constraints $\calD$, the red points are utility vectors $\vw^{1}$ $\vw^{2}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
 \end{figure}
@@
 initially empty set of user responses $\calD$, it repeatedly solves
-the MILP problem above using the set $\calD$ to enforce ranking
+the MILP problem above using $\calD$ to enforce ranking
 constraints on the weight vectors $\{\vw^i\}$. The generated
@@
 new replies are added to $\calD$ and the whole procedure is
-repeated. The algorithm terminates after a fixed number of iterations,
-or once another suitable condition is met.  Termination might be left
-to the user to decide (see for example \cite{Reilly2007});
-alternatively we can terminate when the difference between the utility
-weights is small enough.
+repeated. Termination can be after a fixed number of iterations,
+when the difference between utility vectors is very small, or might be left
+to the user to decide (e.g. \cite{Reilly2007}).
 
@@
 elicitation in large configuration spaces. Our approach relies on an
-extension of max-margin learning to sets of hyperplanes, and is
-effective in the generation of a diverse set of items that can be used
-to ask informative queries to the user.  The main advantages of our
+extension of max-margin learning to sets, and is
+effective in the generation of a diverse set of configurations that can be used
+to ask informative preference queries.  The main advantages of this
 elicitation method are 1) its ability to elicit preferences and
-provide recommendations in large configuration problems 2) its
-robustness with respect to erroneous feedback and 3) its ability to
-encourage sparse utility functions. Experimental comparisons against
+provide recommendations in large configuration problems 2)
+robustness with respect to erroneous feedback and 3)  ability to
+encourage sparse utility functions. 
+Experimental comparisons against
 state-of-the-art Bayesian preference elicitation strategies confirm
 its substantial computational advantages. For sparse utility
-functions, the method compares favourably also in terms of the number
-of queries asked to the user.
+functions, the method compares favorably also in terms of utility loss.
+%the number
+%of queries asked to the user.
 

### 2016-01-28 paper/ijcai16.tex
paper: modified figures to increase font and have better axes

--- 2016-01-28 paper/ijcai16.tex
@@
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
-simulations\footnote{When experimenting with EUS and QI, the user is
-  forced to select its preffered item, so indifference is not
-  possible. While support for indifference could be added, we chose to
-  be consistent with the methods as originally published.}, as
-in~\cite{guo2010real}.
-\andrea{@paolo: e' necessaria la footnote? dobbiamo tagliare..}
+simulations, as in~\cite{guo2010real}.
+% \footnote{When experimenting with EUS and QI, the user is
+%   forced to select its preffered item, so indifference is not
+%   possible. While support for indifference could be added, we chose to
+%   be consistent with the methods as originally published.}, as
 
@@
     \begin{tabular}{cccc}
-        \hline
-        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
-        \multicolumn{2}{c}{{\sc loss/time per query}}
-        \\
-        \multicolumn{4}{c}{{\sc sparse uniform}}
-        \\
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc sparse normal}}
-        \\
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
-        \\
-        \hline
+%        \hline
+%        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
+%        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
+%        \multicolumn{4}{c}{{\sc sparse uniform}}
+        \\
+%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+
+        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        \\
+%        \hline
+        % \multicolumn{4}{c}{{\sc sparse normal}}
+        % \\
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
+        % \\
+        % \hline
     \end{tabular}
@@
     \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
-    dataset.}
+    dataset for the sparse uniform distribution.}
 \end{figure*}

### 2016-01-28 draw.py

@@ -1,5 +1,6 @@
 #!/usr/bin/env python2
 
+import math
 import sys, os
 import cPickle as pickle
 import numpy as np
@@ -103,6 +104,15 @@ def draw_matrices(ax, matrices, cumulative):
 
     return max_x, max_y
 
+def set_offset(max_val):
+    offset = math.floor(max_val / 10)    
+    if offset < 10:
+        return offset
+    if offset < 100:
+        return int(round(offset,-1))
+    if offset < 1000:
+        return int(round(offset,-2))
+
 def draw_groups(basename, groups, upper_max_x, per_query=False):
     loss_matrices, time_matrices = [], []
     for group in groups:
@@ -119,19 +129,23 @@ def draw_groups(basename, groups, upper_max_x, per_query=False):
         loss_matrices.append(loss_matrix[:,:upper_max_x])
         time_matrices.append(time_matrix[:,:upper_max_x])
 
+    # set font size
+    fonts = FontProperties()
+    fonts.set_size('xx-large')
+
     xlabel = "Number of queries" if per_query else "Number of iterations"
 
     # Loss
     fig, ax = plt.subplots(1, 1)
     max_x, max_y = draw_matrices(ax, loss_matrices, False)
 
-    ax.set_xlabel(xlabel)
+    ax.set_xlabel(xlabel, fontproperties=fonts)
     ax.set_xlim([0.0, max_x])
-    ax.set_xticks(np.arange(0, max_x, 10))
+    ax.set_xticks(np.arange(0, max_x, set_offset(max_x)))
 
-    ax.set_ylabel("Utility loss")
+    ax.set_ylabel("Utility loss", fontproperties=fonts)
     ax.set_ylim([0.0, max_y + 0.1])
-    ax.set_yticks(np.arange(0, max_y, 10 if max_y < 130 else 20))
+    ax.set_yticks(np.arange(0, max_y, set_offset(max_y)))
 
     fig.savefig(basename + "_loss.png", bbox_inches="tight")
 
@@ -139,13 +153,13 @@ def draw_groups(basename, groups, upper_max_x, per_query=False):
     fig, ax = plt.subplots(1, 1)
     max_x, max_y = draw_matrices(ax, time_matrices, True)
 
-    ax.set_xlabel(xlabel)
+    ax.set_xlabel(xlabel, fontproperties=fonts)
     ax.set_xlim([0.0, max_x])
[…]

### 2016-01-28 paper/ijcai16.tex
paper: squeezing experiments, adjusting figures

@@
 We adopt the {\em indifference-augmented} Bradley-Terry user response
-model introduced in~\cite{guo2010real}. According to the classical
-(without indifference) Bradley-Terry model~\cite{BraTer52}, when queried about a pair
-of configurations, the probability that a user will prefer
-configuration $\vx^i$ over the alternative $\vx^j$ amounts to:
+model introduced in~\cite{guo2010real}. The probability that a user
+prefers configuration $\vx^i$ over $\vx^j$ is defined according to the
+classical (without indifference) Bradley-Terry model~\cite{BraTer52} as:
 %
@@
 where $\vw$ is the weight vector of the true underlying user utility.
-In other words, the probability grows with the difference between the
-utilities of the two items. Support for indifference has been
-introduced in~\cite{guo2010real} by postulating that the probability
-of the user being indifferent about the two configurations depends on
-how close their utilities are, as in the exponential distribution:
+Support for indifference is modelled as an exponential distribution
+over the closeness of the two utilities:
 %
@@
   SetMargin} through appropriate MILP constraints, while the other
-methods require all datasets to be explicitly grounded.
-
-Users were simulated by drawing $20$ random utility vectors from each
-of four different distributions. The first two mimic those used
+methods require all datasets to be explicitly grounded.  Users were
+simulated by drawing $20$ random utility vectors from each of four
+different distributions. The first two mimic those used
 in~\cite{guo2010real}\footnote{The normal distribution reported
@@
 normal distributions setting to zero $80\%$ of the entries (sampled
-uniformly at random). We set a budget of 100 iterations for all
-methods for simplicity.\andrea{@stefano: vediamo se rimane a hundred}
+uniformly at random). We set a maximum budget of 100 iterations for all
+methods for simplicity.
 
@@
 different competitors on each of the four different utility vector
-distributions. Solution quality is measured in terms of utility loss:
+distributions and datasets $r=3$ and $r=4$. Solution quality is
+measured in terms of utility loss:
 %
@@
 Xeon CPU with 8 cores and 32 GiB of RAM.
-In all figures, thick lines indicate median values over the different
-users, shaded areas indicate standard deviations. Our {\sc SetMargin}
-algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
-in green and {\sc EUS} in gray. For all algorithms, one iteration
-corresponds to a single pairwise query to the user.  Results for $r=3$
-and $r=4$ are reported.  For dense weight vector distributions (first
-two rows), our approach achieves results which are indistinguishable
-from the competitors in a fraction of their time.  Indeed, all
-Bayesian approaches\footnote{Note that for {\sc QI} and {\sc EUS}, the time
-spent for Bayesian inference greatly dominates the time spent for choosing the next
-query to ask.} become quickly intractable for growing values of $r$,
+%In all figures, thick lines indicate median values over the different
+%users, shaded areas indicate standard deviations. Our {\sc SetMargin}
+%algorithm for $k=2$ is reported in orange, {\sc riVOI} in blue, {\sc QI}
+%in green and {\sc EUS} in gray. 
+For all algorithms, one iteration corresponds to a single pairwise (we
+used {\sc SetMargin} with $k=2$).
+%query to the user 
+%Results for $r=3$ and $r=4$ are reported.  
+For dense weight vector distributions (first two rows), our approach
+achieves results which are indistinguishable from the competitors in a
+fraction of their time.  Indeed, all Bayesian approaches\footnote{Note
+  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
+  inference greatly dominates the time spent for choosing the next
+  query to ask.\andrea{@paolo: necessario?}} become quickly intractable for growing values of $r$,
 %cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
@@
 results on $r=4$ and $r=5$ datasets for the uniform and sparse normal
-distributions\footnote{We picked the most difficult dense distribution
-  (uniform) and the simplest sparse distribution (normal) to provide
-  the broadest possible picture, as space constraints prevent us from
-  including all results. Results for the other distributions are
-  however qualitative similar to the ones reported.}, for $k=2$
-(orange), $k=3$ (blue) and $k=4$ (green). The first and third columns
-report results in terms of number of iterations. It can be seen that
-increasing the number of weight vectors tends to favour earlier
-convergence, especially for the more complex dataset ($r=5$). However,
-as in each iteration the user is asked to compare $k$ items, different
-values of $k$ imply a different cognitive effort for the user. The
-second and fourth columns report results in terms of number of
-queries, where we count all $k \choose 2$ pairs of queries when
-comparing $k$ items. In this case, $k=2$ seems to be the best
-option. The cognitive cost for the user will likely lay in between
-these two extremes, but formalizing this concept in an efficient query
-ordering strategy needs to face the effect of noise.  A modified
-sorting algorithm asking only $O(k\log k)$ queries to the user
-resulted in a performance worsening, likely because of a cascading
-effect of inconsistence feedback.
+distributions (the toughest and the simplest, for space limitations).
+% \footnote{We picked the most difficult dense distribution
+%   (uniform) and the simplest sparse distribution (normal) to provide
+%   the broadest possible picture, as space constraints prevent us from
+%   including all results. Results for the other distributions are
+%   however qualitative similar to the ones reported.}, 
+The first and third columns report results in terms of number of
+iterations. It can be seen that increasing the number of weight
+vectors tends to favour earlier convergence, especially for the more
+complex dataset ($r=5$). However, as in each iteration the user is
+asked to compare $k$ items, different values of $k$ imply a different
+cognitive effort for the user. The second and fourth columns report
+results in terms of number of queries, where we count all
+$k \choose 2$ pairs of queries when comparing $k$ items. In this case,
+$k=2$ seems to be the best option. The cognitive cost for the user
+will likely lay in between these two extremes, but formalizing this
+concept in an efficient query ordering strategy needs to face the
+effect of noise.  A modified sorting algorithm asking only
+$O(k\log k)$ queries to the user resulted in a performance worsening,
+likely because of a cascading effect of inconsistent feedback (but
+could be beneficial with different noise levels).
 
@@
 
-Figure~\ref{fig:pc} reports results of {\sc SetMargin} for $k=2$
-(red), $k=3$ (blue) and $k=4$ (green) using the sparse uniform (top
-row) and sparse normal (bottom row) distributions. The first and third
-column report utility loss for increasing number of iterations and
-queries respectively, showing a behaviour which is similar to the one
-in Figure~\ref{fig:selfcomparison}. Overall, between 50 and 70 queries
-on average are needed in order to find out a solution which is only
-10\% worse than the optimal one, out of the more than 700,000
-thousands available. Note that a vendor may ensure a considerably smaller
-number of queries by cleverly constraining the feasible configuration space;
-since our primary aim is benchmarking, we chose not to pursue this direction
-further.
-The second and fourth
+Figure~\ref{fig:pc} reports results of {\sc SetMargin} varying $k$
+using the sparse uniform distribution (the more complex of the sparse
+ones, dense distributions being unrealistic in this scenario). The
+first and third column report utility loss for increasing number of
+iterations and queries respectively, showing a behaviour which is
+similar to the one in Figure~\ref{fig:selfcomparison}. Overall,
+between 50 and 70 queries on average are needed in order to find  a
+solution which is only 10\% worse than the optimal one, out of the
+more than 700,000 thousands available. Note that a vendor may ensure a
+considerably smaller number of queries by cleverly constraining the
+feasible configuration space; since our primary aim is benchmarking,
+we chose not to pursue this direction further.  The second and fourth
 columns report cumulative times. Note that in some cases, standard
@@
     }
-    \caption{\label{fig:pc} Results for {\sc SetMargin} on the constructive PC
+    \caption{\label{fig:pc} Results for {\sc SetMargin} for $k=2$
+      (red), $k=3$ (blue) and $k=4$ (green) on the constructive PC
       dataset for the sparse uniform distribution.}

### 2016-01-28 paper/ijcai16.tex
paper: modifiche all'ultima figura

@@
       (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
-      represents a different distribution for sampling the underlying
-      user utility.  The number of iterations is plotted against the
-      utility loss (first and third columns) and the cumulative time
-      (second and fourth columns). Thick lines indicate median values
-      over users, while standard deviations are shown as shaded
-      areas.}
+      represents a different sampling distribution for user utility.
+      The number of iterations is plotted against the utility loss
+      (first and third columns) and the cumulative time (second and
+      fourth columns). Thick lines indicate median values over users,
+      while standard deviations are shown as shaded areas.}
 \end{figure*}
@@
         \\
-%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-%        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
-
-        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+
+        % \includegraphics[width=..25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
         \\
@@
       (red), $k=3$ (blue) and $k=4$ (green) on the constructive PC
-      dataset for the sparse uniform distribution.}
+      dataset for the sparse uniform distribution. Utility loss and
+      time are plotted against number of iterations (left)
+      and number of queries (right).}
 \end{figure*}

### 2016-01-28 draw.py
paper: ripulito figure

@@ -113,7 +113,7 @@ def set_offset(max_val):
     if offset < 1000:
         return int(round(offset,-2))
 
-def draw_groups(basename, groups, upper_max_x, per_query=False):
+def draw_groups(domsize, basename, groups, upper_max_x, per_query=False):
     loss_matrices, time_matrices = [], []
     for group in groups:
         paths = group.split()
@@ -147,6 +147,9 @@ def draw_groups(basename, groups, upper_max_x, per_query=False):
     ax.set_ylim([0.0, max_y + 0.1])
     ax.set_yticks(np.arange(0, max_y, set_offset(max_y)))
 
+    if domsize:
+        ax.set_title("Synthetic r=%d" %domsize, fontproperties=fonts)
+
     fig.savefig(basename + "_loss.png", bbox_inches="tight")
 
     # Time
@@ -161,10 +164,13 @@ def draw_groups(basename, groups, upper_max_x, per_query=False):
     ax.set_ylim([0.0, max_y + 0.1])
     ax.set_yticks(np.arange(0, max_y, set_offset(max_y)))
 
+    if domsize:
+        ax.set_title("Synthetic r=%d" %domsize, fontproperties=fonts)
+
     fig.savefig(basename + "_time.png", bbox_inches="tight")
 
 if __name__ == "__main__":
-    if len(sys.argv) < 5:
-        print "Usage: {} <per query?> <max x> <output basename> (<group>)+".format(sys.argv[0])
+    if len(sys.argv) < 6:
+        print "Usage: {} <per query?> <domain size> <max x> <output basename> (<group>)+".format(sys.argv[0])
         quit()
-    draw_groups(sys.argv[2], sys.argv[4:], int(sys.argv[3]), per_query=bool(int(sys.argv[1])))
+    draw_groups(int(sys.argv[2]), sys.argv[3], sys.argv[5:], int(sys.argv[4]), per_query=bool(int(sys.argv[1])))

### 2016-01-28 draw_all.sh

@@ -15,7 +15,7 @@ for domain_size in `seq 3 4`; do
         viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
         viappiani_eus_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*EUS*.txt`
         output="paper/figures/synthetic_vs_others_${domain_size}_${sampling_mode}_per_iter"
-        ./draw.py 0 $output $x_max "$setmargin_files" "$guo_files" "$viappiani_qi_files" "$viappiani_eus_files"
+        ./draw.py 0 ${domain_size} $output $x_max "$setmargin_files" "$guo_files" "$viappiani_qi_files" "$viappiani_eus_files"
     done
 done
 
@@ -32,7 +32,7 @@ for domain_size in `seq 3 4`; do
         viappiani_qi_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*QI*.txt`
         viappiani_eus_files=`ls results/ijcai16/viappiani/synthetic_${domain_size}/${sampling_mode}/*EUS*.txt`
         output="paper/figures/synthetic_vs_others_${domain_size}_${sampling_mode}_per_query"
-        ./draw.py 1 $output $x_max "$setmargin_files" "$guo_files" "$viappiani_qi_files" "$viappiani_eus_files"
+        ./draw.py 1 ${domain_size} $output $x_max "$setmargin_files" "$guo_files" "$viappiani_qi_files" "$viappiani_eus_files"
     done
 done
 
@@ -45,7 +45,7 @@ for domain_size in `seq 3 5`; do
         k3_files=`ls results/ijcai16/synthetic_${domain_size}/${sampling_mode}/*k=3*50__300*.pickle`
         k4_files=`ls results/ijcai16/synthetic_${domain_size}/${sampling_mode}/*k=4*50__300*.pickle`
         output="paper/figures/synthetic_vs_self_${domain_size}_${sampling_mode}_per_iter"
-        ./draw.py 0 $output 50 "$k2_files" "$k3_files" "$k4_files"
+        ./draw.py 0 ${domain_size} $output 50 "$k2_files" "$k3_files" "$k4_files"
     done
 done
 
@@ -56,7 +56,7 @@ for domain_size in `seq 3 5`; do
[…]

### 2016-01-28 paper/ijcai16.tex
paper: ripulito figure
paper: adjusted figures

@@
 %\usepackage{latexsym}
-\usepackage{graphicx}
+%\usepackage{graphicx}
+\usepackage{graphbox}
 \usepackage{hyperref}
+\usepackage{multirow}
 
@@
 
+% \begin{figure*}[h!]
+%     \centering
+%     {\footnotesize
+%     \begin{tabular}{cccc}
+%         % \hline
+%         % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
+%         % \\
+%         % \hline \hline
+%         \multicolumn{4}{c}{{\sc Uniform}}
+%         \\
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
+%         \\
+%         \hline
+%         \multicolumn{4}{c}{{\sc Normal}}
+%         \\
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
+%         \\
+%         \hline
+%         \multicolumn{4}{c}{{\sc Sparse Uniform}}
+%         \\
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
+%         \\
+%         \hline
+%         \multicolumn{4}{c}{{\sc Sparse Normal}}
+%         \\
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
+%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
+%         \\
+%         \hline
+%     \end{tabular}
+%     }
+%     \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
+%       (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
+%       (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
+%       represents a different sampling distribution for user utility.
+%       The number of iterations is plotted against the utility loss
+%       (first and third columns) and the cumulative time (second and
+%       fourth columns). Thick lines indicate median values over users,
+%       while standard deviations are shown as shaded areas.}
+% \end{figure*}
+
 \begin{figure*}[h!]
@@
     {\footnotesize
-    \begin{tabular}{cccc}
+    \begin{tabular}{ccccc}
+        % \hline
+        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
+        % \\
+        % \hline \hline
+        % \multicolumn{4}{c}{{\sc Uniform}}
+        \\
+        {\sc Uniform}  &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
+        \\
         \hline
-        {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
-        \\
-        \hline \hline
-        \multicolumn{4}{c}{{\sc Uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
+         % \multicolumn{4}{c}{{\sc Normal}}
+        \\
+         {\sc Normal} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
         \\
         \hline
-        \multicolumn{4}{c}{{\sc Normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
+        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
+        \\
+        {\sc Sparse Uniform} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
         \\
         \hline
-        \multicolumn{4}{c}{{\sc Sparse Uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc Sparse Normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
+        % \multicolumn{4}{c}{{\sc Sparse Normal}}
+        \\
+        {\sc Sparse Normal} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
         \\
@@
     {\footnotesize
-    \begin{tabular}{cccc}
+    \begin{tabular}{ccccc}
+        % \hline
+        % \multicolumn{2}{c}{{\sc synthetic 4}} &
+        % \multicolumn{2}{c}{{\sc synthetic 5}}
+        % \\
+        % {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
+        % \\
+        % \hline \hline
+        {\sc uniform} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
+        \\
         \hline
-        \multicolumn{2}{c}{{\sc synthetic 4}} &
-        \multicolumn{2}{c}{{\sc synthetic 5}}
-        \\
-        {\sc loss per iter.} & {\sc loss per query} & {\sc loss per iter.} & {\sc loss per query}
-        \\
-        \hline \hline
-        \multicolumn{4}{c}{{\sc uniform}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_uniform_per_query_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_uniform_per_query_loss}
-        \\
-        \hline
-        \multicolumn{4}{c}{{\sc sparse normal}}
-        \\
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
+        {\sc sparse normal} &        
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_4_normal_sparse_per_query_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_self_5_normal_sparse_per_query_loss}
         \\
@@
     {\footnotesize
-    \begin{tabular}{cccc}
+    \begin{tabular}{lcccc}
 %        \hline
@@
 %        \multicolumn{4}{c}{{\sc sparse uniform}}
-        \\
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        \includegraphics[width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        {\sc uniform} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
 

### 2016-01-28 paper/ijcai16.tex
paper: figure finally within textpdflatex ijcai16
paper: getting closer to 7 pages..

@@
 \usepackage{multirow}
+\usepackage{placeins}
 
@@
 
-We compare {\sc SetMargin} against the following competitive Bayesian baselines:
-% and \cite{viappiani2010optimal}.
-
-\begin{itemize}
-\item The Bayesian approach from \cite{guo2010real}, selecting queries
-  according to {\em restricted informed VOI} ({\sc riVOI}), a
-  computationally efficient heuristic approximation of
+We compare {\sc SetMargin} against three state-of-the-art Bayesian
+approaches: i) the Bayesian approach from \cite{guo2010real},
+selecting queries according to {\em restricted informed VOI} ({\sc
+  riVOI}), a computationally efficient heuristic approximation of
 value-of-information, and inference using TrueSkill$^{TM}$
 \cite{HerbrichMG06} (an inference scheme based on expectation
-  propagation \cite{Minka01}).
-
- \item The Bayesian framework of \cite{viappiani2010optimal} using Monte Carlo methods
- \footnote{We use 50000 particles to represent the probability distribution.}for Bayesian inference and asking choice queries (asking the user to select its preferred item among a set of diplayed ones); the next query is computed by performing a greedy optimization of {\em Expected Utility of a Selection} (hereafter just called {\sc EUS}), a submodular quantity that can be used as a proxy for value of information (giving strong approximate guarantees).
-
- \item {\em Query Iteration} (referred as {\sc QI} below), also from \cite{viappiani2010optimal}, an even faster query selection method based on sampling a set of utility vectors, selecting the items that maximize utility for each of the vector and forming a first proposal for a query set.
-Then, a new set is computed by considering the optimal items in the associated  beliefs, and the process is repeated until convergence.
-\end{itemize}
+propagation \cite{Minka01}); ii) the Bayesian framework of
+\cite{viappiani2010optimal} using Monte Carlo methods \footnote{We use
+  50000 particles to represent the probability distribution.}for
+Bayesian inference and asking choice queries (asking the user to
+select its preferred item among a set of diplayed ones); the next
+query is computed by performing a greedy optimization of {\em Expected
+  Utility of a Selection} (hereafter just called {\sc EUS}), a
+submodular quantity that can be used as a proxy for value of
+information (giving strong approximate guarantees); iii) {\em Query
+  Iteration} (referred as {\sc QI} below), also from
+\cite{viappiani2010optimal}, an even faster query selection method
+based on sampling a set of utility vectors, selecting the items that
+maximize utility for each of the vector and forming a first proposal
+for a query set. Then, a new set is computed by considering the
+optimal items in the associated beliefs, and the process is repeated
+until convergence.
 
@@
 Eq.~(\ref{eq:wyconstr}) constant, rendering the problem unbounded.}
+
+\begin{figure*}[t]
+    \centering
+    {\footnotesize
+    \begin{tabular}{ccccc}
+        % \hline
+        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
+        % \\
+        % \hline \hline
+        % \multicolumn{4}{c}{{\sc Uniform}}
+        \\
+        {\sc Uniform}  &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
+        \\
+        \hline
+         % \multicolumn{4}{c}{{\sc Normal}}
+        \\
+         {\sc Normal} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
+        \\
+        \hline
+        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
+        \\
+        {\sc Sparse Uniform} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
+        \\
+        \hline
+        % \multicolumn{4}{c}{{\sc Sparse Normal}}
+        \\
+        {\sc Sparse Normal} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
+        \\
+        \hline
+    \end{tabular}
+    }
+    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
+      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
+      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
+      represents a different sampling distribution for user utility.
+      The number of iterations is plotted against the utility loss
+      (first and third columns) and the cumulative time (second and
+      fourth columns). Thick lines indicate median values over users,
+      while standard deviations are shown as shaded areas.}
+\end{figure*}
 
@@
 
-% \begin{figure*}[h!]
-%     \centering
-%     {\footnotesize
-%     \begin{tabular}{cccc}
-%         % \hline
-%         % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
-%         % \\
-%         % \hline \hline
-%         \multicolumn{4}{c}{{\sc Uniform}}
-%         \\
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
-%         \\
-%         \hline
-%         \multicolumn{4}{c}{{\sc Normal}}
-%         \\
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
-%         \\
-%         \hline
-%         \multicolumn{4}{c}{{\sc Sparse Uniform}}
-%         \\
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
-%         \\
-%         \hline
-%         \multicolumn{4}{c}{{\sc Sparse Normal}}
-%         \\
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
-%         \includegraphics[width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
-%         \\
-%         \hline
-%     \end{tabular}
-%     }
-%     \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
-%       (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
-%       (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
-%       represents a different sampling distribution for user utility.
-%       The number of iterations is plotted against the utility loss
-%       (first and third columns) and the cumulative time (second and
-%       fourth columns). Thick lines indicate median values over users,
-%       while standard deviations are shown as shaded areas.}
-% \end{figure*}
-
-\begin{figure*}[h!]
-    \centering
-    {\footnotesize
-    \begin{tabular}{ccccc}
-        % \hline
-        % {\sc synthetic 3 loss} & {\sc synthetic 3 time} & {\sc synthetic 4 loss} & {\sc synthetic 4 time}
-        % \\
-        % \hline \hline
-        % \multicolumn{4}{c}{{\sc Uniform}}
-        \\
-        {\sc Uniform}  &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_per_iter_time} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_per_iter_time}
-        \\
-        \hline
-         % \multicolumn{4}{c}{{\sc Normal}}
-        \\
-         {\sc Normal} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_per_iter_time} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_per_iter_time}
-        \\
-        \hline
-        % \multicolumn{4}{c}{{\sc Sparse Uniform}}
-        \\
-        {\sc Sparse Uniform} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_uniform_sparse_per_iter_time} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_uniform_sparse_per_iter_time}
-        \\
-        \hline
-        % \multicolumn{4}{c}{{\sc Sparse Normal}}
-        \\
-        {\sc Sparse Normal} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_3_normal_sparse_per_iter_time} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/synthetic_vs_others_4_normal_sparse_per_iter_time}
-        \\
-        \hline
-    \end{tabular}
-    }
-    \caption{\label{fig:comparison} Comparison between {\sc SetMargin}
-      (orange), {\sc riVOI} (blue), {\sc QI} (green) and {\sc EUS}
-      (gray) on the $r=3$ (left) and $r=4$ (right) datasets. Each row
-      represents a different sampling distribution for user utility.
-      The number of iterations is plotted against the utility loss
-      (first and third columns) and the cumulative time (second and
-      fourth columns). Thick lines indicate median values over users,
-      while standard deviations are shown as shaded areas.}
-\end{figure*}
-
-%\twocolumn
-
 In Figure~\ref{fig:comparison} we report solution quality and timing
@@
 distributions and datasets $r=3$ and $r=4$. Solution quality is
-measured in terms of utility loss:
-%
-$$ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right) $$
-%
+measured in terms of utility loss $ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right)$,
 where $u(\cdot)$ is the true unknown user utility, and $\vx^*$ is the
@@
 achieves results which are indistinguishable from the competitors in a
-fraction of their time.  Indeed, all Bayesian approaches\footnote{Note
-  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
-  inference greatly dominates the time spent for choosing the next
-  query to ask.\andrea{@paolo: necessario?}} become quickly intractable for growing values of $r$,
+fraction of their time.  Indeed, all Bayesian approaches%
+%\footnote{Note
+%  that for {\sc QI} and {\sc EUS}, the time spent for Bayesian
+%  inference greatly dominates the time spent for choosing the next
+%  query to ask.\andrea{@paolo: necessario?}} 
+ become quickly intractable for growing values of $r$,
 %cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
@@
 
-\begin{figure*}[h!]
+\begin{figure*}[t]
     \centering
@@
 
+\begin{figure*}
+    \centering
+    {\footnotesize
+    \begin{tabular}{lcccc}
+%        \hline
+%        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
+%        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
+%        \multicolumn{4}{c}{{\sc sparse uniform}}
+        {\sc uniform} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+
+        % \includegraphics[width=..25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        \\
+%        \hline
+        % \multicolumn{4}{c}{{\sc sparse normal}}
+        % \\
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
+        % \\
+        % \hline
+    \end{tabular}
+    }
+    \caption{\label{fig:pc} Results for {\sc SetMargin} for $k=2$
+      (red), $k=3$ (blue) and $k=4$ (green) on the constructive PC
+      dataset for the sparse uniform distribution. Utility loss and
+      time are plotted against number of iterations (left)
+      and number of queries (right).}
+\end{figure*}
+
 \paragraph{Constructive dataset.} Next, we tested {\sc SetMargin} on a truly
@@
 constraint might look like this: ``if the manufacturer is Apple, then the
-CPU must be either a PowerPC G3 or a PowerPC G4''\footnote{Despite the components
-included in the dataset, which are a bit outdated, the dataset itself is pretty
-realistic.}, which can be encoded
-into MILP form as
+CPU must be either a PowerPC G3 or a PowerPC G4''%
+%\footnote{Despite the components
+%included in the dataset, which are a bit outdated, the dataset itself is pretty
+%realistic.} \andrea{necessario?}
+, which can be encoded into MILP form as
 %
@@
 
-\begin{figure*}
-    \centering
-    {\footnotesize
-    \begin{tabular}{lcccc}
-%        \hline
-%        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
-%        \multicolumn{2}{c}{{\sc loss/time per query}}
-        \\
-%        \multicolumn{4}{c}{{\sc sparse uniform}}
-        {\sc uniform} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
-
-        % \includegraphics[width=..25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
-        \\
-%        \hline
-        % \multicolumn{4}{c}{{\sc sparse normal}}
-        % \\
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
-        % \\
-        % \hline
-    \end{tabular}
-    }
-    \caption{\label{fig:pc} Results for {\sc SetMargin} for $k=2$
-      (red), $k=3$ (blue) and $k=4$ (green) on the constructive PC
-      dataset for the sparse uniform distribution. Utility loss and
-      time are plotted against number of iterations (left)
-      and number of queries (right).}
-\end{figure*}
-
 \end{document}

### 2016-01-28 paper/ijcai16.tex
paper: 9points for captions

@@
 \usepackage{placeins}
+\usepackage[font=small]{caption}
 
@@
 Both requirements are quite natural\footnote{Utility values are defined on an
-interval scale, therefore the assumption of non negativity is not restrictive as it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
+interval scale, thus it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
 mixed-integer linear problem (as done in Section~\ref{sec:formulation}).
@@
 propagation \cite{Minka01}); ii) the Bayesian framework of
-\cite{viappiani2010optimal} using Monte Carlo methods \footnote{We use
-  50000 particles to represent the probability distribution.}for
+\cite{viappiani2010optimal} using Monte Carlo methods (with 50,000 particles) for
 Bayesian inference and asking choice queries (asking the user to
@@
 different distributions. The first two mimic those used
-in~\cite{guo2010real}\footnote{The normal distribution reported
-  in~\cite{guo2010real} differs slightly from the one implemented in
-  the code of their experimental setting.  In our experiments we use
-  the latter.}: (1) a uniform distribution over $[1, 100]$ for each
+in~\cite{guo2010real}\footnote{We used the parameters implemented in their code, which are slightly different from those reported in the paper.}%
+%\footnote{The normal distribution reported
+%  in~\cite{guo2010real} differs slightly from the one implemented in
+%  the code of their experimental setting.  In our experiments we use
+%  the latter.}
+: (1) a uniform distribution over $[1, 100]$ for each
 individual weight, and (2) a normal distribution with mean $25$
@@
 
-\paolo{any idea about possible future works?}
+%\paolo{any idea about possible future works?}
 % FUTURE WORKS

### 2016-01-28 paper/ijcai16.tex
paper: almost there.. conclusions are out

@@
 prefers configuration $\vx^i$ over $\vx^j$ is defined according to the
-classical (without indifference) Bradley-Terry model~\cite{BraTer52} as:
-%
-$$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $$
-%
+classical (without indifference) Bradley-Terry model~\cite{BraTer52} as
+$ (1 + \exp(-\lambda_1 \langle\vw,\vx^i - \vx^j\rangle))^{-1} $,
 where $\vw$ is the weight vector of the true underlying user utility.
 Support for indifference is modelled as an exponential distribution
-over the closeness of the two utilities:
-%
-$$ \exp(-\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|). $$
-%
+over the closeness of the two utilities, i.e. 
+$ \exp(-\lambda_2 |\langle\vw,\vx^i - \vx^j\rangle|).$
 The parameters $\lambda_1$ and $\lambda_2$ were set to one for all
@@
 
+\begin{figure*}
+    \centering
+    {\footnotesize
+    \begin{tabular}{lcccc}
+%        \hline
+%        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
+%        \multicolumn{2}{c}{{\sc loss/time per query}}
+        \\
+%        \multicolumn{4}{c}{{\sc sparse uniform}}
+        {\sc uniform} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
+
+        % \includegraphics[width=..25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
+        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
+        \\
+%        \hline
+        % \multicolumn{4}{c}{{\sc sparse normal}}
+        % \\
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
+        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
+        % \\
+        % \hline
+    \end{tabular}
+    }
+    \caption{\label{fig:pc} Results for {\sc SetMargin} for $k=2$
+      (red), $k=3$ (blue) and $k=4$ (green) on the constructive PC
+      dataset for the sparse uniform distribution. Utility loss and
+      time are plotted against number of iterations (left)
+      and number of queries (right).}
+\end{figure*}
+
 In order to study the effect of increasing the number of weight
@@
 
-\begin{figure*}
-    \centering
-    {\footnotesize
-    \begin{tabular}{lcccc}
-%        \hline
-%        \multicolumn{2}{c}{{\sc loss/time per iter.}} &
-%        \multicolumn{2}{c}{{\sc loss/time per query}}
-        \\
-%        \multicolumn{4}{c}{{\sc sparse uniform}}
-        {\sc uniform} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_query_time}
-
-        % \includegraphics[width=..25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &
-        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_iter_time} &
-        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_loss} &
-        % \includegraphics[width=.25\textwidth]{figures/pc_with_costs_uniform_sparse_per_query_time}
-        \\
-%        \hline
-        % \multicolumn{4}{c}{{\sc sparse normal}}
-        % \\
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_loss} &
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_iter_time} &
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_loss} &
-        % \includegraphics[width=10em]{figures/pc_with_costs_normal_sparse_per_query_time}
-        % \\
-        % \hline
-    \end{tabular}
-    }
-    \caption{\label{fig:pc} Results for {\sc SetMargin} for $k=2$
-      (red), $k=3$ (blue) and $k=4$ (green) on the constructive PC
-      dataset for the sparse uniform distribution. Utility loss and
-      time are plotted against number of iterations (left)
-      and number of queries (right).}
-\end{figure*}
-
 \paragraph{Constructive dataset.} Next, we tested {\sc SetMargin} on a truly
@@
 %
-Note that mutual exclusivity between all binary variables of the
-one-hot encoding of an attribute ensures that only one of
-$x_\text{G3}$ and $x_\text{G4}$ can be 1.  The dataset includes
-constraints between the following attributes: manufacturer $\to$ type,
-manufacturer $\to$ CPU, type $\to$ RAM amount, type $\to$ storage
-size, type $\to$ monitor size, for 16 Horn constraints total. We do
-not report the full list due to space limitations. Note that the
-search space is of the order of hundreds of thousands of candidate
-configurations, and is far beyond reach of existing Bayesian
-approaches.
+Note that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1 because
+of the mutual exclusivity constraints of the one-hot encoding.  The
+dataset includes constraints between the following attributes:
+manufacturer $\to$ type, manufacturer $\to$ CPU, type $\to$ RAM
+amount, type $\to$ storage size, type $\to$ monitor size, for 16 Horn
+constraints total. We do not report the full list due to space
+limitations. Note that the search space is of the order of hundreds of
+thousands of candidate configurations, and is far beyond reach of
+existing Bayesian approaches.
 
@@
 functions, the method compares favorably also in terms of utility loss.
-%the number
-%of queries asked to the user.
+
+\FloatBarrier
 

### 2016-01-29 paper/ijcai16.tex
paper: almost there, three lines missing

@@
     \text{s.t.}
-        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
-    \\
-        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
+        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \label{eq:wyconstr}
     \\
@@
         & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
-    \\
-        & \;\; \mu \ge 0 \nonumber
+%    \\
+%        & \;\; \mu \ge 0 \nonumber
 \end{align}
@@
 various components. The weight regularization term copes with the
-common scenario in which the user has strong preferences about
-some attributes, but is indifferent to most of
+common scenario in which the user has strong preferences about some
+attributes, but is indifferent to most of
 them. %~\cite{} \stefano{@Paolo: add refs}.
@@
 gains in generalization ability and efficiency, as confirmed by our
-empirical findings (see Section~\ref{sec:experiments}).
-
-The two core constraints are~(\ref{eq:wyconstr}) and~(\ref{eq:wxconstr}),
-which are taken directly from the discussion above. The former enforces the
-correct ranking of the observed user preferences, while the latter ensures that
-the generated configurations are diverse in terms of the weight vectors they
-maximize. Constraint~(\ref{eq:wbounds}), intended to be component-wise, and~(\ref{eq:xbounds})
-ensure that the weights and configurations are feasible, while the
-remaining two guarantee the non-negativity of the slack and of the margin.
-Since we require $\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the weights to be non-negative.
-
-Note that we are choosing the configurations $\{ \vx^i \}$ and the weight vectors $\{
-\vw^i \}$ {\em simultaneously}.
-Look at Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a pair ($k=2$). 
-The constraint \ref{eq:wxconstr} is represented by a red line, that partitions the space of
-feasible utility weights in two parts (in general, there will be $k$ subregions).
-Since we maximize the margin $\mu$, the optimizer
-will prefer a set of configurations $\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space
-in an even way.
-In each subregion, we have corresponding utility weight vector $\vw^{i}$
-lying ``close'' to the centre of each subregion. 
-If we then present $\vx^{1}$ and $\vx^{2}$ to the user, and the user states that he prefers
-the former to the latter, the feasible region for the weights will then become the part of the polytope to the left of the red line; moreover $\vw^{1}$ will maximize the margin in the classic sense ($k=1$).
+empirical findings (see
+Section~\ref{sec:experiments}). Constraint~(\ref{eq:wyconstr})
+enforces the correct ranking of the observed user preferences, while
+constraint~(\ref{eq:wxconstr}) ensures that the generated
+configurations are diverse in terms of the weight vectors they
+maximize. Constraints~(\ref{eq:wbounds}) and~(\ref{eq:xbounds}) ensure
+that the weights and configurations are feasible, while the last one
+guarantees the non-negativity of the slacks.  Since we require
+$\vw^\bot \ge (0,\ldots,0)$, Eq.~(\ref{eq:wbounds}) also enforces the
+weights to be non-negative.
+
+Note that we are choosing the configurations $\{ \vx^i \}$ and the
+weight vectors $\{ \vw^i \}$ {\em simultaneously}.  Look at
+Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a
+pair ($k=2$).  The constraint \ref{eq:wxconstr} is represented by a
+red line, that partitions the space of feasible utility weights in two
+parts (in general, there will be $k$ subregions).  Since we maximize
+the margin $\mu$, the optimizer will prefer a set of configurations
+$\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space in an even
+way.  In each subregion, we have corresponding utility weight vector
+$\vw^{i}$ lying ``close'' to its centre. If the user indicates a
+preference for $\vx^{1}$ over $\vx^{2}$, the feasible region for the
+weights will then become the part of the polytope to the left of the
+red line; moreover $\vw^{1}$ will maximize the margin in the classic
+sense ($k=1$).
 %Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner.
@@
     \text{s.t.}
-        & \;\; \forall \; i \in [k], \forall \; h \in [n] \nonumber
-    \\
-        & \;\; \qquad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
+        & \;\; \forall \; i \in [k], \forall \; h \in [n] \quad \langle \vw^{i}, \vy^{h}_+ - \vy^{h}_- \rangle \ge \mu - \varepsilon^{i}_h \nonumber
     \\
@@
         & \;\; \forall \; i \in [k] \quad \veps^{i}_h \ge 0 \nonumber
-    \\
-        & \;\; \mu \ge 0 \nonumber
+%    \\
+%        & \;\; \mu \ge 0 \nonumber
 \end{align}
@@
 %
-which can be solved by any suitable MILP solver.
+%which can be solved by any suitable MILP solver.
 
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 {\footnotesize
@@
 
-\paragraph{Linearly dependent real attributes.} In many domains of interest,
-items are composed of both Boolean and real-valued attributes, where the latter
-depend linearly  on the former. This is for instance the case for the price,
-weight and power consumption of a laptop model, which depend linearly on the
-choice of components.
-%
-In this setting, item configurations are composed of two parts: $\vx =
-(\vx_B;\vx_R)$, where $\vx_B$ is Boolean and $\vx_R$ is real-valued, and the
-relation between the two is entirely described by an appropriately sized
-non-negative cost matrix $C$, such that $\vx_R = C \vx_B$. It is
-straightforward to extend the MILP formulation to this setting, as follows.
-
-First, we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
+\paragraph{Linearly dependent real attributes.} In many domains of
+interest, items are composed of both Boolean and real-valued
+attributes, where the latter depend linearly on the former. This is
+for instance the case for the price, weight and power consumption of a
+laptop model, which depend linearly on the choice of components.  In
+this setting, item configurations are composed of two parts:
+$\vx = (\vx_B;\vx_R)$, where $\vx_B$ is Boolean and $\vx_R$ is
+real-valued and can be written as $\vx_R = C \vx_B$ for an
+appropriately sized non-negative cost matrix $C$.  It is
+straightforward to extend the MILP formulation to this setting. First,
+we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
 becomes:
@@
 %
-As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
-to be non-negative.
+%As can be seen, non-negativity of $C$ is required for the weights $\{\vv^i\}$
+%to be non-negative.
 
@@
         \\
-        \hline
+        %\hline
     \end{tabular}
@@
 %in green and {\sc EUS} in gray. 
-For all algorithms, one iteration corresponds to a single pairwise (we
+For all algorithms, one iteration corresponds to a single pairwise query (we
 used {\sc SetMargin} with $k=2$).
@@
 achieves results which are indistinguishable from the competitors in a
-fraction of their time.  Indeed, all Bayesian approaches%
+fraction of their time.  Indeed, all Bayesian approaches
 %\footnote{Note
@@
         \\
-        \hline
+        %\hline
      \end{tabular}
@@
 %
-Interactions between attributes are expressed as Horn clauses. An example
-constraint might look like this: ``if the manufacturer is Apple, then the
-CPU must be either a PowerPC G3 or a PowerPC G4''%
-%\footnote{Despite the components
-%included in the dataset, which are a bit outdated, the dataset itself is pretty
-%realistic.} \andrea{necessario?}
-, which can be encoded into MILP form as
-%
-$ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
-%
-Note that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1 because
-of the mutual exclusivity constraints of the one-hot encoding.  The
-dataset includes constraints between the following attributes:
-manufacturer $\to$ type, manufacturer $\to$ CPU, type $\to$ RAM
-amount, type $\to$ storage size, type $\to$ monitor size, for 16 Horn
-constraints total. We do not report the full list due to space
-limitations. Note that the search space is of the order of hundreds of
-thousands of candidate configurations, and is far beyond reach of
-existing Bayesian approaches.
+Interactions between attributes are expressed as Horn clauses (e.g. a certain manufacturer implies a set of possible CPUs).
+% . An example
+% constraint might look like this: ``if the manufacturer is Apple, then the
+% CPU must be either a PowerPC G3 or a PowerPC G4''%
+% %\footnote{Despite the components
+% %included in the dataset, which are a bit outdated, the dataset itself is pretty
+% %realistic.} \andrea{necessario?}
+% , which can be encoded into MILP form as
+% %
+% $ (1 - x_\text{Apple}) + x_\text{G3} + x_\text{G4} \ge 1 $.
+% %
+% Note that only one of $x_\text{G3}$ and $x_\text{G4}$ can be 1 because
+% of the mutual exclusivity constraints of the one-hot encoding.  
+The
+dataset includes 16 Horn constraints (the full list is omitted for
+space limitations).
+% constraints between the following attributes:
+% manufacturer $\to$ type, manufacturer $\to$ CPU, type $\to$ RAM
+% amount, type $\to$ storage size, type $\to$ monitor size, for 16 Horn
+% constraints total. We do not report the full list due to space
+% limitations. 
+Note that the search space is of the order of hundreds of thousands of
+candidate configurations, and is far beyond reach of existing Bayesian
+approaches.
 

### 2016-01-29 paper/ijcai16.tex
paper: donepdflatex ijcai16 still little room for future work

@@
 %
-%which can be solved by any suitable MILP solver.
+which can be solved by any suitable MILP solver.
 
@@
 propagation \cite{Minka01}); ii) the Bayesian framework of
-\cite{viappiani2010optimal} using Monte Carlo methods (with 50,000 particles) for
-Bayesian inference and asking choice queries (asking the user to
-select its preferred item among a set of diplayed ones); the next
-query is computed by performing a greedy optimization of {\em Expected
-  Utility of a Selection} (hereafter just called {\sc EUS}), a
-submodular quantity that can be used as a proxy for value of
-information (giving strong approximate guarantees); iii) {\em Query
-  Iteration} (referred as {\sc QI} below), also from
-\cite{viappiani2010optimal}, an even faster query selection method
-based on sampling a set of utility vectors, selecting the items that
-maximize utility for each of the vector and forming a first proposal
-for a query set. Then, a new set is computed by considering the
-optimal items in the associated beliefs, and the process is repeated
-until convergence.
+\cite{viappiani2010optimal} using Monte Carlo methods (with 50,000
+particles) for Bayesian inference and asking choice queries
+(i.e. selection of the most preferred item in a set) selected
+optimizing {\em Expected Utility of a Selection} (a tight
+approximation for the value of information, hereafter just called {\sc
+  EUS}); iii) {\em Query Iteration} (referred as {\sc QI} below), also
+from \cite{viappiani2010optimal}, an even faster query selection
+method which repeateadly samples a set of utility vectors according to
+their beliefs and forms a query set selecting the items that maximize
+utility for each of the vectors.
 
@@
 
-\FloatBarrier
-
 %\paolo{any idea about possible future works?}

### 2016-01-29 paper/ijcai16.tex
paper: added future work

@@
 elicitation in large configuration spaces. Our approach relies on an
-extension of max-margin learning to sets, and is
-effective in the generation of a diverse set of configurations that can be used
-to ask informative preference queries.  The main advantages of this
-elicitation method are 1) its ability to elicit preferences and
-provide recommendations in large configuration problems 2)
-robustness with respect to erroneous feedback and 3)  ability to
-encourage sparse utility functions. 
-Experimental comparisons against
-state-of-the-art Bayesian preference elicitation strategies confirm
-its substantial computational advantages. For sparse utility
-functions, the method compares favorably also in terms of utility loss.
+extension of max-margin learning to sets, and is effective in the
+generation of a diverse set of configurations that can be used to ask
+informative preference queries.  The main advantages of this
+elicitation method are 1) ability to provide recommendations in large
+configuration problems 2) robustness with respect to erroneous
+feedback and 3) ability to encourage sparse utility functions.
+Experimental comparisons against state-of-the-art Bayesian preference
+elicitation strategies confirm these advantages. Future work includes
+extending the approach to truly hybrid scenarios (where real valued
+attributes do not depend on categorical ones) and studying the
+applicability of the setwise max-margin idea to other problems.
 

### 2016-01-29 paper/ijcai16.tex
paper: s/uniform/sparse uniform/ in PC figure.
paper: add missing 'of' in the intro.

@@
 information about the user's preferences will be available and the
-cost (cognitive or computational) obtaining additional preference information will be high.  
+cost (cognitive or computational) of obtaining additional preference information will be high.  
 %The goal in
@@
 %        \multicolumn{4}{c}{{\sc sparse uniform}}
-        {\sc uniform} &
+        {\sc sparse uniform} &
         \includegraphics[align=c,width=10em]{figures/pc_with_costs_uniform_sparse_per_iter_loss} &

### 2016-01-29 paper/ijcai16.tex
paper: singular.

@@
 
-A PC configurations is defined by eight attributes: computer type (laptop,
+A PC configuration is defined by eight attributes: computer type (laptop,
 desktop, or tower), manufacturer (8 choices), CPU model (37), monitor size (8), RAM amount (10), storage (10)

### 2016-01-29 paper/ijcai16.tex
paper: fix typo.

@@
 from \cite{viappiani2010optimal}, an even faster query selection
-method which repeateadly samples a set of utility vectors according to
+method which repeatedly samples a set of utility vectors according to
 their beliefs and forms a query set selecting the items that maximize

### 2016-01-29 paper/ijcai16.tex
paper: impractical
paper: enforce margin to be non-negative.

@@
 following mixed integer {\em non-linear} optimization problem over the
-variables $\mu \in \bbR$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
+variables $\mu \in \bbR_{\ge 0}$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
 %

### 2016-01-29 paper/ijcai16.tex
minor modifications

--- 2016-01-29 paper/ijcai16.tex
@@
 the margin $\mu$, the optimizer will prefer a set of configurations
-$\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space in an even
+$\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space in an ``even''
 way.  In each subregion, we have corresponding utility weight vector
@@
     \end{center}
-    \caption{\label{fig:setmargin} Optimization of setwise max-margin; the black lines corresponds to preference constraints $\calD$, the red points are utility vectors $\vw^{1}$ $\vw^{2}$, the red line corresponds to the constraint $\langle \vw, \vx^{1} - \vx^{2} \rangle \ge 0$.}
+    \caption{\label{fig:setmargin} Optimization of setwise max-margin; the black lines corresponds to preference constraints $\calD$, the red points are utility vectors $\vw^{1}$ and $\vw^{2}$, the red line corresponds to the hyperplane $\langle \vw, \vx^{1} - \vx^{2} \rangle = 0$.}
 \end{figure}
@@
 %  query to ask.\andrea{@paolo: necessario?}} 
- become quickly intractable for growing values of $r$,
+ become quickly impractical for growing values of $r$,
 %cannot scale over $r=4$ (for $r=5$, {\sc riVOI} did not
@@
 extending the approach to truly hybrid scenarios (where real valued
-attributes do not depend on categorical ones) and studying the
-applicability of the setwise max-margin idea to other problems.
+attributes do not depend on categorical ones) and studying its
+applicability %of the setwise max-margin idea 
+to other problems.
 

### 2016-02-01 paper/ijcai16.tex
explanation intuition setwise margin

@@
 
-In the next Section we describe how informative queries can be generated
-using our setwise maxmargin learning.
+Max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, but with  rudimental methods for query selection.
+%We adopt the idea of max-margin learning (while it has been proposed before \cite{gajos2005}, their methods query generation were rather rudimental) for preference elicitation.
+In the next Section we describe how informative queries can be generated using our setwise maxmargin learning.
 % Queries are typically selected so as
@@
 enforces the correct ranking of the observed user preferences, while
-constraint~(\ref{eq:wxconstr}) ensures that the generated
+%constraint
+~(\ref{eq:wxconstr}) ensures that the generated
 configurations are diverse in terms of the weight vectors they
@@
 Note that we are choosing the configurations $\{ \vx^i \}$ and the
-weight vectors $\{ \vw^i \}$ {\em simultaneously}.  Look at
+weight vectors $\{ \vw^i \}$ {\em simultaneously}.  
+%Note that the methods bears some similarity with volumetric approaches, but  with important differences. 
+We look for $\vw^i$ so that the utility loss (see constraint \ref{eq:wxconstr}) of choosing $\vx^{j}$ instead of $\vx^{i}$, $j \neq i$, is large (at least the shared margin $\mu$).
+Look at
 Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a
@@
 $\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space in an ``even''
-way.  In each subregion, we have corresponding utility weight vector
-$\vw^{i}$ lying ``close'' to its centre. If the user indicates a
+way\footnote{This bears similarity with volumetric approaches \cite{iyengar:acm-ec01}, but there are important differences: first here
+we consider real items to find the best separator, second the margin is also expressed in utility terms, third the query is found via an optimization process.}. 
+In each subregion, we have corresponding utility weight vector
+$\vw^{i}$ lying ``close'' to its centre. 
+If, for example, the user indicates a
 preference for $\vx^{1}$ over $\vx^{2}$, the feasible region for the
 weights will then become the part of the polytope to the left of the
-red line; moreover $\vw^{1}$ will maximize the margin in the classic
-sense ($k=1$).
+red line; moreover the vector $\vw^{1}$ will maximize the margin in the classic ($k=1$) sense in new feasible region.
 %Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner.
@@
 for instance the case for the price, weight and power consumption of a
-laptop model, which depend linearly on the choice of components.  In
-this setting, item configurations are composed of two parts:
+laptop, which depend linearly on the choice of components.  In
+this setting, %item 
+configurations are composed of two parts:
 $\vx = (\vx_B;\vx_R)$, where $\vx_B$ is Boolean and $\vx_R$ is
@@
 appropriately sized non-negative cost matrix $C$.  It is
-straightforward to extend the MILP formulation to this setting. First,
-we rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
+straightforward to extend the MILP formulation to this setting. %First,
+We rewrite the weight vector as $\vw = (\vw_B;\vw_R)$. The utility
 becomes:
@@
 (i.e. selection of the most preferred item in a set) selected
-optimizing {\em Expected Utility of a Selection} (a tight
-approximation for the value of information, hereafter just called {\sc
+using a greedy optimization of {\em Expected Utility of a Selection} (a tight
+approximation of EVOI, hereafter just called {\sc
   EUS}); iii) {\em Query Iteration} (referred as {\sc QI} below), also
 from \cite{viappiani2010optimal}, an even faster query selection
-method which repeatedly samples a set of utility vectors according to
-their beliefs and forms a query set selecting the items that maximize
-utility for each of the vectors.
+method based on sampling sets of utility vectors. % that often converge to query set of high value of information.
 
@@
 while none of the other approaches is %explicitly 
-designed to favour sparse functions.
+designed to do this. %favour sparse functions.
 

### 2016-02-01 paper/ijcai16.tex
accorciato di una riga

@@
 
-Max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, but with  rudimental methods for query selection.
-%We adopt the idea of max-margin learning (while it has been proposed before \cite{gajos2005}, their methods query generation were rather rudimental) for preference elicitation.
 In the next Section we describe how informative queries can be generated using our setwise maxmargin learning.
-% Queries are typically selected so as
-% to maximize some measure of informativeness of the resulting
-% responses.  In the next Section we describe how Details on our actual query selection strategy will be
-% provided in Section~\ref{sec:formulation}.
 
@@
 %Note that the methods bears some similarity with volumetric approaches, but  with important differences. 
-We look for $\vw^i$ so that the utility loss (see constraint \ref{eq:wxconstr}) of choosing $\vx^{j}$ instead of $\vx^{i}$, $j \neq i$, is large (at least the shared margin $\mu$).
+We look for $\vw^i$ so that the utility loss (see constraint \ref{eq:wxconstr}) of choosing $\vx^{j}$ instead of $\vx^{i}$, $j \neq i$, is large (at least $\mu$).
 Look at
 Figure~\ref{fig:setmargin}, where, for simplicity, we need to choose a
-pair ($k=2$).  The constraint \ref{eq:wxconstr} is represented by a
+pair ($k=2$).  Eq. \ref{eq:wxconstr} is represented by a
 red line, that partitions the space of feasible utility weights in two
@@
 we consider real items to find the best separator, second the margin is also expressed in utility terms, third the query is found via an optimization process.}. 
-In each subregion, we have corresponding utility weight vector
+In each subregion, we have corresponding %utility weight vector
 $\vw^{i}$ lying ``close'' to its centre. 
 If, for example, the user indicates a
-preference for $\vx^{1}$ over $\vx^{2}$, the feasible region for the
-weights will then become the part of the polytope to the left of the
+preference for $\vx^{1}$ over $\vx^{2}$, the feasible region %for the weights 
+will then become the part of the polytope to the left of the
 red line; moreover the vector $\vw^{1}$ will maximize the margin in the classic ($k=1$) sense in new feasible region.
@@
 value-of-information, and inference using TrueSkill$^{TM}$
-\cite{HerbrichMG06} (an inference scheme based on expectation
+\cite{HerbrichMG06} (based on expectation
 propagation \cite{Minka01}); ii) the Bayesian framework of
@@
 and %covariance
-standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).
+standard deviation $\frac{25}{3}$ (each attribute is sampled i.i.d).
 We further produced two novel {\em sparse} versions of the uniform and
@@
 different competitors on each of the four different utility vector
-distributions and datasets $r=3$ and $r=4$. Solution quality is
+distributions and datasets $r\!=\!3$ and $r\!=\!4$. Solution quality is
 measured in terms of utility loss $ \max_{\vx\in\calX_\text{feasible}} \left( u(\vx) - u(\vx^*) \right)$,
@@
 parameter $k$. Figure~\ref{fig:selfcomparison} reports utility loss
-results on $r=4$ and $r=5$ datasets for the uniform and sparse normal
+results on $r\!=\!4$ and $r\!=\!5$ datasets for the uniform and sparse normal
 distributions (the toughest and the simplest, for space limitations).
@@
 We presented a max-margin approach for efficient preference
-elicitation in large configuration spaces. Our approach relies on an
+elicitation in large configuration spaces.\footnote{Note that max-margin learning has been proposed before \cite{gajos2005} for preference elicitation, 
+but with  rudimental methods for query selection.}
+Our approach relies on an
 extension of max-margin learning to sets, and is effective in the

### 2016-02-01 paper/ijcai16.tex
typo

@@
 $\{ \vx^{1}, \vx^{2} \}$ that partitions the weight space in an ``even''
-way\footnote{This bears similarity with volumetric approaches \cite{iyengar:acm-ec01}, but there are important differences: first here
-we consider real items to find the best separator, second the margin is also expressed in utility terms, third the query is found via an optimization process.}. 
+way.\footnote{This bears similarity with volumetric approaches \cite{iyengar:acm-ec01}, but there are important differences: first here
+we consider real items to find the best separator, second the margin is also expressed in utility terms, third the query is found via an optimization process.}
 In each subregion, we have corresponding %utility weight vector
@@
 will then become the part of the polytope to the left of the
-red line; moreover the vector $\vw^{1}$ will maximize the margin in the classic ($k=1$) sense in new feasible region.
+red line; moreover the vector $\vw^{1}$ will maximize the margin in the classic ($k=1$) sense in the new feasible region.
 %Therefore, setwise maxmargin can be thought as a ``look-ahead'' model with respect to a standard ($k=1$) max-margin learner.

### 2016-03-11 paper/rebuttal.txt
paper: sketch out the rebuttal.

We would like to thank all reviewers for their feedback. We will buy an
additional page in the camera ready to enrich the related works section and
improve the text where it was previously cut due to space limitations.
Additional replies:

REVIEW 1

% - novelty
%     - certamente le alternative/sets non sono nuovi; la novita' e' la
%       suddivisione dello spazio di ricerca
%     - cercare roba recente che cita Guo/Bonilla/Viappiani e dimostrare che
%       non esiste nulla di piu' sul punto
%     - diversity in IR esiste [ma non ha lo stesso scopo (non dipende da w)] ma
%       per spazio ci siamo concentrati su related work piu' vicini; si potrebbe
%       comprare una pagina ulteriore nel camera ready per spiegare.
% 
% - set search not analyzed thoroughly
%     - il metodo e' giustificato anche con k=2, che funziona pure bene
%     - k>2 e' utile in base al costo cognitivo; per questo abbiamo messo 2
%       grafici (per query e per iteration); qui si vede il vantaggio di k>2
%     - no messaggio definitivo per via del costo cognitivo e del modello utente;
%       si possono esplorare alternative ma non e' il nostro fuoco; o vederlo
%       come recommendation set (di 5 prendi 1), pero' avvantaggerebbe il nostro
%       metodo
% 
% - *
%     - il fatto che i Bayes siano piu' lenti sono comunque lo stato dell'arte:
%       c'e' un tradeoff tra numero di domande e scalare nel bayesiano; noi
%       invece otteniamo le stesse perf ma scaliamo molto oltre.
%     - gli attributi booleani c'e' un misunderstanding: il problema e' ben
%       piu' complesso. spiegare attributi in dettaglio e quanti bits.

- True, usage of sets of candidates to encourage diversity is not new; the
  novelty is that we exploit the candidates to evenly split the search space.
- Diversity in information retrieval has a different goal, unrelated to
  optimization and learning. We will provide details and further references
  in the additional page.
- Recently the focus in preference elicitation shifted a bit. Our searches in
  the literature found nothing more appropriate than the work of Guo & Sanner,
  and Boutilier. We'd be grateful for any other relevant references.
- The analysis of the setmargin procedure could certaily be more theoretical. XXX
  As proven empirically, the message is that our method can provide close
  better performance than state-of-the-art methods at a fraction of the
  computational cost. This is certainly true for pairwise queries (k=2), while
  for set-based queries, where the cognitive load is less well defined, we
  report both per-iteration and per-query results to cover the full spectrum of
  possibilities.
- Bayesian methods are the state-of-the-art. True, they trade off runtime for
  modelling accuracy. But we do provide comparable performance while scaling
  much more.
- There is a misunderstanding: the number of Boolean attributes in the PC
  dataset is 73, not 8, plus a continuous cost. The problem is much more
  complex than that.

REVIEW 2

% R2
% 
% - citare UTA in OR -> citare UTA quando introduciamo le slacks
% - ha ragione, rephrase (primo qualita', secondo diversita', tutti e due diversita' dei buoni)
% - non normalizzando i pesi (che e' solo una possibilita'), la L1 e' lo standard
%   per sparsificare
% - explain that the linearization trick is used in Boutilier et al. 2006
% - chiarire che il num bool e sum domini; ma i problemi sono tranquilli perche'
%   i w utenti sono semplici e gli XOR restringono il problema
% - why should gamma be crossvalidated? intuition:
%     - il vincolo 2 diversifica
%     - il gamma vuole che gli oggetti siano di buona qualita' che in PE servono
%       perche' l'utente non si incazzi; non e' ovvio quale deve essere il suo
%       contributo percio' crossvalidiamo
%     - chiarire i bump a cosa sono dovuti (non a gamma); tagliato per spazio
% - fix footnote 4 only applies to k=1 (chiarisco che chiariro')
% - explain why 80% sparsity mimics a user (typically cares about 3 or 4 features)
%     - siamo d'accordo che con 3-4 features non serve la sparsita'; peccato
%       che dobbiamo simularlo su poche features (synth) for sake of comparisons
%       perche su PC gli altri metodi non funzionano
%     - la sparsita' non l'abbiamo introdotta per far vedere che il nostro sys
%       va meglio -- l'idea e' che scaliamo meglio sia in sparso che in denso.
%       ma usarla e' ragionevole per modellare l'utente. non ci costa nulla
%       toglierla, e non cambia la sua utilita'. il confronto non e' sulla
%       sparsita' ma sul fatto che gli altri non scalano)
%       (infatti la scalabilita' e' il nostro cavallo di battaglia)

- Thanks for the reference, it will be integrated.
- Correct, <wi,xi> encourages the quality of the xi's, diversity is given
  by constraint 2. We will rephrase the sentence.
- Fixing the 1-norm is not the only option. In our case, the norm is variable
  and sparsity is encouraged via LASSO. This is a rather standard approach.
- We will clarify that the linearization trick is taken from elsewhere, and
  refer to the paper by Boutilier et al.
- We plan to extend the work to real variables, as per the Conclusion;
  but this is beyond the scope of this paper.
- User resources strongly limit the overall complexity of the MILP problem.
  Current MILP solvers can easily deal with problems of this magnitude.
- Indeed linearly dependent variables are not very expressive. We are
  interested in extensions to hybrid Boolean/continuous domains, as per the
  Conclusion, and currently investigating two directions: alternating between
  optimization of Boolean and continuous variables, and leveraging non-linear
  solvers.
- Gamma controls the importance of candidate item quality for diversity;
  gamma=0 could be used. Since the correlation is not known beforehand, we
  include gamma in the cross-validation.
- Footnote 4 only applies to k=1 (as constraint 2 vanishes). Will fix.
- NOT SURE HOW TO WRITE IT DOWN

REVIEW 3

% - citeremo e commenteremo, grazie
% - margin must be non-negative, make clearer

- Thanks for the references, we will make sure to include them.
- The margin must be non-negative, as specified in the text above Eq 1; we will
  point it out more clearly.

REVIEW 4

% - BT: forse ci favorisce, ci stiamo solo confrontando con quello che fanno gli
%   altri, preso dai competitors; al contrario, sono i modelli bayesiani che
%   lo sfruttano meglio (o almeno che potrebbero)
% - cognitive load: ha ragione, ma tutti fanno domande a coppie es Guo & Sanner,
%   ci accodiamo. non e' chiaro come formalizzare (chiedere agli psicologi)? e'
%   da investigare. in effetti e' quello che vogliamo fare (trovare query che
%   sono piu' facili per l'utente). ma va oltre lo scopo.
% - scalability: stiamo sempre parlando di interazione con l'utente, quindi siamo
%   lontani dal raggiugnere la complex gestibile da un MILP solver. il PC ad es
%   e' gestibile mentre per l'utente e' complesso (non si va sulle migliaia di
%   var booleane). capire quanti Bools un MILP gestisce (chiedere a Roberto?).
%   la sat del solver sta molto sopra la sat dell'utente.
% - reals: future work di sicuro, infatti l'abbiamo messo li' -- le estensioni
%   sarebbero non-lineari per via di <w,x>, citare ultimo paper sanner. in
%   realta' ongoing su hybrid (alternated o nonlinear).

- We do not think the BT response model favors our method. First, it was
  previously used to evaluate the competitors, we merely took it from there.
  Second, given that Bayesian models are tied to the response model, they are
  more likely to benefit from the additional constraints that the modelling
  provides.
- True, modelling the cognitive load is not easy. We have ongoing work to
  explore this issue further, in collaboration with psychologists. In the
  paper we just followed the common practice of pairwise queries, for
  comparability.
- User resources strongly limit the overall complexity of the MILP problem;
  MILP solvers are well-equipped to solve problems of this magnitude.
- Indeed linearly dependent variables are not very expressive. We are very much
  interested in extending the method to hybrid Boolean/continuous domains, as
  hinted to in the Conclusion. We are currently investigating two directions:
  alternating between optimization of Boolean and continuous variables, and
  leveraging non-linear solvers.
- Thanks for the reference, we'll make sure to include it.

### 2016-03-11 paper/rebuttal.txt
paper: fix various stupidity in the rebuttal.

@@
 We would like to thank all reviewers for their feedback. We will buy an
-additional page in the camera ready to enrich the related works section and
+additional page for the camera ready to enrich the related works section and
 improve the text where it was previously cut due to space limitations.
@@
 - True, usage of sets of candidates to encourage diversity is not new; the
-  novelty is that we exploit the candidates to evenly split the search space.
+  novelty is that we exploit the candidates to rationally split the search space.
 - Diversity in information retrieval has a different goal, unrelated to
@@
   and Boutilier. We'd be grateful for any other relevant references.
-- The analysis of the setmargin procedure could certaily be more theoretical. XXX
+- The analysis of the setmargin procedure could certainly be more theoretical. XXX
   As proven empirically, the message is that our method can provide close
@@
 - Bayesian methods are the state-of-the-art. True, they trade off runtime for
-  modelling accuracy. But we do provide comparable performance while scaling
-  much more.
-- There is a misunderstanding: the number of Boolean attributes in the PC
-  dataset is 73, not 8, plus a continuous cost. The problem is much more
-  complex than that.
+  modelling expressivity. But we do provide comparable performance while
+  scaling much more.
+- There is a misunderstanding. The task is more complex than perceived by
+  the reviewer: the number of Boolean attributes in the PC dataset is 73, not
+  8.
 
@@
 - Fixing the 1-norm is not the only option. In our case, the norm is variable
-  and sparsity is encouraged via LASSO. This is a rather standard approach.
+  and sparsity is encouraged through LASSO. This is a standard approach.
 - We will clarify that the linearization trick is taken from elsewhere, and
-  refer to the paper by Boutilier et al.
+  refer to the paper by Boutilier.
 - We plan to extend the work to real variables, as per the Conclusion;
@@
   Current MILP solvers can easily deal with problems of this magnitude.
-- Indeed linearly dependent variables are not very expressive. We are
-  interested in extensions to hybrid Boolean/continuous domains, as per the
-  Conclusion, and currently investigating two directions: alternating between
-  optimization of Boolean and continuous variables, and leveraging non-linear
-  solvers.
-- Gamma controls the importance of candidate item quality for diversity;
-  gamma=0 could be used. Since the correlation is not known beforehand, we
-  include gamma in the cross-validation.
-- Footnote 4 only applies to k=1 (as constraint 2 vanishes). Will fix.
+- Gamma controls the importance of candidate item quality for ensuring
+  diversity; gamma=0 could be used. Since the correlation is not known
+  beforehand, we include gamma in the cross-validation.
+- Footnote 4 should only apply to k=1 (as constraint 2 vanishes. Will fix.
 - NOT SURE HOW TO WRITE IT DOWN

### 2016-03-11 paper/ijcai16.tex
paper: typo.

@@
 %20 feasible configurations.  
-For larger $r$ the size of the space grows much larger than the one
+For larger $r$ the size of the space grows much larger than the ones
 typically used in the Bayesian preference elicitation literature, and

### 2016-03-11 paper/rebuttal.txt
rebuttal: modifiche a review 1

@@
 
-- True, usage of sets of candidates to encourage diversity is not new; the
-  novelty is that we exploit the candidates to rationally split the search space.
+- We agree that using sets of candidates to encourage diversity is not new. Indeed one of the 
+  competitors we have is a Bayesian approach for optimal recommendation sets. The
+  novelty is in the set-margin formulation adapting the large margin paradigm to this setting.
+
 - Diversity in information retrieval has a different goal, unrelated to
   optimization and learning. We will provide details and further references
-  in the additional page.
-- Recently the focus in preference elicitation shifted a bit. Our searches in
-  the literature found nothing more appropriate than the work of Guo & Sanner,
-  and Boutilier. We'd be grateful for any other relevant references.
-- The analysis of the setmargin procedure could certainly be more theoretical. XXX
-  As proven empirically, the message is that our method can provide close
-  better performance than state-of-the-art methods at a fraction of the
-  computational cost. This is certainly true for pairwise queries (k=2), while
-  for set-based queries, where the cognitive load is less well defined, we
-  report both per-iteration and per-query results to cover the full spectrum of
-  possibilities.
-- Bayesian methods are the state-of-the-art. True, they trade off runtime for
-  modelling expressivity. But we do provide comparable performance while
-  scaling much more.
-- There is a misunderstanding. The task is more complex than perceived by
-  the reviewer: the number of Boolean attributes in the PC dataset is 73, not
-  8.
+  in the additional page. QUESTO LO LEVEREI
+
+- Recently the focus in preference elicitation shifted a bit. Our
+  searches in the literature found nothing more appropriate than the
+  works we compare to. We'd be grateful for any other relevant
+  references.
+
+- The main contribution of the work is that splitting the search space
+  in the fashion proposed achieves state-of-the-art performance at a
+  fraction of the computational cost. This is certainly true for
+  pairwise queries (k=2), where all methods can be compared on a fair
+  basis. For k>2, it is difficult to precisely define the cognitive
+  load.  That's why we report both per-query and per-iteration 
+  results, where the latter show faster convergence for k>2.
+
+- The fact that Bayesian methods are computationally expensive is not
+  sufficient to rule them out, as they tend to be more accurate on
+  small scale problems. What we show is that we are able to get
+  comparable performance on these problems, while scaling much more.
+
+- There is a misunderstanding regarding the complexity of the PC dataset. The
+  task is more complex than perceived by the reviewer: a binary encoding of the 
+  features would produce 73 boolean attributes, not 7. [RIVEDERE]
 

### 2016-03-11 paper/rebuttal.txt
review 4: Bradley Terry comment

@@
 
-- We do not think the BT response model favors our method. First, it was
-  previously used to evaluate the competitors, we merely took it from there.
-  Second, given that Bayesian models are tied to the response model, they are
-  more likely to benefit from the additional constraints that the modelling
-  provides.
+%- We do not think the BT response model favors our method. First, it was
+%  previously used to evaluate the competitors, we merely took it from there.
+%  Second, given that Bayesian models are tied to the response model, they are
+%  more likely to benefit from the additional constraints that the modelling
+%  provides.
+
+- We adopt Bradley Terry (BT) as a response model as it is a common assumption in preference elicitation papers.
+It is true that constraint 2 induce items that are "different" and will often be associated with a low prob of error according to the BT reponse model.
+But we do not think this as giving an unfair advantage to our method; actually it is the Bayesian competitors that exploit BT better, as it is used as an observation model when performing Bayesian updates.
+
 - True, modelling the cognitive load is not easy. We have ongoing work to

### 2016-03-11 paper/rebuttal.txt
paolo: commenti a reviewer 1

@@
 
-- We agree that using sets of candidates to encourage diversity is not new. Indeed one of the 
-  competitors we have is a Bayesian approach for optimal recommendation sets. The
-  novelty is in the set-margin formulation adapting the large margin paradigm to this setting.
+- First of all, there is a misunderstanding regarding the experiment with the PC dataset. 
+The choice space is much larger than perceived by the reviewer: the binary encoding of the 8
+features (7 discrete + 1 real) uses 73 boolean attributes. This corresponds to more than 700K feasible configurations.
+
+- While the idea of using sets of candidates to encourage diversity is not new, the
+  novelty is in extending the large margin paradigm to sets.
+We view this as a significant step forward over previous approaches using max-margin maximization and heuristic methods for uncertainty reduction [for example: Gajos 2005].
 
@@
 
-- Recently the focus in preference elicitation shifted a bit. Our
-  searches in the literature found nothing more appropriate than the
-  works we compare to. We'd be grateful for any other relevant
-  references.
+- Recently the focus in preference elicitation shifted a bit towards sequential decision making and preference-based reinforcement learning.
+Our literature search found nothing more appropriate than the works we compare to.
+We'd be grateful for pointers to other relevant references.
 
@@
   small scale problems. What we show is that we are able to get
-  comparable performance on these problems, while scaling much more.
-
-- There is a misunderstanding regarding the complexity of the PC dataset. The
-  task is more complex than perceived by the reviewer: a binary encoding of the 
-  features would produce 73 boolean attributes, not 7. [RIVEDERE]
+  comparable performance on these problems, while being able to scale much more.
 

### 2016-03-11 paper/rebuttal.txt
rebuttal: changes to rebuttal 2

--- 2016-03-11 paper/rebuttal.txt
@@
 
-- Diversity in information retrieval has a different goal, unrelated to
-  optimization and learning. We will provide details and further references
-  in the additional page. QUESTO LO LEVEREI
+% - Diversity in information retrieval has a different goal, unrelated to
+%  optimization and learning. We will provide details and further references
+%  in the additional page. QUESTO LO LEVEREI
 
@@
   refer to the paper by Boutilier.
-- We plan to extend the work to real variables, as per the Conclusion;
-  but this is beyond the scope of this paper.
-- User resources strongly limit the overall complexity of the MILP problem.
-  Current MILP solvers can easily deal with problems of this magnitude.
+- User resources strongly limit the overall complexity of the MILP
+  problem.  Current MILP solvers can easily deal with problems of this
+  magnitude. Concerning extension to real variables, we do not plan to
+  binarize them, but rather use appropriate solvers like MIQP
 - Gamma controls the importance of candidate item quality for ensuring
@@
   beforehand, we include gamma in the cross-validation.
-- Footnote 4 should only apply to k=1 (as constraint 2 vanishes. Will fix.
-- NOT SURE HOW TO WRITE IT DOWN
+- Footnote 4 should only apply to k=1 (as constraint 2 vanishes). Will fix.
+
+- The intuition for the sparse utility is that when many features are
+  available, a real user will likely use only few of them. We agree
+  with the reviewer that with 3 or 4 features sparsity is not needed,
+  but these were the domain sizes our competitors could afford. Please
+  note that the scalability of our method does not depend on the
+  sparsity assumption, which is rather introduced to account for the
+  bounded rationality of users in large decision spaces.
 

### 2016-03-11 paper/rebuttal.txt
paolo: commenti a reviewer 2

@@
   and sparsity is encouraged through LASSO. This is a standard approach.
-- We will clarify that the linearization trick is taken from elsewhere, and
+- We will clarify that the linearization trick is a classic method in OR, and
   refer to the paper by Boutilier.
-- User resources strongly limit the overall complexity of the MILP
-  problem.  Current MILP solvers can easily deal with problems of this
-  magnitude. Concerning extension to real variables, we do not plan to
+- Computational resources limit the adoption of the MILP
+  optimization.  Current MILP solvers can easily deal with problems of this
+  magnitude. Wrt extension to real variables, we do not plan to
   binarize them, but rather use appropriate solvers like MIQP
-- Gamma controls the importance of candidate item quality for ensuring
-  diversity; gamma=0 could be used. Since the correlation is not known
-  beforehand, we include gamma in the cross-validation.
-- Footnote 4 should only apply to k=1 (as constraint 2 vanishes). Will fix.
+- Gamma controls the importance given to the quality of the candidate, since we do want to favour queries with 'good' items in term of utility; in principle gamma=0 could be used.
+Since the relation between utility of items in a set and the information value of the query  is not known  beforehand (contrary to the settings investigated by Viappiani&Boutilier), we include gamma in the cross-validation.
 
-- The intuition for the sparse utility is that when many features are
+- Footnote 4 is only relevant to the case k=1 (as constraint 2 vanishes). Will fix.
+
+- The intuition for the sparse utility is that when a large number of features are
   available, a real user will likely use only few of them. We agree
   with the reviewer that with 3 or 4 features sparsity is not needed,
-  but these were the domain sizes our competitors could afford. Please
-  note that the scalability of our method does not depend on the
-  sparsity assumption, which is rather introduced to account for the
-  bounded rationality of users in large decision spaces.
+  but these were the domain sizes suitable for the comparison with Bayesian methods.
+Please note that the scalability of our method does not depend on the sparsity assumption, which is rather introduced to account for the bounded rationality of users in large decision spaces.
 

### 2016-03-11 paper/rebuttal.txt
review2 cutting rebuttal

--- 2016-03-11 paper/rebuttal.txt
@@
 
-REVIEW 1
+REVIEW_1
 
@@
 
-- First of all, there is a misunderstanding regarding the experiment with the PC dataset. 
-The choice space is much larger than perceived by the reviewer: the binary encoding of the 8
-features (7 discrete + 1 real) uses 73 boolean attributes. This corresponds to more than 700K feasible configurations.
+First of all, there is a misunderstanding regarding the experiment
+with the PC dataset.  The choice space is much larger than perceived
+by the reviewer: the binary encoding of the 8 features (7 discrete + 1
+real) uses 73 boolean attributes. This corresponds to more than 700K
+feasible configurations.
 
-- While the idea of using sets of candidates to encourage diversity is not new, the
-  novelty is in extending the large margin paradigm to sets.
-We view this as a significant step forward over previous approaches using max-margin maximization and heuristic methods for uncertainty reduction [for example: Gajos 2005].
+While the idea of using sets of candidates to encourage diversity is
+not new, the novelty is in extending the large margin paradigm to
+sets. We view this as a significant step forward over previous
+approaches using max-margin maximization and heuristic methods for
+uncertainty reduction [for example: Gajos 2005].
 
-% - Diversity in information retrieval has a different goal, unrelated to
-%  optimization and learning. We will provide details and further references
-%  in the additional page. QUESTO LO LEVEREI
+Recently the focus in preference elicitation shifted a bit towards
+sequential decision making and preference-based reinforcement
+learning. Our literature search found nothing more appropriate than
+the works we compare to. We'd be grateful for pointers to other
+relevant references.
 
-- Recently the focus in preference elicitation shifted a bit towards sequential decision making and preference-based reinforcement learning.
-Our literature search found nothing more appropriate than the works we compare to.
-We'd be grateful for pointers to other relevant references.
-
-- The main contribution of the work is that splitting the search space
+The main contribution of the work is that splitting the search space
 in the fashion proposed achieves state-of-the-art performance at a
@@
 basis. For k>2, it is difficult to precisely define the cognitive
-  load.  That's why we report both per-query and per-iteration 
-  results, where the latter show faster convergence for k>2.
+load.  That's why we report both per-query and per-iteration results,
+where the latter show faster convergence for k>2.
 
-- The fact that Bayesian methods are computationally expensive is not
-  sufficient to rule them out, as they tend to be more accurate on
-  small scale problems. What we show is that we are able to get
-  comparable performance on these problems, while being able to scale much more.
+The fact that Bayesian methods are computationally expensive is not
+sufficient to rule them out, as they tend to be more accurate on small
+scale problems. What we show is that we are able to get comparable
+performance on these problems, while being able to scale much more.
 
-REVIEW 2
+REVIEW_2
 
@@
 
-- Thanks for the reference, it will be integrated.
-- Correct, <wi,xi> encourages the quality of the xi's, diversity is given
+Thanks for the reference, it will be integrated.
+
+Correct, <wi,xi> encourages the quality of the xi's, diversity is given
 by constraint 2. We will rephrase the sentence.
-- Fixing the 1-norm is not the only option. In our case, the norm is variable
-  and sparsity is encouraged through LASSO. This is a standard approach.
-- We will clarify that the linearization trick is taken from elsewhere, and
-  refer to the paper by Boutilier.
-- User resources strongly limit the overall complexity of the MILP
-  problem.  Current MILP solvers can easily deal with problems of this
-  magnitude. Concerning extension to real variables, we do not plan to
-  binarize them, but rather use appropriate solvers like MIQP
-- Gamma controls the importance of candidate item quality for ensuring
+
+Fixing the 1-norm is not the only option. In our case, the norm can
+vary and sparsity is encouraged through LASSO. This is a standard
+approach.
+
+We will clarify that the linearization trick is taken from elsewhere,
+and refer to the paper by Boutilier.
+
+Our plan for extending the approach real variables is not to binarize
+them, but rather use appropriate solvers like MIQP
+
+Gamma controls the importance of candidate item quality for ensuring
 diversity; gamma=0 could be used. Since the correlation is not known
-  beforehand, we include gamma in the cross-validation.
-- Footnote 4 should only apply to k=1 (as constraint 2 vanishes). Will fix.
+beforehand, we include gamma in the cross-validation.  Footnote 4
+should only apply to k=1 (as constraint 2 vanishes). Will fix.
 
-- The intuition for the sparse utility is that when many features are
-  available, a real user will likely use only few of them. We agree
-  with the reviewer that with 3 or 4 features sparsity is not needed,
-  but these were the domain sizes our competitors could afford. Please
-  note that the scalability of our method does not depend on the
-  sparsity assumption, which is rather introduced to account for the
-  bounded rationality of users in large decision spaces.
+The intuition for the sparse utility is that when many features are
+available, a real user will likely use only few of them. We agree with
+the reviewer that with 3 or 4 features sparsity is not needed, but
+these were the domain sizes our competitors could afford. Please note
+that the scalability of our method does not depend on the sparsity
+assumption, which is rather introduced to account for the bounded
+rationality of users in large decision spaces.
 
-REVIEW 3
+REVIEW_3
 
@@
 
-- Thanks for the references, we will make sure to include them.
-- The margin must be non-negative, as specified in the text above Eq 1; we will
-  point it out more clearly.
+Thanks for the references, we will make sure to include them.
 
-REVIEW 4
+The margin must be non-negative, as specified in the text above Eq 1;
+we will point it out more clearly.
+
+REVIEW_4
 
@@
 
-- We adopt Bradley Terry (BT) as a response model as it is a common assumption in preference elicitation papers.
-It is true that constraint 2 induce items that are "different" and will often be associated with a low prob of error according to the BT reponse model.
-But we do not think this as giving an unfair advantage to our method; actually it is the Bayesian competitors that exploit BT better, as it is used as an observation model when performing Bayesian updates.
+We adopt Bradley Terry (BT) as a response model as it is a common
+assumption in preference elicitation papers. It is true that
+constraint 2 induce items that are "different" and will often be
+associated with a low prob of error according to the BT reponse model.
+But we do not think this as giving an unfair advantage to our method;
+actually it is the Bayesian competitors that exploit BT better, as it
+is used as an observation model when performing Bayesian updates.
 
-- True, modelling the cognitive load is not easy. We have ongoing work to
-  explore this issue further, in collaboration with psychologists. In the
-  paper we just followed the common practice of pairwise queries, for
-  comparability.
-- User resources strongly limit the overall complexity of the MILP problem;
-  MILP solvers are well-equipped to solve problems of this magnitude.
-- Indeed linearly dependent variables are not very expressive. We are very much
-  interested in extending the method to hybrid Boolean/continuous domains, as
-  hinted to in the Conclusion. We are currently investigating two directions:
-  alternating between optimization of Boolean and continuous variables, and
-  leveraging non-linear solvers.
-- Thanks for the reference, we'll make sure to include it.
+Mdelling the cognitive load is indeed not easy. We have ongoing work
+to explore this issue further, in collaboration with psychologists. In
+the paper we just followed the common practice of pairwise queries,
+for comparability.
+
+User resources strongly limit the overall complexity of the MILP
+problem; MILP solvers are well-equipped to solve problems of this
+magnitude.
+
+Indeed linearly dependent variables are not very expressive. We are
+very much interested in extending the method to hybrid
+Boolean/continuous domains, as hinted to in the Conclusion. We are
+currently investigating two directions: alternating between
+optimization of Boolean and continuous variables, and leveraging
+non-linear solvers.
+
+Thanks for the reference, we'll make sure to include it.

### 2016-03-11 paper/rebuttal.txt

@@
 
-Correct, <wi,xi> encourages the quality of the xi's, diversity is given
-by constraint 2. We will rephrase the sentence.
+Correct, <wi,xi> encourages the quality of the xi's, diversity is
+given by constraint 2. We will rephrase the sentence.
 
-Fixing the 1-norm is not the only option. In our case, the norm can
-vary and sparsity is encouraged through LASSO. This is a standard
+Fixing the 1-norm is not the only option. In our case, the norm is
+variable and sparsity is encouraged through LASSO. This is a standard
 approach.
 
-We will clarify that the linearization trick is taken from elsewhere,
-and refer to the paper by Boutilier.
+We will clarify that the linearization trick is a classic method in OR, and
+refer to the paper by Boutilier.
 
@@
 
-Gamma controls the importance of candidate item quality for ensuring
-diversity; gamma=0 could be used. Since the correlation is not known
-beforehand, we include gamma in the cross-validation.  Footnote 4
-should only apply to k=1 (as constraint 2 vanishes). Will fix.
+Gamma controls the importance given to the quality of the candidate,
+since we do want to favour queries with 'good' items in term of
+utility; in principle gamma=0 could be used.  Since the relation
+between utility of items in a set and the information value of the
+query is not known beforehand (contrary to the settings investigated
+by Viappiani&Boutilier), we include gamma in the cross-validation.
 
-The intuition for the sparse utility is that when many features are
-available, a real user will likely use only few of them. We agree with
-the reviewer that with 3 or 4 features sparsity is not needed, but
-these were the domain sizes our competitors could afford. Please note
-that the scalability of our method does not depend on the sparsity
-assumption, which is rather introduced to account for the bounded
-rationality of users in large decision spaces.
+Footnote 4 is only relevant to the case k=1 (as constraint 2 vanishes). Will fix.
+
+The intuition for the sparse utility is that when a large number of
+features are available, a real user will likely use only few of
+them. We agree with the reviewer that with 3 or 4 features sparsity is
+not needed, but these were the domain sizes suitable for the
+comparison with Bayesian methods. Please note that the scalability of
+our method does not depend on the sparsity assumption, which is rather
+introduced to account for the bounded rationality of users in large
+decision spaces.
 

### 2016-03-11 paper/rebuttal.txt
paolo: revisione reviewer 4

@@
 
-User resources strongly limit the overall complexity of the MILP
-problem; MILP solvers are well-equipped to solve problems of this
-magnitude.
+User cognitive resources (rather than computational ones) limit the size (in terms of attributes) of the decision problems; MILP solvers are well-equipped to solve problems of this magnitude.
 
-Indeed linearly dependent variables are not very expressive. We are
-very much interested in extending the method to hybrid
-Boolean/continuous domains, as hinted to in the Conclusion. We are
-currently investigating two directions: alternating between
-optimization of Boolean and continuous variables, and leveraging
-non-linear solvers.
+Indeed linearly dependent variables are not very expressive. We are interested in extending to hybrid Boolean/continuous domains. 
+%We are currently investigating two directions: alternating between
+%optimization of Boolean and continuous variables, and leveraging
+%non-linear solvers.
 
-Thanks for the reference, we'll make sure to include it.
+Thanks for the reference. 
+%, we'll make sure to include it.

### 2016-03-11 paper/rebuttal.txt
paolo: revisione reviewer 4

@@
 constraint 2 induce items that are "different" and will often be
-associated with a low prob of error according to the BT reponse model.
+associated with a low prob-of-error according to the BT reponse model.
 But we do not think this as giving an unfair advantage to our method;

### 2016-03-11 paper/rebuttal.txt
rebuttal 2 cut

--- 2016-03-11 paper/rebuttal.txt
@@
 approaches using max-margin maximization and heuristic methods for
-uncertainty reduction [for example: Gajos 2005].
+uncertainty reduction.
 
@@
 Correct, <wi,xi> encourages the quality of the xi's, diversity is
-given by constraint 2. We will rephrase the sentence.
+given by constraint 2. We will rephrase.
 
@@
 
-Gamma controls the importance given to the quality of the candidate,
-since we do want to favour queries with 'good' items in term of
-utility; in principle gamma=0 could be used.  Since the relation
-between utility of items in a set and the information value of the
-query is not known beforehand (contrary to the settings investigated
-by Viappiani&Boutilier), we include gamma in the cross-validation.
+Gamma controls the importance given to the quality of the candidate
+(we want to suggest 'good' items). This needs to be traded with respect
+to diversity, that's why we include gamma in the cross-validation.
 
@@
 
-The intuition for the sparse utility is that when a large number of
-features are available, a real user will likely use only few of
-them. We agree with the reviewer that with 3 or 4 features sparsity is
-not needed, but these were the domain sizes suitable for the
-comparison with Bayesian methods. Please note that the scalability of
-our method does not depend on the sparsity assumption, which is rather
-introduced to account for the bounded rationality of users in large
-decision spaces.
+The intuition for the sparse utility is based on the bounded
+rationality of humans in large decision spaces. We agree 3 or 4
+features do not require sparsity, but these were the domain sizes
+suitable for the comparison with Bayesian methods. Please note that
+the scalability of our method does not depend on the sparsity
+assumption.
 

### 2016-03-11 paper/rebuttal.txt
cut rebuttal to 598

@@
-We would like to thank all reviewers for their feedback. We will buy an
-additional page for the camera ready to enrich the related works section and
-improve the text where it was previously cut due to space limitations.
-Additional replies:
+We would like to thank all reviewers for their feedback. Detailed replies follow:
 
@@
 
-The main contribution of the work is that splitting the search space
-in the fashion proposed achieves state-of-the-art performance at a
-fraction of the computational cost. This is certainly true for
-pairwise queries (k=2), where all methods can be compared on a fair
-basis. For k>2, it is difficult to precisely define the cognitive
-load.  That's why we report both per-query and per-iteration results,
-where the latter show faster convergence for k>2.
+The main contribution of the work is that our method for splitting the
+search space achieves state-of-the-art performance at a fraction of
+the computational cost. This is certainly true for pairwise queries
+(k=2), where all methods can be compared.  For k>2, it is difficult to
+precisely define the cognitive load.  That's why we report both
+per-query and per-iteration results; the latter show faster
+convergence for k>2.
 
@@
 sufficient to rule them out, as they tend to be more accurate on small
-scale problems. What we show is that we are able to get comparable
-performance on these problems, while being able to scale much more.
+problems. Our method achieves comparable performance on these
+problems, while capable of scaling much more.
 
@@
 
-Correct, <wi,xi> encourages the quality of the xi's, diversity is
+Correct, <wi,xi> encourages the quality of the xi's; diversity is
 given by constraint 2. We will rephrase.
@@
 Fixing the 1-norm is not the only option. In our case, the norm is
-variable and sparsity is encouraged through LASSO. This is a standard
+varying and sparsity is encouraged through LASSO. This is a standard
 approach.
 
-We will clarify that the linearization trick is a classic method in OR, and
-refer to the paper by Boutilier.
+We will clarify that the linearization trick is a classic method in OR.
+%and refer to the paper by Boutilier.
 
@@
 
-The intuition for the sparse utility is based on the bounded
-rationality of humans in large decision spaces. We agree 3 or 4
+The intuition for the sparse utility is based on human's bounded
+rationality in large decision spaces. We agree 3 or 4
 features do not require sparsity, but these were the domain sizes
@@
 
-Thanks for the references, we will make sure to include them.
+Thanks for the references, we'll include them.
 
-The margin must be non-negative, as specified in the text above Eq 1;
+The margin must be non-negative, as specified in the text above Eq.1;
 we will point it out more clearly.
@@
 
-We adopt Bradley Terry (BT) as a response model as it is a common
-assumption in preference elicitation papers. It is true that
-constraint 2 induce items that are "different" and will often be
-associated with a low prob-of-error according to the BT reponse model.
-But we do not think this as giving an unfair advantage to our method;
-actually it is the Bayesian competitors that exploit BT better, as it
-is used as an observation model when performing Bayesian updates.
+Bradley Terry (BT) is a common assumption in preference
+elicitation. It is true that constraint 2 tends to induce items with a
+low prob-of-error according to BT. However we do not see this as
+giving us an unfair advantage; actually it is the Bayesian competitors
+that exploit BT better when using it for performing Bayesian updates.
 
-Mdelling the cognitive load is indeed not easy. We have ongoing work
-to explore this issue further, in collaboration with psychologists. In
-the paper we just followed the common practice of pairwise queries,
-for comparability.
+Mdelling the cognitive load is indeed not easy, we just followed the
+common practice of pairwise queries. We have ongoing work in
+collaboration with psychologists to explore this issue further.
 
-User cognitive resources (rather than computational ones) limit the size (in terms of attributes) of the decision problems; MILP solvers are well-equipped to solve problems of this magnitude.
+User cognitive resources (rather than computational ones) limit the
+size of problems; MILP solvers are well-equipped to solve
+optimizations of this magnitude.
 
-Indeed linearly dependent variables are not very expressive. We are interested in extending to hybrid Boolean/continuous domains. 
+Indeed linearly dependent variables are not very expressive. We are
+interested in extending to hybrid Boolean/continuous domains.  
 %We are currently investigating two directions: alternating between
@@
 
-Thanks for the reference. 
-%, we'll make sure to include it.
+We'll include the suggested reference.

### 2016-03-11 paper/rebuttal.txt
paper: remove comments from rebuttal, fix a couple of things.

@@
 REVIEW_1
-
-% - novelty
-%     - certamente le alternative/sets non sono nuovi; la novita' e' la
-%       suddivisione dello spazio di ricerca
-%     - cercare roba recente che cita Guo/Bonilla/Viappiani e dimostrare che
-%       non esiste nulla di piu' sul punto
-%     - diversity in IR esiste [ma non ha lo stesso scopo (non dipende da w)] ma
-%       per spazio ci siamo concentrati su related work piu' vicini; si potrebbe
-%       comprare una pagina ulteriore nel camera ready per spiegare.
-% 
-% - set search not analyzed thoroughly
-%     - il metodo e' giustificato anche con k=2, che funziona pure bene
-%     - k>2 e' utile in base al costo cognitivo; per questo abbiamo messo 2
-%       grafici (per query e per iteration); qui si vede il vantaggio di k>2
-%     - no messaggio definitivo per via del costo cognitivo e del modello utente;
-%       si possono esplorare alternative ma non e' il nostro fuoco; o vederlo
-%       come recommendation set (di 5 prendi 1), pero' avvantaggerebbe il nostro
-%       metodo
-% 
-% - *
-%     - il fatto che i Bayes siano piu' lenti sono comunque lo stato dell'arte:
-%       c'e' un tradeoff tra numero di domande e scalare nel bayesiano; noi
-%       invece otteniamo le stesse perf ma scaliamo molto oltre.
-%     - gli attributi booleani c'e' un misunderstanding: il problema e' ben
-%       piu' complesso. spiegare attributi in dettaglio e quanti bits.
 
@@
 
-% R2
-% 
-% - citare UTA in OR -> citare UTA quando introduciamo le slacks
-% - ha ragione, rephrase (primo qualita', secondo diversita', tutti e due diversita' dei buoni)
-% - non normalizzando i pesi (che e' solo una possibilita'), la L1 e' lo standard
-%   per sparsificare
-% - explain that the linearization trick is used in Boutilier et al. 2006
-% - chiarire che il num bool e sum domini; ma i problemi sono tranquilli perche'
-%   i w utenti sono semplici e gli XOR restringono il problema
-% - why should gamma be crossvalidated? intuition:
-%     - il vincolo 2 diversifica
-%     - il gamma vuole che gli oggetti siano di buona qualita' che in PE servono
-%       perche' l'utente non si incazzi; non e' ovvio quale deve essere il suo
-%       contributo percio' crossvalidiamo
-%     - chiarire i bump a cosa sono dovuti (non a gamma); tagliato per spazio
-% - fix footnote 4 only applies to k=1 (chiarisco che chiariro')
-% - explain why 80% sparsity mimics a user (typically cares about 3 or 4 features)
-%     - siamo d'accordo che con 3-4 features non serve la sparsita'; peccato
-%       che dobbiamo simularlo su poche features (synth) for sake of comparisons
-%       perche su PC gli altri metodi non funzionano
-%     - la sparsita' non l'abbiamo introdotta per far vedere che il nostro sys
-%       va meglio -- l'idea e' che scaliamo meglio sia in sparso che in denso.
-%       ma usarla e' ragionevole per modellare l'utente. non ci costa nulla
-%       toglierla, e non cambia la sua utilita'. il confronto non e' sulla
-%       sparsita' ma sul fatto che gli altri non scalano)
-%       (infatti la scalabilita' e' il nostro cavallo di battaglia)
-
 Thanks for the reference, it will be integrated.
@@
 We will clarify that the linearization trick is a classic method in OR.
-%and refer to the paper by Boutilier.
 
-Our plan for extending the approach real variables is not to binarize
-them, but rather use appropriate solvers like MIQP
+Our plan for extending the approach to real variables is not to binarize
+them, but rather use appropriate solvers like MIQP.
 
@@
 
-Footnote 4 is only relevant to the case k=1 (as constraint 2 vanishes). Will fix.
+Footnote 4 is only relevant to the case k=1 (as constraint 2 vanishes). Will
+fix.
 
@@
 
-% - citeremo e commenteremo, grazie
-% - margin must be non-negative, make clearer
-
 Thanks for the references, we'll include them.
@@
 
-% - BT: forse ci favorisce, ci stiamo solo confrontando con quello che fanno gli
-%   altri, preso dai competitors; al contrario, sono i modelli bayesiani che
-%   lo sfruttano meglio (o almeno che potrebbero)
-% - cognitive load: ha ragione, ma tutti fanno domande a coppie es Guo & Sanner,
-%   ci accodiamo. non e' chiaro come formalizzare (chiedere agli psicologi)? e'
-%   da investigare. in effetti e' quello che vogliamo fare (trovare query che
-%   sono piu' facili per l'utente). ma va oltre lo scopo.
-% - scalability: stiamo sempre parlando di interazione con l'utente, quindi siamo
-%   lontani dal raggiugnere la complex gestibile da un MILP solver. il PC ad es
-%   e' gestibile mentre per l'utente e' complesso (non si va sulle migliaia di
-%   var booleane). capire quanti Bools un MILP gestisce (chiedere a Roberto?).
-%   la sat del solver sta molto sopra la sat dell'utente.
-% - reals: future work di sicuro, infatti l'abbiamo messo li' -- le estensioni
-%   sarebbero non-lineari per via di <w,x>, citare ultimo paper sanner. in
-%   realta' ongoing su hybrid (alternated o nonlinear).
-
-%- We do not think the BT response model favors our method. First, it was
-%  previously used to evaluate the competitors, we merely took it from there.
-%  Second, given that Bayesian models are tied to the response model, they are
-%  more likely to benefit from the additional constraints that the modelling
-%  provides.
-
 Bradley Terry (BT) is a common assumption in preference
@@
 
-Mdelling the cognitive load is indeed not easy, we just followed the
+Modelling the cognitive load is indeed not easy, we just followed the
 common practice of pairwise queries. We have ongoing work in

### 2016-03-11 paper/rebuttal.txt
removed last comments
rebuttal: removed -

@@
 with the PC dataset.  The choice space is much larger than perceived
-by the reviewer: the binary encoding of the 8 features (7 discrete + 1
+by the reviewer: the binary encoding of the 8 features (7 discrete, 1
 real) uses 73 boolean attributes. This corresponds to more than 700K
@@
 sets. We view this as a significant step forward over previous
-approaches using max-margin maximization and heuristic methods for
+approaches using maxmargin maximization and heuristic methods for
 uncertainty reduction.
@@
 Recently the focus in preference elicitation shifted a bit towards
-sequential decision making and preference-based reinforcement
-learning. Our literature search found nothing more appropriate than
-the works we compare to. We'd be grateful for pointers to other
-relevant references.
+sequential decision making and reinforcement learning. Our literature
+search found nothing more appropriate than the works we compare
+to. We'd be grateful for pointers to other relevant references.
 
@@
 precisely define the cognitive load.  That's why we report both
-per-query and per-iteration results; the latter show faster
+perQuery and perIiteration results; the latter show faster
 convergence for k>2.
@@
 
-Fixing the 1-norm is not the only option. In our case, the norm is
+Fixing the 1_norm is not the only option. In our case, the norm is
 varying and sparsity is encouraged through LASSO. This is a standard
@@
 (we want to suggest 'good' items). This needs to be traded with respect
-to diversity, that's why we include gamma in the cross-validation.
+to diversity, that's why we include gamma in the crossvalidation.
 
@@
 
-The margin must be non-negative, as specified in the text above Eq.1;
+The margin must be nonnegative, as specified in the text above Eq.1;
 we will point it out more clearly.
@@
 elicitation. It is true that constraint 2 tends to induce items with a
-low prob-of-error according to BT. However we do not see this as
+low error probability according to BT. However we do not see this as
 giving us an unfair advantage; actually it is the Bayesian competitors
@@
 User cognitive resources (rather than computational ones) limit the
-size of problems; MILP solvers are well-equipped to solve
-optimizations of this magnitude.
+size of problems; MILP solvers can easily address optimizations of this
+magnitude.
 
@@
 interested in extending to hybrid Boolean/continuous domains.
-%We are currently investigating two directions: alternating between
-%optimization of Boolean and continuous variables, and leveraging
-%non-linear solvers.
 

### 2016-04-13 paper/ijcai16.tex
paper: update following the rebuttal.

@@
 \newcommand{\paolo}[1]{{\bf \textcolor{red}{{\fbox{Paolo:} #1}}}}
+\newcommand{\diff}[1]{{\bf \textcolor{violet}{#1}}}
 
@@
 
-\title{Constructive Preference Elicitation by Setwise Max-margin Learning} %with?
-\author{ID 2243}
+\title{Constructive Preference Elicitation by Setwise Max-margin Learning}
+
+\author{Stefano Teso \\
+University of Trento \\
+Trento, Italy \\
+teso@disi.unitn.it
+\And
+Andrea Passerini \\
+University of Trento \\
+Trento, Italy \\
+passerini@disi.unitn.it
+\And
+Paolo Viappiani \\
+Universit\'e Pierre et Marie Curie \\
+Paris, France \\
+paolo.viappiani@lip6.fr}
 
@@
 
+\paolo{or Andrea: check whether~\cite{greco2008ordinal} is related and comment
+on it, as requested by reviewer 3.}
+
 Preferences \cite{Peintner2008} play an important role in a variety of artificial
@@
 The automated assessment of preferences has received considerable attention,
-starting with pioneering works \cite{White1984,JacquetLagreze95} in the OR community.
+starting with pioneering works \cite{White1984,jaquetlsiskos1982} in the OR community.
+\paolo{or Andrea, check if it is worth mentioning the review in~\cite{jaquetlsiskos2001},
+as highlighted by the fourth reviewer.}
 %In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
@@
 
-The second goal requires to jointly maximize the utility of each
+The second goal requires \stefano{reviewer 2, rephrase: to jointly maximize the utility of each
 $\vx^{i}$ according to its corresponding weight vector $\vw^i$ and its
@@
 the set. We achieve this by maximizing the sum of utilities
-$\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ and adding ranking
+$\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ and} adding ranking
 constraints of the form
@@
 following mixed integer {\em non-linear} optimization problem over the
-variables $\mu \in \bbR_{\ge 0}$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
+variables $\mu \in \bbR_{\ge 0}$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$ \stefano{stress non-zero margin requirement}:
 %
@@
 replace Eq.~(\ref{eq:wxconstr}) with a set of linear constraints through
-a suitable variable transformation.
+a suitable variable transformation; \stefano{this technique has already been
+used in OR~[Boutilier et al., 2006]}.
 
@@
 cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
-x^j_z$.\footnote{Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
+x^j_z$.\footnote{\stefano{This is only relevant for $k=1$.} Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
 cases the $p^{i,j}_z$ variables do not attain the lower bound. As a

### 2016-04-13 paper/ijcai16.tex
paper: fix a point.

@@
 \newcommand{\paolo}[1]{{\bf \textcolor{red}{{\fbox{Paolo:} #1}}}}
-\newcommand{\diff}[1]{{\bf \textcolor{violet}{#1}}}
+\newcommand{\diff}[1]{{\bf \textcolor{yellow}{#1}}}
 
@@
 cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
-x^j_z$.\footnote{\stefano{This is only relevant for $k=1$.} Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
+x^j_z$.\footnote{Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
 cases the $p^{i,j}_z$ variables do not attain the lower bound. As a
@@
     \\
-        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \}
-    \\
-        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \}
+        & \;\; \qquad p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, \, w^{i}_z \} \label{eq:volatile1}
+    \\
+        & \;\; \qquad p^{i,j}_z \ge \max \{ 0, \, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} \label{eq:volatile2}
     \\
@@
 while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
-that $\alpha$ can not be taken to be less than $1$, as in this case, the
+that \diff{for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
+so} $\alpha$ can not be taken to be less than $1$, as in this case, the
 objective can be increased arbitrarily while keeping the right-hand side of

### 2016-04-15 paper/ijcai16.tex
paper: add acks for ST.
paper: attempt to fix another point.

@@
 following mixed integer {\em non-linear} optimization problem over the
-variables $\mu \in \bbR_{\ge 0}$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$ \stefano{stress non-zero margin requirement}:
+\diff{non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$}:
 %
@@
 
-%\section*{Acknowledgments}
-%WRITEME
+\section*{Acknowledgments}
+ST was supported by the Caritro Foundation through project E62I15000530007.
 

### 2016-04-15 paper/ijcai16.tex
paper: slight changes to post-rebuttal version

--- 2016-04-13 paper/ijcai16.tex
@@
 following mixed integer {\em non-linear} optimization problem over the
-variables $\mu \in \bbR_{\ge 0}$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$ \stefano{stress non-zero margin requirement}:
+variables $\mu \in \bbR_{\ge 0}$, $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$ \stefano{stress non-zero margin requirement}\andrea{se non c'entra lascerei stare}:
 %
@@
 
-\paragraph{MILP Formulation.} This initial formulation is problematic to solve,
-as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed continuous integer variables. 
-Here we show how to reformulate the
-problem as a mixed integer linear program (MILP). Our goal is to
-replace Eq.~(\ref{eq:wxconstr}) with a set of linear constraints through
-a suitable variable transformation; \stefano{this technique has already been
-used in OR~[Boutilier et al., 2006]}.
-
-In order to do so, we introduce a set of fresh variables $p^{i,j}_z$ for every
-$i,j\in[k]$ and $z\in[m]$. Assuming for the time being that the new variables
-do satisfy the equation $p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth
-component of the objective function in terms of the new variables as:
+\paragraph{MILP Formulation.} This initial formulation is problematic
+to solve, as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over
+mixed continuous integer variables.  \andrea{However, the problem can
+  be reformulated as a mixed integer linear program (MILP) by a
+  suitable variable transformation. This technique is rather common in
+  operational research, see e.g.~[Boutilier et al., 2006].}
+
+\andrea{Our goal is to replace Eq.~(\ref{eq:wxconstr}) with a set of linear
+constraints}. In order to do so, we introduce a set of fresh variables
+$p^{i,j}_z$ for every $i,j\in[k]$ and $z\in[m]$. Assuming for the time
+being that the new variables do satisfy the equation
+$p^{i,j}_z = w^i_z x^j_z$, we rewrite the fourth component of the
+objective function in terms of the new variables as:
 %
@@
 %
-If $x^j_z = 1$ the lower bound simplifies to $\max \{ 0, w^{i}_z \} = w^{i}_z$,
-due to the non-negativity of $w^i_z$. Otherwise, if $x^j_z = 0$
-then the lower bound becomes $\max \{ 0, w^{i}_z - w_\text{max} \}$, where
-the second term is at most $0$. Since $p^{i,j}_z$ is being minimized, in both
-cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
-x^j_z$.\footnote{\stefano{This is only relevant for $k=1$.} Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
-cases the $p^{i,j}_z$ variables do not attain the lower bound. As a
-consequence, the MILP reformulation of Eq.~(\ref{eq:wxconstr}) is a (tight)
-approximation of the original one. This has no impact on the quality of the
-solutions.}
+If $x^j_z = 1$ the lower bound simplifies to
+$\max \{ 0, w^{i}_z \} = w^{i}_z$, due to the non-negativity of
+$w^i_z$. Otherwise, if $x^j_z = 0$ then the lower bound becomes
+$\max \{ 0, w^{i}_z - w_\text{max} \}$, where the second term is at
+most $0$. Since $p^{i,j}_z$ is being minimized, in both cases it will
+attain the lower bound, and thus satisfy
+$p^{i,j}_z = w^i_z x^j_z$.\footnote{Since $\mu$ is upper-bounded by
+  Eq.~(\ref{eq:wyconstr}), in some cases the $p^{i,j}_z$ variables do
+  not attain the lower bound. As a consequence, the MILP reformulation
+  of Eq.~(\ref{eq:wxconstr}) is a (tight) approximation of the
+  original one. This \andrea{can only happen in the case of $k=1$} and
+  has no impact on the quality of the solutions.}
 

### 2016-04-16 paper/ijcai16.tex

--- 2016-04-15 paper/ijcai16.tex
@@
 %
-If $x^j_z = 1$ the lower bound simplifies to $\max \{ 0, w^{i}_z \} = w^{i}_z$,
-due to the non-negativity of $w^i_z$. Otherwise, if $x^j_z = 0$
-then the lower bound becomes $\max \{ 0, w^{i}_z - w_\text{max} \}$, where
-the second term is at most $0$. Since $p^{i,j}_z$ is being minimized, in both
-cases it will attain the lower bound, and thus satisfy $p^{i,j}_z = w^i_z
-x^j_z$.\footnote{Since $\mu$ is upper-bounded by Eq.~(\ref{eq:wyconstr}), in some
-cases the $p^{i,j}_z$ variables do not attain the lower bound. As a
-consequence, the MILP reformulation of Eq.~(\ref{eq:wxconstr}) is a (tight)
-approximation of the original one. This has no impact on the quality of the
-solutions.}
+If $x^j_z = 1$ the lower bound simplifies to
+$\max \{ 0, w^{i}_z \} = w^{i}_z$, due to the non-negativity of
+$w^i_z$. Otherwise, if $x^j_z = 0$ then the lower bound becomes
+$\max \{ 0, w^{i}_z - w_\text{max} \}$, where the second term is at
+most $0$. Since $p^{i,j}_z$ is being minimized, in both cases it will
+attain the lower bound, and thus satisfy
+$p^{i,j}_z = w^i_z x^j_z$.\footnote{Since $\mu$ is upper-bounded by
+  Eq.~(\ref{eq:wyconstr}), in some cases the $p^{i,j}_z$ variables do
+  not attain the lower bound. As a consequence, the MILP reformulation
+  of Eq.~(\ref{eq:wxconstr}) is a (tight) approximation of the
+  original one. This \andrea{can only happen in the case of $k=1$} and
+  has no impact on the quality of the solutions.}
 

### 2016-04-16 paper/ijcai16.tex
paper: use may name in place of the diff macro.

@@
 \newcommand{\paolo}[1]{{\bf \textcolor{red}{{\fbox{Paolo:} #1}}}}
-\newcommand{\diff}[1]{{\bf \textcolor{yellow}{#1}}}
 
@@
 following mixed integer {\em non-linear} optimization problem over the
-\diff{non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$}:
+\stefano{non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$}:
 %
@@
   suitable variable transformation. This technique is rather common in
-  operational research, see e.g.~[Boutilier et al., 2006].}
+  operational research, see e.g.~\cite{boutilier2006}.}
 
@@
 while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
-that \diff{for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
+that \stefano{for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
 so} $\alpha$ can not be taken to be less than $1$, as in this case, the

### 2016-04-16 paper/ijcai16.tex
paper: one point is bogus IMO.

@@
 $\vx^{i}$ is the ``best'' possible option when evaluated according to
-the corresponding $\vw^{i}$ and configurations are maximally diverse
+the corresponding $\vw^{i}$ {\em and} configurations are maximally diverse
 among each other. These options will be later used to formulate
@@
 
-The second goal requires \stefano{reviewer 2, rephrase: to jointly maximize the utility of each
+The second goal requires \stefano{I think this is correct as-is: the second
+reviewer thinks that the second goal is to maximize diversity; above we clearly
+state that it requires to find diverse {\em and} good configurations} to jointly maximize the utility of each
 $\vx^{i}$ according to its corresponding weight vector $\vw^i$ and its
@@
 the set. We achieve this by maximizing the sum of utilities
-$\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ and} adding ranking
+$\sum_{i=1}^k \langle \vw^{i}, \vx^{i} \rangle$ and adding ranking
 constraints of the form

### 2016-04-18 paper/ijcai16.tex
paolos edits

@@
 \usepackage{graphbox}
-\usepackage{hyperref}
+%\usepackage{hyperref}
 \usepackage{multirow}
@@
 
+% TEMPORARY HACK: redefine href as dummy command to avoid Fatal Error!
+\newcommand{\url}[1]{#1}
 % ---------------------------------------------------------------------------
@@
 Paolo Viappiani \\
-Universit\'e Pierre et Marie Curie \\
+Sorbonne Universit\'es\\
+UPMC Univ Paris 06\\
+CNRS, LIP6 UMR 7606\\
 Paris, France \\
@@
 \section{Introduction}
-
-\paolo{or Andrea: check whether~\cite{greco2008ordinal} is related and comment
-on it, as requested by reviewer 3.}
-
 Preferences \cite{Peintner2008} play an important role in a variety of artificial
@@
 
-The automated assessment of preferences has received considerable attention,
-starting with pioneering works \cite{White1984,jaquetlsiskos1982} in the OR community.
-\paolo{or Andrea, check if it is worth mentioning the review in~\cite{jaquetlsiskos2001},
-as highlighted by the fourth reviewer.}
+\paolo{The automated assessment of preferences has received considerable attention,
+starting with pioneering works in the OR community, such as \cite{White1984} and especially the UTA methodology \cite{jaquetlsiskos1982} giving rise to a wide variety of extensions \cite{jaquetlsiskos2001,greco2008ordinal}.}
+
 %In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
 %While preferences are often assessed using some ad-hoc rules \cite{},
-More recently, a number of AI researchers have proposed interactive methods that elicit
+Within AI, a number researchers have proposed interactive methods that elicit
 preferences in an adaptive way
@@
 
-Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
+Consistently with \paolo{the experimental settings} of previous work~\cite{guo2010real,viappiani2010optimal}, 
 we model users with additive utility functions \cite{keeney1976}; 
@@
 occasional inconsistencies in user feedback, are handled by
-introducing slack variables (whose sum we aim at minimizing). When
+introducing slack variables (whose sum we aim at minimizing)
+\paolo{in a way similar to UTA and its extensions \cite{jaquetlsiskos1982,greco2008ordinal} }. When
 augmented with the slacks, the above inequalities take the form
@@
 6.5.0 for solving the core MILP problem. Both the {\sc SetMargin} source code
-and the full experimental setup are available at \url{https://db.tt/oqY62odW}.
+and the full experimental setup are available at 
+\url{https://db.tt/oqY62odW}.
 
@@
 results in terms of number of queries, where we count all
-$k \choose 2$ pairs of queries when comparing $k$ items. In this case,
+$\binom{k}{2}$
+%k \choose 2$ 
+pairs of queries when comparing $k$ items. In this case,
 $k=2$ seems to be the best option. The cognitive cost for the user
@@
 could be beneficial with different noise levels).
-
 % \paragraph{Cognitive effort.} Contrarily to most other preference elicitation
@@
 % standard query selection strategies \stefano{which ones?}. 
-
 % The simplest solution is to query the user about all $k \choose 2$ pairs of
@@
 % quadratic scheme tends to work reasonably well.
-
 %\paolo{Maybe a plot of an experiment with increased noise in user response? (for example, $\beta=0.1$ instead of $1$.}
@@
 attributes do not depend on categorical ones) and studying its
-applicability %of the setwise max-margin idea 
-to other problems.
-
-%\paolo{any idea about possible future works?}
-% FUTURE WORKS
-
-\section*{Acknowledgments}
+applicability to other problems, \paolo{(as the identification of Choquet models \cite{Pine2013}}.
+
+%\section*{Acknowledgments}
+\paragraph{Acknowledgments}
 ST was supported by the Caritro Foundation through project E62I15000530007.
+PV was supported by the Idex Sorbonne Universit\'{e}s under grant ANR-11-IDEX-0004-02.
+We thank Craig Boutilier for motivating discussion on the topic.
 

### 2016-04-19 paper/ijcai16.tex
paper: add note to self.
paper: finalize first change.

@@
 
-\paolo{The automated assessment of preferences has received considerable attention,
-starting with pioneering works in the OR community, such as \cite{White1984} and especially the UTA methodology \cite{jaquetlsiskos1982} giving rise to a wide variety of extensions \cite{jaquetlsiskos2001,greco2008ordinal}.}
-
+The automated assessment of preferences has received considerable attention,
+starting with pioneering works in the OR community, such as \cite{White1984}
+and especially the UTA methodology \cite{jaquetlsiskos1982} giving rise to a
+wide variety of extensions \cite{jaquetlsiskos2001,greco2008ordinal}.
 %In the past decade a number of researchers have considered elicitation of preferences from an artificial intelligence perspective%; within the machine learning field the term preference learning has been catching up.
@@
 \cite{chajewska2000,boutilier2002,Wang2003,boutilier2006,guo2010real,viappiani2010optimal},
-observing that, by asking informative questions, it is often possible to make near-optimal decisions with only partial preference information.
+observing that, by asking informative questions, it is often possible to make
+near-optimal decisions with only partial preference information.
 
@@
 and the full experimental setup are available at 
-\url{https://db.tt/oqY62odW}.
+\url{https://db.tt/oqY62odW} \stefano{update the link!}.
 

### 2016-04-19 paper/ijcai16.tex
paper: more finalization changes.

@@
 
-Consistently with \paolo{the experimental settings} of previous work~\cite{guo2010real,viappiani2010optimal}, 
+Consistently with the experimental settings of previous work~\cite{guo2010real,viappiani2010optimal},
 we model users with additive utility functions \cite{keeney1976}; 
@@
 introducing slack variables (whose sum we aim at minimizing)
-\paolo{in a way similar to UTA and its extensions \cite{jaquetlsiskos1982,greco2008ordinal} }. When
+in a way similar to UTA and its extensions \cite{jaquetlsiskos1982,greco2008ordinal}. When
 augmented with the slacks, the above inequalities take the form
@@
 
-The second goal requires \stefano{I think this is correct as-is: the second
-reviewer thinks that the second goal is to maximize diversity; above we clearly
-state that it requires to find diverse {\em and} good configurations} to jointly maximize the utility of each
+The second goal requires to jointly maximize the utility of each
 $\vx^{i}$ according to its corresponding weight vector $\vw^i$ and its
@@
 following mixed integer {\em non-linear} optimization problem over the
-\stefano{non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$}:
+non-negative margin $\mu \in \bbR_{\ge 0}$ and vectors $\{ \vw^i \in \bbR^m \}$, $\{ \vx^i \in \{0,1\}^m \}$:
 %

### 2016-04-19 paper/ijcai16.tex
paper: finalize as much as I can.

@@
 \paragraph{MILP Formulation.} This initial formulation is problematic
-to solve, as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over
-mixed continuous integer variables.  \andrea{However, the problem can
-  be reformulated as a mixed integer linear program (MILP) by a
-  suitable variable transformation. This technique is rather common in
-  operational research, see e.g.~\cite{boutilier2006}.}
-
-\andrea{Our goal is to replace Eq.~(\ref{eq:wxconstr}) with a set of linear
-constraints}. In order to do so, we introduce a set of fresh variables
+to solve, as Eq.~(\ref{eq:wxconstr}) involves quadratic terms over mixed
+continuous integer variables. However, the problem can be reformulated as a
+mixed integer linear program (MILP) by a suitable transformation. This
+technique is rather common in operational research, see
+e.g.~\cite{boutilier2006}.
+
+Our goal is to replace Eq.~(\ref{eq:wxconstr}) with a set of linear
+constraints. In order to do so, we introduce a set of fresh variables
 $p^{i,j}_z$ for every $i,j\in[k]$ and $z\in[m]$. Assuming for the time
@@
 $p^{i,j}_z = w^i_z x^j_z$.\footnote{Since $\mu$ is upper-bounded by
-  Eq.~(\ref{eq:wyconstr}), in some cases the $p^{i,j}_z$ variables do
-  not attain the lower bound. As a consequence, the MILP reformulation
-  of Eq.~(\ref{eq:wxconstr}) is a (tight) approximation of the
-  original one. This \andrea{can only happen in the case of $k=1$} and
+Eq.~(\ref{eq:wyconstr}), in some cases the $p^{i,j}_z$ variables do not attain
+the lower bound. As a consequence, the MILP reformulation of
+Eq.~(\ref{eq:wxconstr}) is a (tight) approximation of the original one. This
 has no impact on the quality of the solutions.}
@@
 and the full experimental setup are available at 
-\url{https://db.tt/oqY62odW} \stefano{update the link!}.
+\url{https://github.com/stefanoteso/setmargin}.
 
@@
 while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
-that \stefano{for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
-so} $\alpha$ can not be taken to be less than $1$, as in this case, the
+that for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
+so $\alpha$ can not be taken to be less than $1$, as in this case, the
 objective can be increased arbitrarily while keeping the right-hand side of
@@
 attributes do not depend on categorical ones) and studying its
-applicability to other problems, \paolo{(as the identification of Choquet models \cite{Pine2013}}.
+applicability to other problems, \paolo{as the identification of Choquet models \cite{Pine2013}}.
 

### 2016-04-19 paper/ijcai16.tex
paper: squeeze equations a bit.

@@
     \\
-        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
-    \\
-        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
+        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \;\;, \veps^{i} \ge 0 \label{eq:xbounds}
+%    \\
+%        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \label{eq:xbounds}
+%    \\
+%        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
 %    \\
@@
 maximize. Constraints~(\ref{eq:wbounds}) and~(\ref{eq:xbounds}) ensure
-that the weights and configurations are feasible, while the last one
+that the weights and configurations are feasible
+%, while the last one
+and
 guarantees the non-negativity of the slacks.  Since we require
@@
     \\
-        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
-    \\
-        & \;\; \forall \; i \in [k] \quad \veps^{i}_h \ge 0 \nonumber
+        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \;, \veps^{i} \ge 0 \nonumber
+%    \\
+%        & \;\; \forall \; i \in [k] \quad \vx^{i} \in \calX_{\text{feasible}} \nonumber
+%    \\
+%        & \;\; \forall \; i \in [k] \quad \veps^{i} \ge 0 \nonumber
 %    \\

### 2016-04-19 paper/ijcai16.tex
paper: squeeze the text into 6+1 pages.

@@
 %
-$$ p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \} $$
+$ p^{i,i}_z \le \min \{ w_\text{max} x^{i}_z, w^{i}_z \} $,
 %
@@
 %
-$$ p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} $$
+$ p^{i,j}_z \ge \max \{ 0, w^{i}_z - w_\text{max}(1 - x^{j}_z) \} $.
 %
@@
 
-Substituting the above MILP constraints into the original non-linear
-formulation, we obtain the following mixed-integer linear problem:
+% Substituting the above MILP constraints into the original non-linear
+% formulation,
+We thus obtain the following mixed-integer linear problem:
 %
@@
 while $\beta$ and $\gamma$ are taken in $\{10, 1, 0.1, 0.001\}$.\footnote{Note
-that for $k=1$ Eq.~\ref{eq:volatile1} and Eq.~\ref{eq:volatile2} disappear,
+that for $k=1$ Eq.~(\ref{eq:volatile1}) and Eq.~(\ref{eq:volatile2}) disappear,
 so $\alpha$ can not be taken to be less than $1$, as in this case, the
@@
 different distributions. The first two mimic those used
-in~\cite{guo2010real}\footnote{We used the parameters implemented in their code, which are slightly different from those reported in the paper.}%
+in~\cite{guo2010real}
+%\footnote{We used the parameters implemented in their code, which are slightly different from those reported in the paper.}%
 %\footnote{The normal distribution reported

### 2016-04-19 paper/ijcai16.tex
paper: acks in page 7 are fine, keep the last change.

@@
 attributes do not depend on categorical ones) and studying its
-applicability to other problems, \paolo{as the identification of Choquet models \cite{Pine2013}}.
+applicability to other problems, as the identification of Choquet models
+\cite{Pine2013}.
 
